### 2022-03-16 Manuscript analyses.Rmd
Start Rmd with summary tables

---
title: "Manuscript analyses"
author: "Christelle Colin-Leitzinger"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: united
    highlight: pygments
    df_print: paged
editor_options: 
  chunk_output_type: console
---

<style type= "text/css">

.## Figure {
   margin-top: 25px;
   margin-bottom: 100px;
}

## Table {
    margin-top: 10px;
    margin-bottom: 25px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      fig.align='center'
                      )
options(gtsummary.print_engine = "gt")
options(gtsummary.as_gt.addl_cmds = "gt::tab_options(table.font.size = 14, data_row.padding = gt::px(1))")
```

```{r library}
library(tidyverse)
library(ggplot2)
library(gtsummary)
library(labelled)
# library(survival)
# library(survminer)
```

```{r load}
path <- fs::path("","Volumes","Peres_Research", "Myeloma", "Cytopenias")

Characteristics <-
  readxl::read_xlsx(paste0(path,"/CytopeniaManuscriptData_3.16.22.xlsx"),
                    sheet = "Characteristics", skip = 1, n_max = 54) %>% 
  mutate(MRN = as.character(MRN),
         MRN = coalesce(MRN, PT))
Counts <-
  readxl::read_xlsx(paste0(path,"/CytopeniaManuscriptData_3.16.22.xlsx"),
                    sheet = "Counts") %>% 
  mutate(MRN = as.character(MRN),
         MRN = coalesce(MRN, PT))
Infections <-
  readxl::read_xlsx(paste0(path,"/CytopeniaManuscriptData_3.16.22.xlsx"),
                    sheet = "Infections") %>% 
  mutate(MRN = as.character(MRN),
         MRN = coalesce(MRN, PT))
Support <-
  readxl::read_xlsx(paste0(path,"/CytopeniaManuscriptData_3.16.22.xlsx"),
                    sheet = "Support") %>% 
  mutate(MRN = as.character(MRN),
         MRN = coalesce(MRN, PT))
Pneumo_titers <-
  readxl::read_xlsx(paste0(path,"/CytopeniaManuscriptData_3.16.22.xlsx"),
                    sheet = "Pneumo Titers") %>% 
  mutate(MRN = as.character(MRN))
Other_titers <-
  readxl::read_xlsx(paste0(path,"/CytopeniaManuscriptData_3.16.22.xlsx"),
                    sheet = "Other Titers") %>% 
  mutate(MRN = as.character(MRN))
```

```{r join}
cart <- full_join(Characteristics, Counts, by = c("PT", "MRN")) %>% 
  full_join(., Infections, by = c("PT", "MRN")) %>% 
  full_join(., Support, by = c("PT", "MRN", "Date of CAR T infusion")) %>% 
  full_join(., Pneumo_titers, by = c("PT", "MRN")) %>% 
  select(-`Date of leukapheresis.y`, `Date of leukapheresis` = `Date of leukapheresis.x`) %>% 
  full_join(., Other_titers, by = c("PT", "MRN"))
```

```{r data cleaning}
cart1 <- cart %>% 
  # defining BMI categories
  mutate(bmicat = case_when(
    `BMI at apheresis` < 25             ~ "< 25 kg/m2",
    `BMI at apheresis` >= 25 & 
      cart$`BMI at apheresis` < 30      ~ "25 to < 30 kg/m2",
    `BMI at apheresis` >= 30            ~ "≥ 30 kg/m2"
  ),
  bmicat = factor(bmicat, levels = c("< 25 kg/m2", "25 to < 30 kg/m2", "≥ 30 kg/m2"))
  ) %>% 
  mutate(bmicat2 = case_when(
    `BMI at apheresis` < 25             ~ "< 25 kg/m2",
    `BMI at apheresis` >= 25            ~ "≥ 25 kg/m2"
  ),
  bmicat2 = factor(bmicat2, levels = c("< 25 kg/m2", "≥ 25 kg/m2"))
  ) %>% 
  # age categories
  mutate(agecat = case_when(
    Age < 65                            ~ "< 65 years",
    Age >= 65                           ~ "≥ 65 years"
  ),
  agecat = factor(agecat, levels = c("< 65 years", "≥ 65 years"))
  ) %>% 
  mutate_at(c("Extramedullary Disease? (yes=1, no=0)",
              "ECOG > or = 2? (Yes=1, No=0)",
              "Deletion 17p prior to CAR-T infusion (yes=1, no=0)",
              "t(4:14) prior to CAR-T infusion (yes=1, no=0)",
              "t(14:16) prior to CAR-T infusion (yes=1, no=0)",
              "Bridging Therapy (Yes=1, No=0)",
              "Prior auto SCT (yes=1, no=0)",
              "Did patient meet eligibility criteria for KarMMa 1 on the day of apheresis (Yes=1, No=0)"), 
            ~ case_when(
              . == 1                                              ~ "YES",
              . == 0                                              ~ "NO",
              TRUE                                                ~ NA_character_
            )) %>% 
  # tumor burden
  mutate(tumor_burden = case_when(
    `High Marrow Burden? (yes=1, no=0)` == "NO" ~ "Low",
    `High Marrow Burden? (yes=1, no=0)` == "YES" ~ "High"),
    tumor_burden = factor(tumor_burden, levels = c("Low", "High"))
  ) %>% 
  # ECOG
  mutate(`ECOG at LD (at baseline visit)` = case_when(
    `ECOG at LD (at baseline visit)` == 0 | 
      `ECOG at LD (at baseline visit)` ==1 ~ "0-1"
  )) %>% 
  # cytogenetics
  mutate(cyto = case_when(
    `ECOG > or = 2? (Yes=1, No=0)` == "YES" | 
      `t(4:14) prior to CAR-T infusion (yes=1, no=0)` == "YES" | 
      `t(14:16) prior to CAR-T infusion (yes=1, no=0)` == "YES"      ~ "YES",
    `ECOG > or = 2? (Yes=1, No=0)` == "NO" &
      `t(4:14) prior to CAR-T infusion (yes=1, no=0)` == "NO" &
      `t(14:16) prior to CAR-T infusion (yes=1, no=0)` == "NO"       ~ "NO"
  )) %>% 
  # refractory status
  mutate_at(c("Lenalidomide (exposed=0 vs. refractory=1)",
              "Bortezomib (exposed=0 vs. refractory=1)",
              "Carfilzomib (exposed=0 vs. refractory=1)",
              "Pomalidomide (exposed=0 vs. refractory=1)",
              "Daratumumab (exposed=0 vs. refractory=1)"), 
            ~ case_when(
              . == 1                                              ~ "Refractory",
              . == 0                                              ~ "Exposed",
              TRUE                                                ~ NA_character_
            )) %>% 
  mutate(immuno = case_when(
    `Lenalidomide (exposed=0 vs. refractory=1)` == "Exposed" &
      `Pomalidomide (exposed=0 vs. refractory=1)` == "Exposed"        ~ "NO",
    `Lenalidomide (exposed=0 vs. refractory=1)` == "Refractory" |
      `Pomalidomide (exposed=0 vs. refractory=1)` == "Refractory"     ~ "YES"
  )) %>% 
  mutate(proteasome = case_when(
    `Bortezomib (exposed=0 vs. refractory=1)` == "Exposed" &
      `Bortezomib (exposed=0 vs. refractory=1)` == "Exposed"      ~ "NO",
    `Bortezomib (exposed=0 vs. refractory=1)` == "Refractory" |
      `Bortezomib (exposed=0 vs. refractory=1)` == "Refractory"   ~ "YES"
  )) %>% 
  mutate(dara = case_when(
    `Daratumumab (exposed=0 vs. refractory=1)` == "Refractory"      ~ "YES",
    `Daratumumab (exposed=0 vs. refractory=1)` == "Exposed"             ~ "NO"
  )) %>% 
  mutate(doubleref = case_when(
    proteasome == "NO" | 
      immuno == "NO"                                              ~ "NO",
    proteasome == "YES" &
      immuno == "YES"                                         ~ "YES"
  )) %>% 
  mutate(tripleref = case_when(
    proteasome == "NO" | 
      immuno == "NO" |
      `Daratumumab (exposed=0 vs. refractory=1)` == "Exposed"      ~ "NO",
    proteasome == "YES" &
      immuno == "YES" &
      `Daratumumab (exposed=0 vs. refractory=1)` == "Refractory"        ~ "YES"
  )) %>% 
  mutate(pentaref = case_when(
    `Bortezomib (exposed=0 vs. refractory=1)` == "Exposed" | 
      `Carfilzomib (exposed=0 vs. refractory=1)` == "Exposed" |
      `Lenalidomide (exposed=0 vs. refractory=1)` == "Exposed" | 
      `Pomalidomide (exposed=0 vs. refractory=1)` == "Exposed" | 
      `Daratumumab (exposed=0 vs. refractory=1)` == "Exposed"     ~ "NO",
    `Bortezomib (exposed=0 vs. refractory=1)` == "Refractory" & 
      `Carfilzomib (exposed=0 vs. refractory=1)` == "Refractory" &
      `Lenalidomide (exposed=0 vs. refractory=1)` == "Refractory" & 
      `Pomalidomide (exposed=0 vs. refractory=1)` == "Refractory" & 
      `Daratumumab (exposed=0 vs. refractory=1)` == "Refractory"  ~ "YES"
  )) %>% 
# CRS 
  mutate(CRS_any = ifelse(`Max CRS grade (0-5)` == 0, "NO", "YES"),
         CRS_any2 = ifelse(CRS_any== "NO", 0, 1),
         CRS_gradecat = case_when(
           `Max CRS grade (0-5)` == 0                             ~ "No CRS",
           `Max CRS grade (0-5)` == 1                             ~ "Grade 1 or 2",
           `Max CRS grade (0-5)` == 2                             ~ "Grade 1 or 2",
           `Max CRS grade (0-5)` >= 3                             ~ "Grade ≥3"),
         CRS_gradecat = factor(CRS_gradecat, levels = c("No CRS", "Grade 1 or 2", "Grade ≥3"))
  ) %>% 
  # mutate( = case_when(
  #    == 30                ~ NA,
  #   TRUE                          ~ CRS_fup
  # )) %>% 
  # ICANS 
  mutate(ICANS_any = ifelse(`Max ICANS (relative to infusion)` == 0, "NO", "YES"),
         ICANS_any2 <- ifelse(ICANS_any == "NO", 0, 1),
         ICANS_gradecat = case_when(
           `Max ICANS (relative to infusion)` == 0 ~              "No ICANS",
           `Max ICANS (relative to infusion)` == 1 ~              "Grade 1 or 2",
           `Max ICANS (relative to infusion)` == 2 ~              "Grade 1 or 2",
           `Max ICANS (relative to infusion)` >= 3 ~              "Grade ≥3"),
         ICANS_gradecat = factor(ICANS_gradecat, levels = c("No ICANS", "Grade 1 or 2", "Grade ≥3"))
  )

# # CrCl
# table(cart$`Baseline CrCl`)
# cart$crcl[cart$`Baseline CrCl` == ">70" | cart$`Baseline CrCl` == "60.8" | cart$`Baseline CrCl` == "62" | cart$`Baseline CrCl` == "67"] <- ">60"
# cart$crcl[cart$`Baseline CrCl` == "25" | cart$`Baseline CrCl` == "27" | cart$`Baseline CrCl` == "34" | cart$`Baseline CrCl` == "41" | cart$`Baseline CrCl` == "43" | cart$`Baseline CrCl` == "50" | cart$`Baseline CrCl` == "54" | cart$`Baseline CrCl` == "55"] <- "≤60"
# table(cart$crcl)
# # CRS
# table(cart$`Max CRS grade (0-5)`)
# as.factor(cart$`Max CRS grade (0-5)`)
# table(cart$`Duration of CRS (days)`)
# table(cart$`Duration of ICANS (days)`)
# # steroid use
# table(cart$`Steroid use(Yes, No)`)
# 
# 
# 
# # dose reduction
# cart$`Fludarabine Dose Reduction`[cart$`Fludarabine Dose Reduction` == "YES (HD)"] <- "YES"
# table(cart$`Fludarabine Dose Reduction`)
# 
# # RSS
# table(cart$`R-ISS at CAR-T Infusion`)
# cart$`R-ISS at CAR-T Infusion`[cart$`R-ISS at CAR-T Infusion` == "R-ISS  II"] <- "R-ISS II"

# day 30 + MRD for figure
#table(tct$day30_response_revised, tct$MRD_status, useNA= "ifany")
#table(tct$day30_response_revised, tct$day30_response, useNA= "ifany")
#tct$ORR_MRD <- case_when(#tct$day30_response_revised== "SD" ~ "SD",
                         #tct$day30_response_revised== "PD" ~ "PD",
                         #tct$day30_response_revised== "MR" ~ "MR",
#                         tct$day30_response_revised== "PR" ~ "PR",
                         #tct$day30_response_revised== "PR" & tct$MRD_status== "Positive" ~ "PR and MRD-positive",
                         #tct$day30_response_revised== "PR" & is.na(tct$MRD_status) ~ "PR and MRD unknown",
#                         tct$day30_response_revised== "sCR" & tct$MRD_status== "Negative" ~ "sCR and MRD-negative",
#                         tct$day30_response_revised== "sCR" & tct$MRD_status== "Positive" ~ "sCR and MRD-positive")
#table(tct$ORR_MRD)
## Labeling
var_label(cart1) <- list(Age = "Patient Age", agecat = "Patient Age - Categories", bmicat = "BMI at apheresis - Normal, Overweight, Obese", 
                         tumor_burden = "Tumor burden", proteasome = "Refractory to Proteasome Inhibitors", 
                         immuno = "Refractory to Immunomodulators", doubleref = "Double refractory", 
                         tripleref= "Triple refractory", pentaref= "Penta refractory",
                         bmicat2 = "BMI at apheresis - Normal vs. Overweight/Obese", CRS_any = "Any CRS", CRS_gradecat = "CRS Grade", 
                         ICANS_any = "Any ICANS", ICANS_gradecat = "ICANS Grade", cyto= "High risk cytogenetics", 
                         `ECOG > or = 2? (Yes=1, No=0)` = "Deletion 17p at Infusion", 
                         `t(4:14) prior to CAR-T infusion (yes=1, no=0)` = "t(4:14) at Infusion", `t(14:16) prior to CAR-T infusion (yes=1, no=0)` = "t(4:16) at Infusion", 
                         `Prior auto SCT (yes=1, no=0)` = "Prior auto SCT", `Extramedullary Disease? (yes=1, no=0)` = "Extramedullary disease", dara= "Refractory to Daratumumab")

```

<br>

### Tables

## Table 1. Patient baseline characteristics
```{r}
selected <- cart1 %>% 
  select(Age, agecat, Sex, `BMI at apheresis`, bmicat, bmicat2, `Extramedullary Disease? (yes=1, no=0)`, tumor_burden, `ECOG at LD (at baseline visit)`, `R-ISS at CAR-T Infusion`, `ECOG > or = 2? (Yes=1, No=0)`, `t(4:14) prior to CAR-T infusion (yes=1, no=0)`, `t(14:16) prior to CAR-T infusion (yes=1, no=0)`, cyto, `Bridging Therapy (Yes=1, No=0)`, `Number of prior lines of therapy`, `Prior auto SCT (yes=1, no=0)`, immuno, proteasome, dara, doubleref, tripleref, pentaref, CRS_any, CRS_gradecat, `Duration of CRS (days)`, ICANS_any, ICANS_gradecat, `Duration of ICANS (days)`, `Infection (yes, no)`, `Did patient meet eligibility criteria for KarMMa 1 on the day of apheresis (Yes=1, No=0)`, 
                             # `Baseline CRP`, `Max CRP`, `Baseline Ferritin`, `Max Ferritin`, `Cell Dose (million cells)`, 
                             # `LDH at Baseline`, 
                             `LDH day +30`, `LDH day +90`#, 
                             # `B2 Microglobulin at Baseline`, `Day 30 B2 Microglobulin`, `Day 90 B2 Microglobulin`, `Cyclophosphamide Dose (mg)`, `Cyclophosphamide Actual Dose (mg/m2)`, `Fludarabine Total Dose (mg)`, `Fludarabine Actual Dose (mg/m2)`, `Fludarabine Dose Reduction`
                             )
# write.csv(selected,"/Volumes/Peres_Research/Myeloma/CART/selected_data.csv", row.names = TRUE)
selected %>% 
  tbl_summary(
            statistic=list(all_continuous() ~ "{median} ({min}, {max})"), 
            digits = all_continuous() ~ 1) %>% 
  bold_labels() %>% add_stat_label() %>% 
             #modify_header(update = all_stat_cols() ~ "**{level}**, N = {n}") %>%
  # use `modify_footnote(everything() ~ NA, abbreviation = TRUE)` to delete abbrev. footnotes
  modify_footnote(update = everything() ~ NA)
```

<!-- Age (years)	 -->
<!--      Median (range)	 -->
<!--      60 – n (%)	 -->
<!-- Male sex – n (%)	 -->
<!-- Extramedullary disease – n (%)	 -->
<!-- High marrow burden (≥50%) – n (%)	 -->
<!-- ECOG performance status at lymphodepletion – n (%)	 -->
<!--      0	 -->
<!--      1	 -->
<!--      2	 -->
<!--      3	 -->
<!-- R-ISS stage – n (%)	 -->
<!--      I	 -->
<!--      II	 -->
<!--      III	 -->
<!--      Unknown	 -->
<!-- High-risk cytogenetics – n (%)	 -->
<!--      del(17p)	 -->
<!--      t(4;14)	 -->
<!--      t(14;16)	 -->
<!--      Unknown	 -->
<!-- Bridging therapy – n (%)	 -->
<!-- Prior anti-myeloma regimens – median (range)	 -->
<!--      Prior autoSCT – n (%)	 -->
<!--      Prior alloSCT – n (%)	 -->
<!--      Prior BCMA-directed therapy – n (%)	 -->
<!-- Refractory status – n (%)	 -->
<!--      Immunomodulatory agent	 -->
<!--      Proteasome inhibitor	 -->
<!--      Daratumumab	 -->
<!--      Double-refractory disease	 -->
<!--      Triple-refractory disease	 -->
<!--      Penta-refractory disease	 -->
<!-- Eligible for KarMMa 1 trial at apheresis – n (%)	 -->

## Table 2. Clinical outcomes

<!-- CRS maximum grade – n (%)	 -->
<!--      0	 -->
<!--      1	 -->
<!--      2	 -->
<!--      3	 -->
<!--      4	 -->
<!--      5	 -->
<!-- Time to CRS onset (days) – median (range)	 -->
<!-- Time to maximum severity CRS (days) – median (range)	 -->
<!-- Duration of CRS (days) – median (range)	 -->
<!-- ICANS maximum grade – n (%)	 -->
<!--      0	 -->
<!--      1	 -->
<!--      2	 -->
<!--      3	 -->
<!--      4	 -->
<!--      5	 -->
<!-- Time to ICANS onset (days) – median (range)	 -->
<!-- Time to maximum severity ICANS (days) – median (range)	 -->
<!-- Duration of ICANS (days) – median (range)	 -->
<!-- Treatment for toxicity – n (%)	 -->
<!--      Steroid use	 -->
<!--      Tocilizumab use	 -->
<!--      Anakinra use	 -->
<!-- Day 30 response – n (%)	 -->
<!--      ORR	 -->
<!--      CR or sCR	 -->
<!--           MRD negative	 -->
<!--      VGPR	 -->
<!--      PR	 -->
<!-- Day 90 response – n (%)	 -->
<!--      ORR	 -->
<!--      CR or sCR	 -->
<!--           MRD negative	 -->
<!--      VGPR	 -->
<!--      PR	 -->
<!-- Best response in first 90 days – n (%)	 -->
<!--      CR or sCR	 -->
<!--           MRD negative	 -->
<!--      VGPR	 -->
<!--      PR	 -->

For responses make denominator evaluable patients
Patient who died or was lost to follow-up or had non-secretory disease would not be evaluable

## Table 3. Bone marrow biopsy findings

<!-- 	Baseline (n=52)	Day 30 (n=44)	Day 90 (n=32) -->
<!-- Marrow cellularity – n (%)			 -->
<!--      Hypocellular			 -->
<!--      Normocellular			 -->
<!--      Hypercellular			 -->
<!--      Variable			 -->
<!-- Marrow cellularity (%) – median (range)			 -->
<!-- % CD138 positive by IHC – median (range)			 -->
<!-- Fibrosis grade – median (range)			 -->
<!-- Dysplasia present – n (%)			 -->

## Table 4. Supportive therapies and engraftment

<!-- Granulocyte colony stimulating factor (G-CSF)	 -->
<!--      G-CSF administered – n (%)	 -->
<!--      First day of G-CSF – median (range)	 -->
<!--      Last day of G-CSF – median (range)	 -->
<!-- CD34 stem cell boost	 -->
<!--      Stem cell boost administered – n (%)	 -->
<!--      Day of stem cell boost – median (range)	 -->
<!--      Dose of stem cell boost (million cells) – median (range)	 -->
<!-- Engraftment day – median (range)	 -->
<!-- Transfusions post-CART – n (%)	 -->
<!--      RBC transfusion within 7 days	 -->
<!--      Plt transfusion within 7 days	 -->
<!--      RBC transfusion >7 days	 -->
<!--      Plt transfusion >7 days	 -->
<!-- Intravenous immunoglobulin (IVIG)	 -->
<!--      IVIG administered – n (%)	 -->
<!--      First day of IVIG – median (range)	 -->
<!--      Last day of IVIG – median (range)	 -->

## Table S1. Apheresis and product characteristics

<!-- Collection goal (TBV) – median (range)	 -->
<!-- Whole blood processed (mL) – median (range)	 -->
<!-- Total blood volume processed (mL) – median (range)	 -->
<!-- Access type – n (%)	 -->
<!--      Non-tunneled central venous catheter	 -->
<!--      Peripheral line	 -->
<!--      Other	 -->
<!-- Final collection volume (mL) – median (range)	 -->
<!-- Run time (minutes) – median (range)	 -->
<!-- Time from apheresis to product infusion (days) – median (range)	 -->
<!-- Total bags infused – median (range)	 -->
<!-- Total dose volume (mL) – median (range)	 -->
<!-- Total cell dose (million cells) – median (range)	 -->

## Table S2. Cytopenias following ide-cel

<!-- 	Apheresis	Day -5	Day 0	Day 7	Day 14	Day 21	Day 30	Day 90 -->
<!-- Leukopenia – n (%)								 -->
<!--      Any grade								 -->
<!--      Grade 3+								 -->
<!-- Neutropenia – n (%)								 -->
<!--      Any grade								 -->
<!--      Grade 3+								 -->
<!-- Anemia – n (%)								 -->
<!--      Any grade								 -->
<!--      Grade 3+								 -->
<!-- Thrombocytopenia – n (%)								 -->
<!--      Any grade								 -->
<!--      Grade 3+								 -->

## Table S3. Patient characteristics by grade 3+ cytopenia at Day 30

## Table S3. Patient characteristics by grade 3+ cytopenia at Day 90

(Similar to above)

## Table S4. CRS, ICANS, tocilizumab use, steroid use, anakinra use and bridging therapy by grade 3+ cytopenia at Day 30

## Table S4. Details of infections after ide-cel

## Table S5. Patients with infections in the first 30 days by CRS, ICANS, tocilizumab use, steroid use, anakinra use and bridging therapy

## Table S6. Patients with infections in the first 100 days by cell counts and immunoglobulin levels

### Figures

## Figure 1. Consort diagram

## Figure 2. Inflammatory markers after ide-cel

## Figure 3. Cellular and humoral immune reconstitution after ide-cel

## Figure 4. Cytopenias over time

As discussed at meeting, rather than showing count recovery over time plan to show Grade 3+ cytopenias (Hb, Plt, ANC, WBC) over time (Day –5, Day +30, Day +60, Day +90)

## Figure 5. Cumulative incidence of infection and infection density

## Figure S1. Viral titers after ide-cel

A)	HSV 1/2 IgG

Example from prior cytopenia paper below. We can show this data however you think best
Interpretation: ≤0.89=negative (not immune)
                        0.90-1.09=indeterminate (can consider negative for purposes of analysis)
                        ≥1.10=positive (immune)

B)	VZV IgG (positive=immune vs negative or equivocal=not immune)

C)	CMV IgG (positive=immune vs negative=not immune)

## Figure S2. Pneumococcal titers after ide-cel

Antibody concentration >1.0 is considered long-term protection (immune)
If antibody concentration ≤1.0 consider negative (not immune)