### 2022-03-18 Manuscript analyses.Rmd
Add plots, fix Unk, type, etc

---
title: "Manuscript analyses"
author: "Christelle Colin-Leitzinger"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: united
    highlight: pygments
    df_print: paged
editor_options: 
  chunk_output_type: console
---

<style type= "text/css">

.## Figure {
   margin-top: 25px;
   margin-bottom: 100px;
}

## Table {
    margin-top: 10px;
    margin-bottom: 25px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      fig.align='center'
                      )
options(gtsummary.print_engine = "gt")
options(gtsummary.as_gt.addl_cmds = "gt::tab_options(table.font.size = 14, data_row.padding = gt::px(1))")
```

```{r library}
library(tidyverse)
library(ggplot2)
library(gtsummary)
library(labelled)
library(DescTools)
library(patchwork)
# library(survival)
# library(survminer)
```

```{r load}
# path <- fs::path("","Volumes","Peres_Research", "Myeloma", "Cytopenias")

Characteristics <-
  readxl::read_xlsx("./CytopeniaManuscriptData_3.16.22.xlsx",
                    sheet = "Characteristics", skip = 1, n_max = 54) %>% 
  mutate(MRN = as.character(MRN),
         MRN = coalesce(MRN, PT))
Counts <-
  readxl::read_xlsx("./CytopeniaManuscriptData_3.16.22.xlsx",
                    sheet = "Counts") %>% 
  mutate(MRN = as.character(MRN),
         MRN = coalesce(MRN, PT))
Infections <-
  readxl::read_xlsx("./CytopeniaManuscriptData_3.16.22.xlsx",
                    sheet = "Infections") %>% 
  mutate(MRN = as.character(MRN),
         MRN = coalesce(MRN, PT))
Support <-
  readxl::read_xlsx("./CytopeniaManuscriptData_3.16.22.xlsx",
                    sheet = "Support",
                    range = "A1:U53") %>% 
  mutate(MRN = as.character(MRN),
         MRN = coalesce(MRN, PT))
Pneumo_titers <-
  readxl::read_xlsx("./CytopeniaManuscriptData_3.16.22.xlsx",
                    sheet = "Pneumo Titers") %>% 
  mutate(MRN = as.character(MRN))
Other_titers <-
  readxl::read_xlsx("./CytopeniaManuscriptData_3.16.22.xlsx",
                    sheet = "Other Titers") %>% 
  mutate(MRN = as.character(MRN))
```

```{r join}
cart <- full_join(Characteristics, Counts, by = c("PT", "MRN")) %>% 
  full_join(., Infections, by = c("PT", "MRN")) %>% 
  full_join(., Support, by = c("PT", "MRN", "Date of CAR T infusion")) %>% 
  full_join(., Pneumo_titers, by = c("PT", "MRN")) %>% 
  select(-`Date of leukapheresis.y`, `Date of leukapheresis` = `Date of leukapheresis.x`) %>% 
  full_join(., Other_titers, by = c("PT", "MRN"))
```

```{r data cleaning}
cart1 <- cart %>% 
  mutate(across(where(is.character), ~ str_replace(., "NE", NA_character_))) %>%
  # censored days > 100 to 100
  mutate(across(contains("day") & where(is.numeric), ~ case_when(
    . > 100                                                       ~ 100,
    TRUE                                                          ~ as.numeric(.)
  ))) %>% 
  # defining BMI categories
  mutate(bmicat = case_when(
    `BMI at apheresis` < 25             ~ "< 25 kg/m2",
    `BMI at apheresis` >= 25 & 
      cart$`BMI at apheresis` < 30      ~ "25 to < 30 kg/m2",
    `BMI at apheresis` >= 30            ~ "≥ 30 kg/m2"
  ),
  bmicat = factor(bmicat, levels = c("< 25 kg/m2", "25 to < 30 kg/m2", "≥ 30 kg/m2"))
  ) %>% 
  mutate(bmicat2 = case_when(
    `BMI at apheresis` < 25             ~ "< 25 kg/m2",
    `BMI at apheresis` >= 25            ~ "≥ 25 kg/m2"
  ),
  bmicat2 = factor(bmicat2, levels = c("< 25 kg/m2", "≥ 25 kg/m2"))
  ) %>% 
  # age categories
  mutate(agecat = case_when(
    Age < 65                            ~ "< 65 years",
    Age >= 65                           ~ "≥ 65 years"
  ),
  agecat = factor(agecat, levels = c("< 65 years", "≥ 65 years"))
  ) %>% 
  mutate(male_sex = case_when(
    Sex == "M"                                                    ~ 1,
    TRUE                                                          ~ 0
  )) %>% 
  mutate_at(c("Extramedullary Disease? (yes=1, no=0)",
              "High Marrow Burden? (yes=1, no=0)",
              "ECOG > or = 2? (Yes=1, No=0)",
              "Deletion 17p prior to CAR-T infusion (yes=1, no=0)",
              "t(4:14) prior to CAR-T infusion (yes=1, no=0)",
              "t(14:16) prior to CAR-T infusion (yes=1, no=0)",
              "Steroid use (Yes=1, No=0)", "Tocilizumab (Yes=1, No=0)", "Anakinra (Yes=1, No=0)",
              "Bridging Therapy (Yes=1, No=0)",
              "Prior auto SCT (yes=1, no=0)",
              "Prior Allo SCT (Yes=1, No=0)",
              "Prior treatment with any gene therapy-based therapeutic or investigational cellular therapy or BCMA targeted therapy (Yes=1, No=0)",
              "Need for G-CSF post infus (Yes, No)",
              "Need for IVIG (Yes, No)",
              "CD34 stem cell boost",
              "RBC transfusion within 7d post CART", "Plt transfusion within 7d post CART", 
              "RBC transfusion >7d post CART", "Plt transfusion >7d post CART",
              "Did patient meet eligibility criteria for KarMMa 1 on the day of apheresis (Yes=1, No=0)"), 
            ~ case_when(
              . == 1                                              ~ "YES",
              . == 0                                              ~ "NO",
              TRUE                                                ~ NA_character_
            )) %>% 
  # ECOG
  # mutate(`ECOG at LD (at baseline visit)` = case_when(
  #   `ECOG at LD (at baseline visit)` == 0 | 
  #     `ECOG at LD (at baseline visit)` == 1                       ~ "0-1",
  #   TRUE                                                          ~ `ECOG at LD (at baseline visit)`
  # )) %>% 
  # cytogenetics
  mutate(cyto = case_when(
    `Deletion 17p prior to CAR-T infusion (yes=1, no=0)` == "YES" | 
      `t(4:14) prior to CAR-T infusion (yes=1, no=0)` == "YES" | 
      `t(14:16) prior to CAR-T infusion (yes=1, no=0)` == "YES"   ~ "YES",
    `Deletion 17p prior to CAR-T infusion (yes=1, no=0)` == "NO" &
      `t(4:14) prior to CAR-T infusion (yes=1, no=0)` == "NO" &
      `t(14:16) prior to CAR-T infusion (yes=1, no=0)` == "NO"    ~ "NO"
  )) %>% 
  # refractory status
  mutate_at(c("Lenalidomide (exposed=0 vs. refractory=1)",
              "Bortezomib (exposed=0 vs. refractory=1)",
              "Carfilzomib (exposed=0 vs. refractory=1)",
              "Pomalidomide (exposed=0 vs. refractory=1)",
              "Daratumumab (exposed=0 vs. refractory=1)"), 
            ~ case_when(
              . == 1                                              ~ "Refractory",
              . == 0                                              ~ "Exposed",
              TRUE                                                ~ NA_character_
            )) %>% 
  mutate(immuno = case_when(
    `Lenalidomide (exposed=0 vs. refractory=1)` == "Exposed" &
      `Pomalidomide (exposed=0 vs. refractory=1)` == "Exposed"        ~ "NO",
    `Lenalidomide (exposed=0 vs. refractory=1)` == "Refractory" |
      `Pomalidomide (exposed=0 vs. refractory=1)` == "Refractory"     ~ "YES"
  )) %>% 
  mutate(proteasome = case_when(
    `Bortezomib (exposed=0 vs. refractory=1)` == "Exposed" &
      `Carfilzomib (exposed=0 vs. refractory=1)` == "Exposed"      ~ "NO",
    `Bortezomib (exposed=0 vs. refractory=1)` == "Refractory" |
      `Carfilzomib (exposed=0 vs. refractory=1)` == "Refractory"   ~ "YES"
  )) %>% 
  mutate(dara = case_when(
    `Daratumumab (exposed=0 vs. refractory=1)` == "Refractory"      ~ "YES",
    `Daratumumab (exposed=0 vs. refractory=1)` == "Exposed"             ~ "NO"
  )) %>% 
  mutate(doubleref = case_when(
    proteasome == "NO" | 
      immuno == "NO"                                              ~ "NO",
    proteasome == "YES" &
      immuno == "YES"                                         ~ "YES"
  )) %>% 
  mutate(tripleref = case_when(
    proteasome == "NO" | 
      immuno == "NO" |
      `Daratumumab (exposed=0 vs. refractory=1)` == "Exposed"      ~ "NO",
    proteasome == "YES" &
      immuno == "YES" &
      `Daratumumab (exposed=0 vs. refractory=1)` == "Refractory"        ~ "YES"
  )) %>% 
  mutate(pentaref = case_when(
    `Bortezomib (exposed=0 vs. refractory=1)` == "Exposed" | 
      `Carfilzomib (exposed=0 vs. refractory=1)` == "Exposed" |
      `Lenalidomide (exposed=0 vs. refractory=1)` == "Exposed" | 
      `Pomalidomide (exposed=0 vs. refractory=1)` == "Exposed" | 
      `Daratumumab (exposed=0 vs. refractory=1)` == "Exposed"     ~ "NO",
    `Bortezomib (exposed=0 vs. refractory=1)` == "Refractory" & 
      `Carfilzomib (exposed=0 vs. refractory=1)` == "Refractory" &
      `Lenalidomide (exposed=0 vs. refractory=1)` == "Refractory" & 
      `Pomalidomide (exposed=0 vs. refractory=1)` == "Refractory" & 
      `Daratumumab (exposed=0 vs. refractory=1)` == "Refractory"  ~ "YES"
  )) %>% 
# CRS 
  mutate(CRS_any = ifelse(`Max CRS grade (0-5)` == 0, "NO", "YES"),
         CRS_any2 = ifelse(CRS_any== "NO", 0, 1),
         CRS_gradecat = case_when(
           `Max CRS grade (0-5)` == 0                             ~ "No CRS",
           `Max CRS grade (0-5)` == 1                             ~ "Grade 1 or 2",
           `Max CRS grade (0-5)` == 2                             ~ "Grade 1 or 2",
           `Max CRS grade (0-5)` >= 3                             ~ "Grade ≥3"),
         CRS_gradecat = factor(CRS_gradecat, levels = c("No CRS", "Grade 1 or 2", "Grade ≥3"))
  ) %>% 
  # mutate( = case_when(
  #    == 30                ~ NA,
  #   TRUE                          ~ CRS_fup
  # )) %>% 
  # ICANS 
  mutate(ICANS_any = ifelse(`Max ICANS (relative to infusion)` == 0, "NO", "YES"),
         ICANS_any2 <- ifelse(ICANS_any == "NO", 0, 1),
         ICANS_gradecat = case_when(
           `Max ICANS (relative to infusion)` == 0 ~              "No ICANS",
           `Max ICANS (relative to infusion)` == 1 ~              "Grade 1 or 2",
           `Max ICANS (relative to infusion)` == 2 ~              "Grade 1 or 2",
           `Max ICANS (relative to infusion)` >= 3 ~              "Grade ≥3"),
         ICANS_gradecat = factor(ICANS_gradecat, levels = c("No ICANS", "Grade 1 or 2", "Grade ≥3"))
  ) %>% 
  mutate(`marrow cellularity baseline` = case_when(
    `marrow cellularity baseline` == "\"patchy\""                 ~ "hyper",
    TRUE                                                          ~ `marrow cellularity baseline`
  )) %>% 
  mutate(`marrow cellularity day +30` = case_when(
    `marrow cellularity day +30` == "aplastic"                    ~ "hypo",
    TRUE                                                          ~ `marrow cellularity day +30`
  )) %>% 
  mutate(`Access type` = case_when(
    `Access type` == "Femoral"                                    ~ "Non-tunneled",
    `Access type` == "AVF"                                        ~ "Peripheral",
    TRUE                                                          ~ `Access type`
  )) %>% 
  # mutate(best_response = ) %>% 
  mutate(Leukopenia_apheresis = case_when(
    `WBC apheresis` < 4                                           ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`Leukopenia_-5d` = case_when(
    `WBC day -5 (k/ul)` < 4                                       ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`Leukopenia_0d` = case_when(
    `WBC day 0 (k/uL)` < 4                                        ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`Leukopenia_7d` = case_when(
    `WBC day +7` < 4                                              ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`Leukopenia_14d` = case_when(
    `WBC day +14` < 4                                             ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`Leukopenia_21d` = case_when(
    `WBC day +21` < 4                                             ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`Leukopenia_30d` = case_when(
    `WBC day +30` < 4                                             ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`Leukopenia_90d` = case_when(
    `WBC day +90` < 4                                             ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(gr3_Leukopenia_apheresis = case_when(
    `WBC apheresis` < 2                                           ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`gr3_Leukopenia_-5d` = case_when(
    `WBC day -5 (k/ul)` < 2                                       ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`gr3_Leukopenia_0d` = case_when(
    `WBC day 0 (k/uL)` < 2                                        ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`gr3_Leukopenia_7d` = case_when(
    `WBC day +7` < 2                                              ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`gr3_Leukopenia_14d` = case_when(
    `WBC day +14` < 2                                             ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`gr3_Leukopenia_21d` = case_when(
    `WBC day +21` < 2                                             ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`gr3_Leukopenia_30d` = case_when(
    `WBC day +30` < 2                                             ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`gr3_Leukopenia_90d` = case_when(
    `WBC day +90` < 2                                             ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>%

  mutate(Neutropenia_apheresis = case_when(
    `ANC apheresis` < 1.8                                           ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`Neutropenia_-5d` = case_when(
    `ANC day -5 (k/ul)` < 1.8                                        ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`Neutropenia_0d` = case_when(
    `ANC day 0 (k/uL)` < 1.8                                        ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`Neutropenia_7d` = case_when(
    `ANC day +7` < 1.8                                              ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`Neutropenia_14d` = case_when(
    `ANC day +14` < 1.8                                             ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`Neutropenia_21d` = case_when(
    `ANC day +21` < 1.8                                             ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`Neutropenia_30d` = case_when(
    `ANC day +30` < 1.8                                             ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`Neutropenia_90d` = case_when(
    `ANC day +90` < 1.8                                             ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(gr3_Neutropenia_apheresis = case_when(
    `ANC apheresis` < 1                                           ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`gr3_Neutropenia_-5d` = case_when(
    `ANC day -5 (k/ul)` < 1                                        ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`gr3_Neutropenia_0d` = case_when(
    `ANC day 0 (k/uL)` < 1                                        ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`gr3_Neutropenia_7d` = case_when(
    `ANC day +7` < 1                                              ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`gr3_Neutropenia_14d` = case_when(
    `ANC day +14` < 1                                             ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`gr3_Neutropenia_21d` = case_when(
    `ANC day +21` < 1                                             ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`gr3_Neutropenia_30d` = case_when(
    `ANC day +30` < 1                                             ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`gr3_Neutropenia_90d` = case_when(
    `ANC day +90` < 1                                             ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>%

  mutate(Anemia_apheresis = case_when(
    `Hb apheresis` < 11.4                                           ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`Anemia_-5d` = case_when(
    `Hb day -5 (k/ul)` < 11.4                                        ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`Anemia_0d` = case_when(
    `Hb day 0 (k/uL)` < 11.4                                        ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`Anemia_7d` = case_when(
    `Hb day +7` < 11.4                                              ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`Anemia_14d` = case_when(
    `Hb day +14` < 11.4                                             ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`Anemia_21d` = case_when(
    `Hb day +21` < 11.4                                             ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`Anemia_30d` = case_when(
    `Hb day +30` < 11.4                                             ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`Anemia_90d` = case_when(
    `Hb day +90` < 11.4                                             ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(gr3_Anemia_apheresis = case_when(
    `Hb apheresis` < 8                                           ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`gr3_Anemia_-5d` = case_when(
    `Hb day -5 (k/ul)` < 8                                        ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`gr3_Anemia_0d` = case_when(
    `Hb day 0 (k/uL)` < 8                                        ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`gr3_Anemia_7d` = case_when(
    `Hb day +7` < 8                                              ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`gr3_Anemia_14d` = case_when(
    `Hb day +14` < 8                                             ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`gr3_Anemia_21d` = case_when(
    `Hb day +21` < 8                                             ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`gr3_Anemia_30d` = case_when(
    `Hb day +30` < 8                                             ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`gr3_Anemia_90d` = case_when(
    `Hb day +90` < 8                                             ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>%

  mutate(Thrombocytopenia_apheresis = case_when(
    `Plt apheresis` < 143                                           ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`Thrombocytopenia_-5d` = case_when(
    `Plt day -5 (k/ul)` < 143                                        ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`Thrombocytopenia_0d` = case_when(
    `Plt day 0 (k/uL)` < 143                                        ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`Thrombocytopenia_7d` = case_when(
    `Plt day +7` < 143                                              ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`Thrombocytopenia_14d` = case_when(
    `Plt day +14` < 143                                             ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`Thrombocytopenia_21d` = case_when(
    `Plt day +21` < 143                                             ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`Thrombocytopenia_30d` = case_when(
    `Plt day +30` < 143                                             ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`Thrombocytopenia_90d` = case_when(
    `Plt day +90` < 143                                             ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(gr3_Thrombocytopenia_apheresis = case_when(
    `Plt apheresis` < 50                                           ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`gr3_Thrombocytopenia_-5d` = case_when(
    `Plt day -5 (k/ul)` < 50                                        ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`gr3_Thrombocytopenia_0d` = case_when(
    `Plt day 0 (k/uL)` < 50                                        ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`gr3_Thrombocytopenia_7d` = case_when(
    `Plt day +7` < 50                                              ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`gr3_Thrombocytopenia_14d` = case_when(
    `Plt day +14` < 50                                             ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`gr3_Thrombocytopenia_21d` = case_when(
    `Plt day +21` < 50                                             ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`gr3_Thrombocytopenia_30d` = case_when(
    `Plt day +30` < 50                                             ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`gr3_Thrombocytopenia_90d` = case_when(
    `Plt day +90` < 50                                             ~ "YES",
    TRUE                                                          ~ "NO"
  )) %>%

  mutate(`cytopenia_30d` = case_when(
    `WBC day +30` < 4 | 
      `ANC day +30` < 1.8 | 
      `Hb day +30` < 11.4 | 
      `Plt day +30` < 143                                             ~ "YES",
    TRUE                                                              ~ "NO"
  )) %>% 
  mutate(`gr3_cytopenia_30d` = case_when(
    `WBC day +30` < 2 | 
      `ANC day +30` < 1 | 
      `Hb day +30` < 8 | 
      `Plt day +30` < 50                                              ~ "YES",
    TRUE                                                              ~ "NO"
  )) %>% 
  mutate(`gr3_cytopenia_90d` = case_when(
    `WBC day +90` < 2 | 
      `ANC day +90` < 1 | 
      `Hb day +90` < 8 | 
      `Plt day +90` < 50                                             ~ "YES",
    TRUE                                                              ~ "NO"
  ))

# # CrCl
# table(cart$`Baseline CrCl`)
# cart$crcl[cart$`Baseline CrCl` == ">70" | cart$`Baseline CrCl` == "60.8" | cart$`Baseline CrCl` == "62" | cart$`Baseline CrCl` == "67"] <- ">60"
# cart$crcl[cart$`Baseline CrCl` == "25" | cart$`Baseline CrCl` == "27" | cart$`Baseline CrCl` == "34" | cart$`Baseline CrCl` == "41" | cart$`Baseline CrCl` == "43" | cart$`Baseline CrCl` == "50" | cart$`Baseline CrCl` == "54" | cart$`Baseline CrCl` == "55"] <- "≤60"
# table(cart$crcl)
# # CRS
# table(cart$`Max CRS grade (0-5)`)
# as.factor(cart$`Max CRS grade (0-5)`)
# table(cart$`Duration of CRS (days)`)
# table(cart$`Duration of ICANS (days)`)
# # steroid use
# table(cart$`Steroid use(Yes, No)`)
# 
# 
# 
# # dose reduction
# cart$`Fludarabine Dose Reduction`[cart$`Fludarabine Dose Reduction` == "YES (HD)"] <- "YES"
# table(cart$`Fludarabine Dose Reduction`)
# 
# # RSS
# table(cart$`R-ISS at CAR-T Infusion`)
# cart$`R-ISS at CAR-T Infusion`[cart$`R-ISS at CAR-T Infusion` == "R-ISS  II"] <- "R-ISS II"

# day 30 + MRD for figure
#table(tct$day30_response_revised, tct$MRD_status, useNA= "ifany")
#table(tct$day30_response_revised, tct$day30_response, useNA= "ifany")
#tct$ORR_MRD <- case_when(#tct$day30_response_revised== "SD" ~ "SD",
                         #tct$day30_response_revised== "PD" ~ "PD",
                         #tct$day30_response_revised== "MR" ~ "MR",
#                         tct$day30_response_revised== "PR" ~ "PR",
                         #tct$day30_response_revised== "PR" & tct$MRD_status== "Positive" ~ "PR and MRD-positive",
                         #tct$day30_response_revised== "PR" & is.na(tct$MRD_status) ~ "PR and MRD unknown",
#                         tct$day30_response_revised== "sCR" & tct$MRD_status== "Negative" ~ "sCR and MRD-negative",
#                         tct$day30_response_revised== "sCR" & tct$MRD_status== "Positive" ~ "sCR and MRD-positive")
#table(tct$ORR_MRD)
```

```{r}
## Labeling
var_label(cart1) <- list(Age = "Patient Age", agecat = "Patient Age - Categories", 
                         `Extramedullary Disease? (yes=1, no=0)` = "Extramedullary disease", 
                         `High Marrow Burden? (yes=1, no=0)` = "High marrow burden (≥50%)",
                         # bmicat = "BMI at apheresis - Normal, Overweight, Obese", 
                         proteasome = "Refractory to Proteasome Inhibitors", 
                         immuno = "Refractory to Immunomodulators", doubleref = "Double refractory", 
                         tripleref= "Triple refractory", pentaref= "Penta refractory",
                         # bmicat2 = "BMI at apheresis - Normal vs. Overweight/Obese", 
                         CRS_any = "Any CRS", CRS_gradecat = "CRS Grade", 
                         ICANS_any = "Any ICANS", ICANS_gradecat = "ICANS Grade", cyto= "High risk cytogenetics", 
                         # `ECOG > or = 2? (Yes=1, No=0)` ,
                         `Deletion 17p prior to CAR-T infusion (yes=1, no=0)` = "Deletion 17p at Infusion", 
                         `t(4:14) prior to CAR-T infusion (yes=1, no=0)` = "t(4:14) at Infusion", `t(14:16) prior to CAR-T infusion (yes=1, no=0)` = "t(4:16) at Infusion", 
                         `Prior treatment with any gene therapy-based therapeutic or investigational cellular therapy or BCMA targeted therapy (Yes=1, No=0)` = "Prior BCMA-directed therapy",
                         `Prior auto SCT (yes=1, no=0)` = "Prior autoSCT", 
                         `Prior Allo SCT (Yes=1, No=0)` = "Prior alloSCT",
                         # `Prior treatment with any gene therapy-based therapeutic or investigational cellular therapy or BCMA targeted therapy (Yes=1, No=0)` = "Prior BCMA-directed therapy ",
                         dara= "Refractory to Daratumumab"#,
                         # best_response = "Best response in first 90 days"
)

```

<br>

### Tables

## Table 1. Patient baseline characteristics
**Change Male Sex**
```{r}
cart1 %>% 
  distinct(MRN, .keep_all = TRUE) %>% 
  select(Age, agecat, male_sex, Sex,
         `Extramedullary Disease? (yes=1, no=0)`, 
         `High Marrow Burden? (yes=1, no=0)`, `ECOG at LD (at baseline visit)`, 
         `R-ISS at CAR-T Infusion`, 
         cyto, `Deletion 17p prior to CAR-T infusion (yes=1, no=0)`,
         `t(4:14) prior to CAR-T infusion (yes=1, no=0)`,
         `t(14:16) prior to CAR-T infusion (yes=1, no=0)`,
         `Bridging Therapy (Yes=1, No=0)`,
         `Prior auto SCT (yes=1, no=0)`, `Prior Allo SCT (Yes=1, No=0)`,
         `Prior treatment with any gene therapy-based therapeutic or investigational cellular therapy or BCMA targeted therapy (Yes=1, No=0)`,
         immuno, proteasome, dara, doubleref, tripleref, pentaref, 
         `Did patient meet eligibility criteria for KarMMa 1 on the day of apheresis (Yes=1, No=0)`) %>% 
  tbl_summary(
    type = list(
      c(`Extramedullary Disease? (yes=1, no=0)`, `High Marrow Burden? (yes=1, no=0)`,
        cyto, `Deletion 17p prior to CAR-T infusion (yes=1, no=0)`,
        `t(4:14) prior to CAR-T infusion (yes=1, no=0)`,
        `t(14:16) prior to CAR-T infusion (yes=1, no=0)`,
        `Bridging Therapy (Yes=1, No=0)`, `Prior auto SCT (yes=1, no=0)`,
        immuno, proteasome, dara, doubleref, tripleref, pentaref,
        `Did patient meet eligibility criteria for KarMMa 1 on the day of apheresis (Yes=1, No=0)`)
      ~"categorical"),

    statistic = list(all_continuous() ~ "{median} ({min}, {max})"), 
    digits = all_continuous() ~ 1
  ) %>% 
  bold_labels() %>% add_stat_label() %>% 
  modify_footnote(update = everything() ~ NA)
```

## Table 2. Clinical outcomes
**organize response**
**best response**
```{r}
tbl1 <- cart1 %>% 
  distinct(MRN, .keep_all = TRUE) %>% 
  select(`Max CRS grade (0-5)`,
         `Day of onset of CRS relative to infusion`,
         `Day of Max CRS (relative to infusion)`, `Duration of CRS (days)`,

         `Max ICANS (relative to infusion)`,
         `Day of onset of ICANS relative to infusion`,
         `Day of Max ICANS (relative to infusion)`,
         `Duration of ICANS (days)`
  ) %>% 
  tbl_summary(type = list(
    c(`Day of onset of CRS relative to infusion`, `Day of onset of ICANS relative to infusion`,
      `Day of Max ICANS (relative to infusion)`, `Duration of ICANS (days)`) ~ "continuous"
  ),
  statistic=list(all_continuous() ~ "{median} ({min}, {max})"), 
  digits = all_continuous() ~ 1,
  missing = "no") %>% 
  bold_labels() %>% add_stat_label() %>% 
  modify_footnote(update = everything() ~ NA)

tbl2 <- cart1 %>% 
  distinct(MRN, .keep_all = TRUE) %>% 
  select(`Steroid use (Yes=1, No=0)`, `Tocilizumab (Yes=1, No=0)`, 
         `Anakinra (Yes=1, No=0)`,
         `Day 30 Response (sCR, CR, VGPR, PR and MRD+ vs MRD- )`,
         `Day 90 response (sCR, CR, VGPR, PR and MRD+ vs MRD-)`
  ) %>% 
  tbl_summary(type = list(
    c(`Steroid use (Yes=1, No=0)`, `Tocilizumab (Yes=1, No=0)`, `Anakinra (Yes=1, No=0)`) ~ "categorical"
  ),
  statistic=list(all_continuous() ~ "{median} ({min}, {max})"), 
  digits = all_continuous() ~ 1) %>% 
  bold_labels() %>% add_stat_label() %>% 
  modify_footnote(update = everything() ~ NA)

tbl_stack(list(tbl1, tbl2))
```

For responses make denominator evaluable patients
Patient who died or was lost to follow-up or had non-secretory disease would not be evaluable

## Table 3. Bone marrow biopsy findings
**make sure of the variables used then combine them**
```{r}
tbl1 <- cart1 %>% 
  distinct(MRN, .keep_all = TRUE) %>% 
  select("marrow cellularity" = `marrow cellularity baseline`, 
         "marrow cellularity (%)" = `marrow cellularity baseline (%)`,
         "% CD138 positive by IHC" = `marrow CD138 baseline (%)`, 
         "Fibrosis grade" = `marrow fibrosis baseline`, 
         "Dysplasia present" = `marrow dysplasia baseline`
  ) %>% 
  tbl_summary(type = list(
    c(#`marrow cellularity baseline (%)`, `marrow CD138 baseline (%)`, 
      `Fibrosis grade`) ~ "continuous"
  ),
            statistic=list(all_continuous() ~ "{median} ({min}, {max})"), 
            digits = all_continuous() ~ 1) %>% 
  bold_labels() %>% add_stat_label() %>% 
  modify_footnote(update = everything() ~ NA)

tbl2 <- cart1 %>% 
  distinct(MRN, .keep_all = TRUE) %>% 
  select("marrow cellularity" = `marrow cellularity day +30`, 
         "marrow cellularity (%)" = `marrow % cellularity day +30`,
         "% CD138 positive by IHC" = `marrow %CD138 day +30`, 
         "Fibrosis grade" = `marrow fibrosis day +30`, 
         "Dysplasia present" = `marrow dysplasia day +30`
  ) %>% 
  mutate(`marrow cellularity (%)` = as.numeric(`marrow cellularity (%)`)) %>% 
  mutate(`% CD138 positive by IHC` = as.numeric(`% CD138 positive by IHC`)) %>% 
  mutate(`Fibrosis grade` = as.numeric(`Fibrosis grade`)) %>% 
  mutate(`Dysplasia present` = as.numeric(`Dysplasia present`)) %>% 
  tbl_summary(type = list(
    c(`marrow cellularity (%)`, `% CD138 positive by IHC`, 
      `Fibrosis grade`, `Dysplasia present`) ~ "continuous"
  ),
            statistic=list(all_continuous() ~ "{median} ({min}, {max})"), 
            digits = all_continuous() ~ 1) %>% 
  bold_labels() %>% add_stat_label() %>% 
  modify_footnote(update = everything() ~ NA)

tbl3 <- cart1 %>% 
  distinct(MRN, .keep_all = TRUE) %>% 
  select("marrow cellularity" = `marrow cellularity day +90`, 
         "marrow cellularity (%)" = `marrow % cellularity day +90`,
         "% CD138 positive by IHC" = `marrow %CD138 day +90`, 
         "Fibrosis grade" = `marrow fibrosis day +90`, 
         "Dysplasia present" = `marrow dysplasia day +90`
  ) %>% 
  mutate(`marrow cellularity (%)` = as.numeric(`marrow cellularity (%)`)) %>% 
  mutate(`% CD138 positive by IHC` = as.numeric(`% CD138 positive by IHC`)) %>% 
  mutate(`Fibrosis grade` = as.numeric(`Fibrosis grade`)) %>% 
  mutate(`Dysplasia present` = as.numeric(`Dysplasia present`)) %>% 
  tbl_summary(type = list(
    c(`marrow cellularity (%)`, `% CD138 positive by IHC`, 
      `Fibrosis grade`, `Dysplasia present`) ~ "continuous"
  ),
            statistic=list(all_continuous() ~ "{median} ({min}, {max})"), 
            digits = all_continuous() ~ 1) %>% 
  bold_labels() %>% add_stat_label() %>% 
  modify_footnote(update = everything() ~ NA)

tbl_merge(list(tbl1, tbl2, tbl3), tab_spanner = c("Baseline", "Day 30", "Day 90"))
```

## Table 4. Supportive therapies and engraftment
Jenn

```{r}
# cart1 %>% 
#   distinct(MRN, .keep_all = TRUE) %>% 
#   select(`Need for G-CSF post infus (Yes, No)`, `First Day of GCSF`, `Last Day of GCSF`,
#          `CD34 stem cell boost`, `Day of stem cell boost`, `Boost cell dose (million cells)`,
#          `Engraftment Day`,
#          `RBC transfusion within 7d post CART`, `Plt transfusion within 7d post CART`, 
#          `RBC transfusion >7d post CART`, `Plt transfusion >7d post CART`,
#          `Need for IVIG (Yes, No)`, `First day of IVIG`, `Last day of IVIG`
#   ) %>% 
#   tbl_summary(type = list(
#     c(`Day of stem cell boost`, `Boost cell dose (million cells)`,
#     `Last day of IVIG`) ~ "continuous",
#     c(`Need for G-CSF post infus (Yes, No)`,
#       `CD34 stem cell boost`,
#       `RBC transfusion within 7d post CART`, `Plt transfusion within 7d post CART`, 
#          `RBC transfusion >7d post CART`, `Plt transfusion >7d post CART`,
#       `Need for IVIG (Yes, No)`) ~ "categorical"
#   ),
#             statistic=list(all_continuous() ~ "{median} ({min}, {max})"), 
#             digits = all_continuous() ~ 1) %>% 
#   bold_labels() %>% add_stat_label() %>% 
#   modify_footnote(update = everything() ~ NA)
```

## Table S1. Apheresis and product characteristics
```{r}
cart1 %>% 
  distinct(MRN, .keep_all = TRUE) %>% 
  select(`Final collection volume (mL)`, `Whole blood processed (mL)`,
         `TBV processed`, `Access type`, `Final collection volume (mL)`,
         `Run time (min)`,
         `Time from apheresis to CAR-T infusion (days)`, `Total bags`,
         `Total dose volume (mL)`, `Total cell dose (million cells)`

  ) %>% 
  tbl_summary(type = list(c(`Total bags`) ~ "continuous"),
            statistic=list(all_continuous() ~ "{median} ({min}, {max})"), 
            digits = all_continuous() ~ 1) %>% 
  bold_labels() %>% add_stat_label() %>% 
  modify_footnote(update = everything() ~ NA)
```

## Table S2. Cytopenias following ide-cel
**Change %**
```{r}
tbl1 <- cart1 %>% 
  distinct(MRN, .keep_all = TRUE) %>% 
  select("Leukopenia" = `Leukopenia_apheresis`, 
         "Grade3+ Leukopenia" = `gr3_Leukopenia_apheresis`, 
         "Neutropenia" = `Neutropenia_apheresis`, 
         "Grade3+ Neutropenia" = `gr3_Neutropenia_apheresis`, 
         "Anemia" = `Anemia_apheresis`, 
         "Grade3+ Anemia" = `gr3_Anemia_apheresis`, 
         "Thrombocytopenia" = `Thrombocytopenia_apheresis`, 
         "Grade3+ Thrombocytopenia" = `gr3_Thrombocytopenia_apheresis`
  ) %>% 
  tbl_summary(type = list(everything() ~ "dichotomous"),
            digits = all_continuous() ~ 1,
            missing = "no") %>% 
  bold_labels() %>% add_stat_label() %>% 
  modify_footnote(update = everything() ~ NA)

tbl2 <- cart1 %>% 
  distinct(MRN, .keep_all = TRUE) %>% 
  select("Leukopenia" = `Leukopenia_-5d`, 
         "Grade3+ Leukopenia" = `gr3_Leukopenia_-5d`, 
         "Neutropenia" = `Neutropenia_-5d`, 
         "Grade3+ Neutropenia" = `gr3_Neutropenia_-5d`,
         "Anemia" = `Anemia_-5d`, 
         "Grade3+ Anemia" = `gr3_Anemia_-5d`, 
         "Thrombocytopenia" = `Thrombocytopenia_-5d`, 
         "Grade3+ Thrombocytopenia" = `gr3_Thrombocytopenia_-5d`
  ) %>% 
  tbl_summary(type = list(everything() ~ "dichotomous"),
            digits = all_continuous() ~ 1,
            missing = "no") %>% 
  bold_labels() %>% add_stat_label() %>% 
  modify_footnote(update = everything() ~ NA)

tbl3 <- cart1 %>% 
  distinct(MRN, .keep_all = TRUE) %>% 
  select("Leukopenia" = `Leukopenia_0d`, 
         "Grade3+ Leukopenia" = `gr3_Leukopenia_0d`, 
         "Neutropenia" = `Neutropenia_0d`, 
         "Grade3+ Neutropenia" = `gr3_Neutropenia_0d`, 
         "Anemia" = `Anemia_0d`, 
         "Grade3+ Anemia" = `gr3_Anemia_0d`,
         "Thrombocytopenia" = `Thrombocytopenia_0d`, 
         "Grade3+ Thrombocytopenia" = `gr3_Thrombocytopenia_0d`
  ) %>% 
  tbl_summary(type = list(everything() ~ "dichotomous"),
            digits = all_continuous() ~ 1,
            missing = "no") %>% 
  bold_labels() %>% add_stat_label() %>% 
  modify_footnote(update = everything() ~ NA)

tbl4 <- cart1 %>% 
  distinct(MRN, .keep_all = TRUE) %>% 
  select("Leukopenia" = `Leukopenia_7d`, 
         "Grade3+ Leukopenia" = `gr3_Leukopenia_7d`, 
         "Neutropenia" = `Neutropenia_7d`, 
         "Grade3+ Neutropenia" = `gr3_Neutropenia_7d`, 
         "Anemia" = `Anemia_7d`, 
         "Grade3+ Anemia" = `gr3_Anemia_7d`, 
         "Thrombocytopenia" = `Thrombocytopenia_7d`, 
         "Grade3+ Thrombocytopenia" = `gr3_Thrombocytopenia_7d`
  ) %>% 
  tbl_summary(type = list(everything() ~ "dichotomous"),
            digits = all_continuous() ~ 1,
            missing = "no") %>% 
  bold_labels() %>% add_stat_label() %>% 
  modify_footnote(update = everything() ~ NA)

tbl5 <- cart1 %>% 
  distinct(MRN, .keep_all = TRUE) %>% 
  select("Leukopenia" = `Leukopenia_14d`, 
         "Grade3+ Leukopenia" = `gr3_Leukopenia_14d`, 
         "Neutropenia" = `Neutropenia_14d`, 
         "Grade3+ Neutropenia" = `gr3_Neutropenia_14d`,
         "Anemia" = `Anemia_14d`, 
         "Grade3+ Anemia" = `gr3_Anemia_14d`,
         "Thrombocytopenia" = `Thrombocytopenia_14d`, 
         "Grade3+ Thrombocytopenia" = `gr3_Thrombocytopenia_14d` 
  ) %>% 
  tbl_summary(type = list(everything() ~ "dichotomous"),
            digits = all_continuous() ~ 1,
            missing = "no") %>% 
  bold_labels() %>% add_stat_label() %>% 
  modify_footnote(update = everything() ~ NA)

tbl6 <- cart1 %>% 
  distinct(MRN, .keep_all = TRUE) %>% 
  select("Leukopenia" = `Leukopenia_21d`,
         "Grade3+ Leukopenia" = `gr3_Leukopenia_21d`, 
         "Neutropenia" = `Neutropenia_21d`, 
         "Grade3+ Neutropenia" = `gr3_Neutropenia_21d`, 
         "Anemia" = `Anemia_21d`, 
         "Grade3+ Anemia" = `gr3_Anemia_21d`,
         "Thrombocytopenia" = `Thrombocytopenia_21d`, 
         "Grade3+ Thrombocytopenia" = `gr3_Thrombocytopenia_21d`
  ) %>% 
  tbl_summary(type = list(everything() ~ "dichotomous"),
            digits = all_continuous() ~ 1,
            missing = "no") %>% 
  bold_labels() %>% add_stat_label() %>% 
  modify_footnote(update = everything() ~ NA)

tbl7 <- cart1 %>% 
  distinct(MRN, .keep_all = TRUE) %>% 
  select("Leukopenia" = `Leukopenia_30d`, 
         "Grade3+ Leukopenia" = `gr3_Leukopenia_30d`,
         "Neutropenia" = `Neutropenia_30d`, 
         "Grade3+ Neutropenia" = `gr3_Neutropenia_30d`,
         "Anemia" = `Anemia_30d`, 
         "Grade3+ Anemia" = `gr3_Anemia_30d`, 
         "Thrombocytopenia" = `Thrombocytopenia_30d`,
         "Grade3+ Thrombocytopenia" = `gr3_Thrombocytopenia_30d` 
  ) %>% 
  tbl_summary(type = list(everything() ~ "dichotomous"),
            digits = all_continuous() ~ 1,
            missing = "no") %>% 
  bold_labels() %>% add_stat_label() %>% 
  modify_footnote(update = everything() ~ NA)

tbl8 <- cart1 %>% 
  distinct(MRN, .keep_all = TRUE) %>% 
  select("Leukopenia" = `Leukopenia_90d`, 
         "Grade3+ Leukopenia" = `gr3_Leukopenia_90d`,
         "Neutropenia" = `Neutropenia_90d`, 
         "Grade3+ Neutropenia" = `gr3_Neutropenia_90d`,
         "Anemia" = `Anemia_90d`,
         "Grade3+ Anemia" = `gr3_Anemia_90d`,
         "Thrombocytopenia" = `Thrombocytopenia_90d`, 
         "Grade3+ Thrombocytopenia" = `gr3_Thrombocytopenia_90d`
  ) %>% 
  tbl_summary(type = list(everything() ~ "dichotomous"),
            digits = all_continuous() ~ 1,
            missing = "no") %>% 
  bold_labels() %>% add_stat_label() %>% 
  modify_footnote(update = everything() ~ NA)
tbl_merge(list(tbl1, tbl2, tbl3, tbl4, tbl5, tbl6, tbl7, tbl8), 
          tab_spanner = c("Apheresis", "Day -5",	"Day 0",	"Day 7",	"Day 14",	"Day 21",	"Day 30",	"Day 90"))
```

## Table S3. Patient characteristics by grade 3+ cytopenia at Day 30
**add tab_spanning**
```{r}
cart1 %>% 
  distinct(MRN, .keep_all = TRUE) %>% 
  select(Age, agecat, male_sex, Sex,
         `Extramedullary Disease? (yes=1, no=0)`, 
         `High Marrow Burden? (yes=1, no=0)`, `ECOG at LD (at baseline visit)`, `R-ISS at CAR-T Infusion`, cyto, 
         `Bridging Therapy (Yes=1, No=0)`,
         `Prior auto SCT (yes=1, no=0)`, `Prior Allo SCT (Yes=1, No=0)`,
         `Prior treatment with any gene therapy-based therapeutic or investigational cellular therapy or BCMA targeted therapy (Yes=1, No=0)`,
         immuno, proteasome, dara, doubleref, tripleref, pentaref, 
         `Did patient meet eligibility criteria for KarMMa 1 on the day of apheresis (Yes=1, No=0)`,
         `gr3_cytopenia_30d`) %>% 
  tbl_summary(by = `gr3_cytopenia_30d`,
    type = list(
    `Extramedullary Disease? (yes=1, no=0)` ~ "categorical",
    cyto ~ "categorical",
    `Bridging Therapy (Yes=1, No=0)` ~ "categorical",
    `Prior auto SCT (yes=1, no=0)` ~ "categorical",
    immuno ~ "categorical", proteasome ~ "categorical", dara ~ "categorical",
    doubleref ~ "categorical", tripleref ~ "categorical", pentaref ~ "categorical",
    `Did patient meet eligibility criteria for KarMMa 1 on the day of apheresis (Yes=1, No=0)` ~"categorical"),

    statistic = list(all_continuous() ~ "{median} ({min}, {max})"), 
    digits = all_continuous() ~ 1
    ) %>% 
  bold_labels() %>% add_stat_label() %>% 
  modify_footnote(update = everything() ~ NA) %>%
  modify_spanning_header(all_stat_cols() ~ "**Grade 3+ Cytopenia at days 30?**")
```

## Table S3. Patient characteristics by grade 3+ cytopenia at Day 90
**add tab_spanning**
```{r}
cart1 %>% 
  distinct(MRN, .keep_all = TRUE) %>% 
  select(Age, agecat, male_sex, Sex,
         `Extramedullary Disease? (yes=1, no=0)`, 
         `High Marrow Burden? (yes=1, no=0)`, `ECOG at LD (at baseline visit)`, `R-ISS at CAR-T Infusion`, cyto, 
         `Bridging Therapy (Yes=1, No=0)`,
         `Prior auto SCT (yes=1, no=0)`, `Prior Allo SCT (Yes=1, No=0)`,
         `Prior treatment with any gene therapy-based therapeutic or investigational cellular therapy or BCMA targeted therapy (Yes=1, No=0)`,
         immuno, proteasome, dara, doubleref, tripleref, pentaref, 
         `Did patient meet eligibility criteria for KarMMa 1 on the day of apheresis (Yes=1, No=0)`,
         `gr3_cytopenia_90d`) %>% 
  tbl_summary(by = `gr3_cytopenia_90d`,
    type = list(
    `Extramedullary Disease? (yes=1, no=0)` ~ "categorical",
    cyto ~ "categorical",
    `Bridging Therapy (Yes=1, No=0)` ~ "categorical",
    `Prior auto SCT (yes=1, no=0)` ~ "categorical",
    immuno ~ "categorical", proteasome ~ "categorical", dara ~ "categorical",
    doubleref ~ "categorical", tripleref ~ "categorical", pentaref ~ "categorical",
    `Did patient meet eligibility criteria for KarMMa 1 on the day of apheresis (Yes=1, No=0)` ~"categorical"),

    statistic = list(all_continuous() ~ "{median} ({min}, {max})"), 
    digits = all_continuous() ~ 1
    ) %>% 
  bold_labels() %>% add_stat_label() %>% 
  modify_footnote(update = everything() ~ NA) %>%
  modify_spanning_header(all_stat_cols() ~ "**Grade 3+ Cytopenia at days 90?**")
```

## Table S4. CRS, ICANS, tocilizumab use, steroid use, anakinra use and bridging therapy by grade 3+ cytopenia at Day 30
```{r}
tbl1 <- cart1 %>% 
  distinct(MRN, .keep_all = TRUE) %>% 
  select(`Max CRS grade (0-5)`,
         `Day of onset of CRS relative to infusion`,
         `Day of Max CRS (relative to infusion)`, `Duration of CRS (days)`,

         `Max ICANS (relative to infusion)`,
         `Day of onset of ICANS relative to infusion`,
         `Day of Max ICANS (relative to infusion)`,
         `Duration of ICANS (days)`,
         gr3_cytopenia_30d
  ) %>% 
  tbl_summary(by =`gr3_cytopenia_30d`,
              type = list(
                c(`Day of onset of CRS relative to infusion`, `Day of onset of ICANS relative to infusion`,
                  `Day of Max ICANS (relative to infusion)`, `Duration of ICANS (days)`) ~ "continuous"
              ),
              statistic=list(all_continuous() ~ "{median} ({min}, {max})"), 
              digits = all_continuous() ~ 1,
              missing = "no") %>% 
  bold_labels() %>% add_stat_label() %>% 
  modify_footnote(update = everything() ~ NA) %>%
  modify_spanning_header(all_stat_cols() ~ "**Grade 3+ Cytopenia at days 30?**")

tbl2 <- cart1 %>% 
  distinct(MRN, .keep_all = TRUE) %>% 
  select(`Steroid use (Yes=1, No=0)`, `Tocilizumab (Yes=1, No=0)`, 
         `Anakinra (Yes=1, No=0)`,
         `Day 30 Response (sCR, CR, VGPR, PR and MRD+ vs MRD- )`,
         `Day 90 response (sCR, CR, VGPR, PR and MRD+ vs MRD-)`,
         gr3_cytopenia_30d
  ) %>% 
  tbl_summary(by =`gr3_cytopenia_30d`,
              type = list(
                c(`Steroid use (Yes=1, No=0)`, `Tocilizumab (Yes=1, No=0)`, `Anakinra (Yes=1, No=0)`) ~ "categorical"
              ),
              statistic=list(all_continuous() ~ "{median} ({min}, {max})"), 
              digits = all_continuous() ~ 1) %>% 
  bold_labels() %>% add_stat_label() %>% 
  modify_footnote(update = everything() ~ NA)

tbl_stack(list(tbl1, tbl2))
```

## Table S4. Details of infections after ide-cel
Jenn

## Table S5. Patients with infections in the first 30 days by CRS, ICANS, tocilizumab use, steroid use, anakinra use and bridging therapy

## Table S6. Patients with infections in the first 100 days by cell counts and immunoglobulin levels

### Figures

## Figure 1. Consort diagram

## Figure 2. Inflammatory markers after ide-cel
```{r}
# LDH <- cyto %>% select(`LDH at Pre-Leukapheresis.y`, `LDH day 0`, `LDH day +7`, `LDH day +14`, `LDH day +21`, `LDH day +30`)
# LDH_long <- gather(LDH, day, LDH_value, `LDH at Pre-Leukapheresis.y`:`LDH day +30`)
# LDH_long$day[LDH_long$day=="LDH at Pre-Leukapheresis.y"] <- "Pre"
# LDH_long$day[LDH_long$day=="LDH day 0"] <- "Day 0"
# LDH_long$day[LDH_long$day=="LDH day +7"] <- "Day 7"
# LDH_long$day[LDH_long$day=="LDH day +14"] <- "Day 14"
# LDH_long$day[LDH_long$day=="LDH day +21"] <- "Day 21"
# LDH_long$day[LDH_long$day=="LDH day +30"] <- "Day 30"
# LDH_long$day <- factor(LDH_long$day, levels = c("Pre", "Day 0", "Day 7", "Day 14", "Day 21", "Day 30"))
# ggplot(LDH_long, aes(x=day, y=LDH_value)) + geom_boxplot() + xlab("Days") + ylab("LDH") + theme_classic() + ggtitle("LDH") + expand_limits(x= c(0.5, 6.5), y = c(0, 7000)) + scale_y_break(c(1250, 3500), ticklabels=c(4000, 7000), scales=0.10) 
# 
# cyto$`Ferritin day +7`[cyto$`Ferritin day +7`==">33511"] <- 33511
# cyto$`Ferritin day +7` <- as.numeric(cyto$`Ferritin day +7`)
# Ferritin <- cyto %>% select(MRN, `Ferritin at Pre-Leukapheresis`, `Ferritin day 0`, `Ferritin day +7`, `Ferritin day +14`, `Ferritin day +21`, `Ferritin day +30`)
# Ferritin_long <- gather(Ferritin, day, Ferritin_value, `Ferritin at Pre-Leukapheresis`:`Ferritin day +30`)
# Ferritin_long$day[Ferritin_long$day=="Ferritin at Pre-Leukapheresis"] <- "Pre"
# Ferritin_long$day[Ferritin_long$day=="Ferritin day 0"] <- "Day 0"
# Ferritin_long$day[Ferritin_long$day=="Ferritin day +7"] <- "Day 7"
# Ferritin_long$day[Ferritin_long$day=="Ferritin day +14"] <- "Day 14"
# Ferritin_long$day[Ferritin_long$day=="Ferritin day +21"] <- "Day 21"
# Ferritin_long$day[Ferritin_long$day=="Ferritin day +30"] <- "Day 30"
# Ferritin_long$day <- factor(Ferritin_long$day, levels = c("Pre", "Day 0", "Day 7", "Day 14", "Day 21", "Day 30"))
# ggplot(Ferritin_long, aes(x=day, y=Ferritin_value)) + geom_boxplot() + xlab("Days") + ylab("Ferritin") + theme_classic() + ggtitle("Ferritin") + expand_limits(x= c(0.5, 6.5), y = c(0, 34000)) + scale_y_break(c(11000, 32900), ticklabels=c(33000, 34000), scales=0.05) 
# 
# cyto$`CRP Pre-Leukapheresis`[cyto$`CRP Pre-Leukapheresis`=="<0.03"] <- 0.07
# cyto$`CRP Pre-Leukapheresis` <- as.numeric(cyto$`CRP Pre-Leukapheresis`)
# cyto$`CRP day +14`[cyto$`CRP day +14`=="<0.03"] <- 0.07
# cyto$`CRP day +14` <- as.numeric(cyto$`CRP day +14`)
# cyto$`CRP day +21`[cyto$`CRP day +21`=="<0.03"] <- 0.07
# cyto$`CRP day +21` <- as.numeric(cyto$`CRP day +21`)
# cyto$`CRP day +30`[cyto$`CRP day +30`=="<0.03"] <- 0.07
# cyto$`CRP day +30` <- as.numeric(cyto$`CRP day +30`)
# CRP <- cyto %>% select(`CRP Pre-Leukapheresis`, `CRP day 0`, `CRP day +7`, `CRP day +14`, `CRP day +21`, `CRP day +30`)
# CRP_long <- gather(CRP, day, CRP_value, `CRP Pre-Leukapheresis`:`CRP day +30`)
# CRP_long$day[CRP_long$day=="CRP Pre-Leukapheresis"] <- "Pre"
# CRP_long$day[CRP_long$day=="CRP day 0"] <- "Day 0"
# CRP_long$day[CRP_long$day=="CRP day +7"] <- "Day 7"
# CRP_long$day[CRP_long$day=="CRP day +14"] <- "Day 14"
# CRP_long$day[CRP_long$day=="CRP day +21"] <- "Day 21"
# CRP_long$day[CRP_long$day=="CRP day +30"] <- "Day 30"
# CRP_long$day <- factor(CRP_long$day, levels = c("Pre", "Day 0", "Day 7", "Day 14", "Day 21", "Day 30"))
# ggplot(CRP_long, aes(x=day, y=CRP_value)) + geom_boxplot() + xlab("Days") + ylab("CRP") + theme_classic() + ggtitle("CRP") + expand_limits(x= c(0.5, 6.5), y = c(0, 20)) + scale_y_break(c(5, 6), ticklabels=c(7, 10, 15, 20), scales=0.20) 
```

## Figure 3. Cellular and humoral immune reconstitution after ide-cel
```{r, fig.width=12, fig.height=12}
cart2 <- cart1 %>% 
  distinct(MRN, .keep_all = TRUE)
# WBC
timepoints <- c(-28, -5, 0, 7, 14, 21, 30, 60, 90)
WBC_fig <-
  data.frame(timepoints,
             t(round(
               sapply(cart2[, c(
                 "WBC apheresis",
                 "WBC day -5 (k/ul)",
                 "WBC day 0 (k/uL)",
                 "WBC day +7",
                 "WBC day +14",
                 "WBC day +21",
                 "WBC day +30",
                 "WBC day +60",
                 "WBC day +90"
               )],
               MedianCI,
               conf.level = 0.95,
               na.rm = TRUE), 9
             )))
WBC_fig <- WBC_fig %>%
  ggplot(aes(x = timepoints, y = median)) + 
  geom_line(col = "blue") + 
  xlab("Days") + 
  ylab("Median WBC") + 
  theme_classic() + 
  geom_ribbon(aes(ymin = lwr.ci, ymax = upr.ci), alpha = 0.1) + 
  scale_x_continuous(
    breaks = c(-28, -5, 0, 7, 14, 21, 30, 60, 90),
    labels = c("Apheresis","D-5", "D0", "D7", "D14", "D21", "D30", "D60", "D90")
  ) + ylim(0, 8)

# ANC
ANC_fig <-
  data.frame(timepoints, t(round(
    sapply(cart2[, c(
      "ANC apheresis",
      "ANC day -5 (k/ul)",
      "ANC day 0 (k/uL)",
      "ANC day +7",
      "ANC day +14",
      "ANC day +21",
      "ANC day +30",
      "ANC day +60",
      "ANC day +90"
    )], MedianCI, conf.level = 0.95, na.rm = TRUE), 9
  )))
ANC_fig <- ANC_fig %>% 
  ggplot(aes(x = timepoints, y = median)) + 
  geom_line(col = "blue") + 
  xlab("Days") + 
  ylab("Median ANC") + 
  theme_classic() + 
  geom_ribbon(aes(ymin = lwr.ci, ymax = upr.ci), alpha = 0.1) + 
  scale_x_continuous(
    breaks = c(-28, -5, 0, 7, 14, 21, 30, 60, 90),
    labels = c("Apheresis","D-5", "D0", "D7", "D14", "D21", "D30", "D60", "D90")
  ) + 
  ylim(0, 5)

Hb_fig <-
  data.frame(timepoints, t(round(
    sapply(cart2[, c(
      "Hb apheresis",
      "Hb day -5 (k/ul)",
      "Hb day 0 (k/uL)",
      "Hb day +7",
      "Hb day +14",
      "Hb day +21",
      "Hb day +30",
      "Hb day +60",
      "Hb day +90"
    )], MedianCI, conf.level = 0.95, na.rm = TRUE), 9
  )))
Hb_fig <- Hb_fig %>% 
  ggplot(aes(x = timepoints, y = median)) + 
  geom_line(col = "blue") + 
  xlab("Days") + 
  ylab("Median Hb") + 
  theme_classic() + 
  geom_ribbon(aes(ymin = lwr.ci, ymax = upr.ci), alpha = 0.1) + 
  scale_x_continuous(
    breaks = c(-28, -5, 0, 7, 14, 21, 30, 60, 90),
    labels = c("Apheresis","D-5", "D0", "D7", "D14", "D21", "D30", "D60", "D90")
  ) + 
  ylim(5, 15)

# ALC
ALC_fig <-
  data.frame(timepoints, t(round(
    sapply(cart2[, c(
      "ALC apheresis",
      "ALC day -5 (k/ul)",
      "ALC day 0 (k/uL)",
      "ALC day +7",
      "ALC day +14",
      "ALC day +21",
      "ALC day +30",
      "ALC day +60",
      "ALC day +90"
    )], MedianCI, conf.level = 0.95, na.rm = TRUE), 9
  )))
ALC_fig <- ALC_fig %>% 
  ggplot(aes(x = timepoints, y = median)) + 
  geom_line(col = "blue") + 
  xlab("Days") + 
  ylab("Median ALC") + 
  theme_classic() + 
  geom_ribbon(aes(ymin = lwr.ci, ymax = upr.ci), alpha = 0.1) + 
  scale_x_continuous(
    breaks = c(-28, -5, 0, 7, 14, 21, 30, 60, 90),
    labels = c("Apheresis","D-5", "D0", "D7", "D14", "D21", "D30", "D60", "D90")
  ) + 
  ylim(0, 1.50)

# Plt
Plt_fig <-
  data.frame(timepoints, t(round(
    sapply(cart2[, c(
      "Plt apheresis",
      "Plt day -5 (k/ul)",
      "Plt day 0 (k/uL)",
      "Plt day +7",
      "Plt day +14",
      "Plt day +21",
      "Plt day +30",
      "Plt day +60",
      "Plt day +90"
    )], MedianCI, conf.level = 0.95, na.rm = TRUE), 9
  )))
Plt_fig <- Plt_fig %>% 
  ggplot(aes(x = timepoints, y = median)) + 
  geom_line(col = "blue") + 
  xlab("Days") + 
  ylab("Median Plt") + 
  theme_classic() + 
  geom_ribbon(aes(ymin = lwr.ci, ymax = upr.ci), alpha = 0.1) + 
  scale_x_continuous(
    breaks = c(-28, -5, 0, 7, 14, 21, 30, 60, 90),
    labels = c("Apheresis","D-5", "D0", "D7", "D14", "D21", "D30", "D60", "D90")
  ) +
  ylim(0, 200)

# IgG
#table(cart2$`IgG day +90`)
#table(cart2$`IgG day +180`)
cart2$`IgG day +90`[cart2$`IgG day +90` == "<109"] <- 77
cart2$`IgG day +30`[cart2$`IgG day +30` == "<109"] <- 77
cart2$`IgG day -5`[cart2$`IgG day -5` == "<109"] <- 77
cart2$`IgG apheresis`[cart2$`IgG apheresis` == "<109"] <- 77

cart2$`IgG day +90` <- as.numeric(cart2$`IgG day +90`)
cart2$`IgG day +30` <- as.numeric(cart2$`IgG day +30`)
cart2$`IgG day -5` <- as.numeric(cart2$`IgG day -5`)
cart2$`IgG apheresis` <- as.numeric(cart2$`IgG apheresis`)
timepoints2 <- c(-28, -5, 30, 90)
IgG_fig <-
  data.frame(timepoints2, t(round(
    sapply(cart2[, c(
      "IgG apheresis", 
      "IgG day -5",
      # "IgG day 0 (k/uL)", 
      "IgG day +30", 
      "IgG day +90")], MedianCI, conf.level =
             0.95, na.rm = TRUE), 9
  )))
IgG_fig <- IgG_fig %>% 
  ggplot(aes(x = timepoints2, y = median)) + 
  geom_line(col = "blue") + xlab("Days") + 
  ylab("Median IgG") + 
  theme_classic() + 
  geom_ribbon(aes(ymin = lwr.ci, ymax = upr.ci), alpha = 0.1) + 
  scale_x_continuous(
    breaks = c(-28, -5, 0, 30, 90),
    labels = c("Apheresis", "D-5", "D0", "D30", "D90")) + 
  ylim(0, 2000)

(WBC_fig | ANC_fig) /
  (Hb_fig | ALC_fig) /
  (Plt_fig | IgG_fig) +
  plot_annotation(tag_levels = "A")
```

<span style="color:blue">**IgG I need to fix the other 109**</span>

Values less than the limit of detection (<109) were considered to be 77 which is the lower limit of detection divided by the square root of 2. One patient had an IgG day 0 value of <320. As this is above the limit of detection, I used the midway point between 320 and the lower limit of detection -- 214 -- as the value instead of <320.

<br>

## Figure 4. Cytopenias over time

As discussed at meeting, rather than showing count recovery over time plan to show Grade 3+ cytopenias (Hb, Plt, ANC, WBC) over time (Day –5, Day +30, Day +60, Day +90)

## Figure 5. Cumulative incidence of infection and infection density

## Figure S1. Viral titers after ide-cel

A)	HSV 1/2 IgG

Example from prior cytopenia paper below. We can show this data however you think best
Interpretation: ≤0.89=negative (not immune)
                        0.90-1.09=indeterminate (can consider negative for purposes of analysis)
                        ≥1.10=positive (immune)

B)	VZV IgG (positive=immune vs negative or equivocal=not immune)

C)	CMV IgG (positive=immune vs negative=not immune)

## Figure S2. Pneumococcal titers after ide-cel

Antibody concentration >1.0 is considered long-term protection (immune)
If antibody concentration ≤1.0 consider negative (not immune)