### 2022-05-05 RWE/RWE Manuscript analyses.Rmd
Update exclusion, tables and plots

---
title: "Real World Manuscript Analyses"
author: "Christelle Colin-Leitzinger"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: united
    highlight: pygments
    df_print: paged
editor_options: 
  chunk_output_type: console
---

<style type= "text/css">

. figure {
   margin-top: 25px;
   margin-bottom: 100px;
}

table {
    margin-top: 10px;
    margin-bottom: 25px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      # fig.width = 7, fig.height = 5, 
                      fig.align='center'#, fig.show='hold'
                      )
options(gtsummary.print_engine = "gt")
options(gtsummary.as_gt.addl_cmds = "gt::tab_options(table.font.size = 14, data_row.padding = gt::px(1))")
```

```{r library}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(gtsummary)
library(labelled)
library(DescTools)
library(patchwork)
library(ggbreak)
library(survival)
library(survminer)
```

```{r load}
# path <- fs::path("","Volumes","Peres_Research", "Myeloma", "Cytopenias")
# path <- getwd()
MM_cart <-
  readxl::read_xlsx(
    "/Users/colinccm/Documents/GitHub/Peres/MM_CARTcell_cytopenia/RWE/FINAL 4.13.22 ASCO 2022 Myeloma CAR-T SOC Outcomes Spreadsheet.xlsx",
                    skip = 1, n_max = 198
                    ) %>% 
  rename(patient_id = `PT Identifier (DO NOT USE PHI)`)
```

```{r data cleaning}
cart0 <- MM_cart %>% 
  mutate(across(c(`ANC Day 30`,
                  `ANC Day 90`,
                  `HgB Day 90`,
                  `Platelets Day 90`), ~ as.numeric(.))) %>% 
  mutate(across(c(`Need for GCSF`), ~ str_to_upper(.))) %>% 
  # Keep second infusion date for patients who had failure
  separate(`Date of Apheresis (or # of days from apheresis to LD)`, 
           into= c("first_date_of_apheresis", "date_of_apheresis"), sep = ", 2nd-|; 2nd - ", remove = FALSE) %>% 
  mutate(first_date_of_apheresis = str_remove(first_date_of_apheresis, "1st - |1st-"),
         first_date_of_apheresis = as.Date(as.numeric(first_date_of_apheresis),
                                             origin = "1899-12-30"),
         date_of_apheresis = parse_date_time(date_of_apheresis, "mdy"),
         date_of_apheresis = coalesce(date_of_apheresis, first_date_of_apheresis)
         ) %>% 
  select(-c(`Date of Apheresis (or # of days from apheresis to LD)`, first_date_of_apheresis)) %>%

  mutate(across(c(`Day of Max CRS (relative to infusion)`,
                  `Day of Max ICANS (relative to infusion)`), ~ str_remove(., "Day +"))
         ) %>%

###########################################################################
###########################################################################
###########################################################################
###########################################################################
###########################################################################
###########################################################################

# censored days > 100 to 100
  mutate(across(c(`Day of Max CRS (relative to infusion)`,
                  `Day of Max ICANS (relative to infusion)`), ~ case_when(
    . > 100                                                       ~ 100,
    TRUE                                                          ~ as.numeric(.)
  ))) %>% 
  mutate(`Length of Hospital Stay (Total Days including Readmission)` =
           as.numeric(`Length of Hospital Stay (Total Days including Readmission)`)) %>% 
  # defining BMI categories
  # mutate(bmicat = case_when(
  #   `BMI at apheresis` < 25             ~ "< 25 kg/m2",
  #   `BMI at apheresis` >= 25 & 
  #     cart$`BMI at apheresis` < 30      ~ "25 to < 30 kg/m2",
  #   `BMI at apheresis` >= 30            ~ "≥ 30 kg/m2"
  # ),
  # bmicat = factor(bmicat, levels = c("< 25 kg/m2", "25 to < 30 kg/m2", "≥ 30 kg/m2"))
  # ) %>% 
  # mutate(bmicat2 = case_when(
  #   `BMI at apheresis` < 25             ~ "< 25 kg/m2",
  #   `BMI at apheresis` >= 25            ~ "≥ 25 kg/m2"
  # ),
  # bmicat2 = factor(bmicat2, levels = c("< 25 kg/m2", "≥ 25 kg/m2"))
  # ) %>% 
  # age categories
  mutate(agecat = case_when(
    Age < 65                            ~ "< 65 years",
    Age >= 65                           ~ "≥ 65 years"
  ),
  agecat = factor(agecat, levels = c("< 65 years", "≥ 65 years"))
  ) %>% 
  mutate(Sex = case_when(
    str_detect(Sex, "F")                                          ~ "F",
    str_detect(Sex, "M")                                          ~ "M"
  )) %>% 
  mutate(male_sex = case_when(
    Sex == "M"                                                    ~ 1,
    TRUE                                                          ~ 0
  )) %>% 
  # ECOG
  mutate(`ECOG at LD` = case_when(
    `ECOG at LD` == 0 |
      `ECOG at LD` == 1                                           ~ "0-1",
    `ECOG at LD` %in% c(2:4)                                      ~ "2-4",
    TRUE                                                          ~ NA_character_
  )) %>% 
  # cytogenetics
  mutate(cytogenetics = case_when(
    `Deletion 17p prior to CAR-T infusion (yes, no)` == "YES" | 
      `t(4:14) prior to CAR-T infusion (yes, no)` == "YES" | 
      `t(14:16) prior to CAR-T infusion (yes, no)` == "YES"   ~ "YES",
    `Deletion 17p prior to CAR-T infusion (yes, no)` == "NO" &
      `t(4:14) prior to CAR-T infusion (yes, no)` == "NO" &
      `t(14:16) prior to CAR-T infusion (yes, no)` == "NO"    ~ "NO",
    TRUE                                                      ~ NA_character_
  )) %>% 
  # refractory status
  mutate_at(c("Lenalidomide (exposed vs. refractory)",
              "Bortezomib (exposed vs. refractory)",
              "Carfilzomib (exposed vs. refractory)",
              "Pomalidomide (exposed vs. refractory)",
              "Daratumumab (exposed vs. refractory)"),
            ~ case_when(
              is.na(.)                                            ~ "Not Refractory",
              TRUE                                                ~ as.character(.)
            )) %>%
  mutate(immuno = case_when(
    `Lenalidomide (exposed vs. refractory)` == "Refractory" |
      `Pomalidomide (exposed vs. refractory)` == "Refractory"     ~ "YES",
    TRUE        ~ "NO"
  )) %>% 
  mutate(proteasome = case_when(
    `Bortezomib (exposed vs. refractory)` == "Refractory" |
      `Carfilzomib (exposed vs. refractory)` == "Refractory"   ~ "YES",
    TRUE      ~ "NO"
  )) %>% 
  mutate(dara = case_when(
    `Daratumumab (exposed vs. refractory)` == "Refractory"      ~ "YES",
    TRUE             ~ "NO"
  )) %>% 
  mutate(doubleref = case_when(
    proteasome == "YES" &
      immuno == "YES"                                         ~ "YES",
    TRUE                                              ~ "NO"
  )) %>% 
  mutate(tripleref = case_when(
    proteasome == "YES" &
      immuno == "YES" &
      `Daratumumab (exposed vs. refractory)` == "Refractory"        ~ "YES",
    TRUE      ~ "NO"
  )) %>% 
  mutate(pentaref = case_when(
    `Bortezomib (exposed vs. refractory)` == "Refractory" & 
      `Carfilzomib (exposed vs. refractory)` == "Refractory" &
      `Lenalidomide (exposed vs. refractory)` == "Refractory" & 
      `Pomalidomide (exposed vs. refractory)` == "Refractory" & 
      `Daratumumab (exposed vs. refractory)` == "Refractory"  ~ "YES",
    TRUE     ~ "NO"
  )) %>% 
  mutate(number_prior_lines_therapy = case_when(
    `Number of prior lines of therapy` <= 4           ~ "≤4",
    `Number of prior lines of therapy` > 4            ~ ">4",
  )) %>%
  mutate(`Tocilizumab (Yes, No)` = case_when(
    `Tocilizumab (Yes, No)` == "Siltuximab"                    ~ "YES",
      str_detect(`Tocilizumab (Yes, No)`, "COVID")             ~ "NO",
    TRUE                                                       ~ `Tocilizumab (Yes, No)`
  )) %>% 
  mutate(`Steroid use (Yes, No)` = case_when(
    str_detect(`Steroid use (Yes, No)`, "COVID")               ~ "NO",
    TRUE                                                       ~ `Steroid use (Yes, No)`
  )) %>% 
  mutate(`Disease status at time of referral (Relapsed vs Refractory)` = case_when(
    str_detect(`Disease status at time of referral (Relapsed vs Refractory)`, "Relapse")       ~ "Relapse",
    str_detect(`Disease status at time of referral (Relapsed vs Refractory)`, "Response")      ~ "Response",
    TRUE                                     ~ `Disease status at time of referral (Relapsed vs Refractory)`
  )) %>%

  mutate(exclusion_criteria =  
           rowSums(select(.,`Received investigation agent, plasmapheresis, surgery, radiation, or therapy within 14 days of apheresis (Yes, No)`:`ECOG > or = 2? (Yes, No)`
                          ) == "YES", na.rm = TRUE),
         exclusion_criteria = case_when(
           exclusion_criteria == 0           ~ "0",
           exclusion_criteria == 1           ~ "1",
           exclusion_criteria > 1            ~ "≥2"
         )) %>% 
  mutate(comorbidities = case_when(
    `Renal insufficiency CrCl < 45 mL/min (Yes, No)` == "YES" |
      `H/O Class III or IV CHF or severe nonischemic cardiomyopathy, unstable or poorly controlled angina, myocardial infarction, or ventricular arrhythmia within 6 months (Yes, No)` == "YES" |
      `LVEF <45% (Yes, No)` == "YES" |
      `AST/ALT >2.5 ULN (Yes, No)` == "YES" |
      `Total Bilirubin >1.5ULN (Yes, No)` == "YES" |
      `H/O of malignancies (Yes, No)` == "YES"                    ~ "YES",
    `LVEF <45% (Yes, No)` == "NO" &
      `AST/ALT >2.5 ULN (Yes, No)` == "NO" &
      `Total Bilirubin >1.5ULN (Yes, No)` == "NO" &
      `H/O of malignancies (Yes, No)` == "NO"                     ~ "NO",
    TRUE                                                          ~ NA_character_
  )) %>%

# CRS 
  mutate(CRS_any = ifelse(`Max CRS grade (0-5)` == 0, "NO", "YES"),
         CRS_any2 = ifelse(CRS_any== "NO", 0, 1),
         CRS_gradecat = case_when(
           `Max CRS grade (0-5)` == 0                             ~ "No CRS",
           `Max CRS grade (0-5)` == 1                             ~ "Grade 1 or 2",
           `Max CRS grade (0-5)` == 2                             ~ "Grade 1 or 2",
           `Max CRS grade (0-5)` >= 3                             ~ "Grade ≥3"),
         CRS_gradecat = factor(CRS_gradecat, levels = c("No CRS", "Grade 1 or 2", "Grade ≥3")),
         gr2_CRS = case_when(
           `Max CRS grade (0-5)` %in% c(0:1)                      ~ "Grade <2",
           `Max CRS grade (0-5)` %in% c(2:3)                      ~ "Grade ≥2",
           TRUE                                                   ~ NA_character_
           )) %>% 
  # mutate( = case_when(
  #    == 30                ~ NA,
  #   TRUE                          ~ CRS_fup
  # )) %>% 
  # ICANS 
  mutate(ICANS_any = ifelse(`Max ICANS (relative to infusion)` == 0, "NO", "YES"),
         ICANS_any2 = ifelse(ICANS_any == "NO", 0, 1),
         ICANS_gradecat = case_when(
           `Max ICANS (relative to infusion)` == 0 ~              "No ICANS",
           `Max ICANS (relative to infusion)` == 1 ~              "Grade 1 or 2",
           `Max ICANS (relative to infusion)` == 2 ~              "Grade 1 or 2",
           `Max ICANS (relative to infusion)` >= 3 ~              "Grade ≥3"),
         ICANS_gradecat = factor(ICANS_gradecat, levels = c("No ICANS", "Grade 1 or 2", "Grade ≥3")),
         gr2_ICANS = case_when(
           `Max ICANS (relative to infusion)` %in% c(0:1)         ~ "Grade <2",
           `Max ICANS (relative to infusion)` %in% c(2:3)         ~ "Grade ≥2",
           TRUE                                                   ~ NA_character_
         )) %>%

  mutate(`Neutropenia_7d` = case_when(
    `ANC Day 7` < 1.8                                              ~ "YES",
    is.na(`ANC Day 7`)                                          ~ NA_character_,
    TRUE                                                          ~ "NO"
  )) %>% 
  # mutate(`Neutropenia_14d` = case_when(
  #   `ANC day +14` < 1.8                                             ~ "YES",
  #   is.na(`ANC day +14`)                                          ~ NA_character_,
  #   TRUE                                                          ~ "NO"
  # )) %>% 
  # mutate(`Neutropenia_21d` = case_when(
  #   `ANC day +21` < 1.8                                             ~ "YES",
  #   is.na(`ANC day +21`)                                          ~ NA_character_,
  #   TRUE                                                          ~ "NO"
  # )) %>% 
  mutate(`Neutropenia_30d` = case_when(
    `ANC Day 30` < 1.8                                             ~ "YES",
    is.na(`ANC Day 30`)                                          ~ NA_character_,
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`Neutropenia_60d` = case_when(
    `ANC Day 60` < 1.8                                             ~ "YES",
    is.na(`ANC Day 60`)                                          ~ NA_character_,
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`Neutropenia_90d` = case_when(
    `ANC Day 90` < 1.8                                             ~ "YES",
    is.na(`ANC Day 90`)                                          ~ NA_character_,
    TRUE                                                          ~ "NO"
  )) %>% 
  # mutate(gr3_Neutropenia_apheresis = case_when(
  #   `ANC apheresis` < 1                                           ~ "YES",
  #   is.na(`ANC apheresis`)                                          ~ NA_character_,
  #   TRUE                                                          ~ "NO"
  # )) %>% 
  # mutate(`gr3_Neutropenia_-5d` = case_when(
  #   `ANC day -5 (k/ul)` < 1                                        ~ "YES",
  #   is.na(`ANC day -5 (k/ul)`)                                          ~ NA_character_,
  #   TRUE                                                          ~ "NO"
  # )) %>% 
  # mutate(`gr3_Neutropenia_0d` = case_when(
  #   `ANC day 0 (k/uL)` < 1                                        ~ "YES",
  #   is.na(`ANC day 0 (k/uL)`)                                          ~ NA_character_,
  #   TRUE                                                          ~ "NO"
  # )) %>% 
  mutate(`gr3_Neutropenia_7d` = case_when(
    `ANC Day 7` < 1                                              ~ "YES",
    is.na(`ANC Day 7`)                                          ~ NA_character_,
    TRUE                                                          ~ "NO"
  )) %>% 
  # mutate(`gr3_Neutropenia_14d` = case_when(
  #   `ANC day +14` < 1                                             ~ "YES",
  #   is.na(`ANC day +14`)                                          ~ NA_character_,
  #   TRUE                                                          ~ "NO"
  # )) %>% 
  # mutate(`gr3_Neutropenia_21d` = case_when(
  #   `ANC day +21` < 1                                             ~ "YES",
  #   is.na(`ANC day +21`)                                          ~ NA_character_,
  #   TRUE                                                          ~ "NO"
  # )) %>% 
  mutate(`gr3_Neutropenia_30d` = case_when(
    `ANC Day 30` < 1                                             ~ "YES",
    is.na(`ANC Day 30`)                                          ~ NA_character_,
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`gr3_Neutropenia_60d` = case_when(
    `ANC Day 60` < 1                                             ~ "YES",
    is.na(`ANC Day 60`)                                          ~ NA_character_,
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(`gr3_Neutropenia_90d` = case_when(
    `ANC Day 90` < 1                                             ~ "YES",
    is.na(`ANC Day 90`)                                          ~ NA_character_,
    TRUE                                                          ~ "NO"
  )) %>%

  # mutate(Anemia_apheresis = case_when(
  #   `Hb apheresis` < 11.4                                           ~ "YES",
  #   is.na(`Hb apheresis`)                                          ~ NA_character_,
  #   TRUE                                                          ~ "NO"
  # )) %>% 
  # mutate(`Anemia_-5d` = case_when(
  #   `Hb day -5 (k/ul)` < 11.4                                        ~ "YES",
  #   is.na(`Hb day -5 (k/ul)`)                                          ~ NA_character_,
  #   TRUE                                                          ~ "NO"
  # )) %>% 
  # mutate(`Anemia_0d` = case_when(
  #   `Hb day 0 (k/uL)` < 11.4                                        ~ "YES",
  #   is.na(`Hb day 0 (k/uL)`)                                          ~ NA_character_,
  #   TRUE                                                          ~ "NO"
  # )) %>% 
  mutate(`Anemia_7d` = case_when(
    `HgB Day 7` < 11.4                                              ~ "YES",
    is.na(`HgB Day 7`)                                          ~ NA_character_,
    TRUE                                                          ~ "NO"
  )) %>%
  # mutate(`Anemia_14d` = case_when(
  #   `Hb day +14` < 11.4                                             ~ "YES",
  #   is.na(`Hb day +14`)                                          ~ NA_character_,
  #   TRUE                                                          ~ "NO"
  # )) %>% 
  # mutate(`Anemia_21d` = case_when(
  #   `Hb day +21` < 11.4                                             ~ "YES",
  #   is.na(`Hb day +21`)                                          ~ NA_character_,
  #   TRUE                                                          ~ "NO"
  # )) %>% 
  mutate(`Anemia_30d` = case_when(
    `HgB Day 30` < 11.4                                             ~ "YES",
    is.na(`HgB Day 30`)                                          ~ NA_character_,
    TRUE                                                          ~ "NO"
  )) %>%
  mutate(`Anemia_60d` = case_when(
    `HgB Day 60` < 11.4                                             ~ "YES",
    is.na(`HgB Day 60`)                                          ~ NA_character_,
    TRUE                                                          ~ "NO"
  )) %>%
  mutate(`Anemia_90d` = case_when(
    `HgB Day 90` < 11.4                                             ~ "YES",
    is.na(`HgB Day 90`)                                          ~ NA_character_,
    TRUE                                                          ~ "NO"
  )) %>%
  # mutate(gr3_Anemia_apheresis = case_when(
  #   `Hb apheresis` < 8                                           ~ "YES",
  #   is.na(`Hb apheresis`)                                          ~ NA_character_,
  #   TRUE                                                          ~ "NO"
  # )) %>% 
  # mutate(`gr3_Anemia_-5d` = case_when(
  #   `Hb day -5 (k/ul)` < 8                                        ~ "YES",
  #   is.na(`Hb day -5 (k/ul)`)                                          ~ NA_character_,
  #   TRUE                                                          ~ "NO"
  # )) %>% 
  # mutate(`gr3_Anemia_0d` = case_when(
  #   `Hb day 0 (k/uL)` < 8                                        ~ "YES",
  #   is.na(`Hb day 0 (k/uL)`)                                          ~ NA_character_,
  #   TRUE                                                          ~ "NO"
  # )) %>% 
  mutate(`gr3_Anemia_7d` = case_when(
    `HgB Day 7` < 8                                              ~ "YES",
    is.na(`HgB Day 7`)                                          ~ NA_character_,
    TRUE                                                          ~ "NO"
  )) %>%
  # mutate(`gr3_Anemia_14d` = case_when(
  #   `Hb day +14` < 8                                             ~ "YES",
  #   is.na(`Hb day +14`)                                          ~ NA_character_,
  #   TRUE                                                          ~ "NO"
  # )) %>% 
  # mutate(`gr3_Anemia_21d` = case_when(
  #   `Hb day +21` < 8                                             ~ "YES",
  #   is.na(`Hb day +21`)                                          ~ NA_character_,
  #   TRUE                                                          ~ "NO"
  # )) %>% 
  mutate(`gr3_Anemia_30d` = case_when(
    `HgB Day 30` < 8                                             ~ "YES",
    is.na(`HgB Day 30`)                                          ~ NA_character_,
    TRUE                                                          ~ "NO"
  )) %>%
  mutate(`gr3_Anemia_60d` = case_when(
    `HgB Day 60` < 8                                             ~ "YES",
    is.na(`HgB Day 60`)                                          ~ NA_character_,
    TRUE                                                          ~ "NO"
  )) %>%
  mutate(`gr3_Anemia_90d` = case_when(
    `HgB Day 90` < 8                                             ~ "YES",
    is.na(`HgB Day 90`)                                          ~ NA_character_,
    TRUE                                                          ~ "NO"
  )) %>%
  # 
  # mutate(Thrombocytopenia_apheresis = case_when(
  #   `Plt apheresis` < 143                                           ~ "YES",
  #   is.na(`Plt apheresis`)                                          ~ NA_character_,
  #   TRUE                                                          ~ "NO"
  # )) %>% 
  # mutate(`Thrombocytopenia_-5d` = case_when(
  #   `Plt day -5 (k/ul)` < 143                                        ~ "YES",
  #   is.na(`Plt day -5 (k/ul)`)                                          ~ NA_character_,
  #   TRUE                                                          ~ "NO"
  # )) %>% 
  # mutate(`Thrombocytopenia_0d` = case_when(
  #   `Plt day 0 (k/uL)` < 143                                        ~ "YES",
  #   is.na(`Plt day 0 (k/uL)`)                                          ~ NA_character_,
  #   TRUE                                                          ~ "NO"
  # )) %>% 
  mutate(`Thrombocytopenia_7d` = case_when(
    `Platelets Day 7` < 143                                              ~ "YES",
    is.na(`Platelets Day 7`)                                          ~ NA_character_,
    TRUE                                                          ~ "NO"
  )) %>%
  # mutate(`Thrombocytopenia_14d` = case_when(
  #   `Plt day +14` < 143                                             ~ "YES",
  #   is.na(`Plt day +14`)                                          ~ NA_character_,
  #   TRUE                                                          ~ "NO"
  # )) %>% 
  # mutate(`Thrombocytopenia_21d` = case_when(
  #   `Plt day +21` < 143                                             ~ "YES",
  #   is.na(`Plt day +21`)                                          ~ NA_character_,
  #   TRUE                                                          ~ "NO"
  # )) %>% 
  mutate(`Thrombocytopenia_30d` = case_when(
    `Plateletes Day 30` < 143                                             ~ "YES",
    is.na(`Plateletes Day 30`)                                          ~ NA_character_,
    TRUE                                                          ~ "NO"
  )) %>%
  mutate(`Thrombocytopenia_60d` = case_when(
    `Platelets Day 60` < 143                                             ~ "YES",
    is.na(`Platelets Day 60`)                                          ~ NA_character_,
    TRUE                                                          ~ "NO"
  )) %>%
  mutate(`Thrombocytopenia_90d` = case_when(
    `Platelets Day 90` < 143                                             ~ "YES",
    is.na(`Platelets Day 90`)                                          ~ NA_character_,
    TRUE                                                          ~ "NO"
  )) %>%
  # mutate(gr3_Thrombocytopenia_apheresis = case_when(
  #   `Plt apheresis` < 50                                           ~ "YES",
  #   is.na(`Plt apheresis`)                                          ~ NA_character_,
  #   TRUE                                                          ~ "NO"
  # )) %>% 
  # mutate(`gr3_Thrombocytopenia_-5d` = case_when(
  #   `Plt day -5 (k/ul)` < 50                                        ~ "YES",
  #   is.na(`Plt day -5 (k/ul)`)                                          ~ NA_character_,
  #   TRUE                                                          ~ "NO"
  # )) %>% 
  # mutate(`gr3_Thrombocytopenia_0d` = case_when(
  #   `Plt day 0 (k/uL)` < 50                                        ~ "YES",
  #   is.na(`Plt day 0 (k/uL)`)                                          ~ NA_character_,
  #   TRUE                                                          ~ "NO"
  # )) %>% 
  mutate(`gr3_Thrombocytopenia_7d` = case_when(
    `Platelets Day 7` < 50                                              ~ "YES",
    is.na(`Platelets Day 7`)                                          ~ NA_character_,
    TRUE                                                          ~ "NO"
  )) %>%
  # mutate(`gr3_Thrombocytopenia_14d` = case_when(
  #   `Plt day +14` < 50                                             ~ "YES",
  #   is.na(`Plt day +14`)                                          ~ NA_character_,
  #   TRUE                                                          ~ "NO"
  # )) %>% 
  # mutate(`gr3_Thrombocytopenia_21d` = case_when(
  #   `Plt day +21` < 50                                             ~ "YES",
  #   is.na(`Plt day +21`)                                          ~ NA_character_,
  #   TRUE                                                          ~ "NO"
  # )) %>% 
  mutate(`gr3_Thrombocytopenia_30d` = case_when(
    `Plateletes Day 30` < 50                                             ~ "YES",
    is.na(`Plateletes Day 30`)                                          ~ NA_character_,
    TRUE                                                          ~ "NO"
  )) %>%
  mutate(`gr3_Thrombocytopenia_60d` = case_when(
    `Platelets Day 60` < 50                                             ~ "YES",
    is.na(`Platelets Day 60`)                                          ~ NA_character_,
    TRUE                                                          ~ "NO"
  )) %>%
  mutate(`gr3_Thrombocytopenia_90d` = case_when(
    `Platelets Day 90` < 50                                             ~ "YES",
    is.na(`Platelets Day 90`)                                          ~ NA_character_,
    TRUE                                                          ~ "NO"
  )) %>% 
  mutate(Neutropenia_gr1 = case_when(
    (`ANC Day 7` < 1.8 &
      `ANC Day 7` > 1.5) |
    (`ANC Day 30` < 1.8 &
      `ANC Day 30` > 1.5) |
      (`ANC Day 60` < 1.8 &
      `ANC Day 60` > 1.5) |
      (`ANC Day 90` < 1.8 &
      `ANC Day 90` > 1.5)                                          ~ "YES",
    is.na(`ANC Day 7`) |
      is.na(`ANC Day 30`) |
      is.na(`ANC Day 60`) |
      is.na(`ANC Day 90`)                                          ~ NA_character_,
    TRUE                                          ~ "NO"
  )) %>% 
  mutate(Neutropenia_gr2 = case_when(
    (`ANC Day 7` < 1.5 &
      `ANC Day 7` > 1) |
    (`ANC Day 30` < 1.5 &
      `ANC Day 30` > 1) |
      (`ANC Day 60` < 1.5 &
      `ANC Day 60` > 1) |
      (`ANC Day 90` < 1.5 &
      `ANC Day 90` > 1)                                          ~ "YES",
    is.na(`ANC Day 7`) |
      is.na(`ANC Day 30`) |
      is.na(`ANC Day 60`) |
      is.na(`ANC Day 90`)                                          ~ NA_character_,
    TRUE                                          ~ "NO"
  )) %>% 
  mutate(Neutropenia_gr3 = case_when(
    `ANC Day 7` < 1 |
    `ANC Day 30` < 1 |
      `ANC Day 60` < 1 |
      `ANC Day 90` < 1                                          ~ "YES",
    is.na(`ANC Day 7`) |
      is.na(`ANC Day 30`) |
      is.na(`ANC Day 60`) |
      is.na(`ANC Day 90`)                                          ~ NA_character_,
    TRUE                                          ~ "NO"
  )) %>% 
  # mutate(Neutropenia_gr4 = case_when(
  #   `ANC Day 7` < 0.5 |
  #   `ANC Day 30` < 0.5 |
  #     `ANC Day 60` < 0.5 |
  #     `ANC Day 90` < 0.5                                          ~ "YES",
  #   is.na(`ANC Day 7`) |
  #     is.na(`ANC Day 30`) |
  #     is.na(`ANC Day 60`) |
  #     is.na(`ANC Day 90`)                                          ~ NA_character_,
  #   TRUE                                          ~ "NO"
  # )) %>% 
  mutate(any_neutropenia = case_when(
    Neutropenia_gr1 == "YES" |
      Neutropenia_gr2 == "YES" |
      Neutropenia_gr3 == "YES"                                          ~ "YES",
    is.na(Neutropenia_gr1) |
          is.na(Neutropenia_gr2) |
    is.na(Neutropenia_gr3)                                          ~ NA_character_,
    TRUE                                          ~ "NO"
  )) %>% 
  mutate(severe_neutro_30beyond = case_when(
    `ANC Day 30` < 1 |
      `ANC Day 60` < 1 |
      `ANC Day 90` < 1                                          ~ "YES",
      is.na(`ANC Day 30`) |
      is.na(`ANC Day 60`) |
      is.na(`ANC Day 90`)                                          ~ NA_character_,
    TRUE                                          ~ "NO"
  )) %>%

  mutate(Anemia_gr1 = case_when(
      (`HgB Day 7` < 11.4 &
      `HgB Day 7` > 10) |
      (`HgB Day 30` < 11.4 &
      `HgB Day 30` > 10) |
      (`HgB Day 60` < 11.4 &
      `HgB Day 60` > 10) |
      (`HgB Day 90` < 11.4 &
      `HgB Day 90` > 10)                                        ~ "YES",
      is.na(`HgB Day 7`) |
      is.na(`HgB Day 30`) |
      is.na(`HgB Day 60`) |
      is.na(`HgB Day 90`)                                          ~ NA_character_,
    TRUE                                          ~ "NO"
  )) %>% 
  mutate(Anemia_gr2 = case_when(
      (`HgB Day 7` < 10 &
      `HgB Day 7` > 8) |
      (`HgB Day 30` < 10 &
      `HgB Day 30` > 8) |
      (`HgB Day 60` < 10 &
      `HgB Day 60` > 8) |
      (`HgB Day 90` < 10 &
      `HgB Day 90` > 8)                                        ~ "YES",
      is.na(`HgB Day 7`) |
      is.na(`HgB Day 30`) |
      is.na(`HgB Day 60`) |
      is.na(`HgB Day 90`)                                          ~ NA_character_,
    TRUE                                          ~ "NO"
  )) %>% 
  mutate(Anemia_gr3 = case_when(
      `HgB Day 7` < 8 |
      `HgB Day 30` < 8 |
      `HgB Day 60` < 8 |
      `HgB Day 90` < 8                                        ~ "YES",
      is.na(`HgB Day 7`) |
      is.na(`HgB Day 30`) |
      is.na(`HgB Day 60`) |
      is.na(`HgB Day 90`)                                          ~ NA_character_,
    TRUE                                          ~ "NO"
  )) %>% 
  # mutate(Anemia_gr4 = case_when(
  #     `HgB Day 7` < 0.1 |
  #     `HgB Day 30` < 0.1 |
  #     `HgB Day 60` < 0.1 |
  #     `HgB Day 90` < 0.1                                        ~ "YES",
  #     is.na(`HgB Day 7`) |
  #     is.na(`HgB Day 30`) |
  #     is.na(`HgB Day 60`) |
  #     is.na(`HgB Day 90`)                                          ~ NA_character_,
  #   TRUE                                          ~ "NO"
  # )) %>% 
  mutate(severe_anemia_30beyond = case_when(
    `HgB Day 30` < 8 |
      `HgB Day 60` < 8 |
      `HgB Day 90` < 8                                        ~ "YES",
      is.na(`HgB Day 30`) |
      is.na(`HgB Day 60`) |
      is.na(`HgB Day 90`)                                          ~ NA_character_,
    TRUE                                          ~ "NO"
  )) %>%

  mutate(Thrombocytopenia_gr1 = case_when(
      (`Platelets Day 7` < 143 &
      `Platelets Day 7` > 75) |
      (`Plateletes Day 30` < 143 &
      `Plateletes Day 30` > 75) |
      (`Platelets Day 60` < 143 &
      `Platelets Day 60` > 75) |
      (`Platelets Day 90` < 143 &
      `Platelets Day 90` > 75)                                        ~ "YES",
      is.na(`Platelets Day 7`) |
      is.na(`Plateletes Day 30`) |
      is.na(`Platelets Day 60`) |
      is.na(`Platelets Day 90`)                                          ~ NA_character_,
    TRUE                                          ~ "NO"
  )) %>% 
  mutate(Thrombocytopenia_gr2 = case_when(
      (`Platelets Day 7` < 75 &
      `Platelets Day 7` > 50) |
      (`Plateletes Day 30` < 75 &
      `Plateletes Day 30` > 50) |
      (`Platelets Day 60` < 75 &
      `Platelets Day 60` > 50) |
      (`Platelets Day 90` < 75 &
      `Platelets Day 90` > 50)                                        ~ "YES",
      is.na(`Platelets Day 7`) |
      is.na(`Plateletes Day 30`) |
      is.na(`Platelets Day 60`) |
      is.na(`Platelets Day 90`)                                          ~ NA_character_,
    TRUE                                          ~ "NO"
  )) %>% 
  mutate(Thrombocytopenia_gr3 = case_when(
      `Platelets Day 7` < 50 |
      `Plateletes Day 30` < 50 |
      `Platelets Day 60` < 50 |
      `Platelets Day 90` < 50                                        ~ "YES",
      is.na(`Platelets Day 7`) |
      is.na(`Plateletes Day 30`) |
      is.na(`Platelets Day 60`) |
      is.na(`Platelets Day 90`)                                          ~ NA_character_,
    TRUE                                          ~ "NO"
  )) %>% 
  # mutate(Thrombocytopenia_gr4 = case_when(
  #     `Platelets Day 7` < 25 |
  #     `Plateletes Day 30` < 25 |
  #     `Platelets Day 60` < 25 |
  #     `Platelets Day 90` < 25                                        ~ "YES",
  #     is.na(`Platelets Day 7`) |
  #     is.na(`Plateletes Day 30`) |
  #     is.na(`Platelets Day 60`) |
  #     is.na(`Platelets Day 90`)                                          ~ NA_character_,
  #   TRUE                                          ~ "NO"
  # )) %>% 
  mutate(severe_throm_30beyond = case_when(
    `Plateletes Day 30` < 50 |
      `Platelets Day 60` < 50 |
      `Platelets Day 90` < 50                                        ~ "YES",
    is.na(`Plateletes Day 30`) |
      is.na(`Platelets Day 60`) |
      is.na(`Platelets Day 90`)                                          ~ NA_character_,
    TRUE                                          ~ "NO"
  )) %>%

  mutate(`gr3_cytopenia≥30d` = case_when(
    `ANC Day 30` < 1 |
      `HgB Day 30` < 8 |
      `Plateletes Day 30` < 50 |
      `ANC Day 60` < 1 |
      `HgB Day 60` < 8 |
      `Platelets Day 60` < 50 |
      `ANC Day 90` < 1 |
      `HgB Day 90` < 8 |
      `Platelets Day 90` < 50                                        ~ "YES",
    is.na(`ANC Day 30`) |
      is.na(`HgB Day 30`) |
      is.na(`Plateletes Day 30`) |
      is.na(`ANC Day 60`) |
      is.na(`HgB Day 60`) |
      is.na(`Platelets Day 60`) |
      is.na(`ANC Day 90`) |
      is.na(`HgB Day 90`) |
      is.na(`Platelets Day 90`)                                       ~ NA_character_,
    TRUE                                                              ~ "NO"
  )) %>% 
  mutate(cart_cell_dose = case_when(
    `Cell Dose (million cells)` >= 400                               ~ "≥ 400",
    `Cell Dose (million cells)` >= 300                               ~ "≥ 300",
    TRUE                                                             ~ NA_character_
  )) %>% 
  # mutate(`gr3_cytopenia_30d` = case_when(
  #   `WBC day +30` < 2 | 
  #     `ANC Day 30` < 1 | 
  #     `HgB Day 30` < 8 | 
  #     `Plateletes Day 30` < 50                                              ~ "YES",
  #   is.na(`WBC day +30`) |
  #     is.na(`ANC Day 30`) |
  #     is.na(`HgB Day 30`) |
  #     is.na(`Plateletes Day 30`)                                            ~ NA_character_,
  #   TRUE                                                              ~ "NO"
  # )) %>% 
  # mutate(`gr3_cytopenia_90d` = case_when(
  #   `WBC day +90` < 2 | 
  #     `ANC Day 90` < 1 | 
  #     `HgB Day 90` < 8 | 
  #     `Platelets Day 90` < 50                                             ~ "YES",
  #   is.na(`WBC day +90`) |
  #     is.na(`ANC Day 90`) |
  #     is.na(`HgB Day 90`) |
  #     is.na(`Platelets Day 90`)                                            ~ NA_character_,
  #   TRUE                                                              ~ "NO"
  # )) %>% 
  # mutate(any_infection = case_when(
  #   `Infection (Yes, No)` == 1                                    ~ "Yes",
  #   `Infection (Yes, No)` == 0                                    ~ "No"
  # )) %>%
  # mutate(any_sevinfection = case_when(
  #   `Severe Infection (Yes, No)` == 1                              ~ "Yes",
  #   `Severe Infection (Yes, No)` == 0                              ~ "No"
  # )) %>% 
  mutate(baseline_ferritin = case_when(
    `Baseline Ferritin` >= 400                                    ~ "≥ ULN at LD",
    `Baseline Ferritin` < 400                                    ~ "Normal"
  )) %>% 
  mutate(baseline_crp = case_when(
    `Baseline CRP` >= 0.5                                    ~ "≥ ULN at LD",
    `Baseline CRP` < 0.5                                    ~ "Normal"
  )) %>% 
  mutate(baseline_B2M = case_when(
    `B2M` >= 5.5                                    ~ ">= 5.5",
    `B2M` < 5.5                                    ~ "< 5.5"
  )) %>%

  mutate(`Day 30 Response (sCR, CR, VGPR, PR)` = case_when(
    str_detect(`Day 30 Response (sCR, CR, VGPR, PR)`, "Not")         ~ NA_character_,
    TRUE           ~ `Day 30 Response (sCR, CR, VGPR, PR)`
  )) %>% 
  mutate(ORR_30days = case_when(
    str_detect(`Day 30 Response (sCR, CR, VGPR, PR)`, "CR")   ~ "ORR",
    str_detect(`Day 30 Response (sCR, CR, VGPR, PR)`, "PR")   ~ "ORR",
    is.na(`Day 30 Response (sCR, CR, VGPR, PR)`)              ~ NA_character_,
    TRUE                                                                        ~ "PD/SD/MR"
  )) %>% 
  # mutate(MRDneg_30days = case_when(
  #   str_detect(`Day 30 Response (sCR, CR, VGPR, PR)`, "MRD-")        ~ "Yes",
  #   str_detect(`Day 30 Response (sCR, CR, VGPR, PR)`, "MRD")         ~ "No",
  #   TRUE                             ~ NA_character_
  # )) %>% 
  mutate(day30_response = `Day 30 Response (sCR, CR, VGPR, PR)`
  ) %>% 
  mutate(day30_response_group = case_when(
    str_detect(`Day 30 Response (sCR, CR, VGPR, PR)`, "VGPR")        ~ "VGPR",
    str_detect(`Day 30 Response (sCR, CR, VGPR, PR)`, "PR")          ~ "PR",
    str_detect(`Day 30 Response (sCR, CR, VGPR, PR)`, "CR") &
      str_detect(`Day 30 MRD (positive vs negative)`, "MRD-")        ~ "CR or sCR, MRD-",
    str_detect(`Day 30 Response (sCR, CR, VGPR, PR)`, "CR") &
      str_detect(`Day 30 MRD (positive vs negative)`, "MRD+")        ~ "CR or sCR, MRD+",
    str_detect(`Day 30 Response (sCR, CR, VGPR, PR)`, "CR") &
      is.na(`Day 30 MRD (positive vs negative)`)                     ~ "CR or sCR, MRD unknown",
    is.na(`Day 30 Response (sCR, CR, VGPR, PR)`)                     ~ NA_character_,
    TRUE           ~ `Day 30 Response (sCR, CR, VGPR, PR)`
  )) %>% 
  mutate(day30_response_group1 = case_when(
    str_detect(`Day 30 Response (sCR, CR, VGPR, PR)`, "VGPR")        ~ "VGPR",
    str_detect(`Day 30 Response (sCR, CR, VGPR, PR)`, "PR")          ~ "PR",
    str_detect(`Day 30 Response (sCR, CR, VGPR, PR)`, "CR")          ~ "CR or sCR",
    is.na(`Day 30 Response (sCR, CR, VGPR, PR)`)                     ~ NA_character_,
    TRUE           ~ `Day 30 Response (sCR, CR, VGPR, PR)`
  )) %>% 
  mutate(day30_response_sCRvsOthers = case_when(
    str_detect(`Day 30 Response (sCR, CR, VGPR, PR)`, "CR")	         ~ "CR or sCR",
    is.na(`Day 30 Response (sCR, CR, VGPR, PR)`)                     ~ NA_character_,
    TRUE           ~ "Others"
  )) %>%

  mutate(`3 Month Response (sCR, CR, VGPR, PR)` = case_when(
    str_detect(`3 Month Response (sCR, CR, VGPR, PR)`, "SD")        ~ "SD",
    str_detect(`3 Month Response (sCR, CR, VGPR, PR)`, "evaluable")   ~ NA_character_,
    TRUE           ~ `3 Month Response (sCR, CR, VGPR, PR)`
  )) %>% 
  mutate(ORR_90days = case_when(
    str_detect(`3 Month Response (sCR, CR, VGPR, PR)`, "CR")    ~ "ORR",
    str_detect(`3 Month Response (sCR, CR, VGPR, PR)`, "PR")    ~ "ORR",
    is.na(`3 Month Response (sCR, CR, VGPR, PR)`)               ~ NA_character_,
    TRUE                                                                        ~ "PD/SD/MR"
  )) %>% 
  # mutate(MRDneg_90days = case_when(
  #   str_detect(`3 Month Response (sCR, CR, VGPR, PR)`, "MRD-")        ~ "Yes",
  #   str_detect(`3 Month Response (sCR, CR, VGPR, PR)`, "MRD")         ~ "No",
  #   TRUE                             ~ NA_character_
  # )) %>% 
  mutate(day90_response_group = case_when(
    str_detect(`3 Month Response (sCR, CR, VGPR, PR)`, "VGPR")        ~ "VGPR",
    str_detect(`3 Month Response (sCR, CR, VGPR, PR)`, "PR")          ~ "PR",
    str_detect(`3 Month Response (sCR, CR, VGPR, PR)`, "CR") &
      str_detect(`3 Month MRD (positive vs negative)`, "MRD-")        ~ "CR or sCR, MRD-",
    str_detect(`3 Month Response (sCR, CR, VGPR, PR)`, "CR") &
      str_detect(`3 Month MRD (positive vs negative)`, "MRD+")        ~ "CR or sCR, MRD+",
    str_detect(`3 Month Response (sCR, CR, VGPR, PR)`, "CR") &
      is.na(`3 Month MRD (positive vs negative)`)                     ~ "CR or sCR, MRD unknown",
    is.na(`3 Month Response (sCR, CR, VGPR, PR)`)                     ~ NA_character_,
    TRUE           ~ `3 Month Response (sCR, CR, VGPR, PR)`
  )) %>% 
  mutate(day90_response_group1 = case_when(
    str_detect(`3 Month Response (sCR, CR, VGPR, PR)`, "VGPR")        ~ "VGPR",
    str_detect(`3 Month Response (sCR, CR, VGPR, PR)`, "PR")          ~ "PR",
    str_detect(`3 Month Response (sCR, CR, VGPR, PR)`, "CR")          ~ "CR or sCR",
    is.na(`3 Month Response (sCR, CR, VGPR, PR)`)                     ~ NA_character_,
    TRUE           ~ `3 Month Response (sCR, CR, VGPR, PR)`
  )) %>% 
  mutate(day90_response =`3 Month Response (sCR, CR, VGPR, PR)`
  ) %>% 
  mutate(day90_response_sCRvsOthers = case_when(
    str_detect(`3 Month Response (sCR, CR, VGPR, PR)`, "CR")	         ~ "CR or sCR",
    is.na(`3 Month Response (sCR, CR, VGPR, PR)`)                      ~ NA_character_,
    TRUE           ~ "Others"
  )) %>% 
  mutate(best_response = case_when(
    day30_response_group == "CR or sCR, MRD-" |
      day90_response_group == "CR or sCR, MRD-"                        ~ "CR or sCR, MRD-",
    day30_response_group == "CR or sCR, MRD+" |
      day90_response_group == "CR or sCR, MRD+"                        ~ "CR or sCR, MRD+",
    day30_response_group == "CR or sCR, MRD unknown" |
      day90_response_group == "CR or sCR, MRD unknown"                 ~ "CR or sCR, MRD unknown",
    day30_response_group == "VGPR" |
      day90_response_group == "VGPR"                                   ~ "VGPR",
    day30_response_group == "PR" |
      day90_response_group == "PR"                                     ~ "PR",

    day30_response_group %in% c("MR", "PD", "SD") |
      day90_response_group %in% c("MR", "PD", "SD")                    ~ "Others",

    TRUE                                                               ~ NA_character_
  )) %>% 
  mutate(best_MRD_response = case_when(
    str_detect(`Day 30 MRD (positive vs negative)`, "MRD-") |
      str_detect(`3 Month MRD (positive vs negative)`, "MRD-")         ~ "MRD-",
    str_detect(`Day 30 MRD (positive vs negative)`, "MRD+") |
      str_detect(`3 Month MRD (positive vs negative)`, "MRD+")         ~ "MRD+",
    TRUE                                                               ~ NA_character_
  )) %>% 
  mutate(overall_response = case_when(
    !is.na(best_response)                 ~ "Response",
    is.na(best_response)                  ~ "No response"
  )) %>% 
  mutate(months_at_best_response = case_when(
    best_response == day30_response_group     ~ 1,
    best_response == day90_response_group     ~ 3
  )) %>% 
  mutate(months_at_best_CRresponse = case_when(
    str_detect(best_response, "CR") &
      best_response == day30_response_group     ~ 1,
    str_detect(best_response, "CR") &
      best_response == day90_response_group     ~ 3
  )) %>% 
  # mutate(MRDneg_in_best_response = case_when(
  #   str_detect(`3 Month Response (sCR, CR, VGPR, PR)`, "MRD-") |
  #   str_detect(`Day 30 Response (sCR, CR, VGPR, PR)`, "MRD-")        ~ "Yes",
  #   str_detect(`3 Month Response (sCR, CR, VGPR, PR)`, "MRD") |
  #   str_detect(`Day 30 Response (sCR, CR, VGPR, PR)`, "MRD")         ~ "No",
  #   TRUE                             ~ NA_character_
  # )) %>% 
  # Median follow up
  mutate(time_leukapheresis_fup = 
           interval(
             start = date_of_apheresis, 
             end = `Date of Last Contact`)/
           duration(n=1, units = "days")) %>%
  # Median time from leukapheresis to infusion 
  mutate(time_leukapheresis_infusion = 
           interval(
             start = date_of_apheresis, 
             end = `Date of CAR T infusion (# of days from apharesis to infusion)`)/
           duration(n=1, units = "days")) %>%
  # Code survival
  mutate(pfs_event = case_when(
    is.na(`Date of PD`)                                ~ 0,
    !is.na(`Date of PD`)                               ~ 1,
    TRUE                                               ~ NA_real_
  )) %>% 
  mutate(pfs_time = coalesce(`Date of PD`, `Date of Death`, `Date of Last Contact`)) %>% 
  mutate(days_at_pfs_from_infusion = 
           interval(
             start = date_of_apheresis, 
             end = pfs_time)/
           duration(n=1, units = "months")) %>%

  mutate(os_event = case_when(
    is.na(`Date of Death`)                             ~ 0,
    !is.na(`Date of Death`)                            ~ 1,
    TRUE                                               ~ NA_real_
  )) %>% 
  mutate(os_time = coalesce(`Date of Death`, `Date of Last Contact`)) %>% 
  mutate(days_at_os_from_infusion = 
           interval(
             start = date_of_apheresis, 
             end = os_time)/
           duration(n=1, units = "months"))

  # mutate(`IgG HSV1/2 apheresis` = case_when(
  #   `IgG HSV1/2 apheresis` < 0.9         ~ "Negative",
  #   `IgG HSV1/2 apheresis` >= 0.9 &
  #     `IgG HSV1/2 apheresis` < 1.10      ~ "indeterminate",
  #   `IgG HSV1/2 apheresis` >= 1.10       ~ "Positive",
  # )) %>%
  # mutate(`IgG HSV1/2 D90` = case_when(
  #   `IgG HSV1/2 D90` < 0.9         ~ "Negative",
  #   `IgG HSV1/2 D90` >= 0.9 &
  #     `IgG HSV1/2 D90` < 1.10      ~ "indeterminate",
  #   `IgG HSV1/2 D90` >= 1.10       ~ "Positive",
  # )) %>% 
  # mutate_at(c("VZV IgG apheresis",
  #             "VZV IgG D90 (1=pos, 0=neg)",
  #             "CMV IgG apheresis",
  #             "CMV IgG D90 (1=pos, 0=neg)"), 
  #           ~ case_when(
  #             . == 1                                              ~ "Positive",
  #             . == 0                                              ~ "Negative",
  #             . == "Equivocal"                                    ~ "Negative",
  #             TRUE                                                ~ as.character(.)
  #           )) %>% 
  # mutate_at(c("IgG HSV1/2 apheresis",
  #             "IgG HSV1/2 D90",
  #             "VZV IgG apheresis",
  #             "VZV IgG D90 (1=pos, 0=neg)",
  #             "CMV IgG apheresis",
  #             "CMV IgG D90 (1=pos, 0=neg)"),
  #           ~ factor(., levels = c("Positive", "Negative"))) %>% 
  # mutate(across(starts_with("Pneumo"), ~ case_when(
  #   . <= 1                                                    ~ "Negative",
  #   . > 1                                                     ~ "Positive"
  # ))) %>% 
  # 
  # mutate(igg_hsv_change = case_when(
  #   `IgG HSV1/2 apheresis` == "Positive" &
  #     `IgG HSV1/2 D90` == "Negative"         ~ -1,
  #   `IgG HSV1/2 apheresis` == "Negative" &
  #     `IgG HSV1/2 D90` == "Positive"         ~ 1,
  #   `IgG HSV1/2 apheresis` == "Negative" &
  #     `IgG HSV1/2 D90` == "Negative"         ~ -0.25,
  #   `IgG HSV1/2 apheresis` == "Positive" &
  #     `IgG HSV1/2 D90` == "Positive"         ~ 0.25
  # )) %>% 
  # mutate(igg_vzv_change = case_when(
  #   `VZV IgG apheresis` == "Positive" &
  #     `VZV IgG D90 (1=pos, 0=neg)` == "Negative"               ~ -1,
  #   `VZV IgG apheresis` == "Negative" &
  #     `VZV IgG D90 (1=pos, 0=neg)` == "Positive"               ~ 1,
  #   `VZV IgG apheresis` == "Negative" &
  #     `VZV IgG D90 (1=pos, 0=neg)` == "Negative"               ~ -0.25,
  #   `VZV IgG apheresis` == "Positive" &
  #     `VZV IgG D90 (1=pos, 0=neg)` == "Positive"               ~ 0.25
  # )) %>% 
  # mutate(igg_cmv_change = case_when(
  #   `CMV IgG apheresis` == "Positive" &
  #     `CMV IgG D90 (1=pos, 0=neg)` == "Negative"               ~ -1,
  #   `CMV IgG apheresis` == "Negative" &
  #     `CMV IgG D90 (1=pos, 0=neg)` == "Positive"               ~ 1,
  #   `CMV IgG apheresis` == "Negative" &
  #     `CMV IgG D90 (1=pos, 0=neg)` == "Negative"               ~ -0.25,
  #   `CMV IgG apheresis` == "Positive" &
  #     `CMV IgG D90 (1=pos, 0=neg)` == "Positive"               ~ 0.25
  # ))

# Create Pneumo titers change value
# for (i in colnames(cart0 %>%
#                    select(c(starts_with("Pneumo"))))) {
# 
#   pneumo_type <-
#     str_match(i, ".*(apheresis|D90)")[, 1] %>% 
#     # str_replace_all(., " ", "_") %>% 
#     str_remove(., " apheresis| D90")
#   
#   pneumo_subset <- cart0 %>% select(patient_id, starts_with(pneumo_type))
#   
#   pneumo_type <-
#     str_replace_all(pneumo_type, " ", "_")
#   # day <-
#   #   str_match(i, ".*(apheresis|D90)")[, 2]
#  
#   name <- paste0(pneumo_type, "_", "change")
# 
#   cart0[[paste0(pneumo_type, "_", "change")]] <-  case_when(
#     pneumo_subset[2] == "Positive" &
#       pneumo_subset[3] == "Negative"               ~ -1,
#     pneumo_subset[2] == "Negative" &
#       pneumo_subset[3] == "Positive"               ~ 1,
#     pneumo_subset[2] == "Negative" &
#       pneumo_subset[3] == "Negative"               ~ -0.25,
#     pneumo_subset[2] == "Positive" &
#       pneumo_subset[3] == "Positive"               ~ 0.25
#   )
# }

# # CrCl
# table(cart$`Baseline CrCl`)
# cart$crcl[cart$`Baseline CrCl` == ">70" | cart$`Baseline CrCl` == "60.8" | cart$`Baseline CrCl` == "62" | cart$`Baseline CrCl` == "67"] <- ">60"
# cart$crcl[cart$`Baseline CrCl` == "25" | cart$`Baseline CrCl` == "27" | cart$`Baseline CrCl` == "34" | cart$`Baseline CrCl` == "41" | cart$`Baseline CrCl` == "43" | cart$`Baseline CrCl` == "50" | cart$`Baseline CrCl` == "54" | cart$`Baseline CrCl` == "55"] <- "≤60"
# table(cart$crcl)
# # CRS
# table(cart$`Max CRS grade (0-5)`)
# as.factor(cart$`Max CRS grade (0-5)`)
# table(cart$`Duration of CRS (days)`)
# table(cart$`Duration of ICANS (days)`)
# # steroid use
# table(cart$`Steroid use(Yes, No)`)
# 
# 
#

# 
# # RSS
# table(cart$`R-ISS at CAR-T infusion`)
# cart$`R-ISS at CAR-T infusion`[cart$`R-ISS at CAR-T infusion` == "R-ISS  II"] <- "R-ISS II"

# day 30 + MRD for figure
#table(tct$day30_response_revised, tct$MRD_status, useNA= "ifany")
#table(tct$day30_response_revised, tct$day30_response, useNA= "ifany")
#tct$ORR_MRD <- case_when(#tct$day30_response_revised== "SD" ~ "SD",
                         #tct$day30_response_revised== "PD" ~ "PD",
                         #tct$day30_response_revised== "MR" ~ "MR",
#                         tct$day30_response_revised== "PR" ~ "PR",
                         #tct$day30_response_revised== "PR" & tct$MRD_status== "Positive" ~ "PR and MRD-Positive",
                         #tct$day30_response_revised== "PR" & is.na(tct$MRD_status) ~ "PR and MRD unknown",
#                         tct$day30_response_revised== "sCR" & tct$MRD_status== "Negative" ~ "sCR and MRD-Negative",
#                         tct$day30_response_revised== "sCR" & tct$MRD_status== "Positive" ~ "sCR and MRD-Positive")
#table(tct$ORR_MRD)
```

```{r}
## Labeling
var_label(cart0) <- list(Age = "Patient Age", agecat = "Patient Age - Categories", 
                        # tumor_burden ="Tumor burden", 
                        proteasome = "Refractory to Proteasome Inhibitors",
                       immuno = "Refractory to Immunomodulators", doubleref = "Double refractory",
                       tripleref="Triple refractory", pentaref="Penta refractory",
                       CRS_any = "Any CRS", CRS_gradecat = "CRS Grade",
                       ICANS_any = "Any ICANS", ICANS_gradecat = "ICANS Grade", 
                       `Deletion 17p prior to CAR-T infusion (yes, no)` = "Deletion 17p at Infusion",
                       `t(4:14) prior to CAR-T infusion (yes, no)` = "t(4:14) at Infusion",
                       `t(14:16) prior to CAR-T infusion (yes, no)` = "t(4:16) at Infusion",
                       `Prior auto SCT (yes, no)` = "Prior auto SCT",
                       `Extramedullary Disease? (Yes, No)` = "Extramedullary disease",
                       dara="Refractory to Daratumumab")

```

<br>

### Tables

## Table X. Manufacturing failures
```{r}
MM_cart %>%
  # Exclude patients who had manufacturing cart failure or died prior cart infusion
  select(`First manufacturing failure`,
         `Manufaturing failures with successfull second collection`) %>%
  mutate(across(everything(), ~ replace_na(., 0))) %>%
  tbl_summary(
    statistic = list(all_continuous() ~ "{median} ({min}, {max})"),
    digits = all_continuous() ~ 1,
    missing = "no"
  ) %>%
  bold_labels() %>% add_stat_label() %>%
  modify_footnote(update = everything() ~ NA)
```
<br>

## Table 1. Patient baseline characteristics
```{r}
cart0 %>% 
  select(Age, agecat, male_sex, Sex,
         `Extramedullary Disease? (Yes, No)`, 
         `Plasma Cell Leukemia? (yes, no)`,
         `Amyloid? (yes, no)`,
         POEMS,
         `Non-secretory MM (Yes, No)`, ################??????????
         `High Marrow Burden (>= 50%)`, 
         `ECOG at consult`, `ECOG at LD`, 
         `R-ISS at CAR-T infusion`, 
         cytogenetics,
         `Deletion 17p prior to CAR-T infusion (yes, no)`,
         `t(4:14) prior to CAR-T infusion (yes, no)`,
         `t(14:16) prior to CAR-T infusion (yes, no)`,
         `Bridging Therapy? (Yes, No)`,
         `Number of prior lines of therapy`,
         number_prior_lines_therapy,
         `Disease status at time of referral (Relapsed vs Refractory)`,
         `Prior auto SCT (yes, no)`, 
         `ASCT within 12 weeks prior to apheresis (Yes, No)`,
         `Prior Allo SCT (Yes, No)`,
         `Prior treatment with any gene therapy-based therapeutic or investigational cellular therapy or BCMA targeted therapy (Yes, No)`,
         Blenrep,`BCMA CAR-T on trial`, `Other BCMA trial`,
         immuno, proteasome, dara, doubleref, tripleref, pentaref, 
         `Cell Dose (million cells)`, cart_cell_dose,
         exclusion_criteria,

         `ECOG > or = 2? (Yes, No)`,
         `ANC < 1000/uL (Yes, No)`,
         `Hemoglobin < 8 g/dL (Yes, No)`,
         `Platelets < 50K (Yes, No)`, 
         `H/O or presence of Plasma cell leukemia, Waldenstrom's, POEMS, or amyloidosis (Yes, No)`,
         `Non-secretory MM (Yes, No)`,
         `H/O or presence of CNS pathology (epilepsy, seizure, paresis, aphasia, stroke, CNS bleed, brian injuries, dementia, Parkinson's disease, cerebellar disease, organic brain syndrome, or psychosis) (Yes, No)`,
         `Prior Allo SCT (Yes, No)`,
         `Prior treatment with any gene therapy-based therapeutic or investigational cellular therapy or BCMA targeted therapy (Yes, No)`,
         comorbidities,

         `Did patient met criteria for KarMMa1 pre-CAR-T?`,
         time_leukapheresis_infusion, time_leukapheresis_fup) %>% 
  tbl_summary(
    type = list(
      c(
      `Extramedullary Disease? (Yes, No)`, 
      `Plasma Cell Leukemia? (yes, no)`,
         `Amyloid? (yes, no)`,
         POEMS,
         `Non-secretory MM (Yes, No)`,
      `High Marrow Burden (>= 50%)`,
    cytogenetics,
    `Deletion 17p prior to CAR-T infusion (yes, no)`,
        `t(4:14) prior to CAR-T infusion (yes, no)`,
        `t(14:16) prior to CAR-T infusion (yes, no)`,
    `Prior auto SCT (yes, no)`, 
         `ASCT within 12 weeks prior to apheresis (Yes, No)`,
         `Prior Allo SCT (Yes, No)`,
         `Prior treatment with any gene therapy-based therapeutic or investigational cellular therapy or BCMA targeted therapy (Yes, No)`,
         Blenrep,`BCMA CAR-T on trial`, `Other BCMA trial`,
        immuno, proteasome, dara, doubleref, tripleref, pentaref,
    `ECOG > or = 2? (Yes, No)`,
         `ANC < 1000/uL (Yes, No)`,
         `Hemoglobin < 8 g/dL (Yes, No)`,
         `Platelets < 50K (Yes, No)`, 
         `H/O or presence of Plasma cell leukemia, Waldenstrom's, POEMS, or amyloidosis (Yes, No)`,
    `H/O or presence of CNS pathology (epilepsy, seizure, paresis, aphasia, stroke, CNS bleed, brian injuries, dementia, Parkinson's disease, cerebellar disease, organic brain syndrome, or psychosis) (Yes, No)`,
    comorbidities,
        `Did patient met criteria for KarMMa1 pre-CAR-T?`
    )
      ~"categorical"),

    statistic = list(all_continuous() ~ "{median} ({min}, {max})"), 
    digits = all_continuous() ~ 1
  ) %>% 
  bold_labels() %>% add_stat_label() %>% 
  modify_footnote(update = everything() ~ NA)
```

## From here, we excluded 38 patients
```{r}
cart0 %>% 
  filter(!is.na(exclude_for_manufacturing_cart_failure_or_died_prior_cart_infusion)) %>% 
  select(`Patients excluded` = patient_id)

cart0 <- cart0 %>% 
  # Exclude patients who had manufacturing cart failure or died prior cart infusion
  filter(is.na(exclude_for_manufacturing_cart_failure_or_died_prior_cart_infusion))
```

## Table 2. Safety with Standard of Care Idecabtagene Vicleucel Event and Grade

```{r}
# tbl1 <- 
  cart0 %>% 
  select(`Max CRS grade (0-5)`,
         `Day of Max CRS (relative to infusion)`,
         `Max ICANS (relative to infusion)`,
         `Day of Max ICANS (relative to infusion)`,
         `Length of Hospital Stay (Total Days including Readmission)`,
         `ICU Admission? (Yes, No)`,
         `Tocilizumab (Yes, No)`, `Steroid use (Yes, No)`, `Anakinra (Yes, No)`,
         any_neutropenia,
         Neutropenia_gr1, Neutropenia_gr2, 
         Neutropenia_gr3,
         Anemia_gr1, Anemia_gr2,
         Anemia_gr3,
         Thrombocytopenia_gr1, Thrombocytopenia_gr2,
         Thrombocytopenia_gr3,
         severe_neutro_30beyond, severe_anemia_30beyond,
         severe_throm_30beyond,

         `Need for GCSF`,
         `Need for TPO Agonist`,
         `Stem Cell Boost`
         ) %>% 
  tbl_summary(
    type = list(
      c(`ICU Admission? (Yes, No)`,
        `Anakinra (Yes, No)`,
        any_neutropenia,
         Neutropenia_gr1, Neutropenia_gr2, 
         Neutropenia_gr3,
         Anemia_gr1, Anemia_gr2,
         Anemia_gr3,
         Thrombocytopenia_gr1, Thrombocytopenia_gr2,
         Thrombocytopenia_gr3,
        severe_neutro_30beyond, severe_anemia_30beyond,
         severe_throm_30beyond,
        `Need for GCSF`,
        `Need for TPO Agonist`,
        `Stem Cell Boost`
      ) ~ "categorical"),
    statistic=list(all_continuous() ~ "{median} ({min}, {max}) [{p25}, {p75}]"), 
    digits = all_continuous() ~ 1,
    missing = "no") %>% 
  bold_labels() %>% add_stat_label() %>% 
  modify_footnote(update = everything() ~ NA)
```

## Table 3.Responses
```{r}
cart0 %>% 
  select(day30_response, 
         day30_response_group, day30_response_sCRvsOthers, 
         day90_response, 
         day90_response_group, day90_response_sCRvsOthers) %>% 
  tbl_summary(missing = "no") %>% 
  bold_labels() %>% add_stat_label() %>% 
  modify_footnote(update = everything() ~ NA)
```

left is day 30, top is day 90 - **restricted to patients who have a 30 day AND a 90 day response.**
```{r}
table(cart0$day30_response_sCRvsOthers, cart0$day90_response_sCRvsOthers)

table(cart0$day30_response, cart0$day90_response)
```

```{r}
cart0 %>% 
  select(day90_response_group1, 
         CRS_any) %>% 
  tbl_summary(by= CRS_any,
              missing = "no") %>% 
  bold_labels() %>% add_stat_label() %>% add_p() %>% 
  modify_footnote(update = everything() ~ NA) %>% 
  modify_spanning_header(all_stat_cols() ~ "**Any CRS**")
```

Of the total 141 patients with a response, how many were MRD neg (N and %).
```{r}
cart0 %>% 
  mutate(has_response = case_when(
    !is.na(best_response)     ~ "Patient with responses"
  )) %>% 
  select(has_response, 
         `Day 30 MRD (positive vs negative)`) %>% 
  tbl_summary(by= has_response,
              missing = "no") %>% 
  bold_labels() %>% add_stat_label() %>%
  modify_footnote(update = everything() ~ NA)

cart0 %>% 
  mutate(has_response = case_when(
    !is.na(best_response)     ~ "Patient with responses"
  )) %>% 
  select(has_response, 
         `3 Month MRD (positive vs negative)`) %>% 
  tbl_summary(by= has_response,
              missing = "no") %>% 
  bold_labels() %>% add_stat_label() %>%
  modify_footnote(update = everything() ~ NA)

cart0 %>% 
  mutate(has_response = case_when(
    !is.na(best_response)     ~ "Patient with responses"
  )) %>% 
  select(has_response, 
         best_MRD_response) %>% 
  tbl_summary(by= has_response,
              missing = "no") %>% 
  bold_labels() %>% add_stat_label() %>%
  modify_footnote(update = everything() ~ NA)
```

## Table 4.Characteristics associated with grade ≥ 3 CRS, grade ≥ 3 neurotoxicity, grade ≥ 3 cytopenias ≥ 30 days, best response of ≥ CR at 3 months, best ORR 
```{r}
cart0 %>% 
  select(Age, agecat, male_sex, Sex,
         `Extramedullary Disease? (Yes, No)`, 
         `Disease status at time of referral (Relapsed vs Refractory)`,
         `Plasma Cell Leukemia? (yes, no)`,
         `Amyloid? (yes, no)`,
         POEMS,
         `Non-secretory MM (Yes, No)`, ################??????????
         `High Marrow Burden (>= 50%)`, 
         `ECOG at consult`, `ECOG at LD`, 
         `R-ISS at CAR-T infusion`, 
         cytogenetics,
         `Deletion 17p prior to CAR-T infusion (yes, no)`,
         `t(4:14) prior to CAR-T infusion (yes, no)`,
         `t(14:16) prior to CAR-T infusion (yes, no)`,
         `Bridging Therapy? (Yes, No)`,
         `Number of prior lines of therapy`,
         number_prior_lines_therapy,
         `Prior auto SCT (yes, no)`, 
         `ASCT within 12 weeks prior to apheresis (Yes, No)`,
         `Prior Allo SCT (Yes, No)`,
         Blenrep,
         `Prior treatment with any gene therapy-based therapeutic or investigational cellular therapy or BCMA targeted therapy (Yes, No)`,
         `BCMA CAR-T on trial`, `Other BCMA trial`,
         immuno, proteasome, dara, doubleref, tripleref, pentaref, 
         cart_cell_dose,
         `Infection (Yes, No)`,
         `Tocilizumab (Yes, No)`, `Steroid use (Yes, No)`, `Anakinra (Yes, No)`,
         baseline_ferritin, baseline_crp, baseline_B2M,
         `ECOG > or = 2? (Yes, No)`,
         `Renal insufficiency CrCl < 45 mL/min (Yes, No)`,
         `LVEF <45% (Yes, No)`,
         `ANC < 1000/uL (Yes, No)`,
         `Hemoglobin < 8 g/dL (Yes, No)`,
         `Platelets < 50K (Yes, No)`, 
         `H/O or presence of Plasma cell leukemia, Waldenstrom's, POEMS, or amyloidosis (Yes, No)`,
         `Non-secretory MM (Yes, No)`,
         `Did patient met criteria for KarMMa1 pre-CAR-T?`,
         gr2_CRS
  ) %>% 
  tbl_summary(by = gr2_CRS,
              type = list(
                c(
                  `Extramedullary Disease? (Yes, No)`, 
                  `Plasma Cell Leukemia? (yes, no)`,
                  `Amyloid? (yes, no)`,
                  POEMS,
                  `Non-secretory MM (Yes, No)`,
                  `High Marrow Burden (>= 50%)`,
                  `ECOG at consult`, `ECOG at LD`,
                  cytogenetics,
                  `Deletion 17p prior to CAR-T infusion (yes, no)`,
                  `t(4:14) prior to CAR-T infusion (yes, no)`,
                  `t(14:16) prior to CAR-T infusion (yes, no)`,
                  `Bridging Therapy? (Yes, No)`, `Prior auto SCT (yes, no)`,
                  `ASCT within 12 weeks prior to apheresis (Yes, No)`,
                  `Prior Allo SCT (Yes, No)`,
                  Blenrep,
                  `Prior treatment with any gene therapy-based therapeutic or investigational cellular therapy or BCMA targeted therapy (Yes, No)`,
                  `BCMA CAR-T on trial`, `Other BCMA trial`,
                  immuno, proteasome, dara, doubleref, tripleref, pentaref,
                  `Infection (Yes, No)`,
                  `Tocilizumab (Yes, No)`, `Steroid use (Yes, No)`, `Anakinra (Yes, No)`,
                  `ECOG > or = 2? (Yes, No)`,
                  `Renal insufficiency CrCl < 45 mL/min (Yes, No)`,
                  `LVEF <45% (Yes, No)`,
                  `ANC < 1000/uL (Yes, No)`,
                  `Hemoglobin < 8 g/dL (Yes, No)`,
                  `Platelets < 50K (Yes, No)`, 
                  `H/O or presence of Plasma cell leukemia, Waldenstrom's, POEMS, or amyloidosis (Yes, No)`,
                  `Non-secretory MM (Yes, No)`,

                  `Did patient met criteria for KarMMa1 pre-CAR-T?`
                )
                ~"categorical"),
              statistic = list(all_continuous() ~ "{median} ({min}, {max})"), 
              digits = all_continuous() ~ 1
  ) %>% 
  bold_labels() %>% add_stat_label() %>% 
  add_p() %>% 
  modify_footnote(update = everything() ~ NA) %>%
  modify_spanning_header(all_stat_cols() ~ "**Grade 2 CRS**")
```

```{r}
cart0 %>% 
  select(Age, agecat, male_sex, Sex,
         `Extramedullary Disease? (Yes, No)`, 
         `Disease status at time of referral (Relapsed vs Refractory)`,
         `Plasma Cell Leukemia? (yes, no)`,
         `Amyloid? (yes, no)`,
         POEMS,
         `Non-secretory MM (Yes, No)`, ################??????????
         `High Marrow Burden (>= 50%)`, 
         `ECOG at consult`, `ECOG at LD`, 
         `R-ISS at CAR-T infusion`, 
         cytogenetics,
         `Deletion 17p prior to CAR-T infusion (yes, no)`,
         `t(4:14) prior to CAR-T infusion (yes, no)`,
         `t(14:16) prior to CAR-T infusion (yes, no)`,
         `Bridging Therapy? (Yes, No)`,
         `Number of prior lines of therapy`,
         number_prior_lines_therapy,
         `Prior auto SCT (yes, no)`, 
         `ASCT within 12 weeks prior to apheresis (Yes, No)`,
         `Prior Allo SCT (Yes, No)`,
         Blenrep,
         `Prior treatment with any gene therapy-based therapeutic or investigational cellular therapy or BCMA targeted therapy (Yes, No)`,
         `BCMA CAR-T on trial`, `Other BCMA trial`,
         immuno, proteasome, dara, doubleref, tripleref, pentaref, 
         cart_cell_dose,
         `Infection (Yes, No)`,
         `Tocilizumab (Yes, No)`, `Steroid use (Yes, No)`, `Anakinra (Yes, No)`,
         baseline_ferritin, baseline_crp, baseline_B2M,
         `ECOG > or = 2? (Yes, No)`,
         `Renal insufficiency CrCl < 45 mL/min (Yes, No)`,
         `LVEF <45% (Yes, No)`,
         `ANC < 1000/uL (Yes, No)`,
         `Hemoglobin < 8 g/dL (Yes, No)`,
         `Platelets < 50K (Yes, No)`, 
         `H/O or presence of Plasma cell leukemia, Waldenstrom's, POEMS, or amyloidosis (Yes, No)`,
         `Non-secretory MM (Yes, No)`,
         `Did patient met criteria for KarMMa1 pre-CAR-T?`,
         gr2_ICANS
  ) %>% 
  tbl_summary(by = gr2_ICANS,
              type = list(
                c(
                  `Extramedullary Disease? (Yes, No)`, 
                  `Plasma Cell Leukemia? (yes, no)`,
                  `Amyloid? (yes, no)`,
                  POEMS,
                  `Non-secretory MM (Yes, No)`,
                  `High Marrow Burden (>= 50%)`,
                  `ECOG at consult`, `ECOG at LD`,
                  cytogenetics,
                  `Deletion 17p prior to CAR-T infusion (yes, no)`,
                  `t(4:14) prior to CAR-T infusion (yes, no)`,
                  `t(14:16) prior to CAR-T infusion (yes, no)`,
                  `Bridging Therapy? (Yes, No)`, `Prior auto SCT (yes, no)`,
                  `ASCT within 12 weeks prior to apheresis (Yes, No)`,
                  `Prior Allo SCT (Yes, No)`,
                  Blenrep,
                  `Prior treatment with any gene therapy-based therapeutic or investigational cellular therapy or BCMA targeted therapy (Yes, No)`,
                  `BCMA CAR-T on trial`, `Other BCMA trial`,
                  immuno, proteasome, dara, doubleref, tripleref, pentaref,
                  `Infection (Yes, No)`,
                  `Tocilizumab (Yes, No)`, `Steroid use (Yes, No)`, `Anakinra (Yes, No)`,
                  `ECOG > or = 2? (Yes, No)`,
                  `Renal insufficiency CrCl < 45 mL/min (Yes, No)`,
                  `LVEF <45% (Yes, No)`,
                  `ANC < 1000/uL (Yes, No)`,
                  `Hemoglobin < 8 g/dL (Yes, No)`,
                  `Platelets < 50K (Yes, No)`, 
                  `H/O or presence of Plasma cell leukemia, Waldenstrom's, POEMS, or amyloidosis (Yes, No)`,
                  `Non-secretory MM (Yes, No)`,

                  `Did patient met criteria for KarMMa1 pre-CAR-T?`
                )
                ~"categorical"),
              statistic = list(all_continuous() ~ "{median} ({min}, {max})"), 
              digits = all_continuous() ~ 1
  ) %>% 
  bold_labels() %>% add_stat_label() %>% 
  add_p() %>% 
  modify_footnote(update = everything() ~ NA) %>%
  modify_spanning_header(all_stat_cols() ~ "**Grade 2 ICANS**")
```

```{r}
cart0 %>% 
  select(Age, agecat, male_sex, Sex,
         `Extramedullary Disease? (Yes, No)`, 
         `Disease status at time of referral (Relapsed vs Refractory)`,
         `Plasma Cell Leukemia? (yes, no)`,
         `Amyloid? (yes, no)`,
         POEMS,
         `Non-secretory MM (Yes, No)`, ################??????????
         `High Marrow Burden (>= 50%)`, 
         `ECOG at consult`, `ECOG at LD`, 
         `R-ISS at CAR-T infusion`, 
         cytogenetics,
         `Deletion 17p prior to CAR-T infusion (yes, no)`,
         `t(4:14) prior to CAR-T infusion (yes, no)`,
         `t(14:16) prior to CAR-T infusion (yes, no)`,
         `Bridging Therapy? (Yes, No)`,
         `Number of prior lines of therapy`,
         number_prior_lines_therapy,
         `Prior auto SCT (yes, no)`, 
         `ASCT within 12 weeks prior to apheresis (Yes, No)`,
         `Prior Allo SCT (Yes, No)`,
         Blenrep,
         `Prior treatment with any gene therapy-based therapeutic or investigational cellular therapy or BCMA targeted therapy (Yes, No)`,
         `BCMA CAR-T on trial`, `Other BCMA trial`,
         immuno, proteasome, dara, doubleref, tripleref, pentaref, 
         cart_cell_dose,
         `Infection (Yes, No)`,
         `Tocilizumab (Yes, No)`, `Steroid use (Yes, No)`, `Anakinra (Yes, No)`,
         baseline_ferritin, baseline_crp, baseline_B2M,
         `ECOG > or = 2? (Yes, No)`,
         `Renal insufficiency CrCl < 45 mL/min (Yes, No)`,
         `LVEF <45% (Yes, No)`,
         `ANC < 1000/uL (Yes, No)`,
         `Hemoglobin < 8 g/dL (Yes, No)`,
         `Platelets < 50K (Yes, No)`, 
         `H/O or presence of Plasma cell leukemia, Waldenstrom's, POEMS, or amyloidosis (Yes, No)`,
         `Non-secretory MM (Yes, No)`,
         `Did patient met criteria for KarMMa1 pre-CAR-T?`,
         `gr3_cytopenia≥30d`
  ) %>% 
  tbl_summary(by = `gr3_cytopenia≥30d`,
              type = list(
                c(
                  `Extramedullary Disease? (Yes, No)`, 
                  `Plasma Cell Leukemia? (yes, no)`,
                  `Amyloid? (yes, no)`,
                  POEMS,
                  `Non-secretory MM (Yes, No)`,
                  `High Marrow Burden (>= 50%)`,
                  `ECOG at consult`, `ECOG at LD`,
                  cytogenetics,
                  `Deletion 17p prior to CAR-T infusion (yes, no)`,
                  `t(4:14) prior to CAR-T infusion (yes, no)`,
                  `t(14:16) prior to CAR-T infusion (yes, no)`,
                  `Bridging Therapy? (Yes, No)`, `Prior auto SCT (yes, no)`,
                  `ASCT within 12 weeks prior to apheresis (Yes, No)`,
                  `Prior Allo SCT (Yes, No)`,
                  Blenrep,
                  `Prior treatment with any gene therapy-based therapeutic or investigational cellular therapy or BCMA targeted therapy (Yes, No)`,
                  `BCMA CAR-T on trial`, `Other BCMA trial`,
                  immuno, proteasome, dara, doubleref, tripleref, pentaref,
                  `Infection (Yes, No)`,
                  `Tocilizumab (Yes, No)`, `Steroid use (Yes, No)`, `Anakinra (Yes, No)`,
                  `ECOG > or = 2? (Yes, No)`,
                  `Renal insufficiency CrCl < 45 mL/min (Yes, No)`,
                  `LVEF <45% (Yes, No)`,
                  `ANC < 1000/uL (Yes, No)`,
                  `Hemoglobin < 8 g/dL (Yes, No)`,
                  `Platelets < 50K (Yes, No)`, 
                  `H/O or presence of Plasma cell leukemia, Waldenstrom's, POEMS, or amyloidosis (Yes, No)`,
                  `Non-secretory MM (Yes, No)`,

                  `Did patient met criteria for KarMMa1 pre-CAR-T?`
                )
                ~"categorical"),
              statistic = list(all_continuous() ~ "{median} ({min}, {max})"), 
              digits = all_continuous() ~ 1
  ) %>% 
  bold_labels() %>% add_stat_label() %>% 
  add_p() %>% 
  modify_footnote(update = everything() ~ NA) %>%
  modify_spanning_header(all_stat_cols() ~ "**Grade 3 cytopenia**")
```

```{r}
cart0 %>% 
  select(Age, agecat, male_sex, Sex,
         `Extramedullary Disease? (Yes, No)`, 
         `Disease status at time of referral (Relapsed vs Refractory)`,
         `Plasma Cell Leukemia? (yes, no)`,
         `Amyloid? (yes, no)`,
         POEMS,
         `Non-secretory MM (Yes, No)`, ################??????????
         `High Marrow Burden (>= 50%)`, 
         `ECOG at consult`, `ECOG at LD`, 
         `R-ISS at CAR-T infusion`, 
         cytogenetics,
         `Deletion 17p prior to CAR-T infusion (yes, no)`,
         `t(4:14) prior to CAR-T infusion (yes, no)`,
         `t(14:16) prior to CAR-T infusion (yes, no)`,
         `Bridging Therapy? (Yes, No)`,
         `Number of prior lines of therapy`,
         number_prior_lines_therapy,
         `Prior auto SCT (yes, no)`, 
         `ASCT within 12 weeks prior to apheresis (Yes, No)`,
         `Prior Allo SCT (Yes, No)`,
         Blenrep,
         `Prior treatment with any gene therapy-based therapeutic or investigational cellular therapy or BCMA targeted therapy (Yes, No)`,
         `BCMA CAR-T on trial`, `Other BCMA trial`,
         immuno, proteasome, dara, doubleref, tripleref, pentaref, 
         cart_cell_dose,
         `Infection (Yes, No)`,
         `Tocilizumab (Yes, No)`, `Steroid use (Yes, No)`, `Anakinra (Yes, No)`,
         baseline_ferritin, baseline_crp, baseline_B2M,
         `ECOG > or = 2? (Yes, No)`,
         `Renal insufficiency CrCl < 45 mL/min (Yes, No)`,
         `LVEF <45% (Yes, No)`,
         `ANC < 1000/uL (Yes, No)`,
         `Hemoglobin < 8 g/dL (Yes, No)`,
         `Platelets < 50K (Yes, No)`, 
         `H/O or presence of Plasma cell leukemia, Waldenstrom's, POEMS, or amyloidosis (Yes, No)`,
         `Non-secretory MM (Yes, No)`,
         `Did patient met criteria for KarMMa1 pre-CAR-T?`,
         day90_response_sCRvsOthers
  ) %>% 
  tbl_summary(by = day90_response_sCRvsOthers,
              type = list(
                c(
                  `Extramedullary Disease? (Yes, No)`, 
                  `Plasma Cell Leukemia? (yes, no)`,
                  `Amyloid? (yes, no)`,
                  POEMS,
                  `Non-secretory MM (Yes, No)`,
                  `High Marrow Burden (>= 50%)`,
                  `ECOG at consult`, `ECOG at LD`,
                  cytogenetics,
                  `Deletion 17p prior to CAR-T infusion (yes, no)`,
                  `t(4:14) prior to CAR-T infusion (yes, no)`,
                  `t(14:16) prior to CAR-T infusion (yes, no)`,
                  `Bridging Therapy? (Yes, No)`, `Prior auto SCT (yes, no)`,
                  `ASCT within 12 weeks prior to apheresis (Yes, No)`,
                  `Prior Allo SCT (Yes, No)`,
                  Blenrep,
                  `Prior treatment with any gene therapy-based therapeutic or investigational cellular therapy or BCMA targeted therapy (Yes, No)`,
                  `BCMA CAR-T on trial`, `Other BCMA trial`,
                  immuno, proteasome, dara, doubleref, tripleref, pentaref,
                  `Infection (Yes, No)`,
                  `Tocilizumab (Yes, No)`, `Steroid use (Yes, No)`, `Anakinra (Yes, No)`,
                  `ECOG > or = 2? (Yes, No)`,
                  `Renal insufficiency CrCl < 45 mL/min (Yes, No)`,
                  `LVEF <45% (Yes, No)`,
                  `ANC < 1000/uL (Yes, No)`,
                  `Hemoglobin < 8 g/dL (Yes, No)`,
                  `Platelets < 50K (Yes, No)`, 
                  `H/O or presence of Plasma cell leukemia, Waldenstrom's, POEMS, or amyloidosis (Yes, No)`,
                  `Non-secretory MM (Yes, No)`,

                  `Did patient met criteria for KarMMa1 pre-CAR-T?`
                )
                ~"categorical"),
              statistic = list(all_continuous() ~ "{median} ({min}, {max})"), 
              digits = all_continuous() ~ 1
  ) %>% 
  bold_labels() %>% add_stat_label() %>% 
  add_p() %>% 
  modify_footnote(update = everything() ~ NA) %>%
  modify_spanning_header(all_stat_cols() ~ "**day90_response_sCRvsOthers**")
```

```{r}
cart0 %>% 
  select(Age, agecat, male_sex, Sex,
         `Extramedullary Disease? (Yes, No)`, 
         `Disease status at time of referral (Relapsed vs Refractory)`,
         `Plasma Cell Leukemia? (yes, no)`,
         `Amyloid? (yes, no)`,
         POEMS,
         `Non-secretory MM (Yes, No)`, ################??????????
         `High Marrow Burden (>= 50%)`, 
         `ECOG at consult`, `ECOG at LD`, 
         `R-ISS at CAR-T infusion`, 
         cytogenetics,
         `Deletion 17p prior to CAR-T infusion (yes, no)`,
         `t(4:14) prior to CAR-T infusion (yes, no)`,
         `t(14:16) prior to CAR-T infusion (yes, no)`,
         `Bridging Therapy? (Yes, No)`,
         `Number of prior lines of therapy`,
         number_prior_lines_therapy,
         `Prior auto SCT (yes, no)`, 
         `ASCT within 12 weeks prior to apheresis (Yes, No)`,
         `Prior Allo SCT (Yes, No)`,
         Blenrep,
         `Prior treatment with any gene therapy-based therapeutic or investigational cellular therapy or BCMA targeted therapy (Yes, No)`,
         `BCMA CAR-T on trial`, `Other BCMA trial`,
         immuno, proteasome, dara, doubleref, tripleref, pentaref, 
         cart_cell_dose,
         `Infection (Yes, No)`,
         `Tocilizumab (Yes, No)`, `Steroid use (Yes, No)`, `Anakinra (Yes, No)`,
         baseline_ferritin, baseline_crp, baseline_B2M,
         `ECOG > or = 2? (Yes, No)`,
         `Renal insufficiency CrCl < 45 mL/min (Yes, No)`,
         `LVEF <45% (Yes, No)`,
         `ANC < 1000/uL (Yes, No)`,
         `Hemoglobin < 8 g/dL (Yes, No)`,
         `Platelets < 50K (Yes, No)`, 
         `H/O or presence of Plasma cell leukemia, Waldenstrom's, POEMS, or amyloidosis (Yes, No)`,
         `Non-secretory MM (Yes, No)`,
         `Did patient met criteria for KarMMa1 pre-CAR-T?`,
         ORR_90days
  ) %>% 
  tbl_summary(by = ORR_90days,
              type = list(
                c(
                  `Extramedullary Disease? (Yes, No)`, 
                  `Plasma Cell Leukemia? (yes, no)`,
                  `Amyloid? (yes, no)`,
                  POEMS,
                  `Non-secretory MM (Yes, No)`,
                  `High Marrow Burden (>= 50%)`,
                  `ECOG at consult`, `ECOG at LD`,
                  cytogenetics,
                  `Deletion 17p prior to CAR-T infusion (yes, no)`,
                  `t(4:14) prior to CAR-T infusion (yes, no)`,
                  `t(14:16) prior to CAR-T infusion (yes, no)`,
                  `Bridging Therapy? (Yes, No)`, `Prior auto SCT (yes, no)`,
                  `ASCT within 12 weeks prior to apheresis (Yes, No)`,
                  `Prior Allo SCT (Yes, No)`,
                  Blenrep,
                  `Prior treatment with any gene therapy-based therapeutic or investigational cellular therapy or BCMA targeted therapy (Yes, No)`,
                  `BCMA CAR-T on trial`, `Other BCMA trial`,
                  immuno, proteasome, dara, doubleref, tripleref, pentaref,
                  `Infection (Yes, No)`,
                  `Tocilizumab (Yes, No)`, `Steroid use (Yes, No)`, `Anakinra (Yes, No)`,
                  `ECOG > or = 2? (Yes, No)`,
                  `Renal insufficiency CrCl < 45 mL/min (Yes, No)`,
                  `LVEF <45% (Yes, No)`,
                  `ANC < 1000/uL (Yes, No)`,
                  `Hemoglobin < 8 g/dL (Yes, No)`,
                  `Platelets < 50K (Yes, No)`, 
                  `H/O or presence of Plasma cell leukemia, Waldenstrom's, POEMS, or amyloidosis (Yes, No)`,
                  `Non-secretory MM (Yes, No)`,

                  `Did patient met criteria for KarMMa1 pre-CAR-T?`
                )
                ~"categorical"),
              statistic = list(all_continuous() ~ "{median} ({min}, {max})"), 
              digits = all_continuous() ~ 1
  ) %>% 
  bold_labels() %>% add_stat_label() %>% 
  add_p() %>% 
  modify_footnote(update = everything() ~ NA) %>%
  modify_spanning_header(all_stat_cols() ~ "**ORR_90days**")
```

```{r}
# cart0 %>% 
#   select(Age, agecat, male_sex, Sex,
#          `Extramedullary Disease? (Yes, No)`, 
#          `Disease status at time of referral (Relapsed vs Refractory)`,
#          `Plasma Cell Leukemia? (yes, no)`,
#          `Amyloid? (yes, no)`,
#          POEMS,
#          `Non-secretory MM (Yes, No)`, ################??????????
#          `High Marrow Burden (>= 50%)`, 
#          `ECOG at consult`, `ECOG at LD`, 
#          `R-ISS at CAR-T infusion`, 
#          cytogenetics,
#          `Deletion 17p prior to CAR-T infusion (yes, no)`,
#          `t(4:14) prior to CAR-T infusion (yes, no)`,
#          `t(14:16) prior to CAR-T infusion (yes, no)`,
#          `Bridging Therapy? (Yes, No)`,
#          `Number of prior lines of therapy`,
#          `Prior auto SCT (yes, no)`, 
#          `ASCT within 12 weeks prior to apheresis (Yes, No)`,
#          `Prior Allo SCT (Yes, No)`,
#          Blenrep,
#          `Prior treatment with any gene therapy-based therapeutic or investigational cellular therapy or BCMA targeted therapy (Yes, No)`,
#          `BCMA CAR-T on trial`, `Other BCMA trial`,
#          immuno, proteasome, dara, doubleref, tripleref, pentaref, 
#          cart_cell_dose,
#          `Infection (Yes, No)`,
#          `Tocilizumab (Yes, No)`, `Steroid use (Yes, No)`, `Anakinra (Yes, No)`,
#          baseline_ferritin, baseline_crp, baseline_B2M,
#          `ECOG > or = 2? (Yes, No)`,
#          `Renal insufficiency CrCl < 45 mL/min (Yes, No)`,
#          `LVEF <45% (Yes, No)`,
#          `ANC < 1000/uL (Yes, No)`,
#          `Hemoglobin < 8 g/dL (Yes, No)`,
#          `Platelets < 50K (Yes, No)`, 
#          `H/O or presence of Plasma cell leukemia, Waldenstrom's, POEMS, or amyloidosis (Yes, No)`,
#          `Non-secretory MM (Yes, No)`,
#          `Did patient met criteria for KarMMa1 pre-CAR-T?`,
#          `Number of prior lines of therapy`
#   ) %>% 
#   mutate(`Number of prior lines of therapy` = case_when(
#     `Number of prior lines of therapy` <= 4           ~ "≤4",
#     `Number of prior lines of therapy` > 4            ~ ">4",
#   )) %>% 
#   tbl_summary(by = `Number of prior lines of therapy`,
#               type = list(
#                 c(
#                   `Extramedullary Disease? (Yes, No)`, 
#                   `Plasma Cell Leukemia? (yes, no)`,
#                   `Amyloid? (yes, no)`,
#                   POEMS,
#                   `Non-secretory MM (Yes, No)`,
#                   `High Marrow Burden (>= 50%)`,
#                   `ECOG at consult`, `ECOG at LD`,
#                   cytogenetics,
#                   `Deletion 17p prior to CAR-T infusion (yes, no)`,
#                   `t(4:14) prior to CAR-T infusion (yes, no)`,
#                   `t(14:16) prior to CAR-T infusion (yes, no)`,
#                   `Bridging Therapy? (Yes, No)`, `Prior auto SCT (yes, no)`,
#                   `ASCT within 12 weeks prior to apheresis (Yes, No)`,
#                   `Prior Allo SCT (Yes, No)`,
#                   Blenrep,
#                   `Prior treatment with any gene therapy-based therapeutic or investigational cellular therapy or BCMA targeted therapy (Yes, No)`,
#                   `BCMA CAR-T on trial`, `Other BCMA trial`,
#                   immuno, proteasome, dara, doubleref, tripleref, pentaref,
#                   `Infection (Yes, No)`,
#                   `Tocilizumab (Yes, No)`, `Steroid use (Yes, No)`, `Anakinra (Yes, No)`,
#                   `ECOG > or = 2? (Yes, No)`,
#                   `Renal insufficiency CrCl < 45 mL/min (Yes, No)`,
#                   `LVEF <45% (Yes, No)`,
#                   `ANC < 1000/uL (Yes, No)`,
#                   `Hemoglobin < 8 g/dL (Yes, No)`,
#                   `Platelets < 50K (Yes, No)`, 
#                   `H/O or presence of Plasma cell leukemia, Waldenstrom's, POEMS, or amyloidosis (Yes, No)`,
#                   `Non-secretory MM (Yes, No)`,
#                   
#                   `Did patient met criteria for KarMMa1 pre-CAR-T?`
#                 )
#                 ~"categorical"),
#               statistic = list(all_continuous() ~ "{median} ({min}, {max})"), 
#               digits = all_continuous() ~ 1
#   ) %>% 
#   bold_labels() %>% add_stat_label() %>% 
#   add_p() %>% 
#   modify_footnote(update = everything() ~ NA) %>%
#   modify_spanning_header(all_stat_cols() ~ "**Line of therapy**")
```

## Table 4.Characteristics associated with grade ≥ 3 CRS, grade ≥ 3 neurotoxicity, grade ≥ 3 cytopenias ≥ 30 days, best response of ≥ CR at 3 months, best ORR in Multivariable Models of Idecabtagene vicleucel-treated patients. 
```{r}
cart1 <- cart0 %>% 
  mutate(day90_response_sCRvsOthers = case_when(
    day90_response_sCRvsOthers == "Others"       ~ 0,
    day90_response_sCRvsOthers == "CR or sCR"    ~ 1
  ), 
  day90_response_sCRvsOthers = factor(day90_response_sCRvsOthers)) %>% 
  mutate(ORR_90days = case_when(
    ORR_90days == "PD/SD/MR"    ~ 0,
    ORR_90days == "ORR"         ~ 1
  ), 
  ORR_90days = factor(ORR_90days))
```

### CR or better
```{r}
glm(day90_response_sCRvsOthers ~ `Prior treatment with any gene therapy-based therapeutic or investigational cellular therapy or BCMA targeted therapy (Yes, No)` + cytogenetics + `Extramedullary Disease? (Yes, No)` + cart_cell_dose +
        `ECOG at LD` + pentaref + number_prior_lines_therapy + Age,
                data = cart1,
    family = "binomial"(link='logit')) %>%
            tbl_regression(exponentiate = TRUE) %>%
            modify_spanning_header(everything() ~ "CR or sCR at 90 days vs Others")
```

### best ORR
```{r}
# clogit(ORR_90days ~ `Prior treatment with any gene therapy-based therapeutic or investigational cellular therapy or BCMA targeted therapy (Yes, No)` + cytogenetics + `Extramedullary Disease? (Yes, No)` + `Cell Dose (million cells)` +
#         `ECOG at LD` + pentaref + Age,
#                 data = cart1) #%>%
            # tbl_regression(exponentiate = TRUE) %>%
            # modify_spanning_header(everything() ~ "ORR at 90 days vs PD/SD/MR")
```

```{r}
glm(ORR_90days ~ `Prior treatment with any gene therapy-based therapeutic or investigational cellular therapy or BCMA targeted therapy (Yes, No)` + cytogenetics + `Extramedullary Disease? (Yes, No)` + cart_cell_dose +
        `ECOG at LD` + pentaref + number_prior_lines_therapy + Age,
                data = cart1,
    family = "binomial") %>%
            tbl_regression(exponentiate = TRUE) %>%
            modify_spanning_header(everything() ~ "ORR at 90 days vs PD/SD/MR")
```

### PFS
```{r}
coxph(Surv(time = cart0$days_at_pfs_from_infusion,
           event = cart0$pfs_event) ~ `Prior treatment with any gene therapy-based therapeutic or investigational cellular therapy or BCMA targeted therapy (Yes, No)` + cytogenetics + `Extramedullary Disease? (Yes, No)` + cart_cell_dose +
        `ECOG at LD` + pentaref + number_prior_lines_therapy + Age,
      data =  cart0) %>%
  tbl_regression(exponentiate = TRUE) %>%
  bold_p(t = .05) %>% add_nevent(location = "level") %>% add_n(location = "level")
```

```{r}
# tbl1 <- cart0 %>% 
#   distinct(patient_id, .keep_all = TRUE) %>% 
#   select("marrow cellularity" = `marrow cellularity baseline`, 
#          "marrow cellularity (%)" = `marrow cellularity baseline (%)`,
#          "% CD138 positive by IHC" = `marrow CD138 baseline (%)`, 
#          "Fibrosis grade" = `marrow fibrosis baseline`, 
#          "Dysplasia present" = `marrow dysplasia baseline`
#   ) %>% 
#   tbl_summary(type = list(
#     c(#`marrow cellularity baseline (%)`, `marrow CD138 baseline (%)`, 
#       `Fibrosis grade`) ~ "continuous"
#   ),
#             statistic=list(all_continuous() ~ "{median} ({min}, {max})"), 
#             digits = all_continuous() ~ 1) %>% 
#   bold_labels() %>% add_stat_label() %>% 
#   modify_footnote(update = everything() ~ NA)
# 
# tbl2 <- cart0 %>% 
#   distinct(patient_id, .keep_all = TRUE) %>% 
#   select("marrow cellularity" = `marrow cellularity day +30`, 
#          "marrow cellularity (%)" = `marrow % cellularity day +30`,
#          "% CD138 positive by IHC" = `marrow %CD138 day +30`, 
#          "Fibrosis grade" = `marrow fibrosis day +30`, 
#          "Dysplasia present" = `marrow dysplasia day +30`
#   ) %>% 
#   mutate(`marrow cellularity (%)` = as.numeric(`marrow cellularity (%)`)) %>% 
#   mutate(`% CD138 positive by IHC` = as.numeric(`% CD138 positive by IHC`)) %>% 
#   mutate(`Fibrosis grade` = as.numeric(`Fibrosis grade`)) %>% 
#   mutate(`Dysplasia present` = as.numeric(`Dysplasia present`)) %>% 
#   tbl_summary(type = list(
#     c(`marrow cellularity (%)`, `% CD138 positive by IHC`, 
#       `Fibrosis grade`) ~ "continuous"
#   ),
#             statistic=list(all_continuous() ~ "{median} ({min}, {max})"), 
#             digits = all_continuous() ~ 1) %>% 
#   bold_labels() %>% add_stat_label() %>% 
#   modify_footnote(update = everything() ~ NA)
# 
# tbl3 <- cart0 %>% 
#   distinct(patient_id, .keep_all = TRUE) %>% 
#   select("marrow cellularity" = `marrow cellularity day +90`, 
#          "marrow cellularity (%)" = `marrow % cellularity day +90`,
#          "% CD138 positive by IHC" = `marrow %CD138 day +90`, 
#          "Fibrosis grade" = `marrow fibrosis day +90`, 
#          "Dysplasia present" = `marrow dysplasia day +90`
#   ) %>% 
#   mutate(`marrow cellularity (%)` = as.numeric(`marrow cellularity (%)`)) %>% 
#   mutate(`% CD138 positive by IHC` = as.numeric(`% CD138 positive by IHC`)) %>% 
#   mutate(`Fibrosis grade` = as.numeric(`Fibrosis grade`)) %>% 
#   mutate(`Dysplasia present` = as.numeric(`Dysplasia present`)) %>% 
#   tbl_summary(type = list(
#     c(`marrow cellularity (%)`, `% CD138 positive by IHC`, 
#       `Fibrosis grade`) ~ "continuous"
#   ),
#             statistic=list(all_continuous() ~ "{median} ({min}, {max})"), 
#             digits = all_continuous() ~ 1) %>% 
#   bold_labels() %>% add_stat_label() %>% 
#   modify_footnote(update = everything() ~ NA)
# 
# tbl_merge(list(tbl1, tbl2, tbl3), tab_spanner = c("Baseline", "Day 30", "Day 90"))
```

<!-- ## Table 4. Supportive therapies and engraftment -->
<!-- Jenn -->

```{r}
# cart0 %>% 
#   distinct(patient_id, .keep_all = TRUE) %>% 
#   select(`Need for G-CSF post infus (Yes, No)`, `First Day of GCSF`, `Last Day of GCSF`,
#          `CD34 stem cell boost`, `Day of stem cell boost`, `Boost cell dose (million cells)`,
#          `Engraftment Day`,
#          `RBC transfusion within 7d post CART`, `Plt transfusion within 7d post CART`, 
#          `RBC transfusion >7d post CART`, `Plt transfusion >7d post CART`,
#          `Need for IVIG (Yes, No)`, `First day of IVIG`, `Last day of IVIG`
#   ) %>% 
#   tbl_summary(type = list(
#     c(`Day of stem cell boost`, `Boost cell dose (million cells)`,
#     `Last day of IVIG`) ~ "continuous",
#     c(`Need for G-CSF post infus (Yes, No)`,
#       `CD34 stem cell boost`,
#       `RBC transfusion within 7d post CART`, `Plt transfusion within 7d post CART`, 
#          `RBC transfusion >7d post CART`, `Plt transfusion >7d post CART`,
#       `Need for IVIG (Yes, No)`) ~ "categorical"
#   ),
#             statistic=list(all_continuous() ~ "{median} ({min}, {max})"), 
#             digits = all_continuous() ~ 1) %>% 
#   bold_labels() %>% add_stat_label() %>% 
#   modify_footnote(update = everything() ~ NA)
```

<!-- ## Table S1. Apheresis and product characteristics -->
```{r}
# cart0 %>% 
#   distinct(patient_id, .keep_all = TRUE) %>% 
#   select(TBV, `Whole blood processed (mL)`,
#          `Final collection volume (mL)`, 
#          `TBV processed`, 
#          `Access type`, `Final collection volume (mL)`,
#          `Run time (min)`,
#          `Time from apheresis to CAR-T infusion (days)`, `Total bags`,
#          `Total dose volume (mL)`, `Total cell dose (million cells)`
#          
#   ) %>% 
#   tbl_summary(type = list(c(TBV, `Total bags`) ~ "continuous"),
#             statistic=list(all_continuous() ~ "{median} ({min}, {max})"), 
#             digits = all_continuous() ~ 1) %>% 
#   bold_labels() %>% add_stat_label() %>% 
#   modify_footnote(update = everything() ~ NA)
```

<!-- ## Table S2. Cytopenias following ide-cel -->
<!-- **NAs are the patients who died (black) and test was not done (grey), they are remove from the denom (%) but not from the the N** -->

```{r}
# # tbl1 <- cart0 %>% 
# #   distinct(patient_id, .keep_all = TRUE) %>% 
# #   select("Leukopenia" = `Leukopenia_apheresis`, 
# #          "Grade3+ Leukopenia" = `gr3_Leukopenia_apheresis`, 
# #          "Neutropenia" = `Neutropenia_apheresis`, 
# #          "Grade3+ Neutropenia" = `gr3_Neutropenia_apheresis`, 
# #          "Anemia" = `Anemia_apheresis`, 
# #          "Grade3+ Anemia" = `gr3_Anemia_apheresis`, 
# #          "Thrombocytopenia" = `Thrombocytopenia_apheresis`, 
# #          "Grade3+ Thrombocytopenia" = `gr3_Thrombocytopenia_apheresis`
# #   ) %>% 
# #   tbl_summary(type = list(everything() ~ "dichotomous"),
# #             digits = all_continuous() ~ 1#,
# #             # missing = "no"
# #             ) %>% 
# #   bold_labels() %>% add_stat_label() %>% 
# #   modify_footnote(update = everything() ~ NA)
# # 
# # tbl2 <- cart0 %>% 
# #   distinct(patient_id, .keep_all = TRUE) %>% 
# #   select("Leukopenia" = `Leukopenia_-5d`, 
# #          "Grade3+ Leukopenia" = `gr3_Leukopenia_-5d`, 
# #          "Neutropenia" = `Neutropenia_-5d`, 
# #          "Grade3+ Neutropenia" = `gr3_Neutropenia_-5d`,
# #          "Anemia" = `Anemia_-5d`, 
# #          "Grade3+ Anemia" = `gr3_Anemia_-5d`, 
# #          "Thrombocytopenia" = `Thrombocytopenia_-5d`, 
# #          "Grade3+ Thrombocytopenia" = `gr3_Thrombocytopenia_-5d`
# #   ) %>% 
# #   tbl_summary(type = list(everything() ~ "dichotomous"),
# #             digits = all_continuous() ~ 1#,
# #             # missing = "no"
# #             ) %>% 
# #   bold_labels() %>% add_stat_label() %>% 
# #   modify_footnote(update = everything() ~ NA)
# # 
# # tbl3 <- cart0 %>% 
# #   distinct(patient_id, .keep_all = TRUE) %>% 
# #   select("Leukopenia" = `Leukopenia_0d`, 
# #          "Grade3+ Leukopenia" = `gr3_Leukopenia_0d`, 
# #          "Neutropenia" = `Neutropenia_0d`, 
# #          "Grade3+ Neutropenia" = `gr3_Neutropenia_0d`, 
# #          "Anemia" = `Anemia_0d`, 
# #          "Grade3+ Anemia" = `gr3_Anemia_0d`,
# #          "Thrombocytopenia" = `Thrombocytopenia_0d`, 
# #          "Grade3+ Thrombocytopenia" = `gr3_Thrombocytopenia_0d`
# #   ) %>% 
# #   tbl_summary(type = list(everything() ~ "dichotomous"),
# #             digits = all_continuous() ~ 1#,
# #             # missing = "no"
# #             ) %>% 
# #   bold_labels() %>% add_stat_label() %>% 
# #   modify_footnote(update = everything() ~ NA)
# 
# tbl4 <- cart0 %>% 
#   distinct(patient_id, .keep_all = TRUE) %>% 
#   select(#"Leukopenia" = `Leukopenia_7d`, 
#          # "Grade3+ Leukopenia" = `gr3_Leukopenia_7d`, 
#          "Neutropenia" = `Neutropenia_7d`, 
#          "Grade3+ Neutropenia" = `gr3_Neutropenia_7d`, 
#          # "Anemia" = `Anemia_7d`, 
#          # "Grade3+ Anemia" = `gr3_Anemia_7d`, 
#          "Thrombocytopenia" = `Thrombocytopenia_7d`, 
#          "Grade3+ Thrombocytopenia" = `gr3_Thrombocytopenia_7d`
#   ) %>% 
#   tbl_summary(type = list(everything() ~ "dichotomous"),
#             digits = all_continuous() ~ 1#,
#             # missing = "no"
#             ) %>% 
#   bold_labels() %>% add_stat_label() %>% 
#   modify_footnote(update = everything() ~ NA)
# 
# # tbl5 <- cart0 %>% 
# #   distinct(patient_id, .keep_all = TRUE) %>% 
# #   filter(is.na(Not_evaluable_at_14days)) %>% 
# #   select("Leukopenia" = `Leukopenia_14d`, 
# #          "Grade3+ Leukopenia" = `gr3_Leukopenia_14d`, 
# #          "Neutropenia" = `Neutropenia_14d`, 
# #          "Grade3+ Neutropenia" = `gr3_Neutropenia_14d`,
# #          "Anemia" = `Anemia_14d`, 
# #          "Grade3+ Anemia" = `gr3_Anemia_14d`,
# #          "Thrombocytopenia" = `Thrombocytopenia_14d`, 
# #          "Grade3+ Thrombocytopenia" = `gr3_Thrombocytopenia_14d` 
# #   ) %>% 
# #   tbl_summary(type = list(everything() ~ "dichotomous"),
# #             digits = all_continuous() ~ 1#,
# #             # missing = "no"
# #             ) %>% 
# #   bold_labels() %>% add_stat_label() %>% 
# #   modify_footnote(update = everything() ~ NA)
# # 
# # tbl6 <- cart0 %>% 
# #   distinct(patient_id, .keep_all = TRUE) %>% 
# #   filter(is.na(Not_evaluable_at_21days)) %>% 
# #   select("Leukopenia" = `Leukopenia_21d`,
# #          "Grade3+ Leukopenia" = `gr3_Leukopenia_21d`, 
# #          "Neutropenia" = `Neutropenia_21d`, 
# #          "Grade3+ Neutropenia" = `gr3_Neutropenia_21d`, 
# #          "Anemia" = `Anemia_21d`, 
# #          "Grade3+ Anemia" = `gr3_Anemia_21d`,
# #          "Thrombocytopenia" = `Thrombocytopenia_21d`, 
# #          "Grade3+ Thrombocytopenia" = `gr3_Thrombocytopenia_21d`
# #   ) %>% 
# #   tbl_summary(type = list(everything() ~ "dichotomous"),
# #             digits = all_continuous() ~ 1#,
# #             # missing = "no"
# #             ) %>% 
# #   bold_labels() %>% add_stat_label() %>% 
# #   modify_footnote(update = everything() ~ NA)
# 
# tbl7 <- cart0 %>% 
#   distinct(patient_id, .keep_all = TRUE) %>% 
#   filter(is.na(Not_evaluable_at_30days)) %>% 
#   select(#"Leukopenia" = `Leukopenia_30d`, 
#          # "Grade3+ Leukopenia" = `gr3_Leukopenia_30d`,
#          "Neutropenia" = `Neutropenia_30d`, 
#          "Grade3+ Neutropenia" = `gr3_Neutropenia_30d`,
#          # "Anemia" = `Anemia_30d`, 
#          # "Grade3+ Anemia" = `gr3_Anemia_30d`, 
#          "Thrombocytopenia" = `Thrombocytopenia_30d`,
#          "Grade3+ Thrombocytopenia" = `gr3_Thrombocytopenia_30d` 
#   ) %>% 
#   tbl_summary(type = list(everything() ~ "dichotomous"),
#             digits = all_continuous() ~ 1#,
#             # missing = "no"
#             ) %>% 
#   bold_labels() %>% add_stat_label() %>% 
#   modify_footnote(update = everything() ~ NA)
# 
# tbl8 <- cart0 %>% 
#   distinct(patient_id, .keep_all = TRUE) %>% 
#   filter(is.na(Not_evaluable_at_90days)) %>% 
#   select(#"Leukopenia" = `Leukopenia_90d`, 
#          # "Grade3+ Leukopenia" = `gr3_Leukopenia_90d`,
#          "Neutropenia" = `Neutropenia_90d`, 
#          "Grade3+ Neutropenia" = `gr3_Neutropenia_90d`,
#          # "Anemia" = `Anemia_90d`,
#          # "Grade3+ Anemia" = `gr3_Anemia_90d`,
#          "Thrombocytopenia" = `Thrombocytopenia_90d`, 
#          "Grade3+ Thrombocytopenia" = `gr3_Thrombocytopenia_90d`
#   ) %>% 
#   tbl_summary(type = list(everything() ~ "dichotomous"),
#             digits = all_continuous() ~ 1#,
#             # missing = "no"
#             ) %>% 
#   bold_labels() %>% add_stat_label() %>% 
#   modify_footnote(update = everything() ~ NA)
# 
# tbl_merge(list(tbl1, tbl2, tbl3, tbl4, tbl5, tbl6, tbl7, tbl8), 
#           tab_spanner = c("Apheresis", "Day -5",	"Day 0",	"Day 7",	"Day 14",	"Day 21",	"Day 30",	"Day 90")) %>% as_gt() %>%
#   gt::tab_source_note(gt::md("**Not evaluable patients are removed from the denominator <br>
#                              1 patient has missing 90 days values**"))
```

### Figures

## Figure 1. Response rate bar graph for day 30, day 90, Best ORR
```{r}
cart0 %>% 
  select(day30_response_group, day90_response_group, best_response) %>% 
  pivot_longer(cols = everything()) %>% 
  filter(!is.na(value)) %>% 
  mutate(value = factor(value, levels = c("CR or sCR, MRD-",
                                        "CR or sCR, MRD+",
                                        "CR or sCR, MRD unknown",
                                        "VGPR", "PR", "Others"))) %>% 
  group_by(name, value) %>% 
  summarise(count=n()) %>% 
  mutate(percent=(count/sum(count)*100)
  ) %>% 
  filter(str_detect(value, "CR") | str_detect(value, "PR")) %>% 
  mutate(name = factor(name, levels = c("day30_response_group", "day90_response_group", "best_response"))) %>% 
  group_by(name) %>% 
  mutate(n = sum(count)) %>% 
  ungroup() %>% 
  mutate(name =  paste0(name,'\nN = ',n)) %>% 
  ggplot(aes(x= name, y= percent, fill= value))+
  geom_bar(stat = "identity")+ 
  scale_fill_manual(values = c("lightsteelblue2","steelblue2", "steelblue","thistle", "darkgray"))+
  geom_text(aes(label = paste0("n=", count)), size = 3, position = position_stack(vjust = 0.5))+
  ylim(0,100)+
  labs(x= NULL)+
  theme_classic()

cart0 %>% 
  select(day30_response_group, day90_response_group, best_response) %>% 
  pivot_longer(cols = everything()) %>% 
  filter(!is.na(value)) %>% 
  mutate(value = factor(value, levels = c("CR or sCR, MRD-",
                                        "CR or sCR, MRD+",
                                        "CR or sCR, MRD unknown",
                                        "VGPR", "PR", "Others"))) %>% 
  group_by(name, value) %>% 
  summarise(count=n()) %>% 
  mutate(percent=(count/sum(count)*100)
  ) %>% 
  filter(str_detect(value, "CR") | str_detect(value, "PR")) %>% 
  mutate(name = factor(name, levels = c("day30_response_group", "day90_response_group", "best_response"))) %>% 
  group_by(name) %>% 
  mutate(n = sum(count)) %>% 
  ungroup() %>% 
  mutate(name =  paste0(name,'\nN = ',n)) %>% 
  ggplot(aes(x= name, y= percent, fill= value))+
  geom_bar(stat = "identity")+ 
  scale_fill_manual(values = c("lightsteelblue2","steelblue2", "steelblue","thistle", "darkgray"))+
  ylim(0,100)+
  labs(x= NULL)+
  theme_classic()
```

Panel B: PFS Kaplan Meier (from CAR-T infusion)

```{r, fig.width= 6, fig.height=4}
myplot <-
  ggsurvplot(survfit(Surv(days_at_pfs_from_infusion, pfs_event) ~ 1, 
                   data = cart0),
           title = "PFS from date of infusion", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           legend = "none",
           font.tickslab = c(18, "bold", "black"), 
           size = 1.5,
           xlim = c(0, 12),
           pval=TRUE, 
           risk.table = FALSE,

           ylab = "PFS probability",
           xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))
myplot$plot <- myplot$plot + 
  scale_x_continuous(breaks = c(0, 3, 6, 9, 12))
myplot$table <- myplot$table + 
  scale_x_continuous(breaks = c(0, 3, 6, 9, 12))
myplot

survfit(Surv(days_at_pfs_from_infusion, pfs_event) ~ 1, 
                   data = cart0)
```

Panel C: OS Kaplan Meier (from CAR-T infusion)
```{r, fig.width= 6, fig.height=4}
myplot <- ggsurvplot(survfit(Surv(days_at_os_from_infusion, os_event) ~ 1, 
                   data = cart0),
           title = "OS from date of infusion", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           legend = "none",
           font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           risk.table = FALSE,

           ylab = "OS probability",
           xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 2))
myplot$plot <- myplot$plot + 
  scale_x_continuous(breaks = c(0, 3, 6, 9, 12))
myplot$table <- myplot$table + 
  scale_x_continuous(breaks = c(0, 3, 6, 9, 12))
myplot

survfit(Surv(days_at_os_from_infusion, os_event) ~ 1, 
                   data = cart0)
```

Panel D: Duration of response graph
Waiting Tuesday
```{r}

```

## Figure 2. Consort diagram

## Table 5. Patients’ response to Idecabtagene Vicleucel

```{r}
tbl1 <- cart0 %>%
  select(overall_response,
         best_response) %>%
  tbl_summary(
    label = list(overall_response ~ "overall response within 90 days", 
                 best_response ~ "best response within 90 days"),
    digits = all_continuous() ~ 1,
    missing = "no") %>%
  bold_labels() %>% add_stat_label()

# tbl2 <- cart0 %>%
#   filter(str_detect(best_response, "CR")) %>%
#   select(best_MRD_response) %>%
#   tbl_summary(#type = list(c(best_MRD_response) ~ "categorical"),
#               # label = list(best_MRD_response ~ "best_MRD_response within CR/sCR"),
#               digits = all_continuous() ~ 1,
#               missing = "no") %>%
#   bold_labels() %>% add_stat_label()

tbl3 <- cart0 %>%
  select(months_at_best_response,
         months_at_best_CRresponse) %>%
  tbl_summary(
    type = list(c(months_at_best_response, 
                  months_at_best_CRresponse) ~ "continuous"),
    label = list(months_at_best_response ~ "months at best response within 90 days",
                 months_at_best_CRresponse ~ "months at best CR response within 90 days"),
    digits = all_continuous() ~ 1,
    missing = "no") %>%
  bold_labels() %>% add_stat_label()

tbl_stack(list(tbl1, #tbl2, 
               tbl3)) %>% as_gt()

tbl4 <- cart0 %>%
  filter(str_detect(best_response, "CR")) %>%
  select(best_MRD_response) %>%
  tbl_summary(type = list(c(best_MRD_response) ~ "categorical"),
              label = list(best_MRD_response ~ "best MRD CR response within 90 days"),
              digits = all_continuous() ~ 1,
              missing = "no") %>%
  bold_labels() %>% add_stat_label()

# tbl5 <- cart0 %>%
#   distinct(patient_id, .keep_all = TRUE) %>%
#   select(#MRDneg_90days,
#          ORR_90days,
#          day90_response,
#          day90_response_group) %>%
#   tbl_summary(#type = list(
#     # c(MRDneg_90days) ~ "categorical"),
#     digits = all_continuous() ~ 1,
#     missing = "no") %>%
#   bold_labels() %>% add_stat_label()
# 
# 
# 
# tbl7 <- cart0 %>%
#   distinct(patient_id, .keep_all = TRUE) %>%
#   select(best_response) %>%
#   tbl_summary(#type = list(c(MRDneg_90days) ~ "categorical"),
#               label = list(best_response ~ "best response in the first 90 days"),
#               digits = all_continuous() ~ 1,
#               missing = "no") %>%
#   bold_labels() %>% add_stat_label()
# 
# tbl8 <- cart0 %>%
#   distinct(patient_id, .keep_all = TRUE) %>%
#   filter(best_response == "CR or sCR") %>%
#   select(MRDneg_in_best_response) %>%
#   tbl_summary(type = list(c(MRDneg_in_best_response) ~ "categorical"),
#               label = list(MRDneg_in_best_response ~ "MRDneg_in_best_response within CR/sCR"),
#               statistic=list(all_continuous() ~ "{median} ({min}, {max})"),
#               digits = all_continuous() ~ 1,
#               missing = "no") %>%
#   bold_labels() %>% add_stat_label()
```

## Figure 3: ORR (A) vs. PFS (B) stratified by KarMMa exclusion criteria.

```{r}
cart0 <- cart0 %>% 
  rename(Did_patient_met_criteria_for_KarMMa1_pre_CAR_T = 
           `Did patient met criteria for KarMMa1 pre-CAR-T?`) %>% 
  mutate(Overall = "Overall")
```

```{r, fig.width= 6, fig.height=4}
myplot <- ggsurvplot(survfit(Surv(days_at_pfs_from_infusion, pfs_event) ~ Did_patient_met_criteria_for_KarMMa1_pre_CAR_T, 
                   data = cart0),
           title = "PFS from date of infusion", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,
           xlim = c(0, 12),
           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           risk.table = FALSE,

           ylab = "PFS probability",
           xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 1))
myplot$plot <- myplot$plot + 
  scale_x_continuous(breaks = c(0, 3, 6, 9, 12))
myplot$table <- myplot$table + 
  scale_x_continuous(breaks = c(0, 3, 6, 9, 12))
myplot

survfit(Surv(days_at_pfs_from_infusion, pfs_event) ~ Did_patient_met_criteria_for_KarMMa1_pre_CAR_T, 
                   data = cart0)
```

Panel C: OS Kaplan Meier (from CAR-T infusion)
```{r, fig.width= 6, fig.height=4}
myplot <- ggsurvplot(survfit(Surv(days_at_os_from_infusion, os_event) ~ Did_patient_met_criteria_for_KarMMa1_pre_CAR_T, 
                   data = cart0),
           title = "OS from date of infusion", 
           font.main = c(24, "bold", "black"), 
           font.x = c(20, "bold", "black"), font.y = c(20, "bold", "black"), 
           font.legend = c(14, "bold", "black"), font.tickslab = c(18, "bold", "black"), 
           size = 1.5,

           pval=TRUE, 
           # legend.labs=c(""),
           # palette = c("black", "green","blue", "red"), 
           risk.table = FALSE,

           ylab = "OS probability",
           xlab = "Time (in months)" 
) + guides(colour = guide_legend(ncol = 1))
myplot$plot <- myplot$plot + 
  scale_x_continuous(breaks = c(0, 3, 6, 9, 12))
myplot$table <- myplot$table + 
  scale_x_continuous(breaks = c(0, 3, 6, 9, 12))
myplot

survfit(Surv(days_at_os_from_infusion, os_event) ~ Did_patient_met_criteria_for_KarMMa1_pre_CAR_T, 
                   data = cart0)
```

## Table 6. Cause of death in patients who received Idecabtagene vicleucel
```{r}
cart0 %>%
  filter(!is.na(`Date of Death`)) %>% 
  mutate(`Cardiomyopathy/Acute systolic heart failure (EF 15-25% post CAR-T) (unknown viral illness prior to CAR-T that improved pre-infusion)` = case_when(
    str_detect(`Cause of death (myeloma, toxicitiy,other)`, "Cardiomyopathy")      ~ "YES",
    TRUE      ~ "NO"
  )) %>% 
  mutate(Myeloma = case_when(
    str_detect(`Cause of death (myeloma, toxicitiy,other)`, "Myeloma")      ~ "YES",
    TRUE      ~ "NO"
  )) %>% 
  mutate(`COVID-19` = case_when(
    str_detect(`Cause of death (myeloma, toxicitiy,other)`, "COVID") &
      !str_detect(`Cause of death (myeloma, toxicitiy,other)`, "Myeloma")    ~ "YES",
    TRUE      ~ "NO"
  )) %>% 
  mutate(Toxicity = case_when(
    str_detect(`Cause of death (myeloma, toxicitiy,other)`, "oxicity")      ~ "YES",
    TRUE      ~ "NO"
  )) %>% 
  mutate(HLH = case_when(
    str_detect(`Cause of death (myeloma, toxicitiy,other)`, "HLH")      ~ "YES",
    TRUE      ~ "NO"
  )) %>% 
  select(`Cardiomyopathy/Acute systolic heart failure (EF 15-25% post CAR-T) (unknown viral illness prior to CAR-T that improved pre-infusion)`, `COVID-19`, HLH, Myeloma, Toxicity) %>%
  tbl_summary() %>%
  bold_labels() %>% add_stat_label()
```
1 patients had HLH and Toxicity “Toxicity - grade 5 CRS/HLH with superimposed infection”.  
1 patient in Myeloma is "Myeloma (hypoxic resp failure with multiple lung masses, pleural fluid + PCs, strep due to cytopenias, COVID-19)".  
1 patient is "And Myeloma, CNS".

## Table 7. Multiple Myeloma Efficacy and Toxicity Outcomes following infusion of SOC Idecabtagene vicleucel or in KarMMa (Variables to report on ORR, ≥ CR rate, 3 month PFS rate, any CRS, grade ≥ 3 CRS, any neurotoxicity, grade ≥ 3 neurotoxicity (see lymphoma EXAMPLE table below)

```{r}
tbl0 <- cart0 %>%
  mutate(cohort = "Overall") %>% 
  bind_rows(., cart0 %>% 
              rename(cohort = Did_patient_met_criteria_for_KarMMa1_pre_CAR_T)) %>% 
  mutate(cohort = case_when(
    cohort == "YES"   ~ "Met criteria for KarMMa1 pre-CAR-T",
    cohort == "NO"   ~ "NOT et criteria for KarMMa1 pre-CAR-T",
    TRUE                 ~ cohort
  ),
  cohort = factor(cohort, levels = c("Overall",
                                     "Met criteria for KarMMa1 pre-CAR-T",
                                     "NOT et criteria for KarMMa1 pre-CAR-T"))
  ) %>% 
  mutate(Total = "No of patients") %>% 
  select(Total, cohort) %>%

  tbl_summary(by= Total,
              # type = list(c(Did_patient_met_criteria_for_KarMMa1_pre_CAR_T) ~ "categorical"),
              statistic = all_categorical() ~ "{n}",
              # percent = "row"
              ) %>%
  bold_labels() %>% #add_stat_label() %>% 
  modify_header(all_stat_cols() ~ "**{level}**") %>% 
  modify_footnote(update = everything() ~ NA)

tbl1 <- cart0 %>%
  mutate(cohort = "Overall") %>% 
  bind_rows(., cart0 %>% 
              rename(cohort = Did_patient_met_criteria_for_KarMMa1_pre_CAR_T)) %>% 
  mutate(cohort = case_when(
    cohort == "YES"   ~ "Met criteria for KarMMa1 pre-CAR-T",
    cohort == "NO"   ~ "NOT et criteria for KarMMa1 pre-CAR-T",
    TRUE                 ~ cohort
  ),
  cohort = factor(cohort, levels = c("Overall",
                                     "Met criteria for KarMMa1 pre-CAR-T",
                                     "NOT et criteria for KarMMa1 pre-CAR-T"))) %>% 
  select(ORR_30days, cohort) %>%
  tbl_summary(by= ORR_30days,
              # type = list(c(Did_patient_met_criteria_for_KarMMa1_pre_CAR_T) ~ "categorical"),
              statistic = all_categorical() ~ "{p}%",
              percent = "row") %>%
  bold_labels() %>% add_stat_label() %>% 
  modify_header(all_stat_cols() ~ "**{level}**")

tbl2 <- cart0 %>%
  mutate(cohort = "Overall") %>% 
  bind_rows(., cart0 %>% 
              rename(cohort = Did_patient_met_criteria_for_KarMMa1_pre_CAR_T)) %>% 
  mutate(cohort = case_when(
    cohort == "YES"   ~ "Met criteria for KarMMa1 pre-CAR-T",
    cohort == "NO"   ~ "NOT et criteria for KarMMa1 pre-CAR-T",
    TRUE                 ~ cohort
  ),
  cohort = factor(cohort, levels = c("Overall",
                                     "Met criteria for KarMMa1 pre-CAR-T",
                                     "NOT et criteria for KarMMa1 pre-CAR-T"))) %>% 
  select(day30_response_sCRvsOthers, cohort) %>%
  tbl_summary(by= day30_response_sCRvsOthers,
              # type = list(c(Did_patient_met_criteria_for_KarMMa1_pre_CAR_T) ~ "categorical"),
              statistic = all_categorical() ~ "{p}%",
              percent = "row") %>%
  bold_labels() %>% add_stat_label() %>% 
  modify_header(all_stat_cols() ~ "**{level}**")

tbl3 <- cart0 %>%
  mutate(cohort = "Overall") %>% 
  bind_rows(., cart0 %>% 
              rename(cohort = Did_patient_met_criteria_for_KarMMa1_pre_CAR_T)) %>% 
  mutate(cohort = case_when(
    cohort == "YES"   ~ "Met criteria for KarMMa1 pre-CAR-T",
    cohort == "NO"   ~ "NOT et criteria for KarMMa1 pre-CAR-T",
    TRUE                 ~ cohort
  ),
  cohort = factor(cohort, levels = c("Overall",
                                     "Met criteria for KarMMa1 pre-CAR-T",
                                     "NOT et criteria for KarMMa1 pre-CAR-T"))) %>% 
  select(CRS_any, cohort) %>%
  tbl_summary(by= CRS_any,
              # type = list(c(Did_patient_met_criteria_for_KarMMa1_pre_CAR_T) ~ "categorical"),
              statistic = all_categorical() ~ "{p}%",
              percent = "row") %>%
  bold_labels() %>% add_stat_label() %>% 
  modify_header(all_stat_cols() ~ "**{level}**")

tbl4 <- cart0 %>%
  mutate(cohort = "Overall") %>% 
  bind_rows(., cart0 %>% 
              rename(cohort = Did_patient_met_criteria_for_KarMMa1_pre_CAR_T)) %>% 
  mutate(cohort = case_when(
    cohort == "YES"   ~ "Met criteria for KarMMa1 pre-CAR-T",
    cohort == "NO"   ~ "NOT et criteria for KarMMa1 pre-CAR-T",
    TRUE                 ~ cohort
  ),
  cohort = factor(cohort, levels = c("Overall",
                                     "Met criteria for KarMMa1 pre-CAR-T",
                                     "NOT et criteria for KarMMa1 pre-CAR-T"))) %>% 
  select(gr2_CRS, cohort) %>%
  tbl_summary(by= gr2_CRS,
              # type = list(c(Did_patient_met_criteria_for_KarMMa1_pre_CAR_T) ~ "categorical"),
              statistic = all_categorical() ~ "{p}%",
              percent = "row") %>%
  bold_labels() %>% add_stat_label() %>% 
  modify_header(all_stat_cols() ~ "**{level}**")

tbl5 <- cart0 %>%
  mutate(cohort = "Overall") %>% 
  bind_rows(., cart0 %>% 
              rename(cohort = Did_patient_met_criteria_for_KarMMa1_pre_CAR_T)) %>% 
  mutate(cohort = case_when(
    cohort == "YES"   ~ "Met criteria for KarMMa1 pre-CAR-T",
    cohort == "NO"   ~ "NOT et criteria for KarMMa1 pre-CAR-T",
    TRUE                 ~ cohort
  ),
  cohort = factor(cohort, levels = c("Overall",
                                     "Met criteria for KarMMa1 pre-CAR-T",
                                     "NOT et criteria for KarMMa1 pre-CAR-T"))) %>% 
  select(ICANS_any, cohort) %>%
  tbl_summary(by= ICANS_any,
              # type = list(c(Did_patient_met_criteria_for_KarMMa1_pre_CAR_T) ~ "categorical"),
              statistic = all_categorical() ~ "{p}%",
              percent = "row") %>%
  bold_labels() %>% add_stat_label() %>% 
  modify_header(all_stat_cols() ~ "**{level}**")

tbl6 <- cart0 %>%
  mutate(cohort = "Overall") %>% 
  bind_rows(., cart0 %>% 
              rename(cohort = Did_patient_met_criteria_for_KarMMa1_pre_CAR_T)) %>% 
  mutate(cohort = case_when(
    cohort == "YES"   ~ "Met criteria for KarMMa1 pre-CAR-T",
    cohort == "NO"   ~ "NOT et criteria for KarMMa1 pre-CAR-T",
    TRUE                 ~ cohort
  ),
  cohort = factor(cohort, levels = c("Overall",
                                     "Met criteria for KarMMa1 pre-CAR-T",
                                     "NOT et criteria for KarMMa1 pre-CAR-T"))) %>% 
  select(gr2_ICANS, cohort) %>%
  tbl_summary(by= gr2_ICANS,
              # type = list(c(Did_patient_met_criteria_for_KarMMa1_pre_CAR_T) ~ "categorical"),
              statistic = all_categorical() ~ "{p}%",
              percent = "row") %>%
  bold_labels() %>% add_stat_label() %>% 
  modify_header(all_stat_cols() ~ "**{level}**")

tbl_merge(list(tbl0, tbl1, tbl2, tbl3, tbl4, tbl5, tbl6), 
          tab_spanner = c("", "ORR at 30 days", "sCR/CR vs Others at 30 days",
                          "CRS_any", "CRS_gradecat", "ICANS_any", "ICANS_gradecat"))

# cart0 %>%
#   select(day30_response_sCRvsOthers, CRS_any, CRS_gradecat, ICANS_any, ICANS_gradecat, Did_patient_met_criteria_for_KarMMa1_pre_CAR_T) %>%
#   tbl_summary() %>%
#   bold_labels() %>% add_stat_label()

# tbl1 <- cart0 %>%
#   select(ORR_30days, day30_response_sCRvsOthers) %>%
#   tbl_summary() %>%
#   bold_labels() %>% add_stat_label() %>%
#   modify_spanning_header(all_stat_cols() ~ "**Overall**")
# 
# tbl2 <- cart0 %>%
#   filter(Did_patient_met_criteria_for_KarMMa1_pre_CAR_T == "YES") %>% 
#   select(ORR_30days, day30_response_sCRvsOthers) %>%
#   tbl_summary() %>%
#   bold_labels() %>% add_stat_label() %>%
#   modify_spanning_header(all_stat_cols() ~ "**Met criteria for KarMMa1 pre-CAR-T**")
# 
# tbl3 <- cart0 %>%
#   filter(Did_patient_met_criteria_for_KarMMa1_pre_CAR_T == "NO") %>% 
#   select(ORR_30days, day30_response_sCRvsOthers) %>%
#   tbl_summary() %>%
#   bold_labels() %>% add_stat_label() %>%
#   modify_spanning_header(all_stat_cols() ~ "**NOT met criteria for KarMMa1 pre-CAR-T**")
# 
# tbl_merge(list(tbl1, tbl2, 
#                tbl3), 
#           tab_spanner = c("**Overall**", 
#                           "**Met criteria for KarMMa1 pre-CAR-T**", 
#                           "**NOT met criteria for KarMMa1 pre-CAR-T**")) %>% 
#   as_gt()

# Estimating median survival time
# survfit(Surv(days_at_pfs_from_infusion, pfs_event) ~ 1, 
#                    data = cart0)
tbl1 <- tbl_survfit(
  cart0,
  y = Surv(days_at_pfs_from_infusion, pfs_event),
  include = c(Overall, Did_patient_met_criteria_for_KarMMa1_pre_CAR_T#, 
              # CRS_any, CRS_gradecat, ICANS_any, ICANS_gradecat
              ),
  probs = 0.5,
  label_header = "**Median PFS**"
)
# 3 month pfs rate

tbl2 <- list(
    survfit(Surv(days_at_pfs_from_infusion, pfs_event) ~ 1, 
                   data = cart0),
    survfit(Surv(days_at_pfs_from_infusion, pfs_event) ~ Did_patient_met_criteria_for_KarMMa1_pre_CAR_T, 
                   data = cart0)#,
    # survfit(Surv(days_at_pfs_from_infusion, pfs_event) ~ CRS_any, 
    #                data = cart0),
    # survfit(Surv(days_at_pfs_from_infusion, pfs_event) ~ CRS_gradecat, 
    #                data = cart0),
    # survfit(Surv(days_at_pfs_from_infusion, pfs_event) ~ ICANS_any, 
    #                data = cart0),
    # survfit(Surv(days_at_pfs_from_infusion, pfs_event) ~ ICANS_gradecat, 
    #                data = cart0)
  ) %>%
  tbl_survfit(times = c(3, 6),
              label_header = "**PFS Rate at {time} Months**")

tbl_merge(list(tbl1, tbl2))
```