### 2024-01-12 cilta_cel/Manuscript analyses.Rmd
Update code, add tables and plots

---
title: "Manuscript analyses"
author: "Christelle Colin-Leitzinger"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: united
    highlight: pygments
    df_print: paged
editor_options: 
  chunk_output_type: console
---

<style type= "text/css">

. figure {
   margin-top: 25px;
   margin-bottom: 100px;
}

table {
    margin-top: 10px;
    margin-bottom: 25px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      fig.align='center'
                      )
```

```{r library}
library(here)
library(tidyverse)
library(ggplot2)
library(gtsummary)
library(labelled)
library(DescTools)
library(patchwork)
library(ggbreak)
# library(survival)
library(survminer)
theme_set(theme_classic())
theme_gtsummary_compact()
```

```{r load}
# path <- fs::path("","Volumes","Peres_Research", "Myeloma", "Cytopenias")

Characteristics <-
  readxl::read_xlsx(paste0(here(), "/cilta_cel/Carvykti cellularity_11202023.xlsx"),
                    sheet = "CHARACTERISTICS AND OUTCOMES ", skip = 1)
Counts <-
  readxl::read_xlsx(paste0(here(), "/cilta_cel/Carvykti cellularity_11202023.xlsx"),
                    sheet = "COUNTS")
Infections <-
  readxl::read_xlsx(paste0(here(), "/cilta_cel/Carvykti cellularity_11202023.xlsx"),
                    sheet = "INFECTIONS")
Support <-
  readxl::read_xlsx(paste0(here(), "/cilta_cel/Carvykti cellularity_11202023.xlsx"),
                    sheet = "SUPPORT")
Pneumo_titers <-
  readxl::read_xlsx(paste0(here(), "/cilta_cel/Carvykti cellularity_11202023.xlsx"),
                    sheet = "BACTERIAL AND VIRAL TITERS")
Apheresis <- 
  readxl::read_xlsx(paste0(here(), "/cilta_cel/Carvykti cellularity_11202023.xlsx"),
                    sheet = "APHERESIS")
id_to_modify_neurotox <- 
  read_csv(
    paste0(here(), 
           "/cilta_cel/patient id who had icans and delayed neurotox - need to change neurotox type to PRES.csv"))
```

```{r join}
cart <- full_join(Characteristics, Counts, by = "PT Identifier") %>% 
  full_join(., Infections, by = "PT Identifier") %>% 
  full_join(., Support, by = "PT Identifier") %>% 
  full_join(., Pneumo_titers, by = "PT Identifier") %>% 
  # select(-`Date of leukapheresis.y`, `Date of leukapheresis` = `Date of leukapheresis.x`) %>% 
  full_join(., Apheresis, by = "PT Identifier") %>% 
  janitor::clean_names()
```

```{r exploratory function}
# check_data <- function(data){
#   
#   for(i in 1:length(colnames(data))) {
#     
#     if(class(data[[i]]) == "factor" | class(data[[i]]) == "character") {
#       
#       # print(data[i])
#       print(colnames(data[i]))
#       print(table(data[[i]]))
#     }
#     
#   }
# }
# 
# check_data(cart[2:50])
# check_data(cart[51:100])
# check_data(cart[101:150])
# check_data(cart[151:200]) # has < : "CRP Pre-Leukapheresis", 
# check_data(cart[201:250]) # has < :"CRP day +21",  "CRP day +30", 
# # has > :  "CRP day +90"
# check_data(cart[251:300]) # has < : "CRP day +180"
# check_data(cart[1:44, 301:353]) # need to see patient in Moffitt
```

```{r data cleaning}
cart <- cart %>% 
  # General ----
  # Fix letter case and missing
  mutate(across(where(is.character), ~ str_replace(., "NE|ND|N/A|Not done", NA_character_))) %>%
  mutate(across(c(where(is.character),
                  -c("pt_identifier",
                     "r_iss_at_car_t_infusion")), 
                ~ str_to_sentence(.) )) %>%
  mutate(across(c(contains("vgpr"), "response_to_bridging_chemotherapy"
                  ), 
                ~ str_to_upper(.) )) %>%

  mutate(across(where(is.character), ~ str_replace(., "Unknown|unknown", NA_character_))) %>%
  # Create site variable
  mutate(site = case_when(
    str_detect(pt_identifier, "MCC")               ~ "Moffitt",
    str_detect(pt_identifier, "CCT")               ~ "Stanford",
    str_detect(pt_identifier, "KUMC")              ~ "KUMC",
    str_detect(pt_identifier, "MDA")               ~ "MD Anderson?",
    str_detect(pt_identifier, "Utah")              ~ "Utah"
  )) %>%
  # Suboptimal is NA
  mutate(across(c(#"bm_plasma_cell_percent_at_pre_ld",
                  "high_marrow_burden_50_percent",
                  # contains("normo_hypo_hyper"),
                  contains("fibrosis"),
                  contains("dysplasia"),
                  "b2_microglobulin_day_180"), 
                ~ str_replace(., "Suboptimal", NA_character_))) %>%
  # Are numerical - also remove "Suboptimal"
  mutate(across(c(starts_with("marrow_fibrosis"),
                  starts_with("ldh_")
                  ), 
                ~ as.numeric(.))) %>%
  mutate(across(c("response_to_bridging_chemotherapy"), 
                ~ str_replace(., "Not evaluable", NA_character_))) %>% 
  # Convert 0/1 to No/Yes
  mutate_at(c("infection_yes_1_no_0",
              "severe_infection_need_for_iv_antibiotics_and_or_hospitalization_yes_1_no_0",
              # "bm_plasma_cell_percent_at_pre_ld",
              # "high_marrow_burden_50_percent",
              # "ECOG > or = 2? (Yes, No)",
              # deletion_17p_prior_to_car_t_infusion_yes_no,
              # "t_4_14_prior_to_car_t_infusion_yes_no",
              # "t_14_16_prior_to_car_t_infusion_yes_no",
              # "steroid_use_yes_no", "tocilizumab_yes_no", "anakinra_yes_no",
              # "Bridging Therapy (Yes=1, No=0)",
              # "Received alkylating therapy with bridging (Yes=1, No=0)",
              "need_for_g_csf_post_infus_yes_1_no_0",
              "need_for_ivig_yes_1_no_0",
              "cd34_stem_cell_boost_yes_1_no_0",
              "rbc_transfusion_within_30d_post_cart_yes_1_no_0", "plt_transfusion_within_30d_post_cart_yes_1_no_0", 
              "rbc_transfusion_30d_post_cart_yes_1_no_0", "plt_transfusion_30d_post_cart_yes_1_no_0",
              "tpo_agonist_post_cart_yes_1_no_0"), 
            ~ case_when(
              . == 1                                              ~ "Yes",
              . == 0                                              ~ "No",
              TRUE                                                ~ NA_character_
            )) %>% 
  # Marrow ----
  mutate(marrow_cellularity_baseline_normo_hypo_hyper = case_when(
    marrow_cellularity_baseline_normo_hypo_hyper == "\"patchy\""                 ~ "hyper",
    TRUE                                                          ~ marrow_cellularity_baseline_normo_hypo_hyper
  )) %>%
  mutate(marrow_cellularity_day_30_normo_hypo_hyper = case_when(
    marrow_cellularity_day_30_normo_hypo_hyper == "aplastic"                    ~ "hypo",
    TRUE                                                          ~ marrow_cellularity_day_30_normo_hypo_hyper
  )) %>%
  # Separate variable cellularity
  mutate(across(c(starts_with("marrow_cellularity_") & ends_with("normo_hypo_hyper")), ~ case_when(
    . == "Variable"                         ~ "Yes"
  ), .names = "is_variable_{.col}")) %>% 
  mutate(across(c(starts_with("marrow_cellularity_") & ends_with("normo_hypo_hyper")), ~ case_when(
    . == "Pauci"                            ~ "Hypo",
    . == "Variable"                         ~ NA_character_,
    TRUE                                    ~ .
  ))) %>%

  mutate(across(c(starts_with("marrow_dysplasia_") & ends_with("yes_no")), ~ case_when(
    . == "0"                                ~ "No",
    . == "1"                                ~ "Yes",
    TRUE                                    ~ .
  )))

# Fix the "<" and ">" to take the min or max or each column including numbers next to the signs----
# Also replace "Suboptimal" by NA
vec_col <- c("bm_plasma_cell_percent_at_pre_ld",

             "marrow_cellularity_baseline_percent",
             "marrow_percent_cellularity_day_30",
             "marrow_percent_cellularity_day_90",
             "marrow_percent_cellularity_day_180",

             "marrow_cd138_baseline_percent",
             "marrow_percent_cd138_day_30",
             "marrow_percent_cd138_day_90",
             "marrow_percent_cd138_day_180",

             "wbc_day_7", "plt_day_7",

             "ig_g_apheresis",
             "ig_g_day_5", "ig_g_day_30", "ig_g_day_90", "ig_g_day_180",
             "ig_a_day_5", "ig_a_day_30", "ig_a_day_90", "ig_a_day_180",
             "ig_m_day_5", "ig_m_day_30", "ig_m_day_90", "ig_m_day_180",

             "cd8_apheresis", "cd8_day_30", "cd8_day_90", "cd8_day_180", 
             "cd19_apheresis", "cd19_day_30", "cd19_day_90", "cd19_day_180",

             "ferritin_day_7", 
             "crp_pre_leukapheresis", "crp_day_0",
             "crp_day_7", "crp_day_14", "crp_day_21",
             "crp_day_30", "crp_day_90", "crp_day_180",

             "ig_g_hsv1_2_apheresis", "ig_g_hsv1_2_d90",
             "ig_g_hsv1_2_d180"
             )
class(cart) <- "data.frame"

for (i in vec_col) {

  if (class(cart[, i]) == "character") {

    cart <- cart %>% 
      mutate(wo_sign = as.numeric(str_remove(cart[, i], ">|<"))) %>% 
      mutate(!!i := case_when(
        str_detect(cart[, i], ">")                 ~ max(wo_sign, na.rm = TRUE),
        str_detect(cart[, i], "<")                 ~ min(wo_sign, na.rm = TRUE),
        TRUE                                         ~ as.numeric(wo_sign)
      )) %>% 
      select(-c(wo_sign))
  }
}
```

```{r fix neutropenia values}
cart <- cart %>% 
  mutate(across(c(starts_with("anc_"), 
                  -anc_0_75_x_10_9_l_0_75_k_u_l), 
                ~ . / 1000))
```

```{r data cleaning2}
cart <- cart %>% 
  # Demographics ----
  rename_at(vars(contains("race")), ~ "race") %>% 
  rename_at(vars(contains("ethnicity")), ~ "ethnicity") %>% 
  mutate(ethnicity = case_when(
    str_detect(ethnicity, "Non|Not")            ~ "Non-Hispanic/Latino",
    is.na(ethnicity)                            ~ NA_character_,
    TRUE                                        ~ "Hispanic/Latino"
  )) %>% 
  mutate(race = case_when(
    str_detect(race, "Other")                   ~ "Other",
    str_detect(race, "Black")                   ~ "Black",
    TRUE                                        ~ race
  )) %>% 
  mutate(race_eth = case_when(
    ethnicity == "Hispanic/Latino"                ~ "Hispanic (any race)",
    TRUE ~ race
  )) %>% 
  # Age categories----
  mutate(agecat = case_when(
    age < 65                            ~ "< 65 years",
    age >= 65                           ~ "≥ 65 years"
  ),
  agecat = factor(agecat, levels = c("< 65 years", "≥ 65 years"))
  ) %>% 
  # Sex categories----
  mutate(sex = case_when(
    str_detect(sex, "M")                                          ~ "Male",
    TRUE                                                          ~ "Female"
  )) %>% 
  mutate(male_sex = case_when(
    str_detect(sex, "M")                                          ~ 1,
    TRUE                                                          ~ 0
  )) %>%

  # ECOG
  # mutate(ecog_at_ld = case_when(
  #   ecog_at_ld == 0 | 
  #     ecog_at_ld == 1                       ~ "0-1",
  #   TRUE                                                          ~ ecog_at_ld
  # )) %>%

  # Myeloma
  mutate(oligo_or_non_secretory = case_when(
    oligo_or_non_secretory == "Non-Secretory" ~ "Yes",
    oligo_or_non_secretory == "Oligosecretory" ~ "Yes",
    oligo_or_non_secretory == "Unknown" ~ NA_character_,
    TRUE ~ oligo_or_non_secretory
  )) %>%
  # mutate(oligo_or_non_secretory_yes_no = case_when( # is "oligo_or_non_secretory_yes_no" now
  #   `Yes, Oligo or Yes, Non-Secretory` == "Non-Secretory" ~ "Yes",
  #   `Yes, Oligo or Yes, Non-Secretory` == "Oligosecretory" ~ "Yes",
  #   `Yes, Oligo or Yes, Non-Secretory` == "Oligo-secretory" ~ "Yes",
  #   TRUE ~ `Yes, Oligo or Yes, Non-Secretory`
  # )) %>%
  mutate(subtype = case_when(
    oligo_or_non_secretory_yes_no == "Yes" ~ "Oligo/non-secretory",
    myeloma_involved_heavy_chain == "None (Light Chain Restricted)" | 
     myeloma_involved_light_chain == "Lambda" & 
      is.na(myeloma_involved_heavy_chain) ~ "Light chain only",
    myeloma_involved_heavy_chain != "None (Light Chain Restricted)" 
                   ~ "Intact immunoglobulin (Heavy + Light)",
    TRUE           ~ NA_character_ 
  )) %>%
  # Cytogenetics ----
  mutate(cytogenetics = case_when(
    deletion_17p_prior_to_car_t_infusion_yes_no == "Yes" | 
      t_4_14_prior_to_car_t_infusion_yes_no == "Yes" | 
      t_14_16_prior_to_car_t_infusion_yes_no == "Yes"             ~ "Yes",
    deletion_17p_prior_to_car_t_infusion_yes_no == "No" &
      t_4_14_prior_to_car_t_infusion_yes_no == "No" &
      t_14_16_prior_to_car_t_infusion_yes_no == "No"               ~ "No",
    TRUE                                                           ~ NA_character_
  )) %>% 
  mutate(gain_amp1q21 = case_when(
    gain_amp1q21 == "No"                                           ~ "No",
    !is.na(gain_amp1q21)                                           ~ "Yes"
  )) %>% 
  # Refractory status ----
  mutate(lenalidomide_exposed_vs_refractory_vs_not_exposed = case_when(
    str_detect(
      lenalidomide_exposed_vs_refractory_vs_not_exposed, "Ref")     ~ "Refractory",
    TRUE           ~ lenalidomide_exposed_vs_refractory_vs_not_exposed
  )) %>% 
  mutate(immuno = case_when(
    lenalidomide_exposed_vs_refractory_vs_not_exposed == "Refractory" |
      pomalidomide_exposed_vs_refractory_vs_not_exposed == "Refractory"           ~ "Yes",
    TRUE                                                           ~ "No"
  )) %>% 
  mutate(proteasome = case_when(
    bortezomib_exposed_vs_refractory_vs_not_exposed == "Refractory" |
      carfilzomib_exposed_vs_refractory_vs_not_exposed == "Refractory"            ~ "Yes",
    TRUE                                                           ~ "No"
  )) %>% 
  mutate(dara = case_when(
    daratumumab_or_isatuximab_cd38_m_ab_exposed_vs_refractory_vs_not_exposed == "Refractory"            ~ "Yes",
    TRUE                                                           ~ "No"
  )) %>% 
  mutate(doubleref = case_when(
    proteasome == "Yes" &
      immuno == "Yes"                                              ~ "Yes",
    TRUE                                                           ~ "No"
  )) %>% 
  mutate(tripleref = case_when(
    proteasome == "Yes" &
      immuno == "Yes" &
      daratumumab_or_isatuximab_cd38_m_ab_exposed_vs_refractory_vs_not_exposed == "Refractory"          ~ "Yes",
    TRUE                                                           ~ "No"
  )) %>% 
  mutate(pentaref = case_when(
    bortezomib_exposed_vs_refractory_vs_not_exposed == "Refractory" & 
      carfilzomib_exposed_vs_refractory_vs_not_exposed == "Refractory" &
      lenalidomide_exposed_vs_refractory_vs_not_exposed == "Refractory" & 
      pomalidomide_exposed_vs_refractory_vs_not_exposed == "Refractory" & 
      daratumumab_or_isatuximab_cd38_m_ab_exposed_vs_refractory_vs_not_exposed == "Refractory"          ~ "Yes",
    TRUE                                                           ~ "No"
  )) %>% 
  mutate(number_prior_lines_therapy = case_when(
    number_of_prior_lines_of_therapy <= 4                        ~ "≤ 4",
    number_of_prior_lines_of_therapy > 4                         ~ "> 4"
  ))
```

```{r Baseline labs}
cart1 <- cart %>% 
  # Baseline labs ----
  mutate(baseline_ferritin_cat = case_when(
    baseline_ferritin_pre_ld >= 400                                      ~ "≥ ULN at LD (≥ 400)",
    baseline_ferritin_pre_ld < 400                                       ~ "Normal (< 400)"
  ), baseline_ferritin_cat = 
    factor(baseline_ferritin_cat, levels = c("Normal (< 400)",
                                             "≥ ULN at LD (≥ 400)"))) %>% 
  mutate(baseline_crp_cat = case_when(
    baseline_crp_pre_ld  >= 0.5                                           ~ "≥ ULN at LD (≥ 0.5)",
    baseline_crp_pre_ld  < 0.5                                            ~ "Normal (< 0.5)"
  ), baseline_crp_cat = 
    factor(baseline_crp_cat, levels = c("Normal (< 0.5)",
                                        "≥ ULN at LD (≥ 0.5)"))) %>% 
  mutate(baseline_B2M = case_when(
    baseline_b2m_pre_ld >= 5.5                                    ~ "≥ 5.5",
    baseline_b2m_pre_ld < 5.5                                     ~ "< 5.5"
  )) %>%
  mutate(cart_cell_dose = case_when(
    cell_dose_million_cells >= 0.7                                ~ "≥ 0.7",
    cell_dose_million_cells < 0.7                                 ~ "< 0.7",
    TRUE                                                          ~ NA_character_
  ),cart_cell_dose = 
    factor(cart_cell_dose, levels = c("≥ 0.7", 
                                      "< 0.7"))) %>% 
  mutate(access_type = case_when(
    str_detect(access_type, "^C|catheter")                        ~ "Central venous catheter",
    str_detect(access_type, "^P")                                 ~ "Peripheral",
    TRUE                                                          ~ access_type
  )) %>% 
  # Supportive therapies and neutrophil recovery----
  mutate(neutrophil_recovery_ANC_never_below_500 = case_when(
    day_of_neutrophil_recovery == "Anc never <500"               ~ "Yes",
    day_of_neutrophil_recovery == "No neutropenia"               ~ "Yes"
  )) %>% 
  mutate(neutrophil_recovery_did_not_recover = case_when(
    day_of_neutrophil_recovery == "Deceased"                     ~ "Yes",
    day_of_neutrophil_recovery == "Pd prior to anc recovery"     ~ "Yes"
  )) %>% 
  mutate(day_of_neutrophil_recovery_ = case_when(
    site == "KUMC"                                               ~ as.numeric(day_of_neutrophil_recovery)
  )) %>% 
  mutate(across(c("last_date_of_gcsf", "day_of_neutrophil_recovery",
                  "last_date_of_ivig"), 
         ~ as.Date(as.numeric(.), origin = "1899-12-30"))) %>% 
  mutate(first_day_of_gcsf = interval(
             start = date_of_car_t_infusion, 
             end = first_date_of_gcsf)/
           duration(n=1, units = "days"
  )) %>%
  mutate(last_day_of_gcsf = interval(
             start = date_of_car_t_infusion, 
             end = last_date_of_gcsf)/
           duration(n=1, units = "days"
  )) %>%
  mutate(day_of_stem_cell_boost = interval(
             start = date_of_car_t_infusion, 
             end = date_of_stem_cell_boost)/
           duration(n=1, units = "days"
  )) %>%
  mutate(day_of_neutrophil_recovery = interval(
             start = date_of_car_t_infusion, 
             end = day_of_neutrophil_recovery)/
           duration(n=1, units = "days"
  ), day_of_neutrophil_recovery = coalesce(day_of_neutrophil_recovery_, day_of_neutrophil_recovery)) %>% 
  select(-day_of_neutrophil_recovery_) %>% 
  mutate(transfusion_postCART = case_when(
    rbc_transfusion_within_30d_post_cart_yes_1_no_0 == "Yes" |
      plt_transfusion_within_30d_post_cart_yes_1_no_0 == "Yes" |
      rbc_transfusion_30d_post_cart_yes_1_no_0 == "Yes" |
      plt_transfusion_30d_post_cart_yes_1_no_0 == "Yes"              ~ "Yes"
  )) %>% 
  mutate(first_day_of_tpo_agonist = interval(
             start = date_of_car_t_infusion, 
             end = first_date_of_tpo_agonist)/
           duration(n=1, units = "days"
  )) %>%
  mutate(last_day_of_tpo_agonist = interval(
             start = date_of_car_t_infusion, 
             end = last_date_of_tpo_agonist)/
           duration(n=1, units = "days"
  )) %>% 
  mutate(first_day_of_ivig = interval(
             start = date_of_car_t_infusion, 
             end = first_date_of_ivig)/
           duration(n=1, units = "days"
  )) %>%
  mutate(last_day_of_ivig = interval(
             start = date_of_car_t_infusion, 
             end = last_date_of_ivig)/
           duration(n=1, units = "days"
  ))

```

```{r toxicities}
cart2 <- cart1 %>% 
# CRS ----
  mutate(CRS_any = ifelse(max_crs_grade_0_5 == 0, "No", "Yes"),
       CRS_any2 = ifelse(CRS_any == "No", 0, 1),
       CRS_gradecat = case_when(
         max_crs_grade_0_5 == 0                             ~ "No CRS",
         max_crs_grade_0_5 == 1                             ~ "Grade 1 or 2",
         max_crs_grade_0_5 == 2                             ~ "Grade 1 or 2",
         max_crs_grade_0_5 >= 3                             ~ "Grade ≥3"),
       CRS_gradecat = factor(CRS_gradecat, levels = c("No CRS", "Grade 1 or 2", "Grade ≥3")),
       gr3_CRS = case_when(
         max_crs_grade_0_5 %in% c(0:2)                      ~ "Grade <3",
         max_crs_grade_0_5 %in% c(3:5)                      ~ "Grade ≥3",
         TRUE                                                   ~ NA_character_
       )) %>% 
  mutate(gr2_CRS = case_when(
    max_crs_grade_0_5 %in% c(0:1)                      ~ "Grade <2",
    max_crs_grade_0_5 %in% c(2:5)                      ~ "Grade ≥2",
    TRUE                                               ~ NA_character_
  )) %>%
  mutate(day_of_onset_crs_relative_to_infusion = 
           interval(start = date_of_car_t_infusion, end = date_of_onset_of_crs) /
           duration(n =1, unit = "days")
  ) %>% 
  mutate(day_of_max_crs_relative_to_infusion = 
           interval(start = date_of_car_t_infusion, end = date_of_max_crs) /
           duration(n =1, unit = "days")
  ) %>% 
  mutate(day_of_onset_icans_relative_to_infusion = 
           interval(start = date_of_car_t_infusion, end = date_of_onset_of_icans) /
           duration(n =1, unit = "days")
  ) %>% 
  mutate(day_of_max_icans_relative_to_infusion = 
           interval(start = date_of_car_t_infusion, end = date_of_max_icans) /
           duration(n =1, unit = "days")
  ) %>% 
  # censored days > 100 to 100
  mutate(across(c(day_of_onset_crs_relative_to_infusion,
                  day_of_max_crs_relative_to_infusion,
                  day_of_onset_icans_relative_to_infusion,
                  day_of_max_icans_relative_to_infusion), ~ case_when(
    . > 100                                                       ~ 100,
    TRUE                                                          ~ as.numeric(.)
  ))) %>%

  # ICANS ----
  mutate(ICANS_any = ifelse(max_icans == 0, "No", "Yes"),
       ICANS_any2 = ifelse(ICANS_any == "No", 0, 1),
       ICANS_gradecat = case_when(
         max_icans == 0                                ~ "No ICANS",
         max_icans == 1                                ~ "Grade 1 or 2",
         max_icans == 2                                ~ "Grade 1 or 2",
         max_icans >= 3                                ~ "Grade ≥3"),
       ICANS_gradecat = factor(ICANS_gradecat, levels = c("No ICANS", "Grade 1 or 2", "Grade ≥3")),
       gr3_ICANS = case_when(
         max_icans %in% c(0:2)                         ~ "Grade <3",
         max_icans %in% c(3:5)                         ~ "Grade ≥3",
         TRUE                                          ~ NA_character_
       )) %>% 
  mutate(gr2_ICANS = case_when(
    max_icans %in% c(0:1)                              ~ "Grade <2",
    max_icans %in% c(2:5)                              ~ "Grade ≥2",
    TRUE                                               ~ NA_character_
  )) %>% 
  # BCMA ----
  mutate(Prior_bcma_therapy = case_when(
    str_detect
    (prior_treatment_with_any_gene_therapy_based_therapeutic_or_investigational_cellular_therapy_or_bcma_targeted_therapy_yes_no, "Yes")
                                                                  ~ "Yes",
    TRUE                                                          ~ "No"
  )) %>% 
  # Leukopenia ----
  mutate(Leukopenia_apheresis = case_when(
    wbc_apheresis < 4                                           ~ "Yes",
    is.na(wbc_apheresis)                                        ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(`Leukopenia_-5d` = case_when(
    wbc_day_5 < 4                                               ~ "Yes",
    is.na(wbc_day_5)                                            ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(Leukopenia_0d = case_when(
    wbc_day_0_k_u_l < 4                                         ~ "Yes",
    is.na(wbc_day_0_k_u_l)                                      ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(Leukopenia_7d = case_when(
    wbc_day_7 < 4                                               ~ "Yes",
    is.na(wbc_day_7)                                            ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(Leukopenia_14d = case_when(
    wbc_day_14 < 4                                              ~ "Yes",
    is.na(wbc_day_14)                                           ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(Leukopenia_21d = case_when(
    wbc_day_21 < 4                                              ~ "Yes",
    is.na(wbc_day_21)                                           ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(Leukopenia_30d = case_when(
    wbc_day_30 < 4                                              ~ "Yes",
    is.na(wbc_day_30)                                           ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  # mutate(Leukopenia_60d = case_when(
  #   wbc_day_60 < 4                                             ~ "Yes",
  #   is.na(wbc_day_60)                                          ~ NA_character_,
  #   TRUE                                                          ~ "No"
  # )) %>% 
  mutate(Leukopenia_90d = case_when(
    wbc_day_90 < 4                                              ~ "Yes",
    is.na(wbc_day_90)                                           ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(Leukopenia_180d = case_when(
    wbc_day_180 < 4                                             ~ "Yes",
    is.na(wbc_day_180)                                          ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(gr3_Leukopenia_apheresis = case_when(
    wbc_apheresis < 2                                           ~ "Yes",
    is.na(wbc_apheresis)                                        ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(`gr3_Leukopenia_-5d` = case_when(
    wbc_day_5 < 2                                               ~ "Yes",
    is.na(wbc_day_5)                                            ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(gr3_Leukopenia_0d = case_when(
    wbc_day_0_k_u_l < 2                                         ~ "Yes",
    is.na(wbc_day_0_k_u_l)                                      ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(gr3_Leukopenia_7d = case_when(
    wbc_day_7 < 2                                               ~ "Yes",
    is.na(wbc_day_7)                                            ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(gr3_Leukopenia_14d = case_when(
    wbc_day_14 < 2                                              ~ "Yes",
    is.na(wbc_day_14)                                           ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(gr3_Leukopenia_21d = case_when(
    wbc_day_21 < 2                                              ~ "Yes",
    is.na(wbc_day_21)                                           ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(gr3_Leukopenia_30d = case_when(
    wbc_day_30 < 2                                              ~ "Yes",
    is.na(wbc_day_30)                                           ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  # mutate(gr3_Leukopenia_60d = case_when(
  #   wbc_day_60 < 2                                             ~ "Yes",
  #   is.na(wbc_day_60)                                          ~ NA_character_,
  #   TRUE                                                          ~ "No"
  # )) %>% 
  mutate(gr3_Leukopenia_90d = case_when(
    wbc_day_90 < 2                                              ~ "Yes",
    is.na(wbc_day_90)                                           ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(gr3_Leukopenia_180d = case_when(
    wbc_day_180 < 2                                             ~ "Yes",
    is.na(wbc_day_180)                                          ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>%

  # Neutropenia ----
  mutate(Neutropenia_apheresis = case_when(
    anc_apheresis < 1.8                                         ~ "Yes",
    is.na(anc_apheresis)                                        ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(`Neutropenia_-5d` = case_when(
    anc_day_5_k_ul < 1.8                                        ~ "Yes",
    is.na(anc_day_5_k_ul)                                       ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(Neutropenia_0d = case_when(
    anc_day_0_k_u_l < 1.8                                       ~ "Yes",
    is.na(anc_day_0_k_u_l)                                      ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(Neutropenia_7d = case_when(
    anc_day_7 < 1.8                                             ~ "Yes",
    is.na(anc_day_7)                                            ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(Neutropenia_14d = case_when(
    anc_day_14 < 1.8                                            ~ "Yes",
    is.na(anc_day_14)                                           ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(Neutropenia_21d = case_when(
    anc_day_21 < 1.8                                            ~ "Yes",
    is.na(anc_day_21)                                           ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(Neutropenia_30d = case_when(
    anc_day_30 < 1.8                                            ~ "Yes",
    is.na(anc_day_30)                                           ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  # mutate(Neutropenia_60d = case_when(
  #   anc_day_60 < 1.8                                             ~ "Yes",
  #   is.na(anc_day_60)                                          ~ NA_character_,
  #   TRUE                                                          ~ "No"
  # )) %>% 
  mutate(Neutropenia_90d = case_when(
    anc_day_90 < 1.8                                            ~ "Yes",
    is.na(anc_day_90)                                           ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(Neutropenia_180d = case_when(
    anc_day_180 < 1.8                                           ~ "Yes",
    is.na(anc_day_180)                                          ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(gr3_Neutropenia_apheresis = case_when(
    anc_apheresis < 1                                           ~ "Yes",
    is.na(anc_apheresis)                                        ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(`gr3_Neutropenia_-5d` = case_when(
    anc_day_5_k_ul < 1                                          ~ "Yes",
    is.na(anc_day_5_k_ul)                                       ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(gr3_Neutropenia_0d = case_when(
    anc_day_0_k_u_l < 1                                         ~ "Yes",
    is.na(anc_day_0_k_u_l)                                      ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(gr3_Neutropenia_7d = case_when(
    anc_day_7 < 1                                               ~ "Yes",
    is.na(anc_day_7)                                            ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(gr3_Neutropenia_14d = case_when(
    anc_day_14 < 1                                              ~ "Yes",
    is.na(anc_day_14)                                           ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(gr3_Neutropenia_21d = case_when(
    anc_day_21 < 1                                              ~ "Yes",
    is.na(anc_day_21)                                           ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(gr3_Neutropenia_30d = case_when(
    anc_day_30 < 1                                              ~ "Yes",
    is.na(anc_day_30)                                           ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  # mutate(gr3_Neutropenia_60d = case_when(
  #   anc_day_60 < 1                                             ~ "Yes",
  #   is.na(anc_day_60)                                          ~ NA_character_,
  #   TRUE                                                          ~ "No"
  # )) %>% 
  mutate(gr3_Neutropenia_90d = case_when(
    anc_day_90 < 1                                              ~ "Yes",
    is.na(anc_day_90)                                           ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(gr3_Neutropenia_180d = case_when(
    anc_day_180 < 1                                             ~ "Yes",
    is.na(anc_day_180)                                          ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(Neutropenia_gr1 = case_when(
    (anc_day_7 < 1.8 &
       anc_day_7 > 1.5) |
      (anc_day_30 < 1.8 &
         anc_day_30 > 1.5) |
      # (anc_day_60 < 1.8 &
      #    anc_day_60 > 1.5) |
      (anc_day_180 < 1.8 &
         anc_day_180 > 1.5) |
      (anc_day_90 < 1.8 &
         anc_day_90 > 1.5)                                      ~ "Yes",
    is.na(anc_day_7) |
      is.na(anc_day_30) |
      # is.na(anc_day_60) |
      is.na(anc_day_90)                                         ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(Neutropenia_gr2 = case_when(
    (anc_day_7 < 1.5 &
       anc_day_7 > 1) |
      (anc_day_30 < 1.5 &
         anc_day_30 > 1) |
      (anc_day_180 < 1.5 &
         anc_day_180 > 1) |
      # (anc_day_60 < 1.5 &
      #    anc_day_60 > 1) |
      (anc_day_90 < 1.5 &
         anc_day_90 > 1)                                        ~ "Yes",
    is.na(anc_day_7) |
      is.na(anc_day_30) |
      # is.na(anc_day_60) |
      is.na(anc_day_90)                                         ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(Neutropenia_gr3 = case_when(
    anc_day_7 < 1 |
      anc_day_30 < 1 |
      anc_day_180 < 1 |
      # anc_day_60 < 1 |
      anc_day_90 < 1                                            ~ "Yes",
    is.na(anc_day_7) |
      is.na(anc_day_30) |
      is.na(anc_day_180) |
      # is.na(anc_day_60) |
      is.na(anc_day_90)                                         ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(any_neutropenia = case_when(
    Neutropenia_gr1 == "Yes" |
      Neutropenia_gr2 == "Yes" |
      Neutropenia_gr3 == "Yes"                                  ~ "Yes",
    is.na(Neutropenia_gr1) |
      is.na(Neutropenia_gr2) |
      is.na(Neutropenia_gr3)                                    ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(severe_neutro_30beyond = case_when(
    anc_day_30 < 1 |
    anc_day_180 < 1 |
      # anc_day_60 < 1 |
      anc_day_90 < 1                                            ~ "Yes",
    is.na(anc_day_30) |
    is.na(anc_day_180) |
      # is.na(anc_day_60) |
      is.na(anc_day_90)                                         ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>%

  # Anemia ----
  mutate(Anemia_apheresis = case_when(
    hb_apheresis < 11.4                                         ~ "Yes",
    is.na(hb_apheresis)                                         ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(`Anemia_-5d` = case_when(
    hb_day_5 < 11.4                                             ~ "Yes",
    is.na(hb_day_5)                                             ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(Anemia_0d = case_when(
    hb_day_0_k_u_l < 11.4                                       ~ "Yes",
    is.na(hb_day_0_k_u_l)                                       ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(Anemia_7d = case_when(
    hb_day_7 < 11.4                                             ~ "Yes",
    is.na(hb_day_7)                                             ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>%
  mutate(Anemia_14d = case_when(
    hb_day_14 < 11.4                                            ~ "Yes",
    is.na(hb_day_14)                                            ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(Anemia_21d = case_when(
    hb_day_21 < 11.4                                            ~ "Yes",
    is.na(hb_day_21)                                            ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(Anemia_30d = case_when(
    hb_day_30 < 11.4                                            ~ "Yes",
    is.na(hb_day_30)                                            ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>%
  # mutate(Anemia_60d = case_when(
  #   hb_day_60 < 11.4                                             ~ "Yes",
  #   is.na(hb_day_60)                                          ~ NA_character_,
  #   TRUE                                                          ~ "No"
  # )) %>%
  mutate(Anemia_90d = case_when(
    hb_day_90 < 11.4                                            ~ "Yes",
    is.na(hb_day_90)                                            ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>%             
  mutate(Anemia_180d = case_when(
    hb_day_180 < 11.4                                           ~ "Yes",
    is.na(hb_day_180)                                           ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>%
  mutate(gr3_Anemia_apheresis = case_when(
    hb_apheresis < 8                                            ~ "Yes",
    is.na(hb_apheresis)                                         ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(`gr3_Anemia_-5d` = case_when(
    hb_day_5 < 8                                                ~ "Yes",
    is.na(hb_day_5)                                             ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(gr3_Anemia_0d = case_when(
    hb_day_0_k_u_l < 8                                          ~ "Yes",
    is.na(hb_day_0_k_u_l)                                       ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(gr3_Anemia_7d = case_when(
    hb_day_7 < 8                                                ~ "Yes",
    is.na(hb_day_7)                                             ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>%
  mutate(gr3_Anemia_14d = case_when(
    hb_day_14 < 8                                               ~ "Yes",
    is.na(hb_day_14)                                            ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(gr3_Anemia_21d = case_when(
    hb_day_21 < 8                                               ~ "Yes",
    is.na(hb_day_21)                                            ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(gr3_Anemia_30d = case_when(
    hb_day_30 < 8                                               ~ "Yes",
    is.na(hb_day_30)                                            ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>%
  # mutate(gr3_Anemia_60d = case_when(
  #   hb_day_60 < 8                                             ~ "Yes",
  #   is.na(hb_day_60)                                          ~ NA_character_,
  #   TRUE                                                          ~ "No"
  # )) %>%
  mutate(gr3_Anemia_90d = case_when(
    hb_day_90 < 8                                               ~ "Yes",
    is.na(hb_day_90)                                            ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>%
  mutate(gr3_Anemia_180d = case_when(
    hb_day_180 < 8                                              ~ "Yes",
    is.na(hb_day_180)                                           ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>%
  mutate(Anemia_gr1 = case_when(
    (hb_day_7 < 11.4 &
       hb_day_7 > 10) |
      (hb_day_30 < 11.4 &
         hb_day_30 > 10) |
      # (hb_day_60 < 11.4 &
         # hb_day_60 > 10) |
      (hb_day_180 < 11.4 &
         hb_day_180 > 10) |
      (hb_day_90 < 11.4 &
         hb_day_90 > 10)                                        ~ "Yes",
    is.na(hb_day_7) |
      is.na(hb_day_30) |
      # is.na(hb_day_60) |
      is.na(hb_day_90)                                          ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(Anemia_gr2 = case_when(
    (hb_day_7 < 10 &
       hb_day_7 > 8) |
      (hb_day_30 < 10 &
         hb_day_30 > 8) |
      # (hb_day_60 < 10 &
         # hb_day_60 > 8) |
      (hb_day_180 < 10 &
         hb_day_180 > 8) |
      (hb_day_90 < 10 &
         hb_day_90 > 8)                                         ~ "Yes",
    is.na(hb_day_7) |
      is.na(hb_day_30) |
      # is.na(hb_day_60) |
      is.na(hb_day_90)                                          ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(Anemia_gr3 = case_when(
    hb_day_7 < 8 |
      hb_day_30 < 8 |
      # hb_day_60 < 8 |
      hb_day_180 < 8 |
      hb_day_90 < 8                                             ~ "Yes",
    is.na(hb_day_7) |
      is.na(hb_day_30) |
      # is.na(hb_day_60) |
      is.na(hb_day_180) |
      is.na(hb_day_90)                                          ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(severe_anemia_30beyond = case_when(
    hb_day_30 < 8 |
      # hb_day_60 < 8 |
      hb_day_180 < 8 |
      hb_day_90 < 8                                             ~ "Yes",
    is.na(hb_day_30) |
      # is.na(hb_day_60) |
      is.na(hb_day_180) |
      is.na(hb_day_90)                                          ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(any_anemia = case_when(
    Anemia_gr1 == "Yes" |
      Anemia_gr2 == "Yes" |
      Anemia_gr3 == "Yes"                                       ~ "Yes",
    is.na(Anemia_gr1) |
      is.na(Anemia_gr2) |
      is.na(Anemia_gr3)                                         ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>%

  # Thrombocytopenia----
  mutate(Thrombocytopenia_apheresis = case_when(
    plt_apheresis < 143                                         ~ "Yes",
    is.na(plt_apheresis)                                        ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(`Thrombocytopenia_-5d` = case_when(
    plt_day_5_k_ul < 143                                        ~ "Yes",
    is.na(plt_day_5_k_ul)                                       ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(Thrombocytopenia_0d = case_when(
    plt_day_0_k_u_l < 143                                       ~ "Yes",
    is.na(plt_day_0_k_u_l)                                      ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(Thrombocytopenia_7d = case_when(
    plt_day_7 < 143                                             ~ "Yes",
    is.na(plt_day_7)                                            ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>%
  mutate(Thrombocytopenia_14d = case_when(
    plt_day_14 < 143                                            ~ "Yes",
    is.na(plt_day_14)                                           ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(Thrombocytopenia_21d = case_when(
    plt_day_21 < 143                                            ~ "Yes",
    is.na(plt_day_21)                                           ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(Thrombocytopenia_30d = case_when(
    plt_day_30 < 143                                            ~ "Yes",
    is.na(plt_day_30)                                           ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>%
  # mutate(Thrombocytopenia_60d = case_when(
  #   plt_day_60 < 143                                             ~ "Yes",
  #   is.na(plt_day_60)                                          ~ NA_character_,
  #   TRUE                                                          ~ "No"
  # )) %>%
  mutate(Thrombocytopenia_90d = case_when(
    plt_day_90 < 143                                            ~ "Yes",
    is.na(plt_day_90)                                           ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>%
  mutate(Thrombocytopenia_180d = case_when(
    plt_day_180 < 143                                           ~ "Yes",
    is.na(plt_day_180)                                          ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>%
  mutate(gr3_Thrombocytopenia_apheresis = case_when(
    plt_apheresis < 50                                          ~ "Yes",
    is.na(plt_apheresis)                                        ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(`gr3_Thrombocytopenia_-5d` = case_when(
    plt_day_5_k_ul < 50                                         ~ "Yes",
    is.na(plt_day_5_k_ul)                                       ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(gr3_Thrombocytopenia_0d = case_when(
    plt_day_0_k_u_l < 50                                        ~ "Yes",
    is.na(plt_day_0_k_u_l)                                      ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(gr3_Thrombocytopenia_7d = case_when(
    plt_day_7 < 50                                              ~ "Yes",
    is.na(plt_day_7)                                            ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>%
  mutate(gr3_Thrombocytopenia_14d = case_when(
    plt_day_14 < 50                                             ~ "Yes",
    is.na(plt_day_14)                                           ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(gr3_Thrombocytopenia_21d = case_when(
    plt_day_21 < 50                                             ~ "Yes",
    is.na(plt_day_21)                                           ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(gr3_Thrombocytopenia_30d = case_when(
    plt_day_30 < 50                                             ~ "Yes",
    is.na(plt_day_30)                                           ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>%
  # mutate(gr3_Thrombocytopenia_60d = case_when(
  #   plt_day_60 < 50                                             ~ "Yes",
  #   is.na(plt_day_60)                                          ~ NA_character_,
  #   TRUE                                                          ~ "No"
  # )) %>%
  mutate(gr3_Thrombocytopenia_90d = case_when(
    plt_day_90 < 50                                             ~ "Yes",
    is.na(plt_day_90)                                           ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(gr3_Thrombocytopenia_180d = case_when(
    plt_day_180 < 50                                            ~ "Yes",
    is.na(plt_day_180)                                          ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(Thrombocytopenia_gr1 = case_when(
    (plt_day_7 < 143 &
       plt_day_7 > 75) |
      (plt_day_30 < 143 &
         plt_day_30 > 75) |
      # (plt_day_60 < 143 &
         # plt_day_60 > 75) |
      (plt_day_180 < 143 &
         plt_day_180 > 75) |
      (plt_day_90 < 143 &
         plt_day_90 > 75)                                       ~ "Yes",
    is.na(plt_day_7) |
      is.na(plt_day_30) |
      # is.na(plt_day_60) |
      is.na(plt_day_180) |
      is.na(plt_day_90)                                         ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(Thrombocytopenia_gr2 = case_when(
    (plt_day_7 < 75 &
       plt_day_7 > 50) |
      (plt_day_30 < 75 &
         plt_day_30 > 50) |
      # (plt_day_60 < 75 &
         # plt_day_60 > 50) |
      (plt_day_180 < 75 &
         plt_day_180 > 50) |
      (plt_day_90 < 75 &
         plt_day_90 > 50)                                       ~ "Yes",
    is.na(plt_day_7) |
      is.na(plt_day_30) |
      # is.na(plt_day_60) |
      is.na(plt_day_180) |
      is.na(plt_day_90)                                         ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(Thrombocytopenia_gr3 = case_when(
    plt_day_7 < 50 |
      plt_day_30 < 50 |
      # plt_day_60 < 50 |
      plt_day_180 < 50 |
      plt_day_90 < 50                                           ~ "Yes",
    is.na(plt_day_7) |
      is.na(plt_day_30) |
      is.na(plt_day_180) |
      # is.na(plt_day_60) |
      is.na(plt_day_90)                                         ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(severe_throm_30beyond = case_when(
    plt_day_30 < 50 |
      # plt_day_60 < 50 |
      plt_day_180 < 50 |
      plt_day_90 < 50                                           ~ "Yes",
    is.na(plt_day_30) |
      # is.na(plt_day_60) |
      is.na(plt_day_180) |
      is.na(plt_day_90)                                         ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(any_thrombo = case_when(
    Thrombocytopenia_gr1 == "Yes" |
      Thrombocytopenia_gr2 == "Yes" |
      Thrombocytopenia_gr3 == "Yes"                             ~ "Yes",
    is.na(Thrombocytopenia_gr1) |
      is.na(Thrombocytopenia_gr2) |
      is.na(Thrombocytopenia_gr3)                               ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>%

  # Cytopenia ----
  mutate(cytopenia_apheresis = case_when(
    wbc_apheresis < 4 | 
      anc_apheresis < 1.8 | 
      hb_apheresis < 11.4 | 
      plt_apheresis < 143                                       ~ "Yes",
    is.na(wbc_apheresis) |
      is.na(anc_apheresis) |
      is.na(hb_apheresis) |
      is.na(plt_apheresis)                                      ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(`cytopenia_-5d` = case_when(
    wbc_day_5 < 4 | 
      anc_day_5_k_ul < 1.8 | 
      hb_day_5 < 11.4 | 
      plt_day_5_k_ul < 143                                      ~ "Yes",
    is.na(wbc_day_5) |
      is.na(anc_day_5_k_ul) |
      is.na(hb_day_5) |
      is.na(plt_day_5_k_ul)                                     ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(cytopenia_0d = case_when(
    wbc_day_0_k_u_l < 4 | 
      anc_day_0_k_u_l < 1.8 | 
      hb_day_0_k_u_l < 11.4 | 
      plt_day_0_k_u_l < 143                                     ~ "Yes",
    is.na(wbc_day_0_k_u_l) |
      is.na(anc_day_0_k_u_l) |
      is.na(hb_day_0_k_u_l) |
      is.na(plt_day_0_k_u_l)                                    ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(cytopenia_7d = case_when(
    Anemia_7d == "Yes" | Thrombocytopenia_7d == "Yes" | 
      Neutropenia_7d == "Yes"                                   ~ "Yes",
    is.na(Anemia_7d) | is.na(Thrombocytopenia_7d) | 
      is.na(Neutropenia_7d)                                     ~ NA_character_,
    TRUE ~ "No"
  )) %>%
  mutate(cytopenia_14d = case_when(
    wbc_day_14 < 4 | 
      anc_day_14 < 1.8 | 
      hb_day_14 < 11.4 | 
      plt_day_14 < 143                                          ~ "Yes",
    is.na(wbc_day_14) |
      is.na(anc_day_14) |
      is.na(hb_day_14) |
      is.na(plt_day_14)                                         ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(cytopenia_21d = case_when(
    wbc_day_21 < 4 | 
      anc_day_21 < 1.8 | 
      hb_day_21 < 11.4 | 
      plt_day_21 < 143                                          ~ "Yes",
    is.na(wbc_day_21) |
      is.na(anc_day_21) |
      is.na(hb_day_21) |
      is.na(plt_day_21)                                         ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(cytopenia_30d = case_when(
    Anemia_30d == "Yes" | Thrombocytopenia_30d == "Yes" | 
      Neutropenia_30d == "Yes"                                  ~ "Yes",
    is.na(Anemia_30d) | is.na(Thrombocytopenia_30d) | 
      is.na(Neutropenia_30d)                                    ~ NA_character_,
    TRUE ~ "No"
  )) %>%
  # mutate(cytopenia_d60 = case_when(
  #   Anemia_60d == "Yes" | Thrombocytopenia_60d == "Yes" | Neutropenia_60d == "Yes" ~ "Yes",
  #   is.na(Anemia_60d) | is.na(Thrombocytopenia_60d) | is.na(Neutropenia_60d) ~ NA_character_,
  #   TRUE ~ "No"
  # )) %>%
  mutate(cytopenia_90d = case_when(
    Anemia_90d == "Yes" | Thrombocytopenia_90d == "Yes" | 
      Neutropenia_90d == "Yes"                                  ~ "Yes",
    is.na(Anemia_90d) | is.na(Thrombocytopenia_90d) | 
      is.na(Neutropenia_90d)                                    ~ NA_character_,
    TRUE ~ "No"
  )) %>%
  mutate(cytopenia_180d = case_when(
    Anemia_180d == "Yes" | Thrombocytopenia_180d == "Yes" | 
      Neutropenia_180d == "Yes"                                 ~ "Yes",
    is.na(Anemia_180d) | is.na(Thrombocytopenia_180d) | 
      is.na(Neutropenia_180d)                                   ~ NA_character_,
    TRUE ~ "No"
  )) %>%

  mutate(`gr3_cytopenia_≥30d` = case_when(
    anc_day_30 < 1 |
      hb_day_30 < 8 |
      plt_day_30 < 50 |
      # anc_day_60 < 1 |
      # hb_day_60 < 8 |
      # plt_day_60 < 50 |
      anc_day_180 < 1 |
      hb_day_180 < 8 |
      plt_day_180 < 50 |
      anc_day_90 < 1 |
      hb_day_90 < 8 |
      plt_day_90 < 50                                           ~ "Yes",
    is.na(anc_day_30) |
      is.na(hb_day_30) |
      is.na(plt_day_30) |
      # is.na(anc_day_60) |
      # is.na(hb_day_60) |
      # is.na(plt_day_60) |
      is.na(anc_day_180) |
      is.na(hb_day_180) |
      is.na(plt_day_180) |
      is.na(anc_day_90) |
      is.na(hb_day_90) |
      is.na(plt_day_90)                                         ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
   mutate(gr3_cytopenia_apheresis = case_when(
    wbc_apheresis < 2 | 
      anc_apheresis < 1 | 
      hb_apheresis < 8 | 
      plt_apheresis < 50                                        ~ "Yes",
    is.na(wbc_apheresis) |
      is.na(anc_apheresis) |
      is.na(hb_apheresis) |
      is.na(plt_apheresis)                                      ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(`gr3_cytopenia_-5d` = case_when(
    wbc_day_5 < 2 | 
      anc_day_5_k_ul < 1 | 
      hb_day_5 < 8 | 
      plt_day_5_k_ul < 50                                       ~ "Yes",
    is.na(wbc_day_5) |
      is.na(anc_day_5_k_ul) |
      is.na(hb_day_5) |
      is.na(plt_day_5_k_ul)                                     ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(gr3_cytopenia_0d = case_when(
    wbc_day_0_k_u_l < 2 | 
      anc_day_0_k_u_l < 1 | 
      hb_day_0_k_u_l < 8 | 
      plt_day_0_k_u_l < 50                                      ~ "Yes",
    is.na(wbc_day_0_k_u_l) |
      is.na(anc_day_0_k_u_l) |
      is.na(hb_day_0_k_u_l) |
      is.na(plt_day_0_k_u_l)                                    ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(gr3_cytopenia_7d = case_when(
    gr3_Anemia_7d == "Yes" | 
      gr3_Thrombocytopenia_7d == "Yes" | 
      gr3_Neutropenia_7d == "Yes"                               ~ "Yes",
    is.na(gr3_Anemia_7d) | 
      is.na(gr3_Thrombocytopenia_7d) | 
      is.na(gr3_Neutropenia_7d)                                 ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>%
  mutate(gr3_cytopenia_14d = case_when(
    wbc_day_14 < 2 | 
      anc_day_14 < 1 | 
      hb_day_14 < 8 | 
      plt_day_14 < 50                                           ~ "Yes",
    is.na(wbc_day_14) |
      is.na(anc_day_14) |
      is.na(hb_day_14) |
      is.na(plt_day_14)                                         ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(gr3_cytopenia_21d = case_when(
    wbc_day_21 < 2 | 
      anc_day_21 < 1 | 
      hb_day_21 < 8 | 
      plt_day_21 < 50                                           ~ "Yes",
    is.na(wbc_day_21) |
      is.na(anc_day_21) |
      is.na(hb_day_21) |
      is.na(plt_day_21)                                         ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(gr3_cytopenia_30d = case_when(
    gr3_Anemia_30d == "Yes" | 
      gr3_Thrombocytopenia_30d == "Yes" | 
      gr3_Neutropenia_30d == "Yes"                              ~ "Yes",
    is.na(gr3_Anemia_30d) | 
      is.na(gr3_Thrombocytopenia_30d) | 
      is.na(gr3_Neutropenia_30d)                                ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>%
  # mutate(gr3_cytopenia_60d = case_when(
  #   gr3_Anemia_60d == "Yes" | gr3_Thrombocytopenia_60d == "Yes" | gr3_Neutropenia_60d == "Yes" ~ "Yes",
  #   is.na(gr3_Anemia_60d) | is.na(gr3_Thrombocytopenia_60d) | is.na(gr3_Neutropenia_60d) ~ NA_character_,
  #   TRUE ~ "No"
  # )) %>%
  mutate(gr3_cytopenia_90d = case_when(
    gr3_Anemia_90d == "Yes" | 
      gr3_Thrombocytopenia_90d == "Yes" | 
      gr3_Neutropenia_90d == "Yes"                              ~ "Yes",
    is.na(gr3_Anemia_90d) | 
      is.na(gr3_Thrombocytopenia_90d) | 
      is.na(gr3_Neutropenia_90d)                                ~ NA_character_,
    TRUE                                                        ~ "No"
  )) %>% 
  mutate(gr3_cytopenia_180d = case_when(
    gr3_Anemia_180d == "Yes" | 
      gr3_Thrombocytopenia_180d == "Yes" |
      gr3_Neutropenia_180d == "Yes"                             ~ "Yes",
    is.na(gr3_Anemia_180d) | 
      is.na(gr3_Thrombocytopenia_180d) |
      is.na(gr3_Neutropenia_180d)                               ~ NA_character_,
    TRUE                                                        ~ "No"
  ))
```

```{r survival}
cart3 <- cart2 %>% 
  # infusion
  mutate(notinfused = case_when(
    !is.na(date_of_car_t_infusion) | 
      !is.na(time_from_apheresis_to_car_t_infusion_days)          ~ "Yes",
    TRUE                                                          ~ "No"
  )) #%>% 
  # mutate(time_aph_infusion = 
  #          interval(
  #            start = date_of_apheresis_or_number_of_days_from_apheresis_to_ld, 
  #            end = time_from_apheresis_to_car_t_infusion_days)/
  #          duration(n=1, units = "days"))

  # Day 30 response ----
# cart3 <- cart2 %>% 
#   select(pt_identifier, starts_with("date"))
cart4 <- cart3 %>% 
  mutate(date_of_onset_of_delayed_neurotoxicity1 = 
           str_extract(date_of_onset_of_delayed_neurotoxicity, "([:digit:]*)/([:digit:]*)/([:digit:]*)")) %>% 
  mutate(date_of_pd = as.Date(as.numeric(date_of_pd),
                                             origin = "1899-12-30")

  )

cart5 <- cart4 %>% 
  mutate(across(c(starts_with("date")#,
                  # starts_with("day_of_max_"), contains("_dt"), 
                  # ends_with("date")
  ),
  ~ as.Date(., format = "%m/%d/%y")
  )) %>% 
  # pfs
  mutate(pfs_event = case_when(
    !is.na(date_of_pd) | !is.na(date_of_death)         ~ 1,
    TRUE                                               ~ 0
  )) %>% 
  mutate(pfs_date = coalesce(date_of_pd, date_of_death, date_of_last_contact)) %>% 
  mutate(pfs_from_infusion_months = 
           interval(
             start = date_of_car_t_infusion, 
             end = pfs_date)/
           duration(n=1, units = "months")) %>% 
  # os 
  mutate(os_event = case_when(
    !is.na(date_of_death)                              ~ 1,
    TRUE                                               ~ 0
  )) %>% 
  mutate(os_date = coalesce(date_of_death, date_of_last_contact)) %>% 
  mutate(os_from_infusion_months = 
           interval(
             start = date_of_car_t_infusion, 
             end = os_date)/
           duration(n=1, units = "months"))
```

```{r response}
cart6 <- cart5 %>%   
  # mutate(day_30_response_s_cr_cr_vgpr_pr_sd_pd = case_when(
  #   str_detect(day_30_response_s_cr_cr_vgpr_pr_sd_pd, "evaluable")   ~ NA_character_,
  #   TRUE           ~ day_30_response_s_cr_cr_vgpr_pr_sd_pd
  # )) %>% 
  mutate(day30_response = case_when(
    (os_event == 1 | pfs_event == 1) & 
      pfs_from_infusion_months < 1                                   ~ "Died or progressed before day 30",
    str_detect(day_30_response_s_cr_cr_vgpr_pr_sd_pd, "EVAL")        ~ NA_character_,
    str_detect(day_30_response_s_cr_cr_vgpr_pr_sd_pd, "VGPR")        ~ "VGPR",
    str_detect(day_30_response_s_cr_cr_vgpr_pr_sd_pd, "PR")          ~ "PR",
    # str_detect(day_30_response_s_cr_cr_vgpr_pr_sd_pd, "UCR")	       ~ "UCR",
    str_detect(day_30_response_s_cr_cr_vgpr_pr_sd_pd, "CR")          ~ "sCR or CR",
    str_detect(day_30_response_s_cr_cr_vgpr_pr_sd_pd, "MR") |
      str_detect(day_30_response_s_cr_cr_vgpr_pr_sd_pd, "SD")        ~ "SD",
    str_detect(day_30_response_s_cr_cr_vgpr_pr_sd_pd, "PD")          ~ "PD",
    is.na(day_30_response_s_cr_cr_vgpr_pr_sd_pd)                     ~ NA_character_,
    TRUE           ~ day_30_response_s_cr_cr_vgpr_pr_sd_pd
  )) %>% 
  mutate(day30_response_v2 = case_when(
    str_detect(day30_response, "Died")                               ~ "PD",
    is.na(day30_response)                                            ~ NA_character_,
    TRUE           ~ day30_response
  )) %>% 
  mutate(ORR_30days = case_when(
    str_detect(day30_response_v2, "CR|PR")                           ~ "ORR",
    is.na(day30_response_v2)                                         ~ NA_character_,
    TRUE                                                             ~ "PD/SD/MR"
  )) %>% 
  mutate(CRorbetter_30days = case_when(
    str_detect(day30_response_v2, "CR")	                             ~ "sCR or CR",
    is.na(day30_response_v2)                                         ~ NA_character_,
    TRUE                                                             ~ "Others"
  )) %>% 
  # Day 90 response ----
  mutate(mon3_response = case_when(
    day30_response_v2 == "PD"                                        ~ "PD",
    (os_event == 1 | pfs_event == 1) & 
      pfs_from_infusion_months < 3                                   ~ "Died or progressed before 3 months",
    str_detect(x3_month_response_s_cr_cr_vgpr_pr_sd_pd, "EVAL")      ~ NA_character_,
    str_detect(x3_month_response_s_cr_cr_vgpr_pr_sd_pd, "VGPR")      ~ "VGPR",
    str_detect(x3_month_response_s_cr_cr_vgpr_pr_sd_pd, "PR")        ~ "PR",
    str_detect(x3_month_response_s_cr_cr_vgpr_pr_sd_pd, "CR")	       ~ "sCR or CR",
    str_detect(x3_month_response_s_cr_cr_vgpr_pr_sd_pd, "MR") |
      str_detect(x3_month_response_s_cr_cr_vgpr_pr_sd_pd, "SD")      ~ "SD",
    str_detect(x3_month_response_s_cr_cr_vgpr_pr_sd_pd, "PD")       ~ "PD",
    is.na(x3_month_response_s_cr_cr_vgpr_pr_sd_pd)                     ~ NA_character_,
    TRUE           ~ x3_month_response_s_cr_cr_vgpr_pr_sd_pd
  )) %>% 
  mutate(mon3_response_v2 = case_when(
    str_detect(mon3_response, "Died")                              ~ "PD",
    is.na(mon3_response)                                           ~ NA_character_,
    TRUE           ~ mon3_response
  )) %>% 
  mutate(ORR_mon3 = case_when(
    str_detect(mon3_response_v2, "CR|PR")                          ~ "ORR",
    is.na(mon3_response_v2)                                        ~ NA_character_,
    TRUE                                                            ~ "PD/SD/MR"
  )) %>% 
  mutate(CRorbetter_mon3 = case_when(
    mon3_response_v2 == "sCR or CR"                        	      ~ "sCR or CR",
    is.na(mon3_response_v2)                                        ~ NA_character_,
    TRUE                                                            ~ "Others"
  )) %>% 
    # Month 6 response ----
  mutate(mon6_response = case_when(
    mon3_response_v2 == "PD"                                        ~ "PD",
    (os_event == 1 | pfs_event == 1) & 
      pfs_from_infusion_months < 6                                     ~ "Died or progressed before 3 months",
    str_detect(x6_month_response_s_cr_cr_vgpr_pr_sd_pd, "EVAL")        ~ NA_character_,
    str_detect(x6_month_response_s_cr_cr_vgpr_pr_sd_pd, "VGPR")        ~ "VGPR",
    str_detect(x6_month_response_s_cr_cr_vgpr_pr_sd_pd, "PR")          ~ "PR",
    str_detect(x6_month_response_s_cr_cr_vgpr_pr_sd_pd, "CR")	         ~ "sCR or CR",
    str_detect(x6_month_response_s_cr_cr_vgpr_pr_sd_pd, "MR") |
      str_detect(x6_month_response_s_cr_cr_vgpr_pr_sd_pd, "SD")        ~ "SD",
    str_detect(x6_month_response_s_cr_cr_vgpr_pr_sd_pd, "PD")          ~ "PD",
    is.na(x6_month_response_s_cr_cr_vgpr_pr_sd_pd)                     ~ NA_character_,
    TRUE           ~ x6_month_response_s_cr_cr_vgpr_pr_sd_pd
  )) %>% 
  mutate(mon6_response_v2 = case_when(
    str_detect(mon6_response, "Died")                                 ~ "PD",
    is.na(mon6_response)                                              ~ NA_character_,
    TRUE           ~ mon6_response
  )) %>% 
  mutate(ORR_mon6 = case_when(
    str_detect(mon6_response_v2, "CR|PR")                             ~ "ORR",
    is.na(mon6_response_v2)                                           ~ NA_character_,
    TRUE                                                               ~ "PD/SD/MR"
  )) %>% 
  mutate(CRorbetter_mon6 = case_when(
    mon6_response_v2 == "sCR or CR"                        	       ~ "sCR or CR",
    is.na(mon6_response_v2)                                         ~ NA_character_,
    TRUE                                                             ~ "Others"
  )) %>% 
    # Month 9 response ----
  mutate(mon9_response = case_when(
    mon6_response_v2 == "PD"                                        ~ "PD",
    (os_event == 1 | pfs_event == 1) & 
      pfs_from_infusion_months < 9                                     ~ "Died or progressed before 9 months",
    str_detect(x9_month_response_s_cr_cr_vgpr_pr_sd_pd, "EVAL")        ~ NA_character_,
    str_detect(x9_month_response_s_cr_cr_vgpr_pr_sd_pd, "VGPR")        ~ "VGPR",
    str_detect(x9_month_response_s_cr_cr_vgpr_pr_sd_pd, "PR")          ~ "PR",
    str_detect(x9_month_response_s_cr_cr_vgpr_pr_sd_pd, "CR")	         ~ "sCR or CR",
    str_detect(x9_month_response_s_cr_cr_vgpr_pr_sd_pd, "MR") |
      str_detect(x9_month_response_s_cr_cr_vgpr_pr_sd_pd, "SD")        ~ "SD",
    str_detect(x9_month_response_s_cr_cr_vgpr_pr_sd_pd, "PD")          ~ "PD",
    is.na(x9_month_response_s_cr_cr_vgpr_pr_sd_pd)                     ~ NA_character_,
    TRUE           ~ x9_month_response_s_cr_cr_vgpr_pr_sd_pd
  )) %>% 
    mutate(mon9_response_v2 = case_when(
      str_detect(mon9_response, "Died")                                 ~ "PD",
      is.na(mon9_response)                                              ~ NA_character_,
      TRUE           ~ mon9_response
    )) %>% 
    mutate(ORR_mon9 = case_when(
      str_detect(mon9_response_v2, "CR|PR")                             ~ "ORR",
      is.na(mon9_response_v2)                                           ~ NA_character_,
      TRUE                                                               ~ "PD/SD/MR"
    )) %>% 
    mutate(CRorbetter_mon9 = case_when(
      mon9_response_v2 == "sCR or CR"                        	       ~ "sCR or CR",
      is.na(mon9_response_v2)                                         ~ NA_character_,
      TRUE                                                             ~ "Others"
    )) %>% 
    # Month 12 response ----
  mutate(mon12_response = case_when(
    mon9_response_v2 == "PD"                                        ~ "PD",
    (os_event == 1 | pfs_event == 1) & 
      pfs_from_infusion_months < 12                                     ~ "Died or progressed before 12 months",
    str_detect(x12_month_response_s_cr_cr_vgpr_pr_sd_pd, "EVAL")        ~ NA_character_,
    str_detect(x12_month_response_s_cr_cr_vgpr_pr_sd_pd, "VGPR")        ~ "VGPR",
    str_detect(x12_month_response_s_cr_cr_vgpr_pr_sd_pd, "PR")          ~ "PR",
    str_detect(x12_month_response_s_cr_cr_vgpr_pr_sd_pd, "CR")	         ~ "sCR or CR",
    str_detect(x12_month_response_s_cr_cr_vgpr_pr_sd_pd, "MR") |
      str_detect(x12_month_response_s_cr_cr_vgpr_pr_sd_pd, "SD")        ~ "SD",
    str_detect(x12_month_response_s_cr_cr_vgpr_pr_sd_pd, "PD")          ~ "PD",
    is.na(x12_month_response_s_cr_cr_vgpr_pr_sd_pd)                     ~ NA_character_,
    TRUE           ~ x12_month_response_s_cr_cr_vgpr_pr_sd_pd
  )) %>% 
    mutate(mon12_response_v2 = case_when(
      str_detect(mon12_response, "Died")                                 ~ "PD",
      is.na(mon12_response)                                              ~ NA_character_,
      TRUE           ~ mon12_response
    )) %>% 
    mutate(ORR_mon12 = case_when(
      str_detect(mon12_response_v2, "CR|PR")                             ~ "ORR",
      is.na(mon12_response_v2)                                           ~ NA_character_,
      TRUE                                                               ~ "PD/SD/MR"
    )) %>% 
    mutate(CRorbetter_mon12 = case_when(
      mon12_response_v2 == "sCR or CR"                        	       ~ "sCR or CR",
      is.na(mon12_response_v2)                                         ~ NA_character_,
      TRUE                                                             ~ "Others"
    )) %>% 
    # Best response
  mutate(ORR_as_best_response = case_when(
    ORR_30days == "ORR" |
    ORR_mon3 == "ORR" |
    ORR_mon6 == "ORR" |
    ORR_mon9 == "ORR" |
    ORR_mon12 == "ORR"                                            ~ "ORR",
    is.na(ORR_30days) & 
      is.na(ORR_mon3) & 
      is.na(ORR_mon6) & 
      is.na(ORR_mon9) &
      is.na(ORR_mon12)                                            ~ NA_character_, 
     TRUE                                                         ~ "SD/PD"
  )) %>% 
  mutate(best_CRorbetter = case_when(
     CRorbetter_30days == "sCR or CR" | 
       CRorbetter_mon3 == "sCR or CR" | 
       CRorbetter_mon6 == "sCR or CR" | 
       CRorbetter_mon9 == "sCR or CR" | 
       CRorbetter_mon12 == "sCR or CR"                            ~ "sCR or CR",
     is.na(CRorbetter_30days) & 
       is.na(CRorbetter_mon3) & 
       is.na(CRorbetter_mon6) & 
       is.na(CRorbetter_mon9) & 
       is.na(CRorbetter_mon12) ~ NA_character_, 
     TRUE                                                         ~ "< CR"
   ))  %>%
  mutate(best_response = case_when(
    day30_response_v2 == "sCR or CR" |
      mon3_response_v2 == "sCR or CR" |
      mon6_response_v2 == "sCR or CR" |
      mon9_response_v2 == "sCR or CR" |
      mon12_response_v2 == "sCR or CR"                            ~ "sCR or CR",
    day30_response_v2 == "VGPR" |
      mon3_response_v2 == "VGPR" |
      mon6_response_v2 == "VGPR" |
      mon9_response_v2 == "VGPR" |
      mon12_response_v2 == "VGPR"                                 ~ "VGPR",
    day30_response_v2 == "PR" |
      mon3_response_v2 == "PR" |
      mon6_response_v2 == "PR" |
      mon9_response_v2 == "PR" |
      mon12_response_v2 == "PR"                                   ~ "PR",
    day30_response_v2 == "SD" |
      mon3_response_v2 == "SD" |
      mon6_response_v2 == "SD" |
      mon9_response_v2 == "SD" |
      mon12_response_v2 == "SD"                                   ~ "SD",
    day30_response_v2 == "PD" |
      mon3_response_v2 == "PD" |
      mon6_response_v2 == "PD" |
      mon9_response_v2 == "PD" |
      mon12_response_v2 == "PD"                                   ~ "PD",
    TRUE                                                          ~ NA_character_
  )) %>% 
  mutate(best_ORRcollapsed = case_when(
    str_detect(best_response, "CR|PR")                            ~ "PR or better", 
    str_detect(best_response, "SD|PD")                            ~ "< PR",
    TRUE                                                          ~ NA_character_
  )) %>% 
  # MRD----
  mutate(across(c(contains("mrd_by_")),
            ~ case_when(
              . == "Positive"                                     ~ "Positive",
              . == "Negative"                                     ~ "Negative",
              TRUE                                                ~ NA_character_
            ))) %>%
  mutate(MRD_day30 = coalesce(day_30_mrd_by_clono_seq_positive_vs_negative, 
                               day_30_mrd_by_flow_positive_vs_negative
  )) %>%
  mutate(MRD_mon3 = coalesce(x3_month_mrd_by_clono_seq_positive_vs_negative,
                               x3_month_mrd_by_flow_positive_vs_negative
  )) %>% 
  mutate(MRD_mon6 = coalesce(x6_month_mrd_by_clono_seq_positive_vs_negative,
                               x6_month_mrd_by_flow_positive_vs_negative
  )) %>%
  mutate(MRD_mon9 = coalesce(x9_month_mrd_by_clono_seq_positive_vs_negative,
                               x9_month_mrd_by_flow_positive_vs_negative
  )) %>% 
  mutate(MRD_mon12 = coalesce(x12_month_mrd_by_clono_seq_positive_vs_negative,
                               x12_month_mrd_by_flow_positive_vs_negative
  )) %>% 
  mutate(response_MRD_day30 = case_when(
    day30_response_v2 == "sCR or CR" & 
      MRD_day30 == "Negative"                                     ~ "sCR or CR, MRD-",
    day30_response_v2 == "sCR or CR" & 
      MRD_day30 == "Positive"                                     ~ "sCR or CR, MRD+",
    day30_response_v2 == "sCR or CR" & 
      is.na(MRD_day30)                                            ~ "sCR or CR, MRD unknown",
    TRUE ~ day30_response_v2
  )) %>%
  mutate(response_MRD_mon3 = case_when(
    mon3_response_v2 == "sCR or CR" & 
      MRD_mon3 == "Negative"                                      ~ "sCR or CR, MRD-",
    mon3_response_v2 == "sCR or CR" & 
      MRD_mon3 == "Positive"                                      ~ "sCR or CR, MRD+",
    mon3_response_v2 == "sCR or CR" & 
      is.na(MRD_mon3)                                             ~ "sCR or CR, MRD unknown",
    TRUE ~ mon3_response_v2
  )) %>%
  mutate(response_MRD_mon6 = case_when(
    mon6_response_v2 == "sCR or CR" & 
      MRD_mon6 == "Negative"                                      ~ "sCR or CR, MRD-",
    mon6_response_v2 == "sCR or CR" & 
      MRD_mon6 == "Positive"                                      ~ "sCR or CR, MRD+",
    mon6_response_v2 == "sCR or CR" & 
      is.na(MRD_mon6)                                             ~ "sCR or CR, MRD unknown",
    TRUE ~ mon6_response_v2
  )) %>%
  mutate(response_MRD_mon9 = case_when(
    mon9_response_v2 == "sCR or CR" & 
      MRD_mon9 == "Negative"                                      ~ "sCR or CR, MRD-",
    mon9_response_v2 == "sCR or CR" & 
      MRD_mon9 == "Positive"                                      ~ "sCR or CR, MRD+",
    mon9_response_v2 == "sCR or CR" & 
      is.na(MRD_mon9)                                             ~ "sCR or CR, MRD unknown",
    TRUE ~ mon9_response_v2
  )) %>%
  mutate(response_MRD_mon12 = case_when(
    mon12_response_v2 == "sCR or CR" & 
      MRD_mon12 == "Negative"                                      ~ "sCR or CR, MRD-",
    mon12_response_v2 == "sCR or CR" & 
      MRD_mon12 == "Positive"                                      ~ "sCR or CR, MRD+",
    mon12_response_v2 == "sCR or CR" & 
      is.na(MRD_mon12)                                             ~ "sCR or CR, MRD unknown",
    TRUE ~ mon12_response_v2
  )) %>% 
  mutate(best_response_MRD = case_when (
    response_MRD_day30 == "sCR or CR, MRD-" | 
      response_MRD_mon3 == "sCR or CR, MRD-" | 
      response_MRD_mon6 == "sCR or CR, MRD-" | 
      response_MRD_mon9 == "sCR or CR, MRD-" | 
      response_MRD_mon12 == "sCR or CR, MRD-"                      ~ "sCR or CR, MRD-",
     response_MRD_day30 == "sCR or CR, MRD+" | 
      response_MRD_mon3 == "sCR or CR, MRD+" | 
      response_MRD_mon6 == "sCR or CR, MRD+" | 
      response_MRD_mon9 == "sCR or CR, MRD+" | 
      response_MRD_mon12 == "sCR or CR, MRD+"                      ~ "sCR or CR, MRD+",
     response_MRD_day30 == "sCR or CR, MRD unknown" | 
      response_MRD_mon3 == "sCR or CR, MRD unknown" | 
      response_MRD_mon6 == "sCR or CR, MRD unknown" | 
      response_MRD_mon9 == "sCR or CR, MRD unknown" | 
      response_MRD_mon12 == "sCR or CR, MRD unknown"               ~ "sCR or CR, MRD unknown",
    TRUE ~ best_response
  )) %>%
  # mutate(MRDneg_in_best_response = case_when(
  #   MRD_day30 ==  "Negative" |
  #     MRD_mon3 ==  "Negative" |
  #     MRD_mon6 ==  "Negative" |
  #     MRD_mon9 ==  "Negative" |
  #     MRD_mon12 ==  "Negative"                                    ~ "Negative",
  #   MRD_day30 ==  "Positive" |
  #   MRD_mon3 ==  "Positive" |
  #   MRD_mon6 ==  "Positive" |
  #   MRD_mon9 ==  "Positive" |
  #     MRD_mon12 ==  "Positive"                                    ~ "Positive",
#   TRUE                                                          ~ NA_character_
# )) %>% 
  mutate(bridging_response = case_when(
    response_to_bridging_chemotherapy == "MR" |
      response_to_bridging_chemotherapy == "SD/MR" |
      response_to_bridging_chemotherapy == "SD" |
      response_to_bridging_chemotherapy == "PD" |
      response_to_bridging_chemotherapy == "No" ~ "SD/PD",
    response_to_bridging_chemotherapy == "CR" |
      response_to_bridging_chemotherapy == "PR" |
      response_to_bridging_chemotherapy == "VGPR" ~ "PR or better",
    response_to_bridging_chemotherapy == "Not Evaluable" |
      response_to_bridging_chemotherapy == "ND" |
      response_to_bridging_chemotherapy == "Unknown" ~ NA_character_,
    TRUE ~ NA_character_ 
  ))
```

```{r infection}
cart6 <- cart6 %>%
  # Infection ----
  mutate(any_infection = infection_yes_1_no_0
  ) %>%
  mutate(any_sevinfection = severe_infection_need_for_iv_antibiotics_and_or_hospitalization_yes_1_no_0
  ) %>% 
  mutate(time_from_infusion_to_infection_days = 
           interval(
             start = date_of_car_t_infusion, 
             end = date_of_infection)/
           duration(n=1, units = "days"))

# cart7 <- cart6 %>% 
#   mutate(infection = case_when(
#     notinfused=="YES" ~ NA_character_,
#     notinfused=="NO" & is.na(`Infection Type`) & is.na(`Infection (Yes, No)`) ~ NA_character_,
#     notinfused=="NO" & `Infection Type` == "No" ~ "No", 
#     notinfused=="NO" & `Infection Type` != "No" ~ "Yes",
#     notinfused=="NO" & is.na(`Infection Type`) ~ "No",
#     TRUE ~ NA_character_
#   )) %>%
#   mutate(HLH = case_when(
#     notinfused=="YES" ~ NA_character_,
#     notinfused=="NO" & `HLH/HLH-Like Toxicities?`=="Yes" ~ "Yes",
#     TRUE ~ "No"
#   )) %>%
#   mutate(infection_type = case_when(
#     infection=="No" | is.na(infection) ~ NA_character_,
#     infection=="Yes" & `Infection Details (specific infection, ie. E.coli bacteremia, etc.)` == "fungus was unidentified" ~ "Fungal",
#     infection== "Yes" & `Infection Details (specific infection, ie. E.coli bacteremia, etc.)` == "E. coli bacteremia and Salmonella GI infection treated with ceftriaxone x 7 days" ~ "Bacterial",
#     infection=="Yes" & `Infection Type`=="bacterial + viral" ~ "Bacterial + Viral",
#     infection=="Yes" & `Infection Type`=="viral" ~ "Viral",
#     infection=="Yes" & `Infection Type`=="All of the above" ~ "Bacterial, Viral, and Fungal",
#     infection=="Yes" & `Infection Type`=="Bacterial; Viral; Bacterial" ~ "Bacterial + Viral",
#     infection=="Yes" & `Infection Type`=="Viral; Bacterial; Bacterial" ~ "Bacterial + Viral",
#     TRUE ~ `Infection Type`
#   )) %>% 
#   mutate(infection_severity = case_when(
#      infection=="No" | is.na(infection) ~ NA_character_,
#      infection=="Yes" & is.na(`Infection Severity (Severe = hospitalization or requiring IV antibiotics)`) ~ NA_character_,
#      infection=="Yes" & `Infection Severity (Severe = hospitalization or requiring IV antibiotics)`=="severe" ~ "Severe",
#      infection=="Yes" & `Infection Severity (Severe = hospitalization or requiring IV antibiotics)`=="Yes" ~ "Severe",
#      infection=="Yes" & `Infection Severity (Severe = hospitalization or requiring IV antibiotics)`=="non-severe" ~ "Not severe",
#     infection=="Yes" & `Infection Severity (Severe = hospitalization or requiring IV antibiotics)`=="Non-Severe" ~ "Not severe",
#     TRUE ~ `Infection Severity (Severe = hospitalization or requiring IV antibiotics)`
#   )) %>%
#   mutate(infection_type = factor(infection_type, levels = c("Bacterial", "Fungal", "Viral", "Bacterial + Fungal", "Bacterial + Viral", "Bacterial, Viral, and Fungal")))
```

```{r delayed NT}
cilta <- cart6 %>%
  mutate(what_non_parkinsonian_neurotoxicity = case_when(
    what_non_parkinsonian_neurotoxicity == "Delayed icans, hippocampal sclerosis/gliosis"
                                                                  ~ "cerebral edema/PRES",
    TRUE                                                          ~ what_non_parkinsonian_neurotoxicity
  )) %>% 
  mutate(other_delayed_non_parkinsonian_neurologic_toxicity_yes_no = case_when(
    is.na(other_delayed_non_parkinsonian_neurologic_toxicity_yes_no) 
                                                                  ~ "No",
    TRUE           ~ other_delayed_non_parkinsonian_neurologic_toxicity_yes_no
  )) %>%
  mutate(parkinsonian_neurotoxicity_yes_no = case_when(
    str_detect(pt_identifier, id_to_modify_neurotox$id)           ~ "No",
    parkinsonian_neurotoxicity_yes_no == "Yes" & 
      what_non_parkinsonian_neurotoxicity == "Bell's palsy"       ~ "No",
    is.na(parkinsonian_neurotoxicity_yes_no)                      ~ "No",
    TRUE                                                          ~ parkinsonian_neurotoxicity_yes_no
  )) %>%
  mutate(delayed_neuro = case_when(
    other_delayed_non_parkinsonian_neurologic_toxicity_yes_no == "Yes" |
      parkinsonian_neurotoxicity_yes_no == "Yes"                  ~ "Yes",
    TRUE                                                          ~ "No"
  )) %>%
  mutate(delayed_neuro_type = case_when(
    str_detect(pt_identifier, id_to_modify_neurotox$id)           ~ "PRES",
    what_non_parkinsonian_neurotoxicity == "cerebral edema/PRES"
                                                                  ~ "PRES",
    what_non_parkinsonian_neurotoxicity == "Aidp"                 ~ "Polyneuropathy",
    what_non_parkinsonian_neurotoxicity == 
      "Posterior reversible encephalopathy syndrome (pres) and parkinsonism" 
                                                                  ~ "Parkinsonian/PRES",
    what_non_parkinsonian_neurotoxicity == 
      "Posterior reversible encephalopathy syndrome (pres); progressive multifocal leukoencephalopathy (pml)" 
                                                                  ~ "PRES/PML",
    parkinsonian_neurotoxicity_yes_no == "Yes"                    ~ "Parkinsonian",
    other_delayed_non_parkinsonian_neurologic_toxicity_yes_no == "Yes" &
      str_detect(what_non_parkinsonian_neurotoxicity, "alsy")     ~ "Bell's palsy",
    other_delayed_non_parkinsonian_neurologic_toxicity_yes_no == "Yes" &
      str_detect(what_non_parkinsonian_neurotoxicity, "alsey")    ~ "Bell's palsy", 
    other_delayed_non_parkinsonian_neurologic_toxicity_yes_no == "Yes" &
      str_detect(what_non_parkinsonian_neurotoxicity, "droop")    ~ "Bell's palsy", 
    other_delayed_non_parkinsonian_neurologic_toxicity_yes_no == "Yes" &
      str_detect(what_non_parkinsonian_neurotoxicity, "PRES")    ~ "PRES", 
    other_delayed_non_parkinsonian_neurologic_toxicity_yes_no == "Yes" &
      str_detect(what_non_parkinsonian_neurotoxicity, "dysautonomia") 
                                                                 ~ "Dysautonomia",
    other_delayed_non_parkinsonian_neurologic_toxicity_yes_no == "Yes" &
      str_detect(what_non_parkinsonian_neurotoxicity, "plopia")    
                                                                  ~ "Diplopia",
    other_delayed_non_parkinsonian_neurologic_toxicity_yes_no == "Yes" &
      str_detect(what_non_parkinsonian_neurotoxicity, "Polyneuropathy") 
                                                                  ~ "Polyneuropathy",
    other_delayed_non_parkinsonian_neurologic_toxicity_yes_no == "Yes" &
      is.na(what_non_parkinsonian_neurotoxicity)                  ~ "Unknown type",
    TRUE                                                          ~ what_non_parkinsonian_neurotoxicity
  )) 
```

```{r titers}
# Titers ----
cilta <- cilta %>% 
  mutate(across(c(starts_with("ig_g_hsv1_2")), ~ case_when(
    . < 0.9         ~ "Negative",
    . >= 0.9 &
      . < 1.10      ~ "Indeterminate",
    . >= 1.10       ~ "Positive"
  ))) %>%

  # mutate(ig_g_hsv1_2_apheresis = case_when(
  #   ig_g_hsv1_2_apheresis < 0.9         ~ "Negative",
  #   ig_g_hsv1_2_apheresis >= 0.9 &
  #     ig_g_hsv1_2_apheresis < 1.10      ~ "Indeterminate",
  #   ig_g_hsv1_2_apheresis >= 1.10       ~ "Positive"
  # )) %>%
  # mutate(ig_g_hsv1_2_d90 = case_when(
  #   ig_g_hsv1_2_d90 < 0.9         ~ "Negative",
  #   ig_g_hsv1_2_d90 >= 0.9 &
  #     ig_g_hsv1_2_d90 < 1.10      ~ "Indeterminate",
  #   ig_g_hsv1_2_d90 >= 1.10       ~ "Positive"
  # )) %>%
  # mutate(ig_g_hsv1_2_d180 = case_when(
  #   ig_g_hsv1_2_d180 < 0.9         ~ "Negative",
  #   ig_g_hsv1_2_d180 >= 0.9 &
  #     ig_g_hsv1_2_d180 < 1.10      ~ "Indeterminate",
  #   ig_g_hsv1_2_d180 >= 1.10       ~ "Positive"
  # )) %>%
  mutate_at(c("ig_g_hsv1_2_apheresis",
              "ig_g_hsv1_2_d90",
              "ig_g_hsv1_2_d180"),
            ~ factor(., levels = c("Positive", "Indeterminate", "Negative"))) %>%

  # mutate_at(c("vzv_ig_g_apheresis",
  #             "vzv_ig_g_d90",
  #             "cmv_ig_g_apheresis",
  #             "cmv_ig_g_d90"),
  #           ~ case_when(
  #             . == 1                                              ~ "Positive",
  #             . == 0                                              ~ "Negative",
  #             . == "Equivocal"                                    ~ "Equivocal"
  #           )) %>%
  # mutate_at(c("vzv_ig_g_apheresis",
  #             "vzv_ig_g_d90",
  #             "cmv_ig_g_apheresis",
  #             "cmv_ig_g_d90"),
  #           ~ factor(., levels = c("Positive", "Negative", "Equivocal"))) %>%
  mutate(across(starts_with("pneumo"), ~ case_when(
    . <= 1                                                    ~ "Negative",
    . > 1                                                     ~ "Positive"
  ))) %>%
  mutate(igg_hsv_change = case_when(
    ig_g_hsv1_2_apheresis == "Positive" &
      ig_g_hsv1_2_d90 == "Negative"         ~ -1,
    ig_g_hsv1_2_apheresis == "Negative" &
      ig_g_hsv1_2_d90 == "Positive"         ~ 1,
    ig_g_hsv1_2_apheresis == "Negative" &
      ig_g_hsv1_2_d90 == "Negative"         ~ -0.25,
    ig_g_hsv1_2_apheresis == "Positive" &
      ig_g_hsv1_2_d90 == "Positive"         ~ 0.25
  )) %>%
  mutate(igg_vzv_change = case_when(
    vzv_ig_g_apheresis == "Positive" &
      vzv_ig_g_d90 == "Negative"               ~ -1,
    vzv_ig_g_apheresis == "Negative" &
      vzv_ig_g_d90 == "Positive"               ~ 1,
    vzv_ig_g_apheresis == "Negative" &
      vzv_ig_g_d90 == "Negative"               ~ -0.25,
    vzv_ig_g_apheresis == "Positive" &
      vzv_ig_g_d90 == "Positive"               ~ 0.25
  )) %>%
  mutate(igg_cmv_change = case_when(
    cmv_ig_g_apheresis == "Positive" &
      cmv_ig_g_d90 == "Negative"               ~ -1,
    cmv_ig_g_apheresis == "Negative" &
      cmv_ig_g_d90 == "Positive"               ~ 1,
    cmv_ig_g_apheresis == "Negative" &
      cmv_ig_g_d90 == "Negative"               ~ -0.25,
    cmv_ig_g_apheresis == "Positive" &
      cmv_ig_g_d90 == "Positive"               ~ 0.25
  ))

# Create Pneumo titers change value
for (i in colnames(cilta %>%
                   select(c(starts_with("pneumo") & 
                            (ends_with("apheresis") | ends_with("d90"))
                   )))) {

  pneumo_type <-
    str_match(i, "(.*)(apheresis|d90)")[, 1] %>%
    str_remove(., "_apheresis|_d90")

  pneumo_subset <- cilta %>% select(pt_identifier, starts_with(pneumo_type) & 
                                      (ends_with("apheresis") | ends_with("d90")))

  name <- paste0(pneumo_type, "_", "change")

  cilta <- cilta %>% 
    mutate(!!paste0(pneumo_type, "_", "change") := case_when(
      pneumo_subset[2] == "Positive" &
        pneumo_subset[3] == "Negative"               ~ -1,
      pneumo_subset[2] == "Negative" &
        pneumo_subset[3] == "Positive"               ~ 1,
      pneumo_subset[2] == "Negative" &
        pneumo_subset[3] == "Negative"               ~ -0.25,
      pneumo_subset[2] == "Positive" &
        pneumo_subset[3] == "Positive"               ~ 0.25
    ))

}
```

```{r Labeling}
## Labeling ----
var_label(cilta) <- list(age = "Patient age", agecat = "Patient age - Categories", 
                         male_sex = "Male Sex", sex = "Sex",
                         race_eth = "Race and ethnicity",
                         os_from_infusion_months = "Follow-up time from infusion (months)",
                         extramedullary_disease_yes_no = "Extramedullary disease",
                         bm_plasma_cell_percent_at_pre_ld = "Circulating Plasma Cells?",
                         high_marrow_burden_50_percent = "High marrow burden (≥50%)",
                         ecog_at_ld = "ECOG at LD",
                         # `ECOG > or = 2? (Yes, No)` ,
                         r_iss_at_car_t_infusion = "R-ISS at cart infusion",
                         cytogenetics = "High risk cytogenetics (not include gain 1q)", 
                         deletion_17p_prior_to_car_t_infusion_yes_no = "Deletion 17p at Infusion", 
                         t_4_14_prior_to_car_t_infusion_yes_no = "t(4:14) at Infusion", 
                         t_14_16_prior_to_car_t_infusion_yes_no = "t(4:16) at Infusion", 
                         gain_amp1q21 = "Gain/amp 1q21",
                         number_of_prior_lines_of_therapy = "Number of prior lines of therapy",
                         number_prior_lines_therapy = "Number of prior lines of therapy (cat)",
                         bridging_therapy_yes_no = "Bridging Therapy",
                         prior_auto_sct_yes_no = "Prior autoSCT", 
                         prior_allo_sct_within_6_months_before_apheresis = "Prior alloSCT",
                         # prior_treatment_with_any_gene_therapy_based_therapeutic_or_investigational_cellular_therapy_or_bcma_targeted_therapy_yes_no = "Prior BCMA-directed therapy",
                         Prior_bcma_therapy = "Prior BCMA-directed therapy",
                         need_for_g_csf_post_infus_yes_1_no_0 = "Granulocyte colony stimulating factor (G-CSF)",
                         need_for_ivig_yes_1_no_0 = "Intravenous immunoglobulin (IVIG)",
                         cd34_stem_cell_boost_yes_1_no_0 = "CD34 stem cell boost",
                         day_of_neutrophil_recovery = "Day of neutrophil recovery",
                         transfusion_postCART = "Transfusion post-CAR T",
                         tpo_agonist_post_cart_yes_1_no_0 = "Thrombopoietin (TPO) agonist",
                         # bmicat = "BMI at apheresis - Normal, Overweight, Obese", 
                         steroid_use_yes_no = "Steroid Use", tocilizumab_yes_no = "Tocilizumab Use", 
                         anakinra_yes_no = "Anakinra Use",
                         proteasome = "Refractory to Proteasome Inhibitors", 
                         lenalidomide_exposed_vs_refractory_vs_not_exposed = "Refractory to Lenalidomide",
                         bortezomib_exposed_vs_refractory_vs_not_exposed = "Refractory to Bortezomib", 
                         carfilzomib_exposed_vs_refractory_vs_not_exposed = "Refractory to Carfilzomib", 
                         pomalidomide_exposed_vs_refractory_vs_not_exposed = "Refractory to Pomalidomide",
                         dara = "Refractory to Daratumumab or Isatuximab",
                         immuno = "Refractory to Immunomodulators", doubleref = "Double refractory", 
                         tripleref= "Triple refractory", pentaref= "Penta refractory",
                         # bmicat2 = "BMI at apheresis - Normal vs. Overweight/Obese", 
                         CRS_any = "Any CRS", CRS_gradecat = "CRS Grade", 
                         ICANS_any = "Any ICANS", ICANS_gradecat = "ICANS Grade", 
                         day_of_onset_crs_relative_to_infusion = "Day of onset of CRS relative to infusion",
                         day_of_max_crs_relative_to_infusion = "Day of Max CRS (relative to infusion)",
                         duration_of_crs_days = "Duration of CRS (days)", 
                         day_of_onset_icans_relative_to_infusion = "Day of onset of ICANS relative to infusion",
                         day_of_max_icans_relative_to_infusion = "Day of Max ICANS (relative to infusion)",
                         duration_of_icans_days = "Duration of ICANS (days)", 
                         delayed_neuro = "Delayed Neurotoxicity",
                         parkinsonian_neurotoxicity_yes_no = "Parkinsonian Neurotoxicity",
                         other_delayed_non_parkinsonian_neurologic_toxicity_yes_no = 
                           "Other/Delayed Non-Parkinsonian Neurologic Toxicity",
                         delayed_neuro_type = "Delayed Neurotoxicity Type",

                         tbv = "TBV", whole_blood_processed_m_l = "Whole blood processed (mL)",
                         final_collection_volume_m_l = "Final collection volume (mL)", 
                         tbv_processed = "TBV processed", 
                         access_type = "Access type", total_bags = "Total bags",
                         run_time_min = "Run time (min)",
                         time_from_apheresis_to_car_t_infusion_days = "Time from apheresis to CAR-T infusion (days)", 
                         total_dose_volume_m_l = "Total dose volume (mL)", 
                         cell_dose_million_cells = "Total cell dose (million cells)",
                         best_ORRcollapsed = "Best ORR at ≤ 12 months",
                         did_patient_meet_eligibility_criteria_for_cartitude_1_on_the_day_of_apheresis_yes_no =
                           "Did patient meet cartitude criteria"#,
                         # best_response = "Best response in first 90 days"
)

```
<br>

### Tables
Patients evaluable at each time point
```{r evaluable}
day7 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  filter(os_from_infusion_months > (7 / 30.417)) %>% 
  select(pt_identifier) %>% 
  tbl_summary(missing = "no") %>% 
  bold_labels() %>% add_stat_label()
day14 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  filter(os_from_infusion_months > (14 / 30.417)) %>% 
  select(pt_identifier) %>% 
  tbl_summary(missing = "no") %>% 
  bold_labels() %>% add_stat_label()
day21 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  filter(os_from_infusion_months > (21 / 30.417)) %>% 
  select(pt_identifier) %>% 
  tbl_summary(missing = "no") %>% 
  bold_labels() %>% add_stat_label()
day30 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  filter(os_from_infusion_months > (30 / 30.417)) %>% 
  select(pt_identifier) %>% 
  tbl_summary(missing = "no") %>% 
  bold_labels() %>% add_stat_label()
day90 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  filter(os_from_infusion_months > 3) %>% 
  select(pt_identifier) %>% 
  tbl_summary(missing = "no") %>% 
  bold_labels() %>% add_stat_label()
day180 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  filter(os_from_infusion_months > 6) %>% 
  select(pt_identifier) %>% 
  tbl_summary(missing = "no") %>% 
  bold_labels() %>% add_stat_label()
tbl_merge(list(day7, day14, day21, day30, day90, day180), 
          tab_spanner = c("day7", "day14", "day21", "day30", "day90", "day180"))
```

```{r infused}
cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  select(`infused?` = notinfused) %>% 
  tbl_summary(missing = "no") %>% 
  bold_labels() %>% add_stat_label()
```

## Table 1. Patient baseline characteristics
```{r Table 1}
cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  select(os_from_infusion_months, age, agecat, male_sex, sex,
         extramedullary_disease_yes_no, 
         high_marrow_burden_50_percent, ecog_at_ld, 
         r_iss_at_car_t_infusion, 
         cytogenetics, deletion_17p_prior_to_car_t_infusion_yes_no,
         t_4_14_prior_to_car_t_infusion_yes_no,
         t_14_16_prior_to_car_t_infusion_yes_no, gain_amp1q21,
         bridging_therapy_yes_no,
         # `Received alkylating therapy with bridging (Yes=1, No=0)`,
         prior_auto_sct_yes_no, prior_allo_sct_within_6_months_before_apheresis,
         Prior_bcma_therapy,
         immuno, proteasome, dara, doubleref, tripleref, pentaref, 
         did_patient_meet_eligibility_criteria_for_cartitude_1_on_the_day_of_apheresis_yes_no) %>% 
  tbl_summary(
    type = list(
      c(agecat, sex, 
        extramedullary_disease_yes_no, high_marrow_burden_50_percent,
        cytogenetics, deletion_17p_prior_to_car_t_infusion_yes_no,
        t_4_14_prior_to_car_t_infusion_yes_no,
        t_14_16_prior_to_car_t_infusion_yes_no, gain_amp1q21,
        bridging_therapy_yes_no, prior_auto_sct_yes_no,
        Prior_bcma_therapy,
        # `Received alkylating therapy with bridging (Yes=1, No=0)`,
        immuno, proteasome, dara, doubleref, tripleref, pentaref,
        did_patient_meet_eligibility_criteria_for_cartitude_1_on_the_day_of_apheresis_yes_no)
      ~"categorical"),

    statistic = list(all_continuous() ~ "{median} ({min}, {max})"), 
    digits = all_continuous() ~ 1
  ) %>% 
  bold_labels() %>% add_stat_label() %>% 
  modify_footnote(update = everything() ~ NA)
```

## Table 2. Clinical outcomes
```{r}
tbl1 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  select(max_crs_grade_0_5, CRS_gradecat,
         day_of_onset_crs_relative_to_infusion,
         day_of_max_crs_relative_to_infusion,
         duration_of_crs_days, 
         max_icans, ICANS_gradecat,
         day_of_onset_icans_relative_to_infusion,
         day_of_max_icans_relative_to_infusion,
         duration_of_icans_days,
         delayed_neuro,
         parkinsonian_neurotoxicity_yes_no,
         other_delayed_non_parkinsonian_neurologic_toxicity_yes_no,
         delayed_neuro_type) %>% 
  tbl_summary(type = list(
    c(day_of_onset_crs_relative_to_infusion,
         day_of_max_crs_relative_to_infusion,
         duration_of_crs_days,
      day_of_onset_icans_relative_to_infusion,
         day_of_max_icans_relative_to_infusion,
         duration_of_icans_days) ~ "continuous",
    c(delayed_neuro,
         parkinsonian_neurotoxicity_yes_no,
         other_delayed_non_parkinsonian_neurologic_toxicity_yes_no) 
    ~ "categorical"),
    statistic=list(all_continuous() ~ "{median} ({min}, {max})"),
    digits = all_continuous() ~ 1,
    missing = "no"
    ) %>% 
  bold_labels() %>% add_stat_label()

tbl2 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  select(steroid_use_yes_no, tocilizumab_yes_no, 
         anakinra_yes_no) %>% 
  tbl_summary(type = list(
    c(steroid_use_yes_no, tocilizumab_yes_no, 
      anakinra_yes_no) ~ "categorical"),
    statistic=list(all_continuous() ~ "{median} ({min}, {max})"), 
    digits = all_continuous() ~ 1
    ) %>% 
  bold_labels() %>% add_stat_label()

tbl3 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  select(MRD_day30,
         ORR_30days,
         day30_response,
         day30_response_v2) %>% 
  tbl_summary(type = list(
    c(MRD_day30) ~ "categorical"),
    digits = all_continuous() ~ 1,
    missing = "no") %>% 
  bold_labels() %>% add_stat_label()

tbl4 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  filter(day30_response_v2 == "sCR or CR") %>% 
  select(MRD_day30) %>% 
  tbl_summary(type = list(
    c(MRD_day30) ~ "categorical"),
    label = list(MRD_day30 ~ 
                   "MRD_day30 within CR/sCR"),
    digits = all_continuous() ~ 1,
    missing = "no") %>% 
  bold_labels() %>% add_stat_label()

tbl5 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  select(MRD_mon3,
         ORR_mon3,
         mon3_response,
         mon3_response_v2) %>% 
  tbl_summary(type = list(
    c(MRD_mon3) ~ "categorical"),
    digits = all_continuous() ~ 1,
    missing = "no") %>% 
  bold_labels() %>% add_stat_label()

tbl6 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  filter(mon3_response_v2 == "sCR or CR") %>% 
  select(MRD_mon3) %>% 
  tbl_summary(type = list(c(MRD_mon3) ~ "categorical"),
              label = list(MRD_mon3 ~ "MRD_mon3 within CR/sCR"),
              digits = all_continuous() ~ 1,
              missing = "no") %>% 
  bold_labels() %>% add_stat_label()

tbl7 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  select(ORR_as_best_response, best_ORRcollapsed,
         best_response, best_response_MRD) %>% 
  tbl_summary(#type = list(c(MRD_mon3) ~ "categorical"),
              # label = list(best_response ~ "best response in the first 90 days"),
              digits = all_continuous() ~ 1,
              missing = "no") %>% 
  bold_labels() %>% add_stat_label()

tbl8 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  filter(best_response == "sCR or CR") %>% 
  select(best_response_MRD) %>% 
  tbl_summary(type = list(c(best_response_MRD) ~ "categorical"),
              label = list(best_response_MRD ~ "best_response_MRD within CR/sCR"),
              statistic=list(all_continuous() ~ "{median} ({min}, {max})"), 
              digits = all_continuous() ~ 1,
              missing = "no") %>% 
  bold_labels() %>% add_stat_label()

tbl_stack(list(tbl1, tbl2, tbl3, tbl4, tbl5, tbl6, tbl7, tbl8)) %>% as_gt() #%>%
  # gt::tab_source_note(gt::md("**At 30days: <br>
  #                            1 patient is not evaluable by IMWG - non-secretory disease, including no bone marrow involvement; <br>
  #                            1 patients was not evaluated. <br>
  #                            At 90days: <br>
  #                            1 patient is not evaluable by IMWG - non-secretory disease, including no bone marrow involvement; <br>
  #                            1 patient is not Heme not evaluable; <br>
  #                            2 patients were not evaluated; <br>
  #                            1 patients is not evaluable by death**"))

```

<!-- For responses make denominator evaluable patients -->
<!-- Patient who died or was lost to follow-up or had non-secretory disease would not be evaluable -->

## Table S3. Apheresis and product characteristics
```{r Apheresis and product characteristics}
cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  select(tbv, whole_blood_processed_m_l,
         final_collection_volume_m_l, 
         tbv_processed, 
         access_type,
         run_time_min,
         time_from_apheresis_to_car_t_infusion_days, total_bags,
         total_dose_volume_m_l, cell_dose_million_cells
  ) %>% 
  tbl_summary(type = list(
    c(tbv, total_bags) ~ "continuous"),
    statistic=list(all_continuous() ~ "{median} ({min}, {max})"), 
    digits = all_continuous() ~ 1,
  missing = "no") %>% 
  bold_labels() %>% add_stat_label()
```

## Table 3. Cytopenias following ide-cel
```{r Table 3}
tbl1 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  select("Leukopenia" = Leukopenia_apheresis, 
         "Grade3+ Leukopenia" = gr3_Leukopenia_apheresis, 
         "Neutropenia" = Neutropenia_apheresis, 
         "Grade3+ Neutropenia" = gr3_Neutropenia_apheresis, 
         "Anemia" = Anemia_apheresis, 
         "Grade3+ Anemia" = gr3_Anemia_apheresis, 
         "Thrombocytopenia" = Thrombocytopenia_apheresis, 
         "Grade3+ Thrombocytopenia" = gr3_Thrombocytopenia_apheresis,
         "Any Cytopenia" = cytopenia_apheresis,
         "Grade3+ Cytopenia" = gr3_cytopenia_apheresis
  ) %>% 
  tbl_summary(type = list(everything() ~ "dichotomous"),
              digits = all_continuous() ~ 1,
              missing = "no"
  ) %>% 
  bold_labels() %>% add_stat_label()

tbl2 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  select("Leukopenia" = `Leukopenia_-5d`, 
         "Grade3+ Leukopenia" = `gr3_Leukopenia_-5d`, 
         "Neutropenia" = `Neutropenia_-5d`, 
         "Grade3+ Neutropenia" = `gr3_Neutropenia_-5d`,
         "Anemia" = `Anemia_-5d`, 
         "Grade3+ Anemia" = `gr3_Anemia_-5d`, 
         "Thrombocytopenia" = `Thrombocytopenia_-5d`, 
         "Grade3+ Thrombocytopenia" = `gr3_Thrombocytopenia_-5d`,
         "Any Cytopenia" = `cytopenia_-5d`,
         "Grade3+ Cytopenia" = `gr3_cytopenia_-5d`
  ) %>% 
  tbl_summary(type = list(everything() ~ "dichotomous"),
              digits = all_continuous() ~ 1,
              missing = "no"
  ) %>% 
  bold_labels() %>% add_stat_label()

tbl3 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  select("Leukopenia" = Leukopenia_0d, 
         "Grade3+ Leukopenia" = gr3_Leukopenia_0d, 
         "Neutropenia" = Neutropenia_0d, 
         "Grade3+ Neutropenia" = gr3_Neutropenia_0d, 
         "Anemia" = Anemia_0d, 
         "Grade3+ Anemia" = gr3_Anemia_0d,
         "Thrombocytopenia" = Thrombocytopenia_0d, 
         "Grade3+ Thrombocytopenia" = gr3_Thrombocytopenia_0d,
         "Any Cytopenia" = cytopenia_0d,
         "Grade3+ Cytopenia" = gr3_cytopenia_0d
  ) %>% 
  tbl_summary(type = list(everything() ~ "dichotomous"),
              digits = all_continuous() ~ 1,
              missing = "no"
  ) %>% 
  bold_labels() %>% add_stat_label()

tbl4 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  select("Leukopenia" = Leukopenia_7d, 
         "Grade3+ Leukopenia" = gr3_Leukopenia_7d, 
         "Neutropenia" = Neutropenia_7d, 
         "Grade3+ Neutropenia" = gr3_Neutropenia_7d, 
         "Anemia" = Anemia_7d, 
         "Grade3+ Anemia" = gr3_Anemia_7d, 
         "Thrombocytopenia" = Thrombocytopenia_7d, 
         "Grade3+ Thrombocytopenia" = gr3_Thrombocytopenia_7d,
         "Any Cytopenia" = cytopenia_7d,
         "Grade3+ Cytopenia" = gr3_cytopenia_7d
  ) %>% 
  tbl_summary(type = list(everything() ~ "dichotomous"),
              digits = all_continuous() ~ 1,
              missing = "no"
  ) %>% 
  bold_labels() %>% add_stat_label()

tbl5 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  # filter(is.na(Not_evaluable_at_14days)) %>% ######################### Fix evaluable---------- here
  filter(os_from_infusion_months > (14 / 30.417)) %>% 
select("Leukopenia" = Leukopenia_14d, 
       "Grade3+ Leukopenia" = gr3_Leukopenia_14d, 
       "Neutropenia" = Neutropenia_14d, 
       "Grade3+ Neutropenia" = gr3_Neutropenia_14d,
       "Anemia" = Anemia_14d, 
       "Grade3+ Anemia" = gr3_Anemia_14d,
       "Thrombocytopenia" = Thrombocytopenia_14d, 
       "Grade3+ Thrombocytopenia" = gr3_Thrombocytopenia_14d,
       "Any Cytopenia" = cytopenia_14d,
       "Grade3+ Cytopenia" = gr3_cytopenia_14d
) %>% 
  tbl_summary(type = list(everything() ~ "dichotomous"),
              digits = all_continuous() ~ 1,
              missing = "no"
  ) %>% 
  bold_labels() %>% add_stat_label()

tbl6 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  # filter(is.na(Not_evaluable_at_21days)) %>% 
  filter(os_from_infusion_months > (21 / 30.417)) %>% 
  select("Leukopenia" = Leukopenia_21d,
         "Grade3+ Leukopenia" = gr3_Leukopenia_21d, 
         "Neutropenia" = Neutropenia_21d, 
         "Grade3+ Neutropenia" = gr3_Neutropenia_21d, 
         "Anemia" = Anemia_21d, 
         "Grade3+ Anemia" = gr3_Anemia_21d,
         "Thrombocytopenia" = Thrombocytopenia_21d, 
         "Grade3+ Thrombocytopenia" = gr3_Thrombocytopenia_21d,
         "Any Cytopenia" = cytopenia_21d,
         "Grade3+ Cytopenia" = gr3_cytopenia_21d
  ) %>% 
  tbl_summary(type = list(everything() ~ "dichotomous"),
              digits = all_continuous() ~ 1,
              missing = "no"
  ) %>% 
  bold_labels() %>% add_stat_label()

tbl7 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  # filter(is.na(Not_evaluable_at_30days)) %>% 
  filter(os_from_infusion_months > (30 / 30.417)) %>% 
  select("Leukopenia" = Leukopenia_30d, 
         "Grade3+ Leukopenia" = gr3_Leukopenia_30d,
         "Neutropenia" = Neutropenia_30d, 
         "Grade3+ Neutropenia" = gr3_Neutropenia_30d,
         "Anemia" = Anemia_30d, 
         "Grade3+ Anemia" = gr3_Anemia_30d, 
         "Thrombocytopenia" = Thrombocytopenia_30d,
         "Grade3+ Thrombocytopenia" = gr3_Thrombocytopenia_30d,
         "Any Cytopenia" = cytopenia_30d,
         "Grade3+ Cytopenia" = gr3_cytopenia_30d
  ) %>% 
  tbl_summary(type = list(everything() ~ "dichotomous"),
              digits = all_continuous() ~ 1,
              missing = "no"
  ) %>% 
  bold_labels() %>% add_stat_label()

tbl8 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  # filter(is.na(Not_evaluable_at_90days)) %>% 
  filter(os_from_infusion_months > 3) %>% 
  select("Leukopenia" = Leukopenia_90d, 
         "Grade3+ Leukopenia" = gr3_Leukopenia_90d,
         "Neutropenia" = Neutropenia_90d, 
         "Grade3+ Neutropenia" = gr3_Neutropenia_90d,
         "Anemia" = Anemia_90d,
         "Grade3+ Anemia" = gr3_Anemia_90d,
         "Thrombocytopenia" = Thrombocytopenia_90d, 
         "Grade3+ Thrombocytopenia" = gr3_Thrombocytopenia_90d,
         "Any Cytopenia" = cytopenia_90d,
         "Grade3+ Cytopenia" = gr3_cytopenia_90d
  ) %>% 
  tbl_summary(type = list(everything() ~ "dichotomous"),
              digits = all_continuous() ~ 1,
              missing = "no"
  ) %>% 
  bold_labels() %>% add_stat_label()

tbl9 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  # filter(is.na(Not_evaluable_at_180days)) %>% 
  filter(os_from_infusion_months > 6) %>% 
  select("Leukopenia" = Leukopenia_180d, 
         "Grade3+ Leukopenia" = gr3_Leukopenia_180d,
         "Neutropenia" = Neutropenia_180d, 
         "Grade3+ Neutropenia" = gr3_Neutropenia_180d,
         "Anemia" = Anemia_180d,
         "Grade3+ Anemia" = gr3_Anemia_180d,
         "Thrombocytopenia" = Thrombocytopenia_180d, 
         "Grade3+ Thrombocytopenia" = gr3_Thrombocytopenia_180d,
         "Any Cytopenia" = cytopenia_180d,
         "Grade3+ Cytopenia" = gr3_cytopenia_180d
  ) %>% 
  tbl_summary(type = list(everything() ~ "dichotomous"),
              digits = all_continuous() ~ 1,
              missing = "no"
  ) %>% 
  bold_labels() %>% add_stat_label()

tbl_merge(list(tbl1, tbl2, tbl3, tbl4, tbl5, tbl6, tbl7, tbl8, tbl9), 
          tab_spanner = c("Apheresis", "Day -5", "Day 0", 
                          "Day 7", "Day 14", "Day 21",
                          "Day 30", "Day 90", "Day 180")) %>% as_gt() #%>%
# gt::tab_source_note(gt::md("**Not evaluable patients are removed from the denominator <br>
#                            1 patient has missing 90 days values**"))
```

## Table 4. Bone marrow biopsy findings
```{r Table 4}
tbl1 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  select("marrow cellularity" = marrow_cellularity_baseline_normo_hypo_hyper, 
         "marrow cellularity (%)" = marrow_cellularity_baseline_percent,
         "% CD138 positive by IHC" = marrow_cd138_baseline_percent,
         "is cellularity variable?"= contains("marrow_cellularity_baseline") & contains("is_variable_"),
         "Fibrosis grade" = marrow_fibrosis_baseline_0_1_2_3, 
         "Dysplasia present" = marrow_dysplasia_baseline_yes_no
  ) %>% 
  tbl_summary(type = list(
    c("marrow cellularity (%)", "% CD138 positive by IHC",
      "Fibrosis grade") ~ "continuous",
    c("Dysplasia present") ~ "categorical"
  ),
  statistic=list(all_continuous() ~ "{median} ({min}, {max})"), 
  digits = all_continuous() ~ 1,
  missing = "no") %>% 
  bold_labels() %>% add_stat_label()

tbl2 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  filter(os_from_infusion_months > (30 / 30.417)) %>% 
  select("marrow cellularity" = marrow_cellularity_day_30_normo_hypo_hyper, 
         "marrow cellularity (%)" = marrow_percent_cellularity_day_30,
         "% CD138 positive by IHC" = marrow_percent_cd138_day_30, 
         "is cellularity variable?"= contains("marrow_cellularity_day_30") & contains("is_variable_"),
         "Fibrosis grade" = marrow_fibrosis_day_30_0_1_2_3, 
         "Dysplasia present" = marrow_dysplasia_day_30_yes_no
  ) %>% 
  tbl_summary(type = list(
    c("marrow cellularity (%)", "% CD138 positive by IHC",
      "Fibrosis grade") ~ "continuous",
    c("Dysplasia present") ~ "categorical"
  ),
  statistic=list(all_continuous() ~ "{median} ({min}, {max})"), 
  digits = all_continuous() ~ 1,
  missing = "no") %>% 
  bold_labels() %>% add_stat_label()

tbl3 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  filter(os_from_infusion_months > 3) %>% 
  select("marrow cellularity" = marrow_cellularity_day_90_normo_hypo_hyper, 
         "marrow cellularity (%)" = marrow_percent_cellularity_day_90,
         "% CD138 positive by IHC" = marrow_percent_cd138_day_90, 
         "is cellularity variable?"= contains("marrow_cellularity_day_90") & contains("is_variable_"),
         "Fibrosis grade" = marrow_fibrosis_day_90_0_1_2_3, 
         "Dysplasia present" = marrow_dysplasia_day_90_yes_no
  ) %>% 
  tbl_summary(type = list(
    c("marrow cellularity (%)", "% CD138 positive by IHC",
      "Fibrosis grade") ~ "continuous",
    c("Dysplasia present") ~ "categorical"
  ),
  statistic=list(all_continuous() ~ "{median} ({min}, {max})"), 
  digits = all_continuous() ~ 1,
  missing = "no") %>% 
  bold_labels() %>% add_stat_label()

tbl4 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  filter(os_from_infusion_months > 6) %>% 
  select("marrow cellularity" = marrow_cellularity_day_180_normo_hypo_hyper, 
         "marrow cellularity (%)" = marrow_percent_cellularity_day_180,
         "% CD138 positive by IHC" = marrow_percent_cd138_day_180, 
         "is cellularity variable?"= contains("marrow_cellularity_day_180") & contains("is_variable_"),
         "Fibrosis grade" = marrow_fibrosis_day_180_0_1_2_3, 
         "Dysplasia present" = marrow_dysplasia_day_180_yes_no
  ) %>% 
  tbl_summary(type = list(
    c("marrow cellularity (%)", "% CD138 positive by IHC",
      "Fibrosis grade") ~ "continuous",
    c("Dysplasia present") ~ "categorical"
  ),
  statistic=list(all_continuous() ~ "{median} ({min}, {max})"), 
  digits = all_continuous() ~ 1,
  missing = "no") %>% 
  bold_labels() %>% add_stat_label()

tbl_merge(list(tbl1, tbl2, tbl3, tbl4), 
          tab_spanner = c("Baseline", "Day 30", "Day 90", "Day 180"))
```

## Table 5. Supportive therapies and engraftment

```{r Table 5. Supportive therapies and engraftment}
tbl1 <- cilta %>%
  distinct(pt_identifier, .keep_all = TRUE) %>%
  select(need_for_g_csf_post_infus_yes_1_no_0) %>%
  tbl_summary() %>%
  bold_labels() %>% add_stat_label()

tbl1_1 <- cilta %>%
  distinct(pt_identifier, .keep_all = TRUE) %>%
  select(first_day_of_gcsf, last_day_of_gcsf) %>%
  tbl_summary(type = list(
    everything() ~ "continuous"),
    statistic=list(all_continuous() ~ "{median} ({min}, {max})"),
    digits = all_continuous() ~ 1,
    missing = "no") %>%
  add_stat_label()

tbl2 <- cilta %>%
  distinct(pt_identifier, .keep_all = TRUE) %>%
  select(cd34_stem_cell_boost_yes_1_no_0) %>%
  tbl_summary() %>%
  bold_labels() %>% add_stat_label()

tbl2_1 <- cilta %>%
  distinct(pt_identifier, .keep_all = TRUE) %>%
  select(day_of_stem_cell_boost, boost_cell_dose_million_cells) %>%
  tbl_summary(type = list(
    everything() ~ "continuous"),
    statistic=list(all_continuous() ~ "{median} ({min}, {max})"),
    digits = all_continuous() ~ 1,
    missing = "no") %>%
  add_stat_label()

tbl3 <- cilta %>%
  distinct(pt_identifier, .keep_all = TRUE) %>%
  select(day_of_neutrophil_recovery) %>%
  tbl_summary(statistic=list(all_continuous() ~ "{median} ({min}, {max})"),
    digits = all_continuous() ~ 1,
    missing = "no") %>%
  bold_labels() %>% add_stat_label()

tbl3_1 <- cilta %>%
  distinct(pt_identifier, .keep_all = TRUE) %>%
  select(neutrophil_recovery_ANC_never_below_500, neutrophil_recovery_did_not_recover) %>%
  tbl_summary(type = list(
    everything() ~ "dichotomous"),
    missing = "no") %>%
  add_stat_label()

tbl4 <- cilta %>%
  distinct(pt_identifier, .keep_all = TRUE) %>%
  select(transfusion_postCART) %>%
  tbl_summary(type = list(everything() ~ "dichotomous"),
    missing = "no") %>%
  bold_labels() %>% add_stat_label()

tbl4_1 <- cilta %>%
  distinct(pt_identifier, .keep_all = TRUE) %>%
  select(rbc_transfusion_within_30d_post_cart_yes_1_no_0, plt_transfusion_within_30d_post_cart_yes_1_no_0,
         rbc_transfusion_30d_post_cart_yes_1_no_0, plt_transfusion_30d_post_cart_yes_1_no_0) %>%
  tbl_summary(
    missing = "no") %>%
  add_stat_label()

tbl5 <- cilta %>%
  distinct(pt_identifier, .keep_all = TRUE) %>%
  select(tpo_agonist_post_cart_yes_1_no_0) %>%
  tbl_summary(
    missing = "no") %>%
  bold_labels() %>% add_stat_label()

tbl5_1 <- cilta %>%
  distinct(pt_identifier, .keep_all = TRUE) %>%
  select(first_day_of_tpo_agonist, last_day_of_tpo_agonist) %>%
  tbl_summary(type = list(
    everything() ~ "continuous"),
    statistic=list(all_continuous() ~ "{median} ({min}, {max})"),
    digits = all_continuous() ~ 1,
    missing = "no") %>%
  add_stat_label()

tbl6 <- cilta %>%
  distinct(pt_identifier, .keep_all = TRUE) %>%
  select(need_for_ivig_yes_1_no_0) %>%
  tbl_summary() %>%
  bold_labels() %>% add_stat_label()

tbl6_1 <- cilta %>%
  distinct(pt_identifier, .keep_all = TRUE) %>%
  select(first_day_of_ivig, last_day_of_ivig) %>%
  tbl_summary(type = list(
    everything() ~ "continuous"),
    statistic=list(all_continuous() ~ "{median} ({min}, {max})"),
    digits = all_continuous() ~ 1,
    missing = "no") %>%
  add_stat_label()

tbl_stack(list(tbl1, tbl1_1, tbl2, tbl2_1, tbl3, tbl3_1,
               tbl4, tbl4_1, tbl5, tbl5_1, tbl6, tbl6_1))
```

## Patient characteristics by grade 3+ cytopenia at Day 30
```{r Patient characteristics by grade 3+ cytopenia at Day 30}
cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  # filter(is.na(Not_evaluable_at_30days)) %>% ######################### Fix evaluable----------
  select(bm_plasma_cell_percent_at_pre_ld,
         number_of_prior_lines_of_therapy,
         number_prior_lines_therapy,
         # `Days from last bridging to CAR-T infusion`,         ################# Not in data?
         lenalidomide_exposed_vs_refractory_vs_not_exposed,
         bortezomib_exposed_vs_refractory_vs_not_exposed,
         carfilzomib_exposed_vs_refractory_vs_not_exposed,
         pomalidomide_exposed_vs_refractory_vs_not_exposed,
         # daratumumab_or_isatuximab_cd38_m_ab_exposed_vs_refractory_vs_not_exposed,
         immuno, proteasome, dara, doubleref, tripleref, pentaref, 
         "Marrow Cellularity (%) at Day30" = marrow_percent_cellularity_day_30,
         os_from_infusion_months, age, agecat, male_sex, sex,
         extramedullary_disease_yes_no, 
         high_marrow_burden_50_percent, ecog_at_ld, 
         r_iss_at_car_t_infusion, cytogenetics, 
         bridging_therapy_yes_no,
         # `Received alkylating therapy with bridging (Yes=1, No=0)`,
         prior_auto_sct_yes_no, prior_allo_sct_within_6_months_before_apheresis,
         # prior_treatment_with_any_gene_therapy_based_therapeutic_or_investigational_cellular_therapy_or_bcma_targeted_therapy_yes_no,
         Prior_bcma_therapy,
         immuno, proteasome, dara, doubleref, tripleref, pentaref, 
         did_patient_meet_eligibility_criteria_for_cartitude_1_on_the_day_of_apheresis_yes_no,
         cytopenia_apheresis, Neutropenia_apheresis,
         Anemia_apheresis,Thrombocytopenia_apheresis,
         `cytopenia_-5d`, `Neutropenia_-5d`,
         `Anemia_-5d`, `Thrombocytopenia_-5d`,
         gr3_cytopenia_apheresis, gr3_Neutropenia_apheresis,
         gr3_Anemia_apheresis, gr3_Thrombocytopenia_apheresis,
         `gr3_cytopenia_-5d`, `gr3_Neutropenia_-5d`,
         `gr3_Anemia_-5d`, `gr3_Thrombocytopenia_-5d`,
         gr3_cytopenia_30d) %>% 
  tbl_summary(by = gr3_cytopenia_30d,
              type = list(
                c(extramedullary_disease_yes_no,
                  bm_plasma_cell_percent_at_pre_ld,
                  cytopenia_apheresis,`cytopenia_-5d`,
                  cytogenetics,
                  bridging_therapy_yes_no,
                  # `Received alkylating therapy with bridging (Yes=1, No=0)`,
                  prior_auto_sct_yes_no,
                  Prior_bcma_therapy,
                  immuno, proteasome, dara,
                  doubleref, tripleref, pentaref,
                  cytopenia_apheresis, Neutropenia_apheresis,
                  Anemia_apheresis,Thrombocytopenia_apheresis,
                  `cytopenia_-5d`, `Neutropenia_-5d`,
                  `Anemia_-5d`, `Thrombocytopenia_-5d`,
                  gr3_cytopenia_apheresis, gr3_Neutropenia_apheresis,
                  gr3_Anemia_apheresis, gr3_Thrombocytopenia_apheresis,
                  `gr3_cytopenia_-5d`, `gr3_Neutropenia_-5d`,
                  `gr3_Anemia_-5d`, `gr3_Thrombocytopenia_-5d`,
                  did_patient_meet_eligibility_criteria_for_cartitude_1_on_the_day_of_apheresis_yes_no)
                ~"categorical",
                c(bm_plasma_cell_percent_at_pre_ld,
                  number_of_prior_lines_of_therapy, `Marrow Cellularity (%) at Day30`)
                ~ "continuous"),
              statistic = list(all_continuous() ~ "{median} ({min}, {max})"), 
              digits = all_continuous() ~ 1
  ) %>% 
  bold_labels() %>% add_stat_label() %>% add_p() %>% 
  modify_spanning_header(all_stat_cols() ~ "**Grade 3+ Cytopenia at days 30?**")
```

## Patient characteristics by grade 3+ cytopenia at Day 90
```{r Patient characteristics by grade 3+ cytopenia at Day 90}
cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  # filter(is.na(Not_evaluable_at_90days)) %>%  ######################### Fix evaluable----------
  select(bm_plasma_cell_percent_at_pre_ld,
         number_of_prior_lines_of_therapy,
         number_prior_lines_therapy,
         # `Days from last bridging to CAR-T infusion`,
         lenalidomide_exposed_vs_refractory_vs_not_exposed,
         bortezomib_exposed_vs_refractory_vs_not_exposed,
         carfilzomib_exposed_vs_refractory_vs_not_exposed,
         pomalidomide_exposed_vs_refractory_vs_not_exposed,
         # daratumumab_or_isatuximab_cd38_m_ab_exposed_vs_refractory_vs_not_exposed,
         immuno, proteasome, dara, doubleref, tripleref, pentaref, 
         "Marrow Cellularity (%) at Day90" = marrow_percent_cellularity_day_90,
         os_from_infusion_months,
         age, agecat, male_sex, sex,
         extramedullary_disease_yes_no, 
         high_marrow_burden_50_percent, ecog_at_ld, r_iss_at_car_t_infusion, cytogenetics, 
         bridging_therapy_yes_no,
         # `Received alkylating therapy with bridging (Yes=1, No=0)`,
         prior_auto_sct_yes_no, prior_allo_sct_within_6_months_before_apheresis,
         # prior_treatment_with_any_gene_therapy_based_therapeutic_or_investigational_cellular_therapy_or_bcma_targeted_therapy_yes_no,
         Prior_bcma_therapy,
         did_patient_meet_eligibility_criteria_for_cartitude_1_on_the_day_of_apheresis_yes_no,
         cytopenia_apheresis, Neutropenia_apheresis,
         Anemia_apheresis,Thrombocytopenia_apheresis,
         `cytopenia_-5d`, `Neutropenia_-5d`,
         `Anemia_-5d`, `Thrombocytopenia_-5d`,
         gr3_cytopenia_apheresis, gr3_Neutropenia_apheresis,
         gr3_Anemia_apheresis, gr3_Thrombocytopenia_apheresis,
         `gr3_cytopenia_-5d`, `gr3_Neutropenia_-5d`,
         `gr3_Anemia_-5d`, `gr3_Thrombocytopenia_-5d`,
         gr3_cytopenia_90d) %>% 
  tbl_summary(by = gr3_cytopenia_90d,
              type = list(
                c(extramedullary_disease_yes_no,
                  bm_plasma_cell_percent_at_pre_ld,
                  cytopenia_apheresis,`cytopenia_-5d`,
                  cytogenetics,
                  bridging_therapy_yes_no,
                  # `Received alkylating therapy with bridging (Yes=1, No=0)`,
                  prior_auto_sct_yes_no,
                  immuno, proteasome, dara,
                  doubleref, tripleref, pentaref,
                  cytopenia_apheresis, Neutropenia_apheresis,
                  Anemia_apheresis,Thrombocytopenia_apheresis,
                  `cytopenia_-5d`, `Neutropenia_-5d`,
                  `Anemia_-5d`, `Thrombocytopenia_-5d`,
                  gr3_cytopenia_apheresis, gr3_Neutropenia_apheresis,
                  gr3_Anemia_apheresis, gr3_Thrombocytopenia_apheresis,
                  `gr3_cytopenia_-5d`, `gr3_Neutropenia_-5d`,
                  `gr3_Anemia_-5d`, `gr3_Thrombocytopenia_-5d`,
                  did_patient_meet_eligibility_criteria_for_cartitude_1_on_the_day_of_apheresis_yes_no)
                ~ "categorical",
                c(bm_plasma_cell_percent_at_pre_ld,
                  number_of_prior_lines_of_therapy, `Marrow Cellularity (%) at Day90`)
                ~ "continuous"),
              statistic = list(all_continuous() ~ "{median} ({min}, {max})"), 
              digits = all_continuous() ~ 1
  ) %>% 
  bold_labels() %>% add_stat_label() %>%  add_p() %>% 
  modify_spanning_header(all_stat_cols() ~ "**Grade 3+ Cytopenia at days 90?**")
```

## CRS, ICANS, tocilizumab use, steroid use, anakinra use and bridging therapy by grade 3+ cytopenia at Day 30
```{r CRS, ICANS, tocilizumab use, steroid use, anakinra use and bridging therapy by grade 3+ cytopenia at Day 30}
tbl1 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  # filter(is.na(Not_evaluable_at_30days)) %>% ######################### Fix evaluable----------
  select(cell_dose_million_cells,
         CRS_gradecat,
         day_of_onset_crs_relative_to_infusion,
         day_of_max_crs_relative_to_infusion,
         duration_of_crs_days, 
         ICANS_gradecat,
         day_of_onset_icans_relative_to_infusion,
         day_of_max_icans_relative_to_infusion,
         duration_of_icans_days, 
         delayed_neuro,
         parkinsonian_neurotoxicity_yes_no,
         other_delayed_non_parkinsonian_neurologic_toxicity_yes_no,
         delayed_neuro_type,
         gr3_cytopenia_30d) %>% 
  tbl_summary(by = gr3_cytopenia_30d,
              type = list(
                c(day_of_onset_crs_relative_to_infusion,
                  day_of_max_crs_relative_to_infusion,
                  duration_of_crs_days,
                  day_of_onset_icans_relative_to_infusion,
                  day_of_max_icans_relative_to_infusion,
                  duration_of_icans_days) ~ "continuous",
                c(delayed_neuro,
                  other_delayed_non_parkinsonian_neurologic_toxicity_yes_no,
                  parkinsonian_neurotoxicity_yes_no) ~ "categorical"),
              statistic=list(all_continuous() ~ "{median} ({min}, {max})"), 
    digits = all_continuous() ~ 1,
    missing = "no") %>% 
  bold_labels() %>% add_stat_label() %>% add_p() %>% 
  modify_spanning_header(all_stat_cols() ~ "**Grade 3+ Cytopenia at days 30?**")

tbl2 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  # filter(is.na(Not_evaluable_at_30days)) %>% ######################### Fix evaluable----------
  select(steroid_use_yes_no, tocilizumab_yes_no, 
         anakinra_yes_no, gr3_cytopenia_30d) %>% 
  tbl_summary(by = gr3_cytopenia_30d,
              type = list(
    c(steroid_use_yes_no, tocilizumab_yes_no, 
      anakinra_yes_no) ~ "categorical"),
    statistic=list(all_continuous() ~ "{median} ({min}, {max})"), 
    digits = all_continuous() ~ 1) %>% 
  bold_labels() %>% add_stat_label() %>% add_p()

tbl3 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  # filter(is.na(Not_evaluable_at_30days)) %>% ######################### Fix evaluable----------
  select(#day_30_mrd_by_clono_seq_positive_vs_negative,
         ORR_30days, CRorbetter_30days,
         # day30_response_v2, 
         gr3_cytopenia_30d) %>% 
  tbl_summary(by = gr3_cytopenia_30d,
    #           type = list(
    # c(day_30_mrd_by_clono_seq_positive_vs_negative) ~ "categorical"),
    digits = all_continuous() ~ 1,
    missing = "no") %>% 
  bold_labels() %>% add_stat_label() %>% add_p()

# tbl4 <- cart1 %>% 
#   distinct(pt_identifier, .keep_all = TRUE) %>% 
#   filter(is.na(Not_evaluable_at_30days)) %>% 
#   filter(day30_response_v2 == "sCR or CR") %>% 
#   select(day_30_mrd_by_clono_seq_positive_vs_negative, gr3_cytopenia_30d) %>% 
#   tbl_summary(by =gr3_cytopenia_30d,
#               type = list(c(day_30_mrd_by_clono_seq_positive_vs_negative) ~ "categorical"),
#               label = list(day_30_mrd_by_clono_seq_positive_vs_negative ~ "day_30_mrd_by_clono_seq_positive_vs_negative within CR/sCR"),
#               digits = all_continuous() ~ 1,
#               missing = "no") %>% 
#   bold_labels() %>% add_stat_label() %>% add_p()

tbl5 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  # filter(is.na(Not_evaluable_at_30days)) %>% ######################### Fix evaluable----------
  select(#MRD_mon3,
         ORR_mon3, CRorbetter_mon3,
         # mon3_response_v2, 
         gr3_cytopenia_30d) %>% 
  tbl_summary(by = gr3_cytopenia_30d,
    #           type = list(
    # c(MRD_mon3) ~ "categorical"),
    digits = all_continuous() ~ 1,
    missing = "no") %>% 
  bold_labels() %>% add_stat_label() %>% add_p()

# tbl6 <- cart1 %>% 
#   distinct(pt_identifier, .keep_all = TRUE) %>% 
#   filter(is.na(Not_evaluable_at_30days)) %>% 
#   filter(mon3_response_v2 == "sCR or CR") %>% 
#   select(MRD_mon3, gr3_cytopenia_30d) %>% 
#   tbl_summary(by =gr3_cytopenia_30d,
#               type = list(c(MRD_mon3) ~ "categorical"),
#               label = list(MRD_mon3 ~ "MRD_mon3 within CR/sCR"),
#               digits = all_continuous() ~ 1,
#               missing = "no") %>% 
#   bold_labels() %>% add_stat_label() %>% add_p()

tbl7 <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  # filter(is.na(Not_evaluable_at_30days)) %>% ######################### Fix evaluable----------
  select(best_response, best_ORRcollapsed,
         gr3_cytopenia_30d) %>% 
  tbl_summary(by = gr3_cytopenia_30d,
              #type = list(c(MRD_mon3) ~ "categorical"),
              label = list(best_response ~ "best response in the first 90 days"),
              digits = all_continuous() ~ 1,
              missing = "no") %>% 
  bold_labels() %>% add_stat_label() %>% add_p()

# tbl8 <- cart1 %>% 
#   distinct(pt_identifier, .keep_all = TRUE) %>% 
#   filter(is.na(Not_evaluable_at_30days)) %>% 
#   filter(best_response == "sCR or CR") %>% 
#   select(MRDneg_in_best_response, gr3_cytopenia_30d) %>% 
#   tbl_summary(by =gr3_cytopenia_30d,
#               type = list(c(MRDneg_in_best_response) ~ "categorical"),
#               label = list(MRDneg_in_best_response ~ "MRDneg_in_best_response within CR/sCR"),
#               statistic=list(all_continuous() ~ "{median} ({min}, {max})"), 
#               digits = all_continuous() ~ 1,
#               missing = "no") %>% 
#   bold_labels() %>% add_stat_label() %>% add_p()

tbl_stack(list(tbl1, tbl2, tbl3, tbl5, tbl7))
```

## Table S8.Details of infections after ide-cel
```{r infection details 30 first days}
tbl1 <- cilta %>% 
  filter(time_from_infusion_to_infection_days <= 30) %>% 
  select(infection_type_bacterial_viral_fungal_parasitic) %>% 
  tbl_summary() %>%
  remove_row_type(variables = infection_type_bacterial_viral_fungal_parasitic,
                  type= "level",
                  level_value = c("Fungal", "Viral")) %>%
  remove_row_type(infection_type_bacterial_viral_fungal_parasitic, type = "header") %>%
  add_stat_label(label = infection_type_bacterial_viral_fungal_parasitic ~ c("n (%)"), location = "column") %>%
  modify_column_indent(columns = label, undo = TRUE) %>% 
  bold_levels() %>% 
  modify_header(update = all_stat_cols() ~ "**{level}**, N = {n}") %>%
  modify_footnote(update = everything() ~ NA) %>%
  modify_spanning_header(all_stat_cols() ~ "**Infections in the first 30 days**")

tbl1_1 <- cilta %>% 
  filter(time_from_infusion_to_infection_days <= 30) %>% 
  filter(infection_type_bacterial_viral_fungal_parasitic == "Bacterial") %>%
  select(infection_type_details) %>% 
  tbl_summary(statistic = infection_type_details ~ c("{n}")) %>% 
  remove_row_type(infection_type_details, type = "header")

tbl2 <- cilta %>% 
  filter(time_from_infusion_to_infection_days <= 30) %>% 
  select(infection_type_bacterial_viral_fungal_parasitic) %>% 
  tbl_summary() %>%
  remove_row_type(variables = infection_type_bacterial_viral_fungal_parasitic,
                  type= "level",
                  level_value = c("Bacterial", "Viral")) %>% 
  remove_row_type(infection_type_bacterial_viral_fungal_parasitic, type = "header") %>% 
  add_stat_label(label = infection_type_bacterial_viral_fungal_parasitic ~ c("n (%)"), location = "column") %>%
  bold_levels() %>%
  modify_column_indent(columns = label, undo = TRUE)

tbl2_1 <- cilta %>% 
  filter(time_from_infusion_to_infection_days <= 30) %>% 
  filter(infection_type_bacterial_viral_fungal_parasitic == "Fungal") %>%
  select(infection_type_details) %>% 
  tbl_summary(statistic = infection_type_details ~ c("{n}")) %>% 
  remove_row_type(infection_type_details, type = "header")

tbl3 <- cilta %>% 
  filter(time_from_infusion_to_infection_days <= 30) %>% 
  select(infection_type_bacterial_viral_fungal_parasitic) %>% 
  tbl_summary() %>%
  remove_row_type(variables = infection_type_bacterial_viral_fungal_parasitic,
                  type= "level",
                  level_value = c("Bacterial", "Fungal")) %>% 
  remove_row_type(infection_type_bacterial_viral_fungal_parasitic, type = "header") %>% 
  add_stat_label(label = infection_type_bacterial_viral_fungal_parasitic ~ c("n (%)"), location = "column") %>%
  bold_levels() %>%
  modify_column_indent(columns = label, undo = TRUE)

tbl3_1 <- cilta %>% 
  filter(time_from_infusion_to_infection_days <= 30) %>% 
  filter(infection_type_bacterial_viral_fungal_parasitic == "Viral") %>%
  select(infection_type_details) %>% 
  tbl_summary(statistic = infection_type_details ~ c("{n}")) %>% 
  remove_row_type(infection_type_details, type = "header")

tbl_stack(list(tbl1, tbl1_1, tbl2, tbl2_1, tbl3, tbl3_1))
```

```{r infection details between 31 - 100 days}
tbl1 <- cilta %>% 
  filter(time_from_infusion_to_infection_days > 30 &
           time_from_infusion_to_infection_days <= 100) %>% 
  select(infection_type_bacterial_viral_fungal_parasitic) %>% 
  tbl_summary() %>%
  remove_row_type(variables = infection_type_bacterial_viral_fungal_parasitic,
                  type= "level",
                  level_value = c("Fungal", "Viral")) %>%
  remove_row_type(infection_type_bacterial_viral_fungal_parasitic, type = "header") %>%
  add_stat_label(label = infection_type_bacterial_viral_fungal_parasitic ~ c("n (%)"), location = "column") %>%
  modify_column_indent(columns = label, undo = TRUE) %>% 
  bold_levels() %>% 
  modify_header(update = all_stat_cols() ~ "**{level}**, N = {n}") %>%
  modify_footnote(update = everything() ~ NA) %>%
  modify_spanning_header(all_stat_cols() ~ "**Infections day 31 – 100**")

tbl1_1 <- cilta %>% 
  filter(time_from_infusion_to_infection_days > 30 &
           time_from_infusion_to_infection_days <= 100) %>% 
  filter(infection_type_bacterial_viral_fungal_parasitic == "Bacterial") %>%
  select(infection_type_details) %>% 
  tbl_summary(statistic = infection_type_details ~ c("{n}")) %>% 
  remove_row_type(infection_type_details, type = "header")

tbl2 <- cilta %>% 
  filter(time_from_infusion_to_infection_days > 30 &
           time_from_infusion_to_infection_days <= 100) %>% 
  select(infection_type_bacterial_viral_fungal_parasitic) %>% 
  tbl_summary() %>%
  remove_row_type(variables = infection_type_bacterial_viral_fungal_parasitic,
                  type= "level",
                  level_value = c("Bacterial", "Viral")) %>% 
  remove_row_type(infection_type_bacterial_viral_fungal_parasitic, type = "header") %>% 
  add_stat_label(label = infection_type_bacterial_viral_fungal_parasitic ~ c("n (%)"), location = "column") %>%
  bold_levels() %>%
  modify_column_indent(columns = label, undo = TRUE)

# tbl2_1 <- cilta %>% 
#   filter(time_from_infusion_to_infection_days > 30 &
#            time_from_infusion_to_infection_days <= 100) %>% 
#   filter(infection_type_bacterial_viral_fungal_parasitic == "Fungal") %>%
#   select(infection_type_details) %>% 
#   tbl_summary(statistic = infection_type_details ~ c("{n}")) %>% 
#   remove_row_type(infection_type_details, type = "header")

tbl3 <- cilta %>% 
  filter(time_from_infusion_to_infection_days > 30 &
           time_from_infusion_to_infection_days <= 100) %>% 
  select(infection_type_bacterial_viral_fungal_parasitic) %>% 
  tbl_summary() %>%
  remove_row_type(variables = infection_type_bacterial_viral_fungal_parasitic,
                  type= "level",
                  level_value = c("Bacterial", "Fungal")) %>% 
  remove_row_type(infection_type_bacterial_viral_fungal_parasitic, type = "header") %>% 
  add_stat_label(label = infection_type_bacterial_viral_fungal_parasitic ~ c("n (%)"), location = "column") %>%
  bold_levels() %>%
  modify_column_indent(columns = label, undo = TRUE)

tbl3_1 <- cilta %>% 
  filter(time_from_infusion_to_infection_days > 30 &
           time_from_infusion_to_infection_days <= 100) %>% 
  filter(infection_type_bacterial_viral_fungal_parasitic == "Viral") %>%
  select(infection_type_details) %>% 
  tbl_summary(statistic = infection_type_details ~ c("{n}")) %>% 
  remove_row_type(infection_type_details, type = "header")

tbl_stack(list(tbl1, tbl1_1, #tbl2,# tbl2_1, 
               tbl3, tbl3_1))
```

```{r infection details after 100 days}
tbl1 <- cilta %>% 
  filter(time_from_infusion_to_infection_days > 100) %>% 
  select(infection_type_bacterial_viral_fungal_parasitic) %>% 
  tbl_summary() %>%
  remove_row_type(variables = infection_type_bacterial_viral_fungal_parasitic,
                  type= "level",
                  level_value = c("Fungal", "Viral")) %>%
  remove_row_type(infection_type_bacterial_viral_fungal_parasitic, type = "header") %>%
  add_stat_label(label = infection_type_bacterial_viral_fungal_parasitic ~ c("n (%)"), location = "column") %>%
  modify_column_indent(columns = label, undo = TRUE) %>% 
  bold_levels() %>% 
  modify_header(update = all_stat_cols() ~ "**{level}**, N = {n}") %>%
  modify_footnote(update = everything() ~ NA) %>%
  modify_spanning_header(all_stat_cols() ~ "**Infections after day 100**")

tbl1_1 <- cilta %>% 
  filter(time_from_infusion_to_infection_days > 100) %>% 
  filter(infection_type_bacterial_viral_fungal_parasitic == "Bacterial") %>%
  select(infection_type_details) %>% 
  tbl_summary(statistic = infection_type_details ~ c("{n}")) %>% 
  remove_row_type(infection_type_details, type = "header")

# tbl2 <- cilta %>% 
#   filter(time_from_infusion_to_infection_days > 100) %>% 
#   select(infection_type_bacterial_viral_fungal_parasitic) %>% 
#   tbl_summary() %>%
#   remove_row_type(variables = infection_type_bacterial_viral_fungal_parasitic,
#                   type= "level",
#                   level_value = c("Bacterial", "Viral")) %>% 
#   remove_row_type(infection_type_bacterial_viral_fungal_parasitic, type = "header") %>% 
#   add_stat_label(label = infection_type_bacterial_viral_fungal_parasitic ~ c("n (%)"), location = "column") %>%
#   bold_levels() %>%
#   modify_column_indent(columns = label, undo = TRUE)
# 
# tbl2_1 <- cilta %>% 
#   filter(time_from_infusion_to_infection_days > 100) %>% 
#   filter(infection_type_bacterial_viral_fungal_parasitic == "Fungal") %>%
#   select(infection_type_details) %>% 
#   tbl_summary(statistic = infection_type_details ~ c("{n}")) %>% 
#   remove_row_type(infection_type_details, type = "header")

tbl3 <- cilta %>% 
  filter(time_from_infusion_to_infection_days > 100) %>% 
  select(infection_type_bacterial_viral_fungal_parasitic) %>% 
  tbl_summary() %>%
  remove_row_type(variables = infection_type_bacterial_viral_fungal_parasitic,
                  type= "level",
                  level_value = c("Bacterial", "Fungal")) %>% 
  remove_row_type(infection_type_bacterial_viral_fungal_parasitic, type = "header") %>% 
  add_stat_label(label = infection_type_bacterial_viral_fungal_parasitic ~ c("n (%)"), location = "column") %>%
  bold_levels() %>%
  modify_column_indent(columns = label, undo = TRUE)

tbl3_1 <- cilta %>% 
  filter(time_from_infusion_to_infection_days > 100) %>% 
  filter(infection_type_bacterial_viral_fungal_parasitic == "Viral") %>%
  select(infection_type_details) %>% 
  tbl_summary(statistic = infection_type_details ~ c("{n}")) %>% 
  remove_row_type(infection_type_details, type = "header")

tbl_stack(list(tbl1, tbl1_1, #tbl2, tbl2_1, 
               tbl3, tbl3_1))
```

## Patients with infections by CRS, ICANS, tocilizumab use, steroid use, anakinra use and bridging therapy
```{r Patients with infections by CRS, ICANS, tocilizumab use, steroid use, anakinra use and bridging therapy}
cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  select(number_prior_lines_therapy,
         number_of_prior_lines_of_therapy,
         prior_auto_sct_yes_no,
         # prior_treatment_with_any_gene_therapy_based_therapeutic_or_investigational_cellular_therapy_or_bcma_targeted_therapy_yes_no,
         Prior_bcma_therapy,
         # `Received alkylating therapy with bridging (Yes=1, No=0)`,
         # `Days from last bridging to CAR-T infusion`,
         cell_dose_million_cells,
         need_for_g_csf_post_infus_yes_1_no_0,
         cd34_stem_cell_boost_yes_1_no_0,
         need_for_ivig_yes_1_no_0,

         CRS_gradecat, ICANS_gradecat,
         delayed_neuro,
         parkinsonian_neurotoxicity_yes_no,
         other_delayed_non_parkinsonian_neurologic_toxicity_yes_no,
         delayed_neuro_type,
         steroid_use_yes_no, 
         tocilizumab_yes_no, anakinra_yes_no,
         bridging_therapy_yes_no, any_infection) %>% 
  tbl_summary(by= any_infection, 
              statistic=list(all_continuous() ~ "{median} ({min}, {max})"), 
              type = list(all_dichotomous() ~ "categorical", 
                          number_of_prior_lines_of_therapy ~ "continuous")) %>%
  add_p() %>%  bold_labels() %>% add_stat_label() %>% 
  modify_header(update = all_stat_cols() ~ "**{level}**, N = {n}") %>%
  modify_footnote(update = everything() ~ NA) %>%
  modify_spanning_header(all_stat_cols() ~ "**Any infection**")

cilta %>% 
  arrange(pt_identifier, desc(any_infection), desc(any_sevinfection)) %>%
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  select(number_prior_lines_therapy,
         number_of_prior_lines_of_therapy,
         prior_auto_sct_yes_no,
         Prior_bcma_therapy,
         # `Received alkylating therapy with bridging (Yes=1, No=0)`,
         # `Days from last bridging to CAR-T infusion`,
         cell_dose_million_cells,
         need_for_g_csf_post_infus_yes_1_no_0,
         cd34_stem_cell_boost_yes_1_no_0,
         need_for_ivig_yes_1_no_0,

         CRS_gradecat, ICANS_gradecat,
         delayed_neuro,
         parkinsonian_neurotoxicity_yes_no,
         other_delayed_non_parkinsonian_neurologic_toxicity_yes_no,
         delayed_neuro_type,
         steroid_use_yes_no, 
         tocilizumab_yes_no, anakinra_yes_no,
         bridging_therapy_yes_no, any_sevinfection) %>% 
  tbl_summary(by=any_sevinfection, 
              statistic=list(all_continuous() ~ "{median} ({min}, {max})"), 
              type = list(all_dichotomous() ~ "categorical", 
                          number_of_prior_lines_of_therapy ~ "continuous")) %>%
  add_p() %>%  bold_labels() %>% add_stat_label() %>% 
  modify_header(update = all_stat_cols() ~ "**{level}**, N = {n}") %>%
  modify_footnote(update = everything() ~ NA) %>%
  modify_spanning_header(all_stat_cols() ~ "**Any severe infection**")
```

## Patients with infections by cell counts and immunoglobulin levels
```{r Patients with infections by cell counts and immunoglobulin levels}
cilta %>%
  filter(any_infection == "No") %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  select(ig_g_day_5, ig_a_day_5, ig_m_day_5,
         wbc_day_5, anc_day_5_k_ul,
         # `Eos day -5 (k/ul)`, `Monos day -5 (k/ul)`, 
         # `CD3 apheresis`, `CD56 apheresis`, 
         cd4_apheresis, cd8_apheresis,
         cd19_apheresis, 
         ig_g_day_90, ig_a_day_90, ig_m_day_90,
         wbc_day_90, anc_day_90,
         # `Eos day +90`, `Monos day +90`,
         # `CD3 day +90`, `CD56 day +90`, 
         cd4_day_90, cd8_day_90,
         cd19_day_90,
         any_infection) %>% 
  tbl_summary(by = any_infection,
              type = list(everything() ~ "continuous"),
              statistic=list(all_continuous() ~ "{median} ({min}, {max})"), 
              digits = all_continuous() ~ 1) %>% 
  add_p() %>%  bold_labels() %>% add_stat_label() %>% 
  modify_header(update = all_stat_cols() ~ "**{level}**, N = {n}") %>%
  modify_footnote(update = everything() ~ NA) %>%
  modify_spanning_header(all_stat_cols() ~ "**Any infection**")

cilta %>% 
  filter(any_sevinfection == "No") %>% 
  arrange(pt_identifier, desc(any_infection), desc(any_sevinfection)) %>%
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  select(ig_g_day_5, ig_a_day_5, ig_m_day_5,
         wbc_day_5, anc_day_5_k_ul,
         # `Eos day -5 (k/ul)`, `Monos day -5 (k/ul)`, 
         # `CD3 apheresis`, `CD56 apheresis`, 
         cd4_apheresis, cd8_apheresis,
         cd19_apheresis, 
         ig_g_day_90, ig_a_day_90, ig_m_day_90,
         wbc_day_90, anc_day_90,
         # `Eos day +90`, `Monos day +90`,
         # `CD3 day +90`, `CD56 day +90`, 
         cd4_day_90, cd8_day_90,
         cd19_day_90,
         any_sevinfection) %>% 
  tbl_summary(by = any_sevinfection,
              type = list(everything() ~ "continuous"),
              statistic=list(all_continuous() ~ "{median} ({min}, {max})"), 
              digits = all_continuous() ~ 1) %>% 
  add_p() %>%  bold_labels() %>% add_stat_label() %>% 
  modify_header(update = all_stat_cols() ~ "**{level}**, N = {n}") %>%
  modify_footnote(update = everything() ~ NA) %>%
  modify_spanning_header(all_stat_cols() ~ "**Any severe infection**")
```

## Immunoglobulin levels excluding patients with that type of myeloma 
(myeloma_involved_heavy_chain != "Igg")
```{r Immunoglobulin levels excluding patients with that type of myeloma}
cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  filter(myeloma_involved_heavy_chain != "Igg" | is.na(myeloma_involved_heavy_chain)) %>% 
  select(ig_g_day_5, ig_g_day_30,
         ig_g_day_90, any_infection) %>% 
  tbl_summary(by = any_infection,
              statistic = list(all_continuous() ~ "{median} ({min}, {max})"), 
              digits = all_continuous() ~ 1
  ) %>% 
  add_p() %>% bold_labels() %>% add_stat_label() %>% 
  add_overall() %>% 
  modify_footnote(update = everything() ~ NA) %>%
  modify_spanning_header(all_stat_cols() ~ "**Any infection**")

cilta %>% 
  arrange(pt_identifier, desc(any_infection), desc(any_sevinfection)) %>%
  distinct(pt_identifier, .keep_all = TRUE) %>% 
  filter(myeloma_involved_heavy_chain != "Igg" | is.na(myeloma_involved_heavy_chain)) %>% 
  select(ig_g_day_5, ig_g_day_30,
         ig_g_day_90, any_sevinfection) %>% 
  tbl_summary(by = any_sevinfection,
              statistic = list(all_continuous() ~ "{median} ({min}, {max})"), 
              digits = all_continuous() ~ 1
  ) %>% 
  add_p() %>% bold_labels() %>% add_stat_label() %>% 
  add_overall() %>% 
  modify_footnote(update = everything() ~ NA) %>%
  modify_spanning_header(all_stat_cols() ~ "**Any severe infection**")
```

### Figures
```{r fig_dat}
fig_dat <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE)
```

## Figure 1. Consort diagram

## Figure 2. Inflammatory markers after ide-cel

```{r Figure 2. Inflammatory markers after ide-cel, fig.height=12}
LDH <- fig_dat %>% 
  select(pt_identifier, starts_with("ldh_")) %>% 
  pivot_longer(cols = -c(pt_identifier)) %>% 
  mutate(name = str_remove(name, "_k_ul|_k_u_l"),
         day = case_when(
           str_detect(name, "apheresis")            ~ "-28",
           str_detect(name, "pretreatment")         ~ "-28",
           str_detect(name, "day")                  ~ ((str_match(name, "([0-9]+)"))[,2])
         ), 
         day = case_when(
           day == 5                                 ~ "-5",
           TRUE                                     ~ day
         ),
         day = factor(day, levels = c("-28", "-5", "0", 
                                      "7", "14", "21", "30", "90", "180"))) %>% 
  ggplot(aes(x=day, y=value)) + 
  geom_boxplot() + 
  xlab("Days") + ylab("LDH represented with a log2 scale") + 
  theme_classic() + ggtitle("LDH") + 
  scale_y_continuous(trans = "log2") +
  stat_compare_means(label.x = 5 
                     # aes(label = paste0("p = ", ..p.format..))
  #                  aes(
  # label=ifelse(..p.format.. == 1e-04, "p<0.001", "Other")
  )

# cart1$`Ferritin day +7`[cart1$`Ferritin day +7`==">33511"] <- 33511
# cart1$`Ferritin day +7` <- as.numeric(cart1$`Ferritin day +7`)
Ferritin <- fig_dat %>% 
  select(pt_identifier, starts_with("ferritin_")) %>% 
  pivot_longer(cols = -c(pt_identifier)) %>% 
  mutate(name = str_remove(name, "_k_ul|_k_u_l"),
         day = case_when(
           str_detect(name, "apheresis")            ~ "-28",
           str_detect(name, "pretreatment")         ~ "-28",
           str_detect(name, "day")                  ~ ((str_match(name, "([0-9]+)"))[,2])
         ), 
         day = case_when(
           day == 5                                 ~ "-5",
           TRUE                                     ~ day
         ),
         day = factor(day, levels = c("-28", "-5", "0", 
                                      "7", "14", "21", "30", "90", "180"))) %>% 
  ggplot(aes(x=day, y=value)) + 
  geom_boxplot() + 
  xlab("Days") + ylab("Ferritin represented with a log2 scale") + 
  theme_classic() + ggtitle("Ferritin") + 
  scale_y_continuous(trans = "log2") +
  stat_compare_means(label.x = 5 
                     # aes(label = paste0("p = ", ..p.format..))
  #                  aes(
  # label=ifelse(..p.format.. == 1e-04, "p<0.001", "Other")
  )

# cart1$`CRP Pre-Leukapheresis`[cart1$`CRP Pre-Leukapheresis`=="<0.03"] <- 0.07
# cart1$`CRP Pre-Leukapheresis` <- as.numeric(cart1$`CRP Pre-Leukapheresis`)
# cart1$`CRP day +14`[cart1$`CRP day +14`=="<0.03"] <- 0.07
# cart1$`CRP day +14` <- as.numeric(cart1$`CRP day +14`)
# cart1$`CRP day +21`[cart1$`CRP day +21`=="<0.03"] <- 0.07
# cart1$`CRP day +21` <- as.numeric(cart1$`CRP day +21`)
# cart1$`CRP day +30`[cart1$`CRP day +30`=="<0.03"] <- 0.07
# cart1$`CRP day +30` <- as.numeric(cart1$`CRP day +30`)
CRP <- fig_dat %>% 
  select(pt_identifier, starts_with("crp_")) %>% 
  pivot_longer(cols = -c(pt_identifier)) %>% 
  mutate(name = str_remove(name, "_k_ul|_k_u_l"),
         day = case_when(
           str_detect(name, "apheresis")            ~ "-28",
           str_detect(name, "pretreatment")         ~ "-28",
           str_detect(name, "day")                  ~ ((str_match(name, "([0-9]+)"))[,2])
         ), 
         day = case_when(
           day == 5                                 ~ "-5",
           TRUE                                     ~ day
         ),
         day = factor(day, levels = c("-28", "-5", "0", 
                                      "7", "14", "21", "30", "90", "180"))) %>% 
  ggplot(aes(x=day, y=value)) + 
  geom_boxplot() + 
  xlab("Days") + ylab("CRP represented with a log2 scale") + 
  theme_classic() + ggtitle("CRP") + 
  scale_y_continuous(trans = "log2") +
  stat_compare_means(label.x = 5 
                     # aes(label = paste0("p = ", ..p.format..))
  #                  aes(
  # label=ifelse(..p.format.. == 1e-04, "p<0.001", "Other")
  )

(LDH) /
(Ferritin) /
(CRP)+
plot_annotation(tag_levels = "A")
```

## Figure 3. Cellular and humoral immune reconstitution after cilta-cel
```{r timeserie, fig.width=12, fig.height=12}
# WBC
val <- fig_dat %>% reframe(wbc_apheresis = MedianCI(wbc_apheresis, conf.level = 0.95, na.rm = TRUE),
                  wbc_day_5 = MedianCI(wbc_day_5, conf.level = 0.95, na.rm = TRUE),
                  wbc_day_0_k_u_l = MedianCI(wbc_day_0_k_u_l, conf.level = 0.95, na.rm = TRUE),
                  wbc_day_7 = MedianCI(wbc_day_7, conf.level = 0.95, na.rm = TRUE),
                  wbc_day_14 = MedianCI(wbc_day_14, conf.level = 0.95, na.rm = TRUE),
                  wbc_day_21 = MedianCI(wbc_day_21, conf.level = 0.95, na.rm = TRUE),
                  wbc_day_30 = MedianCI(wbc_day_30, conf.level = 0.95, na.rm = TRUE),
                  wbc_day_90 = MedianCI(wbc_day_90, conf.level = 0.95, na.rm = TRUE),
                  wbc_day_180 = MedianCI(wbc_day_180, conf.level = 0.95, na.rm = TRUE))
timepoints <- c(-28, -5, 0, 7, 14, 21, 30, 90, 180)
WBC_fig <-
  data.frame(timepoints,
             t(val))

WBC_fig <-
  WBC_fig %>%
  ggplot(aes(x = timepoints, y = X1, group = 1)) + 
  geom_line(col = "blue") + 
  xlab("Days relative to infusion") + 
  ylab("Median WBC") + 
  theme_classic() + 
  geom_ribbon(aes(ymin = X2, ymax = X3), alpha = 0.1) + 
  scale_x_continuous(
    breaks = c(-28, -5, 0, 7, 14, 21, 30, 90, 180),
    labels = c("Apheresis","-5", "0", "7", "14", "21", "30", "90", "180")
  ) #+ ylim(0, 6)

# ANC
val <- fig_dat %>% reframe(anc_apheresis = MedianCI(anc_apheresis, conf.level = 0.95, na.rm = TRUE),
                  anc_day_5_k_ul = MedianCI(anc_day_5_k_ul, conf.level = 0.95, na.rm = TRUE),
                  anc_day_0_k_u_l = MedianCI(anc_day_0_k_u_l, conf.level = 0.95, na.rm = TRUE),
                  anc_day_7 = MedianCI(anc_day_7, conf.level = 0.95, na.rm = TRUE),
                  anc_day_14 = MedianCI(anc_day_14, conf.level = 0.95, na.rm = TRUE),
                  anc_day_21 = MedianCI(anc_day_21, conf.level = 0.95, na.rm = TRUE),
                  anc_day_30 = MedianCI(anc_day_30, conf.level = 0.95, na.rm = TRUE),
                  anc_day_90 = MedianCI(anc_day_90, conf.level = 0.95, na.rm = TRUE),
                  anc_day_180 = MedianCI(anc_day_180, conf.level = 0.95, na.rm = TRUE))
timepoints <- c(-28, -5, 0, 7, 14, 21, 30, 90, 180)
ANC_fig <-
  data.frame(timepoints,
             t(val))

ANC_fig <- ANC_fig %>% 
  ggplot(aes(x = timepoints, y = X1, group = 1)) + 
  geom_line(col = "blue") + 
  xlab("Days relative to infusion") + 
  ylab("Median ANC") + 
  theme_classic() + 
  geom_ribbon(aes(ymin = X2, ymax = X3), alpha = 0.1) + 
  scale_x_continuous(
    breaks = c(-28, -5, 0, 7, 14, 21, 30, 90, 180),
    labels = c("Apheresis","-5", "0", "7", "14", "21", "30", "90", "180")
  ) #+ ylim(0, 6)

# HB
val <- fig_dat %>% reframe(hb_apheresis = MedianCI(hb_apheresis, conf.level = 0.95, na.rm = TRUE),
                  hb_day_5 = MedianCI(hb_day_5, conf.level = 0.95, na.rm = TRUE),
                  hb_day_0_k_u_l = MedianCI(hb_day_0_k_u_l, conf.level = 0.95, na.rm = TRUE),
                  hb_day_7 = MedianCI(hb_day_7, conf.level = 0.95, na.rm = TRUE),
                  hb_day_14 = MedianCI(hb_day_14, conf.level = 0.95, na.rm = TRUE),
                  hb_day_21 = MedianCI(hb_day_21, conf.level = 0.95, na.rm = TRUE),
                  hb_day_30 = MedianCI(hb_day_30, conf.level = 0.95, na.rm = TRUE),
                  hb_day_90 = MedianCI(hb_day_90, conf.level = 0.95, na.rm = TRUE),
                  hb_day_180 = MedianCI(hb_day_180, conf.level = 0.95, na.rm = TRUE))
timepoints <- c(-28, -5, 0, 7, 14, 21, 30, 90, 180)
Hb_fig <-
  data.frame(timepoints,
             t(val))

Hb_fig <- Hb_fig %>% 
  ggplot(aes(x = timepoints, y = X1, group = 1)) + 
  geom_line(col = "blue") + 
  xlab("Days relative to infusion") + 
  ylab("Median Hb") + 
  theme_classic() + 
  geom_ribbon(aes(ymin = X2, ymax = X3), alpha = 0.1) + 
  scale_x_continuous(
    breaks = c(-28, -5, 0, 7, 14, 21, 30, 90, 180),
    labels = c("Apheresis","-5", "0", "7", "14", "21", "30", "90", "180")
  ) # +ylim(7.5, 12)

# ALC
val <- fig_dat %>% reframe(alc_apheresis = MedianCI(alc_apheresis, conf.level = 0.95, na.rm = TRUE),
                  alc_day_5_k_ul = MedianCI(alc_day_5_k_ul, conf.level = 0.95, na.rm = TRUE),
                  alc_day_0_k_u_l = MedianCI(alc_day_0_k_u_l, conf.level = 0.95, na.rm = TRUE),
                  alc_day_7 = MedianCI(alc_day_7, conf.level = 0.95, na.rm = TRUE),
                  alc_day_14 = MedianCI(alc_day_14, conf.level = 0.95, na.rm = TRUE),
                  alc_day_21 = MedianCI(alc_day_21, conf.level = 0.95, na.rm = TRUE),
                  alc_day_30 = MedianCI(alc_day_30, conf.level = 0.95, na.rm = TRUE),
                  alc_day_90 = MedianCI(alc_day_90, conf.level = 0.95, na.rm = TRUE),
                  alc_day_180 = MedianCI(alc_day_180, conf.level = 0.95, na.rm = TRUE))
timepoints <- c(-28, -5, 0, 7, 14, 21, 30, 90, 180)
ALC_fig <-
  data.frame(timepoints,
             t(val))

ALC_fig <- ALC_fig %>% 
  ggplot(aes(x = timepoints, y = X1, group = 1)) + 
  geom_line(col = "blue") + 
  xlab("Days relative to infusion") + 
  ylab("Median ALC") + 
  theme_classic() + 
  geom_ribbon(aes(ymin = X2, ymax = X3), alpha = 0.1) + 
  scale_x_continuous(
    breaks = c(-28, -5, 0, 7, 14, 21, 30, 90, 180),
    labels = c("Apheresis","-5", "0", "7", "14", "21", "30", "90", "180")
  ) # + ylim(0, 1.1)

# Plt
val <- fig_dat %>% reframe(plt_apheresis = MedianCI(plt_apheresis, conf.level = 0.95, na.rm = TRUE),
                  plt_day_5_k_ul = MedianCI(plt_day_5_k_ul, conf.level = 0.95, na.rm = TRUE),
                  plt_day_0_k_u_l = MedianCI(plt_day_0_k_u_l, conf.level = 0.95, na.rm = TRUE),
                  plt_day_7 = MedianCI(plt_day_7, conf.level = 0.95, na.rm = TRUE),
                  plt_day_14 = MedianCI(plt_day_14, conf.level = 0.95, na.rm = TRUE),
                  plt_day_21 = MedianCI(plt_day_21, conf.level = 0.95, na.rm = TRUE),
                  plt_day_30 = MedianCI(plt_day_30, conf.level = 0.95, na.rm = TRUE),
                  plt_day_90 = MedianCI(plt_day_90, conf.level = 0.95, na.rm = TRUE),
                  plt_day_180 = MedianCI(plt_day_180, conf.level = 0.95, na.rm = TRUE))
timepoints <- c(-28, -5, 0, 7, 14, 21, 30, 90, 180)
Plt_fig <-
  data.frame(timepoints,
             t(val))

Plt_fig <- Plt_fig %>% 
  ggplot(aes(x = timepoints, y = X1, group = 1)) + 
  geom_line(col = "blue") + 
  xlab("Days relative to infusion") + 
  ylab("Median Plt") + 
  theme_classic() + 
  geom_ribbon(aes(ymin = X2, ymax = X3), alpha = 0.1) + 
  scale_x_continuous(
    breaks = c(-28, -5, 0, 7, 14, 21, 30, 90, 180),
    labels = c("Apheresis","-5", "0", "7", "14", "21", "30", "90", "180")
  ) # + ylim(20, 170)

# IgG
timepoints2 <- c(-28, -5, 30, 90, 180)
val <- fig_dat %>% reframe(ig_g_apheresis = MedianCI(ig_g_apheresis, conf.level = 0.95, na.rm = TRUE),
                  ig_g_day_5 = MedianCI(ig_g_day_5, conf.level = 0.95, na.rm = TRUE),
                  ig_g_day_30 = MedianCI(ig_g_day_30, conf.level = 0.95, na.rm = TRUE),
                  ig_g_day_90 = MedianCI(ig_g_day_90, conf.level = 0.95, na.rm = TRUE),
                  ig_g_day_180 = MedianCI(ig_g_day_180, conf.level = 0.95, na.rm = TRUE))
IgG_fig <-
  data.frame(timepoints2,
             t(val))

IgG_fig <- IgG_fig %>% 
  ggplot(aes(x = timepoints2, y = X1, group = 1)) + 
  geom_line(col = "blue") +
  xlab("Days relative to infusion") + 
  ylab("Median IgG") + 
  theme_classic() + 
  geom_ribbon(aes(ymin = X2, ymax = X3), alpha = 0.1) + 
  scale_x_continuous(
    breaks = c(-28, -5, 30, 90, 180),
    labels = c("Apheresis", "-5", "30", "90", "180"))# + 
  # ylim(0, 2000)

# CD4
timepoints3 <- c(-28, 30, 90, 180)
val <- fig_dat %>% reframe(cd4_apheresis = MedianCI(cd4_apheresis, conf.level = 0.95, na.rm = TRUE),
                  cd4_day_30 = MedianCI(cd4_day_30, conf.level = 0.95, na.rm = TRUE),
                  cd4_day_90 = MedianCI(cd4_day_90, conf.level = 0.95, na.rm = TRUE),
                  cd4_day_180 = MedianCI(cd4_day_180, conf.level = 0.95, na.rm = TRUE))
CD4_fig <-
  data.frame(timepoints3,
             t(val))

CD4_fig <- CD4_fig %>% 
  ggplot(aes(x = timepoints3, y = X1, group = 1)) + 
  geom_line(col = "blue") + 
  xlab("Days relative to infusion") + 
  ylab("Median CD4") + 
  theme_classic() + 
  geom_ribbon(aes(ymin = X2, ymax = X3), alpha = 0.1) + 
  scale_x_continuous(
    breaks = c(-28, 30, 90, 180),
    labels = c("Apheresis", "30", "90", "180")
  )

#CD8
timepoints3 <- c(-28, 30, 90, 180)
val <- fig_dat %>% reframe(cd8_apheresis = MedianCI(cd8_apheresis, conf.level = 0.95, na.rm = TRUE),
                  cd8_day_30 = MedianCI(cd8_day_30, conf.level = 0.95, na.rm = TRUE),
                  cd8_day_90 = MedianCI(cd8_day_90, conf.level = 0.95, na.rm = TRUE),
                  cd8_day_180 = MedianCI(cd8_day_180, conf.level = 0.95, na.rm = TRUE))
CD8_fig <-
  data.frame(timepoints3,
             t(val))

CD8_fig <- CD8_fig %>% 
  ggplot(aes(x = timepoints3, y = X1, group = 1)) + 
  geom_line(col = "blue") + 
  xlab("Days relative to infusion") + 
  ylab("Median CD8") + 
  theme_classic() + 
  geom_ribbon(aes(ymin = X2, ymax = X3), alpha = 0.1) + 
  scale_x_continuous(
    breaks = c(-28, 30, 90, 180),
    labels = c("Apheresis", "30", "90", "180")
  )

#CD19
timepoints3 <- c(-28, 30, 90, 180)
val <- fig_dat %>% reframe(cd19_apheresis = MedianCI(cd19_apheresis, conf.level = 0.95, na.rm = TRUE),
                  cd19_day_30 = MedianCI(cd19_day_30, conf.level = 0.95, na.rm = TRUE),
                  cd19_day_90 = MedianCI(cd19_day_90, conf.level = 0.95, na.rm = TRUE),
                  cd19_day_180 = MedianCI(cd19_day_180, conf.level = 0.95, na.rm = TRUE))
CD19_fig <-
  data.frame(timepoints3,
             t(val))

CD19_fig <- CD19_fig %>% 
  ggplot(aes(x = timepoints3, y = X1, group = 1)) + 
  geom_line(col = "blue") + 
  xlab("Days relative to infusion") + 
  ylab("Median CD19") + 
  theme_classic() + 
  geom_ribbon(aes(ymin = X2, ymax = X3), alpha = 0.1) + 
  scale_x_continuous(
    breaks = c(-28, 30, 90, 180),
    labels = c("Apheresis", "30", "90", "180")
  )

(WBC_fig | ANC_fig | CD4_fig) /
  (Hb_fig | ALC_fig | CD8_fig) /
  (Plt_fig | IgG_fig | CD19_fig) +
  plot_annotation(tag_levels = "A")
```
<br>

<!-- Aggregated median -->
<!-- ```{r timeserie aggreagate med, fig.width=12, fig.height=12} -->
<!-- fig_dat <- cilta %>%  -->
<!--   distinct(pt_identifier, .keep_all = TRUE) -->

<!-- timepoints <- c(-28, -5, 0, 7, 14, 21, 30, 90, 180) -->
<!-- WBC_agg <- fig_dat %>%  -->
<!--   select(site, starts_with("wbc_")) %>%  -->
<!--   group_by(site) %>%  -->
<!--   reframe(across(.cols = everything(), \(x) MedianCI(x, na.rm = TRUE))) %>%  -->
<!--   mutate(type = rep(c("Median", "lwr.ci", "upr.ci"), 5), .before = 1) %>%  -->
<!--   ungroup() %>% select(-site) %>%  -->
<!--   group_by(type) %>%  -->
<!--   summarise(across(.cols = everything(), list(median = median) , na.rm = TRUE)) %>%  -->
<!--   ungroup() %>%  -->
<!--   pivot_longer(cols = -type, names_to = 'Type') %>%  -->
<!--   pivot_wider(names_from = type, values_from = value) %>%  -->
<!--   mutate(timepoints = timepoints) %>%  -->
<!--   ggplot(aes(x = timepoints, y = Median)) +  -->
<!--   geom_line(col = "blue") +  -->
<!--   xlab("Days relative to infusion") +  -->
<!--   ylab("Median WBC") +  -->
<!--   theme_classic() +  -->
<!--   geom_ribbon(aes(ymin = lwr.ci, ymax = upr.ci), alpha = 0.1) +  -->
<!--   scale_x_continuous( -->
<!--     breaks = c(-28, -5, 0, 7, 14, 21, 30, 90, 180), -->
<!--     labels = c("Apheresis","-5", "0", "7", "14", "21", "30", "90", "180") -->
<!--   ) -->

<!-- ANC_agg <- fig_dat %>%  -->
<!--   select(site, starts_with("anc_"), -anc_0_75_x_10_9_l_0_75_k_u_l) %>%  -->
<!--   group_by(site) %>%  -->
<!--   reframe(across(.cols = everything(), \(x) MedianCI(x, na.rm = TRUE))) %>%  -->
<!--   mutate(type = rep(c("Median", "lwr.ci", "upr.ci"), 5), .before = 1) %>%  -->
<!--   ungroup() %>% select(-site) %>%  -->
<!--   group_by(type) %>%  -->
<!--   summarise(across(.cols = everything(), list(median = median) , na.rm = TRUE)) %>%  -->
<!--   ungroup() %>%  -->
<!--   pivot_longer(cols = -type, names_to = 'Type') %>%  -->
<!--   pivot_wider(names_from = type, values_from = value) %>%  -->
<!--   mutate(timepoints = timepoints) %>%  -->
<!--   ggplot(aes(x = timepoints, y = Median)) +  -->
<!--   geom_line(col = "blue") +  -->
<!--   xlab("Days relative to infusion") +  -->
<!--   ylab("Median ANC") +  -->
<!--   theme_classic() +  -->
<!--   geom_ribbon(aes(ymin = lwr.ci, ymax = upr.ci), alpha = 0.1) +  -->
<!--   scale_x_continuous( -->
<!--     breaks = c(-28, -5, 0, 7, 14, 21, 30, 90, 180), -->
<!--     labels = c("Apheresis","-5", "0", "7", "14", "21", "30", "90", "180") -->
<!--   )  -->

<!-- Hb_agg <- fig_dat %>%  -->
<!--   select(site, starts_with("hb_")) %>%  -->
<!--   group_by(site) %>%  -->
<!--   reframe(across(.cols = everything(), \(x) MedianCI(x, na.rm = TRUE))) %>%  -->
<!--   mutate(type = rep(c("Median", "lwr.ci", "upr.ci"), 5), .before = 1) %>%  -->
<!--   ungroup() %>% select(-site) %>%  -->
<!--   group_by(type) %>%  -->
<!--   summarise(across(.cols = everything(), list(median = median) , na.rm = TRUE)) %>%  -->
<!--   ungroup() %>%  -->
<!--   pivot_longer(cols = -type, names_to = 'Type') %>%  -->
<!--   pivot_wider(names_from = type, values_from = value) %>%  -->
<!--   mutate(timepoints = timepoints) %>%  -->
<!--   ggplot(aes(x = timepoints, y = Median)) +  -->
<!--   geom_line(col = "blue") +  -->
<!--   xlab("Days relative to infusion") +  -->
<!--   ylab("Median Hb") +  -->
<!--   theme_classic() +  -->
<!--   geom_ribbon(aes(ymin = lwr.ci, ymax = upr.ci), alpha = 0.1) +  -->
<!--   scale_x_continuous( -->
<!--     breaks = c(-28, -5, 0, 7, 14, 21, 30, 90, 180), -->
<!--     labels = c("Apheresis","-5", "0", "7", "14", "21", "30", "90", "180") -->
<!--   )  -->

<!-- ALC_agg <- fig_dat %>%  -->
<!--   select(site, starts_with("alc_"), -alc_0_3_x_10_9_l_0_3k_u_l) %>%  -->
<!--   group_by(site) %>%  -->
<!--   reframe(across(.cols = everything(), \(x) MedianCI(x, na.rm = TRUE))) %>%  -->
<!--   mutate(type = rep(c("Median", "lwr.ci", "upr.ci"), 5), .before = 1) %>%  -->
<!--   ungroup() %>% select(-site) %>%  -->
<!--   group_by(type) %>%  -->
<!--   summarise(across(.cols = everything(), list(median = median) , na.rm = TRUE)) %>%  -->
<!--   ungroup() %>%  -->
<!--   pivot_longer(cols = -type, names_to = 'Type') %>%  -->
<!--   pivot_wider(names_from = type, values_from = value) %>%  -->
<!--   mutate(timepoints = timepoints) %>%  -->
<!--   ggplot(aes(x = timepoints, y = Median)) +  -->
<!--   geom_line(col = "blue") +  -->
<!--   xlab("Days relative to infusion") +  -->
<!--   ylab("Median ALC") +  -->
<!--   theme_classic() +  -->
<!--   geom_ribbon(aes(ymin = lwr.ci, ymax = upr.ci), alpha = 0.1) +  -->
<!--   scale_x_continuous( -->
<!--     breaks = c(-28, -5, 0, 7, 14, 21, 30, 90, 180), -->
<!--     labels = c("Apheresis","-5", "0", "7", "14", "21", "30", "90", "180") -->
<!--   )  -->

<!-- Plt_agg <- fig_dat %>%  -->
<!--   select(site, starts_with("plt_"), -c(starts_with("plt_transfusion"))) %>%  -->
<!--   group_by(site) %>%  -->
<!--   reframe(across(.cols = everything(), \(x) MedianCI(x, na.rm = TRUE))) %>%  -->
<!--   mutate(type = rep(c("Median", "lwr.ci", "upr.ci"), 5), .before = 1) %>%  -->
<!--   ungroup() %>% select(-site) %>%  -->
<!--   group_by(type) %>%  -->
<!--   summarise(across(.cols = everything(), list(median = median) , na.rm = TRUE)) %>%  -->
<!--   ungroup() %>%  -->
<!--   pivot_longer(cols = -type, names_to = 'Type') %>%  -->
<!--   pivot_wider(names_from = type, values_from = value) %>%  -->
<!--   mutate(timepoints = timepoints) %>%  -->
<!--   ggplot(aes(x = timepoints, y = Median)) +  -->
<!--   geom_line(col = "blue") +  -->
<!--   xlab("Days relative to infusion") +  -->
<!--   ylab("Median Plt") +  -->
<!--   theme_classic() +  -->
<!--   geom_ribbon(aes(ymin = lwr.ci, ymax = upr.ci), alpha = 0.1) +  -->
<!--   scale_x_continuous( -->
<!--     breaks = c(-28, -5, 0, 7, 14, 21, 30, 90, 180), -->
<!--     labels = c("Apheresis","-5", "0", "7", "14", "21", "30", "90", "180") -->
<!--   ) -->

<!-- timepoints2 <- c(-28, -5, 30, 90, 180) -->
<!-- IgG_agg <- fig_dat %>%  -->
<!--   select(site, ig_g_apheresis, starts_with("ig_g_day")) %>%  -->
<!--   group_by(site) %>%  -->
<!--   reframe(across(.cols = everything(), \(x) MedianCI(x, na.rm = TRUE))) %>%  -->
<!--   mutate(type = rep(c("Median", "lwr.ci", "upr.ci"), 5), .before = 1) %>%  -->
<!--   ungroup() %>% select(-site) %>%  -->
<!--   group_by(type) %>%  -->
<!--   summarise(across(.cols = everything(), list(median = median) , na.rm = TRUE)) %>%  -->
<!--   ungroup() %>%  -->
<!--   pivot_longer(cols = -type, names_to = 'Type') %>%  -->
<!--   pivot_wider(names_from = type, values_from = value) %>%  -->
<!--   mutate(timepoints = timepoints2) %>%  -->
<!--   ggplot(aes(x = timepoints, y = Median)) +  -->
<!--   geom_line(col = "blue") +  -->
<!--   xlab("Days relative to infusion") +  -->
<!--   ylab("Median Plt") +  -->
<!--   theme_classic() +  -->
<!--   geom_ribbon(aes(ymin = lwr.ci, ymax = upr.ci), alpha = 0.1) +  -->
<!--   scale_x_continuous( -->
<!--     breaks = c(-28, -5, 0, 30, 90, 180), -->
<!--     labels = c("Apheresis", "-5", "0", "30", "90", "180") -->
<!--   ) -->

<!-- timepoints3 <- c(-28, 30, 90, 180) -->
<!-- CD4_fig <- fig_dat %>%  -->
<!--   select(site, starts_with("cd4_")) %>%  -->
<!--   group_by(site) %>%  -->
<!--   reframe(across(.cols = everything(), \(x) MedianCI(x, na.rm = TRUE))) %>%  -->
<!--   mutate(type = rep(c("Median", "lwr.ci", "upr.ci"), 5), .before = 1) %>%  -->
<!--   ungroup() %>% select(-site) %>%  -->
<!--   group_by(type) %>%  -->
<!--   summarise(across(.cols = everything(), list(median = median) , na.rm = TRUE)) %>%  -->
<!--   ungroup() %>%  -->
<!--   pivot_longer(cols = -type, names_to = 'Type') %>%  -->
<!--   pivot_wider(names_from = type, values_from = value) %>%  -->
<!--   mutate(timepoints = timepoints3) %>%  -->
<!--   ggplot(aes(x = timepoints, y = Median)) +  -->
<!--   geom_line(col = "blue") +  -->
<!--   xlab("Days relative to infusion") +  -->
<!--   ylab("Median CD4") +  -->
<!--   theme_classic() +  -->
<!--   geom_ribbon(aes(ymin = lwr.ci, ymax = upr.ci), alpha = 0.1) +  -->
<!--   scale_x_continuous( -->
<!--     breaks = c(-28, 30, 90, 180), -->
<!--     labels = c("Apheresis", "30", "90", "180") -->
<!--   ) -->
<!-- CD8_fig <- fig_dat %>%  -->
<!--   select(site, starts_with("cd8_")) %>%  -->
<!--   group_by(site) %>%  -->
<!--   reframe(across(.cols = everything(), \(x) MedianCI(x, na.rm = TRUE))) %>%  -->
<!--   mutate(type = rep(c("Median", "lwr.ci", "upr.ci"), 5), .before = 1) %>%  -->
<!--   ungroup() %>% select(-site) %>%  -->
<!--   group_by(type) %>%  -->
<!--   summarise(across(.cols = everything(), list(median = median) , na.rm = TRUE)) %>%  -->
<!--   ungroup() %>%  -->
<!--   pivot_longer(cols = -type, names_to = 'Type') %>%  -->
<!--   pivot_wider(names_from = type, values_from = value) %>%  -->
<!--   mutate(timepoints = timepoints3) %>%  -->
<!--   ggplot(aes(x = timepoints, y = Median)) +  -->
<!--   geom_line(col = "blue") +  -->
<!--   xlab("Days relative to infusion") +  -->
<!--   ylab("Median CD8") +  -->
<!--   theme_classic() +  -->
<!--   geom_ribbon(aes(ymin = lwr.ci, ymax = upr.ci), alpha = 0.1) +  -->
<!--   scale_x_continuous( -->
<!--     breaks = c(-28, 30, 90, 180), -->
<!--     labels = c("Apheresis", "30", "90", "180") -->
<!--   ) -->
<!-- CD19_fig <- fig_dat %>%  -->
<!--   select(site, starts_with("cd19_")) %>%  -->
<!--   group_by(site) %>%  -->
<!--   reframe(across(.cols = everything(), \(x) MedianCI(x, na.rm = TRUE))) %>%  -->
<!--   mutate(type = rep(c("Median", "lwr.ci", "upr.ci"), 5), .before = 1) %>%  -->
<!--   ungroup() %>% select(-site) %>%  -->
<!--   group_by(type) %>%  -->
<!--   summarise(across(.cols = everything(), list(median = median) , na.rm = TRUE)) %>%  -->
<!--   ungroup() %>%  -->
<!--   pivot_longer(cols = -type, names_to = 'Type') %>%  -->
<!--   pivot_wider(names_from = type, values_from = value) %>%  -->
<!--   mutate(timepoints = timepoints3) %>%  -->
<!--   ggplot(aes(x = timepoints, y = Median)) +  -->
<!--   geom_line(col = "blue") +  -->
<!--   xlab("Days relative to infusion") +  -->
<!--   ylab("Median CD19") +  -->
<!--   theme_classic() +  -->
<!--   geom_ribbon(aes(ymin = lwr.ci, ymax = upr.ci), alpha = 0.1) +  -->
<!--   scale_x_continuous( -->
<!--     breaks = c(-28, 30, 90, 180), -->
<!--     labels = c("Apheresis", "30", "90", "180") -->
<!--   ) -->

<!-- (WBC_agg | ANC_agg | CD4_fig) / -->
<!--   (Hb_agg | ALC_agg | CD8_fig) / -->
<!--   (Plt_agg | IgG_agg | CD19_fig) + -->
<!--   plot_annotation(tag_levels = "A") -->
<!-- ``` -->

<!-- <span style="color:blue">**IgG I need to fix the other 109**</span>  -->

<!-- Values less than the limit of detection (<109) were considered to be 77 which is the lower limit of detection divided by the square root of 2. One patient had an IgG day 0 value of <320. As this is above the limit of detection, I used the midway point between 320 and the lower limit of detection -- 214 -- as the value instead of <320.  -->

<br>

## Figure 4. Cytopenias over time

```{r Figure 4. Cytopenias over time, fig.width=12, fig.height=7}
# fig_dat %>% 
#   select(PT,
#          `gr3_Leukopenia_-5d`, gr3_Leukopenia_30d, gr3_Leukopenia_60d, gr3_Leukopenia_90d,
#          `gr3_Neutropenia_-5d`, gr3_Neutropenia_30d, gr3_Neutropenia_60d, gr3_Neutropenia_90d,
#          `gr3_Anemia_-5d`, gr3_Anemia_30d, gr3_Anemia_60d, gr3_Anemia_90d,
#          `gr3_Thrombocytopenia_-5d`, gr3_Thrombocytopenia_30d, gr3_Thrombocytopenia_60d, gr3_Thrombocytopenia_90d
#   ) %>% 
#   pivot_longer(cols = `gr3_Leukopenia_-5d`: gr3_Thrombocytopenia_90d) %>% 
#   group_by(name, value) %>% 
#   filter(!is.na(value)) %>% 
#   summarise(count=n()) %>% 
#   mutate(perc=(count/sum(count)*100)
#   ) %>% 
#   separate("name", into = c("gr3", "name", "day"), sep = "_") %>% 
#   filter(value == "Yes") %>% 
#   # unite(time, c("day":"me"), sep = " ") %>% 
#   ggplot(aes(x= day, y= perc, fill= name))+
#   geom_bar(stat = "identity",
#            position=position_dodge(width=0.9))+
#   scale_fill_viridis_d() +
#   labs(fill = NULL, y = "%", title = "Grade 3+")+
#   theme_classic()
# 
# fig_dat %>% 
#   select(PT,
#          `Leukopenia_-5d`, Leukopenia_30d, Leukopenia_60d, Leukopenia_90d,
#          `Neutropenia_-5d`, Neutropenia_30d, Neutropenia_60d, Neutropenia_90d,
#          `Anemia_-5d`, Anemia_30d, Anemia_60d, Anemia_90d,
#          `Thrombocytopenia_-5d`, Thrombocytopenia_30d, Thrombocytopenia_60d, Thrombocytopenia_90d,
#          
#          `gr3_Leukopenia_-5d`, gr3_Leukopenia_30d, gr3_Leukopenia_60d, gr3_Leukopenia_90d,
#          `gr3_Neutropenia_-5d`, gr3_Neutropenia_30d, gr3_Neutropenia_60d, gr3_Neutropenia_90d,
#          `gr3_Anemia_-5d`, gr3_Anemia_30d, gr3_Anemia_60d, gr3_Anemia_90d,
#          `gr3_Thrombocytopenia_-5d`, gr3_Thrombocytopenia_30d, gr3_Thrombocytopenia_60d, gr3_Thrombocytopenia_90d
#   ) %>% 
#   pivot_longer(cols = `Leukopenia_-5d`: gr3_Thrombocytopenia_90d) %>% 
#   mutate(grade = str_match(name, "(gr3_)(.*)")[,2],
#          grade = case_when(
#            grade == "gr3_"       ~ "Grade 3+",
#            TRUE                  ~ "Any grade"
#          )) %>% 
#   mutate(name1 = str_match(name, "(gr3_)(.*)")[,3],
#          name2 = coalesce(name1, name)) %>% 
#   group_by(name2, value, grade) %>% 
#   filter(!is.na(value)) %>% 
#   summarise(count=n()) %>%
#   group_by(name2, grade) %>% 
#   mutate(perc=(count/sum(count)*100)
#   ) %>%
#   separate("name2", into = c("name", "day"), sep = "_") %>% 
#   filter(value == "Yes") %>% 
#   # unite(time, c("day":"me"), sep = " ") %>% 
#   ggplot(aes(x= day, y= perc, fill= name, alpha= grade))+
#   geom_bar(stat = "identity",
#            position=position_dodge(width=0.9))+
#   scale_fill_viridis_d() + 
#   scale_alpha_ordinal(range = c(0.5, 1))+
#   labs(fill = NULL, y = "%")+
#   theme_classic()

a <- fig_dat %>% 
  select(pt_identifier,
         Leukopenia_apheresis,
         `Leukopenia_-5d`, Leukopenia_30d, #Leukopenia_60d,
         Leukopenia_90d, Leukopenia_180d,
         Neutropenia_apheresis,
         `Neutropenia_-5d`, Neutropenia_30d, #Neutropenia_60d,
         Neutropenia_90d, Neutropenia_180d,
         Anemia_apheresis,
         `Anemia_-5d`, Anemia_30d, #Anemia_60d, 
         Anemia_90d, Anemia_180d,
         Thrombocytopenia_apheresis,
         `Thrombocytopenia_-5d`, Thrombocytopenia_30d, #Thrombocytopenia_60d, 
         Thrombocytopenia_90d, Thrombocytopenia_180d,
         cytopenia_apheresis, `cytopenia_-5d`, cytopenia_30d,# cytopenia_60d, 
         cytopenia_90d, cytopenia_180d,

         gr3_Leukopenia_apheresis,
         `gr3_Leukopenia_-5d`, gr3_Leukopenia_30d, #gr3_Leukopenia_60d, 
         gr3_Leukopenia_90d, gr3_Leukopenia_180d,
         gr3_Neutropenia_apheresis,
         `gr3_Neutropenia_-5d`, gr3_Neutropenia_30d, #gr3_Neutropenia_60d, 
         gr3_Neutropenia_90d, gr3_Neutropenia_180d,
         gr3_Anemia_apheresis,
         `gr3_Anemia_-5d`, gr3_Anemia_30d, #gr3_Anemia_60d, 
         gr3_Anemia_90d, gr3_Anemia_180d,
         gr3_Thrombocytopenia_apheresis,
         `gr3_Thrombocytopenia_-5d`, gr3_Thrombocytopenia_30d, #gr3_Thrombocytopenia_60d, 
         gr3_Thrombocytopenia_90d, gr3_Thrombocytopenia_180d,
         gr3_cytopenia_apheresis, `gr3_cytopenia_-5d`, gr3_cytopenia_30d, #gr3_cytopenia_60d, 
         gr3_cytopenia_90d, gr3_cytopenia_180d
  ) %>% 
  pivot_longer(cols = Neutropenia_apheresis: gr3_cytopenia_180d) %>% 
  mutate(grade = str_match(name, "(gr3_)(.*)")[,2],
         grade = case_when(
           grade == "gr3_"       ~ "Grade 3+",
           TRUE                  ~ "Any grade"
         )) %>% 
  mutate(name1 = str_match(name, "(gr3_)(.*)")[,3],
         name2 = coalesce(name1, name)) %>% 
  group_by(name2, value, grade) %>% 
  filter(!is.na(value)) %>%
  summarise(count=n()) %>%
  group_by(name2, grade) %>% 
  mutate(perc=(count/sum(count)*100)
  ) %>%
  separate("name2", into = c("name", "day"), sep = "_") %>% 
  ungroup()

# We want to plot the Yes patients 
# but filtering filter(value == "Yes") leave blanks if all patients are NO
# Code to fix it is :
c <- a %>% 
  arrange(name, day, grade, desc(value)) %>%  # desc to have yes first
  distinct(name, day, grade, .keep_all = TRUE) %>% 
  mutate(count = case_when(
    value == "No"           ~ 0,
    TRUE                    ~ count
  ), perc = case_when(
    value == "No"           ~ 0,
    TRUE                    ~ perc
  )) %>%

  # filter(value == "Yes") %>% 
  # add_row("name"="Leukopenia",  "day"="apheresis", "value"= "Yes", "grade"="Any grade", "count"=0, "perc"=0) %>% 
  mutate(day = factor(day, levels = c("apheresis", "-5d", "30d", "60d", "90d", "180d"))) %>% 
  mutate(name = str_replace(name, "^cytopenia", "Any Cytopenia")) %>% 
  mutate(name = factor(name, levels = c("Anemia", "Neutropenia", "Thrombocytopenia", "Any Cytopenia"))) %>% 
  filter(!is.na(name))

b <- c %>% filter(grade == "Any grade")
c %>% filter(grade == "Grade 3+") %>% 
  ggplot(aes(x= day, 
             y= perc, 
             fill= name, alpha= 1
             ))+
  geom_bar(stat = "identity",
           position=position_dodge(width=0.9))+
  geom_bar(stat = "identity",
           position=position_dodge(width=0.9), data= b,
           aes(x= day, y= perc, fill= name), alpha= 0.4)+
  scale_fill_brewer(palette = "Spectral")+
  labs(fill = NULL, alpha=NULL, x= NULL, y = "%")+
  theme_classic()+
  scale_alpha_continuous(limits = c(0.4,1), breaks = c(0.4,1),
                         labels= c("Any Grade", "Grade 3+"))

c %>% filter(grade == "Grade 3+") %>% 
  ggplot(aes(x= day, 
             y= perc, 
             fill= name, 
             ))+
  geom_bar(stat = "identity",
           position=position_dodge(width=0.9))+
  geom_text(aes(label = paste0(round(perc, 0), "%"), y=perc-2), size = 5,
            position = position_dodge(width=0.9), vjust=-1.5
            )+

  scale_fill_brewer(palette = "Spectral")+
  ylim(0, 60)+
  labs(fill = NULL, x= NULL, y = "%")+
  theme_classic(base_size = 14)+
  theme(legend.position="bottom")
```

## Figure 5. Cumulative incidence of infection and infection density
```{r Figure 5}
library(tidycmprsk)
library(ggsurvfit)

risk_comp <- cilta %>%
  mutate(os_from_infusion_days = os_from_infusion_months * 30.417) %>% 
  mutate(event_type = case_when(
    infection_type_bacterial_viral_fungal_parasitic == "Bacterial"      ~ "Bacterial",
    infection_type_bacterial_viral_fungal_parasitic == "Fungal"         ~ "Fungal",
    infection_type_bacterial_viral_fungal_parasitic == "Viral"          ~ "Viral",
    is.na(infection_type_bacterial_viral_fungal_parasitic)              ~ NA_character_
  )) %>%
  mutate(event_type =
           factor(event_type, levels= c("Bacterial","Fungal", "Viral"))) %>%
  mutate(infection_event = case_when(
    is.na(event_type)                        ~ 0,
    !is.na(event_type)                       ~ 1
  )) %>% 
  # mutate(event_type = case_when(
  #   is.na(event_type)                        ~ "N",
  #   TRUE ~ event_type
  # )) %>% 
  mutate(date_of_event = interval(start = date_of_car_t_infusion,
                                  end = date_of_infection) /
           duration(n = 1, units = "days"),
         date_of_event = case_when(
      is.na(date_of_event)             ~ os_from_infusion_days,
      TRUE                             ~ date_of_event)) %>% 
  select(pt_identifier, date_of_event, infection_event, event_type, os_from_infusion_days, os_event, race) 
# 
# cum_inc <- cuminc(risk_comp$date_of_event, risk_comp$event_type)
# timepoints(cum_inc, times = c(0, 100, 200, 300, 400))$est
# 
# ggcompetingrisks(fit = cum_inc, multiple_panels = F, xlab = "Days", ylab = "Cumulative incidence of event",title = "Competing Risks Analysis"
#                  )

# cum_inc <- cuminc(Surv(date_of_event, infection_event) ~ 1, data = b) %>%
#   ggcuminc(outcome = c("Bacterial", "Fungal", "Viral"),
#            geom_step(linewidth=10)) +
#   ylim(c(0, 1)) +
#   labs(
#     x = "Days"
#   )+
#   add_risktable()
# cum_inc
# 
# survfit2(Surv(date_of_event, event_type) ~ 1, data = a) %>%
#   ggcuminc(outcome = c("Viral", "Fungal")#,
#            # geom_step(linewidth=10)
#            ) +
#   # ylim(c(0, 0.25)) +
#   labs(
#     x = "Days"
#   )+
#   add_risktable()
```

```{r ggsurvplot combine}
# risk_comp <- risk_comp %>% 
#   filter(date_of_event > 0) 
# a <- risk_comp %>% 
#   arrange(pt_identifier, desc(infection_event)) %>% 
#   distinct(pt_identifier, .keep_all = TRUE)
# 
# # OS
# os <- survfit(Surv(os_from_infusion_days, os_event) ~ 1, data=a)
# ggsurvplot(survfit(Surv(os_from_infusion_days, os_event) ~ 1, 
#                    data=a),
#            title = "OS Analysis",
#            fun = "event",
#            
#            font.main = c(20, "bold", "black"),
#            font.x = c(18, "bold", "black"),
#            font.y = c(18, "bold", "black"),
#            font.legend = c(16, "black"),
#            font.tickslab = c(16, "bold", "black"),
#            size = 1.5,
# 
#            xlab = "Time in days", 
#            palette = "orange",
#            legend = "top",
#            legend.title = "",
#            # pval = TRUE,
#            conf.int = FALSE,
#            # Censor
#            censor = TRUE
# )
# 
# # Overall infections
# ov_inf <- survfit(Surv(date_of_event, infection_event) ~ 1, data=a)
# ggsurvplot(survfit(Surv(date_of_event, infection_event) ~ 1, 
#                    data=a),
#            title = "Cumulative Incidence Analysis",
#            fun = "event",
#            
#            font.main = c(20, "bold", "black"),
#            font.x = c(18, "bold", "black"),
#            font.y = c(18, "bold", "black"),
#            font.legend = c(16, "black"),
#            font.tickslab = c(16, "bold", "black"),
#            size = 1.5,
# 
#            xlab = "Time in days",
#            legend = "top",
#            legend.title = "",
#            legend.labs = c("Overall infections"),
#            # pval = TRUE,
#            conf.int = FALSE,
#            # Censor
#            censor = TRUE
# )
# 
# # By infection types
# ggsurvplot(survfit(Surv(date_of_event, infection_event) ~ event_type, 
#                    data=a),
#            title = "Cumulative Incidence Analysis",
#            fun = "event",
#            
#            font.main = c(20, "bold", "black"),
#            font.x = c(18, "bold", "black"),
#            font.y = c(18, "bold", "black"),
#            font.legend = c(16, "black"),
#            font.tickslab = c(16, "bold", "black"),
#            size = 1.5,
# 
#            xlab = "Time in days",
#            legend = "top",
#            legend.title = "",
#            # legend.labs = c("Bact", "Fungal"),
#            # palette = c("blue", "green", "purple", "red"),
#            pval = TRUE,
#            conf.int = FALSE,
#            # Censor
#            censor = TRUE
# ) + guides(colour = guide_legend(ncol = 1))
# 
# 
# 
# 
# 
# b <- risk_comp %>% 
#   mutate(event_type =
#            factor(event_type, levels= c("Bacterial"))) %>%
#   mutate(infection_event = case_when(
#     event_type == "Bacterial"     ~ 1,
#     TRUE                          ~ 0
#   )) %>% 
#   arrange(pt_identifier, desc(event_type)) %>% 
#   distinct(pt_identifier, .keep_all = TRUE)
# 
# bact_inf <- survfit(Surv(date_of_event, infection_event) ~ 1, data=b)
# ggsurvplot(survfit(Surv(date_of_event, infection_event) ~ 1, 
#                    data=b),
#            title = "Cumulative Incidence Analysis",
#            fun = "event",
#            
#            font.main = c(20, "bold", "black"),
#            font.x = c(18, "bold", "black"),
#            font.y = c(18, "bold", "black"),
#            font.legend = c(16, "black"),
#            font.tickslab = c(16, "bold", "black"),
#            size = 1.5,
# 
#            xlab = "Time in days",
#            legend = "top",
#            legend.title = "",
#            legend.labs = c("Bacterial"),
#            # palette = c("blue", "green", "purple", "red"),
#            # pval = TRUE,
#            conf.int = FALSE,
#            # Censor
#            censor = TRUE
# ) + guides(colour = guide_legend(ncol = 1))
# 
# b <- risk_comp %>% 
#   mutate(event_type =
#            factor(event_type, levels= c("Viral"))) %>%
#   # mutate(date_of_event = case_when(
#     # pt_identifier == ""   ~ 385,
#     # TRUE                          ~ date_of_event
#     # )) %>%
#   mutate(infection_event = case_when(
#     # pt_identifier == "" ~ 0,
#     event_type == "Viral"     ~ 1,
#     TRUE                          ~ 0
#   )) %>% 
#   arrange(pt_identifier, desc(event_type)) %>% 
#   distinct(pt_identifier, .keep_all = TRUE)
# 
# viral_inf <- survfit(Surv(date_of_event, infection_event) ~ 1, data=b)
# ggsurvplot(survfit(Surv(date_of_event, infection_event) ~ 1, 
#                    data=b),
#            title = "Cumulative Incidence Analysis",
#            fun = "event",
#            
#            font.main = c(20, "bold", "black"),
#            font.x = c(18, "bold", "black"),
#            font.y = c(18, "bold", "black"),
#            font.legend = c(16, "black"),
#            font.tickslab = c(16, "bold", "black"),
#            size = 1.5,
# 
#            xlab = "Time in days",
#            legend = "top",
#            legend.title = "",
#            legend.labs = c("Viral"),
#            # palette = c("blue", "green", "purple", "red"),
#            # pval = TRUE,
#            cumevents = TRUE,
#            conf.int = FALSE,
#            # Censor
#            censor = TRUE
# ) + guides(colour = guide_legend(ncol = 1))
# 
# ggsurvfit(viral_inf, type = "risk") +
#   labs(
#     x = "Days",
#     y = "Event-free proportion"
#   ) +
#   add_censor_mark()
# 
# survfit2(Surv(date_of_event, infection_event) ~ 1, b) %>%
#   ggcuminc(outcome = "Viral") +
#   add_confidence_interval() +
#   add_risktable() +
#   scale_ggsurvfit()
# 
# 
# risk_comp %>% 
#   filter(event_type == "Viral") %>%
#   arrange(date_of_event) %>% 
#   distinct(pt_identifier, .keep_all = TRUE) %>% 
#   mutate(n = cumsum(infection_event)) %>% 
#   mutate(n2 = n / 105) %>%
#   ggplot(aes(x = date_of_event, y = n2))+
#   geom_step()+
#   ylim(0,1)
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# # Fungal
# b <- risk_comp %>% 
#   mutate(event_type =
#            factor(event_type, levels= c("Fungal"))) %>%
#   mutate(infection_event = case_when(
#     event_type == "Fungal"     ~ 1,
#     TRUE                          ~ 0
#   )) %>% 
#   arrange(pt_identifier, desc(event_type)) %>% 
#   distinct(pt_identifier, .keep_all = TRUE)
# 
# fungal_inf <- survfit(Surv(date_of_event, infection_event) ~ 1, data=b)
# ggsurvplot(survfit(Surv(date_of_event, infection_event) ~ 1, 
#                    data=b),
#            title = "Cumulative Incidence Analysis",
#            fun = "event",
#            
#            font.main = c(20, "bold", "black"),
#            font.x = c(18, "bold", "black"),
#            font.y = c(18, "bold", "black"),
#            font.legend = c(16, "black"),
#            font.tickslab = c(16, "bold", "black"),
#            size = 1.5,
# 
#            xlab = "Time in days",
#            legend = "top",
#            legend.title = "",
#            legend.labs = c("Fungal"),
#            # palette = c("blue", "green", "purple", "red"),
#            # pval = TRUE,
#            conf.int = FALSE,
#            # Censor
#            censor = TRUE
# ) + guides(colour = guide_legend(ncol = 1))
# 
# # inf_by_type <- survfit(Surv(date_of_event, infection_event) ~ 1, data=risk_comp)
# 
# fit <- list(Overall_infections = ov_inf, OS = os, 
#             Bacterial = bact_inf,
#             Viral = viral_inf,
#             Fungal = fungal_inf
#             )
# ggsurvplot_combine(fit, risk_comp, combine = TRUE,
#                    fun = "event",
#                    legend.title = "",
#                    ylim = c(0,1)
#                    ) + 
#   guides(colour = guide_legend(ncol = 1))
# 
# 
# fit <- list(By_type = inf_by_type, OS = os)
# ggsurvplot_combine(fit, risk_comp, combine = TRUE,
#                    fun = "cumhaz") + 
#   guides(colour = guide_legend(ncol = 1))
# 
# 
# fit <- list(Overall_infections = ov_inf, By_type = inf_by_type, OS = os)
# ggsurvplot_combine(fit, risk_comp, combine = TRUE,
#                    fun = "cumhaz") + 
#   guides(colour = guide_legend(ncol = 1))
```

## Figure S1. Viral titers after ide-cel
Focus on MCC patients / I didn't use day 180 as they are less values but I can add it if you want.
```{r Figure S1. Viral titers after ide-cel}
fig_dat <- cilta %>% 
  distinct(pt_identifier, .keep_all = TRUE)
moffitt_subset <- fig_dat %>% filter(site == "Moffitt")
```

### A)	HSV 1/2 IgG
See below
```{r HSV 1/2 IgG}
# Alluvial
library(ggalluvial)

alluvial_A <- moffitt_subset %>% 
  select(pt_identifier, igg_hsv_change,
         ig_g_hsv1_2_apheresis, ig_g_hsv1_2_d90) %>%
  drop_na() %>%
  group_by(ig_g_hsv1_2_apheresis, ig_g_hsv1_2_d90) %>% 
  summarise(count=n()) %>%

  ggplot(aes(axis1 = ig_g_hsv1_2_apheresis, axis2 = ig_g_hsv1_2_d90,
             y = count)) +
  geom_alluvium(aes(fill = ig_g_hsv1_2_apheresis)
  )+
  geom_stratum()+
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  theme_classic()+
  ggtitle("Evolution of virus titers")
```

### B)	VZV IgG (positive=immune vs negative or equivocal=not immune)
See below
```{r VZV IgG}
# Alluvial
alluvial_B <- moffitt_subset %>% 
  select(pt_identifier, igg_vzv_change,
         vzv_ig_g_apheresis, vzv_ig_g_d90
  ) %>% 
  drop_na() %>% 
  group_by(vzv_ig_g_apheresis, vzv_ig_g_d90) %>% 
  summarise(count=n()) %>%

  ggplot(aes(axis1 = vzv_ig_g_apheresis, axis2= vzv_ig_g_d90,
             y = count)) +
  geom_alluvium(aes(fill = vzv_ig_g_apheresis), 
                color = c("transparent", "transparent", "white", "transparent", "transparent", "transparent")
  )+
  geom_stratum()+
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  # geom_hline(yintercept = 5, color = "blue")+
  theme_classic()+
  ggtitle("Evolution of virus titers")
```

### C)	CMV IgG (positive=immune vs negative=not immune)
```{r CMV IgG}
alluvial_C <- moffitt_subset %>% 
  select(pt_identifier, igg_cmv_change,
         cmv_ig_g_apheresis, cmv_ig_g_d90
  ) %>% 
  drop_na() %>% 
  group_by(cmv_ig_g_apheresis, cmv_ig_g_d90) %>% 
  summarise(count=n()) %>%

  ggplot(aes(axis1 = cmv_ig_g_apheresis, axis2= cmv_ig_g_d90,
             y = count)) +
  geom_alluvium(aes(fill = cmv_ig_g_apheresis)
  )+
  geom_stratum()+
  # geom_segment(aes(x=1.17, y=12, xend=1.83, yend=12), 
  #              color= "white", size=0.3)+
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  theme_classic()+
  ggtitle("Evolution of virus titers")
```

```{r alluvial combine, fig.height=12}
(alluvial_A) /
  (alluvial_B) + #/
  # (alluvial_C) +
  plot_annotation(tag_levels = "A")
```

### A)	HSV 1/2 IgG with day 180
See below
```{r HSV 1/2 IgG with day 180}
# Alluvial
library(ggalluvial)

alluvial_A <- moffitt_subset %>% 
  select(pt_identifier, igg_hsv_change,
         ig_g_hsv1_2_apheresis, ig_g_hsv1_2_d90, ig_g_hsv1_2_d180) %>%
  drop_na() %>%
  group_by(ig_g_hsv1_2_apheresis, ig_g_hsv1_2_d90, ig_g_hsv1_2_d180) %>% 
  summarise(count=n()) %>%

  ggplot(aes(axis1 = ig_g_hsv1_2_apheresis, axis2 = ig_g_hsv1_2_d90, axis3 = ig_g_hsv1_2_d180,
             y = count)) +
  geom_alluvium(aes(fill = ig_g_hsv1_2_apheresis)
  )+
  geom_stratum()+
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  theme_classic()+
  ggtitle("Evolution of virus titers")
```
12 patients don't have data at day 180.

### B)	VZV IgG (positive=immune vs negative or equivocal=not immune) with day 180
See below
```{r VZV IgG with day 180}
# Alluvial
alluvial_B <- moffitt_subset %>% 
  select(pt_identifier, igg_vzv_change,
         vzv_ig_g_apheresis, vzv_ig_g_d90, vzv_ig_g_d180
  ) %>% 
  drop_na() %>% 
  group_by(vzv_ig_g_apheresis, vzv_ig_g_d90, vzv_ig_g_d180) %>% 
  summarise(count=n()) %>%

  ggplot(aes(axis1 = vzv_ig_g_apheresis, axis2= vzv_ig_g_d90, axis3 = vzv_ig_g_d180,
             y = count)) +
  geom_alluvium(aes(fill = vzv_ig_g_apheresis)
  )+
  geom_stratum()+
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  # geom_hline(yintercept = 5, color = "blue")+
  theme_classic()+
  ggtitle("Evolution of virus titers")
```
11 patients don't have data at day 180.

```{r alluvial day 180 combine, fig.height=12}
(alluvial_A) /
  (alluvial_B) + #/
  # (alluvial_C) 
  plot_annotation(tag_levels = "A")
```

## Figure S2. Pneumococcal titers after ide-cel

Antibody concentration >1.0 is considered long-term protection (immune)
If antibody concentration ≤1.0 consider negative (not immune)
```{r, fig.width=12, fig.height=12}
# Histogram
# fig_dat %>% 
#   select(PT, 
#          starts_with("Pneumo"), -ends_with("change")) %>% 
#   pivot_longer(cols = -PT) %>% 
#   filter(!is.na(value)) %>% 
#   mutate(day = str_match(name, ".*(apheresis|D90)")[,2]) %>% 
#   mutate(name = str_remove(name, " apheresis| D90")) %>% 
#   
#   group_by(name, day, value) %>% 
#   summarise(count=n()) %>% 
#   mutate(perc=(count/sum(count)*100)
#   ) %>% 
# 
#   ggplot(aes(x= day, y= perc, fill= value))+
#   # geom_bar(stat = "identity",
#   #          position=position_dodge2(width=0.9, preserve = "single"))+
#   geom_bar(stat = "identity")+
#   labs(x= "Days relative to infusion", y= "Proportion immunity (%)",
#        fill= "Immunity")+
#   scale_y_continuous(expand = c(0, 0))+
#   theme_classic()+
#   facet_wrap(.~ name)
```

```{r, fig.width=12, fig.height=14}
# Tile
# fig_dat %>% 
#   select(PT, 
#          starts_with("Pneumo"), -ends_with("change")) %>% 
#   pivot_longer(cols = -PT) %>% 
#   mutate(day = str_match(name, ".*(apheresis|D90)")[,2]) %>% 
#   mutate(name = str_remove(name, " apheresis| D90")) %>% 
#   
#   
#   filter(!is.na(value)) %>% 
#   arrange(value) %>% 
#   ggplot(aes(x= day, y= fct_reorder2(PT, value, day), fill= value))+
#   geom_tile(color = "white")+
#   labs(fill = NULL, x = NULL, y= NULL)+
#   theme_classic()+
#   facet_wrap(.~ name)
```

```{r Figure S2. Pneumococcal titers after ide-cel, fig.width=15, fig.height=12}
library(ComplexHeatmap)
mat <- moffitt_subset %>% 
  select(pt_identifier, 
         intersect(starts_with("pneumo"), ends_with("change"))) %>% 
  drop_na() %>% 
  mutate(new_sort =
           c(4,3,12,31,13,14,9,15,5,6,8,10,32,16,18,17,19,30,2,1,20,21,40,22),
         .before = 2) %>%

  arrange(new_sort) %>%
    select(-new_sort)

df5 <- t((as.matrix(mat)))

df <- df5 %>% `colnames<-`(c(.[1,])) %>% 
  .[-1,]

df <- df %>% 
sub(pattern = " 1.00", replacement = "Acquired immunity") %>% 
sub(pattern = " 0.25", replacement = "Positive overtime") %>% 
sub(pattern = "-0.25", replacement = "Negative overtime") %>% 
sub(pattern = "-1.00", replacement = "Lost immunity")

colors = structure(c("blue", "darkgrey", "lightgrey", "red"), 
                   names = c("Lost immunity", "Negative overtime", "Positive overtime", "Acquired immunity"))

ht_list <- Heatmap(df, name = "change", 
        rect_gp = gpar(col = "white", lwd = 0.1),
        column_title = "Change in titers",
        # show_column_names =FALSE,
        row_names_side = "left",
        row_names_gp = gpar(fontsize = 11),
        col = colors)
draw(ht_list, heatmap_legend_side = "bottom")
```

## MVA and forest plot
Limit to evaluable patients at day 90
```{r MVA and forest plot}
mva_dat <- fig_dat %>% 
  filter(os_from_infusion_months > 3) %>% 
  mutate(gr3_cytopenia_30d = case_when(
    gr3_cytopenia_30d == "No"    ~ 0,
    gr3_cytopenia_30d == "Yes"   ~ 1
  )) %>% 
  mutate(gr3_cytopenia_90d = case_when(
    gr3_cytopenia_90d == "No"    ~ 0,
    gr3_cytopenia_90d == "Yes"   ~ 1
  )) %>% 
  mutate(CRS_grade2 = case_when(
    max_crs_grade_0_5 < 2                              ~ "No CRS or Grade < 2",
    max_crs_grade_0_5 >= 2                             ~ "Grade ≥2"),
    CRS_grade2 = factor(CRS_grade2, levels = c("No CRS or Grade < 2", "Grade ≥2"))
  )

# glm(gr3_cytopenia_30d ~ high_marrow_burden_50_percent + `gr3_cytopenia_-5d` + CRS_grade2 + steroid_use_yes_no + number_of_prior_lines_of_therapy,
#     data = mva_dat,
#     family = "binomial") %>%
#   tbl_regression(exponentiate = TRUE, 
#                  intercept = TRUE) %>%
#   add_nevent(location = "level") %>% add_n(location = "level") %>%
#   modify_spanning_header(everything() ~ "**Grade 3+ Cytopenia at days 30**") %>%
#   bold_p(t = .05) 
# 
# glm(gr3_cytopenia_90d ~ high_marrow_burden_50_percent + `gr3_cytopenia_-5d` + CRS_grade2 + steroid_use_yes_no + number_of_prior_lines_of_therapy,
#     data = mva_dat,
#     family = "binomial") %>%
#   tbl_regression(exponentiate = TRUE, 
#                  intercept = TRUE) %>%
#   add_nevent(location = "level") %>% add_n(location = "level") %>%
#   modify_spanning_header(everything() ~ "**Grade 3+ Cytopenia at days 90**") %>%
#   bold_p(t = .05) 
# 
# glm(gr3_cytopenia_30d ~ high_marrow_burden_50_percent + CRS_grade2 + steroid_use_yes_no + number_of_prior_lines_of_therapy,
#     data = mva_dat,
#     family = "binomial") %>%
#   tbl_regression(exponentiate = TRUE, 
#                  intercept = TRUE) %>%
#   add_nevent(location = "level") %>% add_n(location = "level") %>%
#   modify_spanning_header(everything() ~ "**Grade 3+ Cytopenia at days 30**") %>%
#   bold_p(t = .05)

glm(gr3_cytopenia_90d ~ high_marrow_burden_50_percent + `gr3_cytopenia_-5d` + 
      steroid_use_yes_no + number_of_prior_lines_of_therapy + baseline_ferritin_cat,
    data = mva_dat,
    family = "binomial") %>%
  tbl_regression(exponentiate = TRUE, 
                 intercept = TRUE) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  modify_spanning_header(everything() ~ "**Grade 3+ Cytopenia at days 90**") %>%
  bold_p(t = .05) 
```

```{r forest plot}
forest_p <- 
  glm(gr3_cytopenia_90d ~ high_marrow_burden_50_percent + `gr3_cytopenia_-5d` +
        steroid_use_yes_no + number_of_prior_lines_of_therapy + baseline_ferritin_cat,
    data = mva_dat,
    family = "binomial") %>%
  tbl_regression(exponentiate = TRUE) %>%
  as_tibble() %>% 
  `colnames<-`(str_remove_all(colnames(.), "\\*")) %>% 
  fill(OR, `95% CI`, .direction = "up") %>% 
  separate(`95% CI`, into = c("lower", "upper"), sep = ", ") %>% 
  filter(Characteristic %in% c("High marrow burden (≥50%)",
                 "gr3_cytopenia_-5d",
                 "baseline_ferritin_cat",
                 "Steroid Use",
                 "Number of prior lines of therapy") ) %>% 
  mutate_at(c("OR", "lower", "upper"), ~ as.numeric(.)) %>% 
  mutate(name= case_when(
    Characteristic == "High marrow burden (≥50%)"      ~ "High marrow burden (≥50%)",
    Characteristic == "gr3_cytopenia_-5d"      ~ "Grade ≥ 3 Cytopenia at Day -5",
    Characteristic == "Steroid Use"      ~ "Steroid use",
    Characteristic == "Number of prior lines of therapy"      ~ "Number of prior lines of therapy",
    Characteristic == "baseline_ferritin_cat"      ~ "Baseline ferritin"
  ))

ggplot(data=forest_p, aes(y=name, x=OR, xmin=lower, xmax=upper)) +
  geom_point() + 
  geom_errorbarh(height=.1) +
  geom_vline(xintercept=1, color='black', linetype='dashed', alpha=.5) +
  theme_classic()+
  scale_x_log10(breaks = c(0, 0.25, 0.5, 1, 2.5, 5, 10, 40))+
  labs(y="")
```

## Survival analysis - KM by gr3_cytopenia_90d
```{r Survival analysis - KM by gr3_cytopenia_90d}
surv_dat <- fig_dat %>% 
  mutate(ORR_as_best_response = case_when(
    ORR_as_best_response == "ORR"         ~ "ORR",
    is.na(ORR_as_best_response)           ~ "never ORR"
  ))

library(survival)
ggsurvplot(survfit(Surv(pfs_from_infusion_months, pfs_event) ~ gr3_cytopenia_90d, 
                   data=surv_dat),
           title = "PFS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time in months",
           legend = "top",
           legend.title = "",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(ncol = 1))

coxph(Surv(time = surv_dat$pfs_from_infusion_months, 
                   event = surv_dat$pfs_event) ~ gr3_cytopenia_90d, 
              data = surv_dat)  %>% 
  tbl_regression(exponentiate = TRUE) %>% 
  bold_p(t = .05) %>% 
  add_nevent(location = "level") %>% add_n(location = "level")

ggsurvplot(survfit(Surv(os_from_infusion_months, os_event) ~ gr3_cytopenia_90d, 
                   data=surv_dat),
           title = "OS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time in months",
           legend = "top",
           legend.title = "",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(ncol = 1))

coxph(Surv(time = surv_dat$os_from_infusion_months, 
                   event = surv_dat$os_event) ~ gr3_cytopenia_90d, 
              data = surv_dat)  %>% 
  tbl_regression(exponentiate = TRUE) %>% 
  bold_p(t = .05) %>% 
  add_nevent(location = "level") %>% add_n(location = "level")
```

## Survival analysis - KM by ORR_as_best_response
```{r Survival analysis - KM by ORR_as_best_response}
ggsurvplot(survfit(Surv(pfs_from_infusion_months, pfs_event) ~ ORR_as_best_response, 
                   data=surv_dat),
           title = "PFS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time in months",
           legend = "top",
           legend.title = "",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(ncol = 1))

coxph(Surv(time = surv_dat$pfs_from_infusion_months, 
                   event = surv_dat$pfs_event) ~ ORR_as_best_response, 
              data = surv_dat)  %>% 
  tbl_regression(exponentiate = TRUE) %>% 
  bold_p(t = .05) %>% 
  add_nevent(location = "level") %>% add_n(location = "level")

ggsurvplot(survfit(Surv(os_from_infusion_months, os_event) ~ ORR_as_best_response, 
                   data=surv_dat),
           title = "OS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time in months",
           legend = "top",
           legend.title = "",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) + guides(colour = guide_legend(ncol = 1))

coxph(Surv(time = surv_dat$os_from_infusion_months, 
                   event = surv_dat$os_event) ~ ORR_as_best_response, 
              data = surv_dat)  %>% 
  tbl_regression(exponentiate = TRUE) %>% 
  bold_p(t = .05) %>% 
  add_nevent(location = "level") %>% add_n(location = "level")
```

```{r glm}
mva_dat <- surv_dat %>% 
  filter(os_from_infusion_months > 3) %>% 
  mutate(ORR_as_best_response = case_when(
    ORR_as_best_response == "ORR"         ~ 1,
    ORR_as_best_response == "never ORR"   ~ 0
  ))

glm(ORR_as_best_response ~ gr3_cytopenia_90d,
    data = mva_dat,
    family = "binomial") %>%
  tbl_regression(exponentiate = TRUE, 
                 intercept = TRUE) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  modify_spanning_header(everything() ~ "**ORR_as_best_response**") %>%
  bold_p(t = .05) 
```