### 2025-06-12 MM_teclistamab/Rmd/teclistamab_analyses.Rmd
First analysis

---
title: "Teclistamab analyses"
author: "Christelle Colin-Leitzinger"
date: '`r Sys.Date()`'
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: united
    highlight: pygments
    df_print: paged
editor_options:
  chunk_output_type: console
---

<style type= "text/css">

. figure {
   margin-top: 25px;
   margin-bottom: 100px;
}

table {
    margin-top: 10px;
    margin-bottom: 25px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      fig.align='center'
                      )
```

```{r library}
library(here)
library(tidyverse)
library(ggplot2)
library(gtsummary)
library(labelled)
library(DescTools)
library(patchwork)
library(ggbreak)
library(survival)
library(survminer)
theme_set(theme_classic())
theme_gtsummary_compact()
```

```{r load}
# path <- fs::path("","Volumes","Peres_Research", "Myeloma", "Analyses", "Janssen Teclistamab")

data <- 
  read_csv(locale = locale(encoding="latin1"),
    paste0(#path, 
           here(), 
           "/Tec_05222025.csv")#,
    # na = "N/A"
    ) %>% 
  janitor::clean_names()
```

```{r column name}
data <- data %>% 
  `colnames<-`(paste0(colnames(.), "~", .[1,])) %>% 
  `colnames<-`(str_remove(colnames(.), "~NA")) %>% 
  janitor::clean_names() %>% 
  rename(pt_identifier = patient_identifier_use_same_id_for_cns_amyloid_sheets_if_appropriate,
         if_ineligible_for_bs_ab_trial_why_not_be_specific_and_can_separate_reasons_with_commas = 
           if_ineligible_for_bs_ab_trial_why_not_be_specific_and_can_separate_reasons_with_commas_no_0_y_1, 
         date_of_infection_1st = date_of_infection_74,
         infection_severity_1st_severe_hospitalization_or_requiring_iv_antibiotics = infection_severity_73_severe_hospitalization_or_requiring_iv_antibiotics,
         date_of_infection_2nd = date_of_infection_77,
         infection_severity_2nd_severe_hospitalization_or_requiring_iv_antibiotics = infection_severity_76_severe_hospitalization_or_requiring_iv_antibiotics,
         date_of_infection_3rd = date_of_infection_80,
         infection_severity_3rd_severe_hospitalization_or_requiring_iv_antibiotics = infection_severity_79_severe_hospitalization_or_requiring_iv_antibiotics,
         date_of_infection_4th = date_of_infection_83,
         infection_severity_4th_severe_hospitalization_or_requiring_iv_antibiotics = infection_severity_82_severe_hospitalization_or_requiring_iv_antibiotics
         ) %>% 
  slice(-1)

data1 <- type.convert(data, as.is = TRUE)
```

# Data

What we did:

- The start date to calculate survival and days of onset is date_of_bs_ab_step_up_d1.   
- There was text into the variable date_of_first_bs_ab_full_dose. The text was transferred into reason_no_date_of_first_bs_ab_full_dose.  
  Patient with text have progressed.   
- max_crp has a date, now it is NA.   
- date_of_max_crs has 0/1.   
- Difference between prior_cart_0_no_1_yes and Prior CART (yes/no)? The later has 0/1 values so I am using prior_cart_0_no_1_yes.   
- baseline_crp_prior_to_bs_ab_initiation_mg_l has a "4.4." value. It was replaced by "4.4".   
- ALC Day 15 has a value of “-85“ and was replaced by 85.   
- in_out_patient was coded "In" for patient with hospital length stay values and "Out" when no values.

Issues in data with no impact for now:

- Wrong date of Discontinuation "1/0/1900" or should be NA as the value in Discontinuation (yes/no) is "No". If the date is wrong, then Discontinuation (yes/no) should be "Yes".    
- Wrong date of MM diagnosis "10/1/20218".   
- For treatment_delays_during_first_cycle_infection_1_cytoepnias_2_crs_3_icans_4_other_5_explain_leave_blank_if_no_delays,
should I combine "CRS and Infection" and "Infection and CRS"? others? or is there an time order to consider?   
- date_of_max_ferritin has a date "20-Sep".   
- date of max crp.   
- Prior BsAb  (yes/no) has a “0” value.   
- difference between prior_bcma_directed_therapy_no_0_y_1 and prior_bcma_directed_therapy_yes_no. Looks none others that the yes/no is not recoded well so I am using 0/1 that I recoded myself.   
- #Infection_1 (Bacterial, Viral fungal) has numbers instead of “Bacterial”, “Viral” etc .

```{r temp_fix}
temp_fix <- data1 %>% 
  mutate(baseline_crp_prior_to_bs_ab_initiation_mg_l = case_when(
    baseline_crp_prior_to_bs_ab_initiation_mg_l == "4.4."        ~ 4.4,
    TRUE                  ~ as.numeric(baseline_crp_prior_to_bs_ab_initiation_mg_l)
  )) %>% 
  mutate(alc_day_15 = case_when(
    alc_day_15 == "85-"   ~ 85,
    TRUE                  ~ as.numeric(alc_day_15)
  )) %>% 
  mutate_at(c(   ############################### Remove when fix temp un necessary
              "max_crp"
  ), ~ as.numeric(.)
  )
```

```{r data cleaning}
data1 <- temp_fix %>% 
  # General----
  # Fix letter case and missing
  # mutate(across(where(is.character), ~ str_replace(., "NE|ND|N/A|Not done", NA_character_))) %>%
  mutate(across(c(where(is.character),
                  -c("pt_identifier",
                     "myeloma_involved_heavy_chain_ig_g_ig_a_ig_m_ig_d_ig_e_or_blank",
                     "best_documented_emd_response"
                     )),
                ~ str_to_sentence(.) )) %>%
  mutate(across(c(contains("vgpr"), #"response_to_bridging_chemotherapy"
                  ),
                ~ str_to_upper(.) )) %>%
  # Recode 0/1
  mutate_at(c("bs_ab_served_only_as_pre_car_t_bridging_0_no_1_yes",
              "pfs_with_1st_line_mm_diagnosis_18_months_0_no_1_yes",
              "oligo_or_non_secretory_at_bs_ab_initiation_0_no_1_yes", 
              "deletion_17p_prior_to_bs_ab_0_no_1_yes",
              "t_4_14_prior_to_bs_ab_0_no_1_yes",
              "t_14_16_prior_to_bs_ab_0_no_1_yes",
              "gain_amp1q21_prior_to_bs_ab_0_no_1_yes",
              "amp_1_q_only_0_no_1_yes",
              "deletion_1p_0_no_1_yes",
              "prior_auto_sct_no_0_yes_1",
              "prior_cart_0_no_1_yes",
              "amyloidosis_if_yes_fill_out_amyloid_sheet_as_well_please_once_youre_done_with_this_one_0_no_1_yes",
              "active_cns_mm_if_yes_fill_out_cns_sheet_as_well_once_done_with_this_one_0_no_1_yes",
              "active_pcl_5_percent_cp_cs_within_30_days_prior_or_after_bs_ab_initiation_0_no_1_yes",
              "prior_bcma_directed_therapy_no_0_y_1",
              "triple_class_refractory_i_mi_d_pi_anti_cd38_no_0_y_1",
              "penta_refractory_2_p_is_2_i_mi_ds_anti_cd38_no_0_y_1",
              "eligible_for_corresponding_bs_ab_trial_on_first_dose_use_trial_elig_sheet_to_help_you_no_0_y_1",
              "high_marrow_burden_50_percent_p_cs_if_checked_within_30_days_before_bs_ab_initiation_n_0_y_1",
              "prophylactic_tocilizumab_n_0_y_1",
              "recurrent_crs_n_0_y_1",
              "tocilizumab_as_tx_n_0_y_1",
              "steroid_use_as_tx_n_0_y_1",
              "anakinra_n_0_y_1",
              "infection_n_0_y_1",
              "icu_admission_during_first_cycle_n_0_y_1",
              "need_for_gcsf", "need_for_tpo_agonist",
              "ivig_use_yes_1_no_0", 
              "hemodialysis_pre_bs_ab_y_1_n_0",
              "hemodialysis_post_bs_ab_y_1_n_0",
              "modification_in_schedule_to_monthly_0_no_1_yes"
  ), 
  ~ case_when(
    . == 1                                              ~ "Yes",
    . == 0                                              ~ "No",
    TRUE                                                ~ as.character(.)
  )) %>% 
  # Recode 1/2
  mutate_at(c("any_invasive_second_primary_malignancy_exclude_non_melanoma_skin_cancer_dcis_etc_1_yes_2_no"
  ), 
  ~ case_when(
    . == 1                                              ~ "Yes",
    . == 2                                              ~ "No",
    TRUE                                                ~ NA_character_
  )) %>% 
  # Exposed, refractory
  mutate_at(c("bortezomib_status_exposed_0_vs_refractory_1_vs_naive_2",
              "carfilzomib_status_exposed_0_vs_refractory_1_vs_naive_2",
              "lenalidomide_status_exposed_0_vs_refractory_1_vs_naive_2",
              "pomalidomide_status_exposed_0_vs_refractory_1_vs_naive_2",
              "daratumumab_or_isatuximab_status_exposed_0_vs_refractory_1_vs_naive_2"
  ), 
  ~ case_when(
    . == 0                                              ~ "Exposed",
    . == 1                                              ~ "Refractory",
    . == 2                                              ~ "Naive",
    TRUE                                                ~ NA_character_
  )) %>% 
  # mutate_at(c("baseline_crp_prior_to_bs_ab_initiation_mg_l"#,   ############################### Remove when fix temp un necessary
  #             # "max_crp"
  # ), ~ as.numeric(.)
  # ) %>%
  mutate_at(c("ecog_ps_at_bs_ab_initiation_0_5"#,
              # "max_crs_grade_0_5",
              # "max_icans_grade_grade"
  ), ~ factor(.)
  ) %>%
  mutate(reason_no_date_of_first_bs_ab_full_dose = case_when(
    date_of_first_bs_ab_full_dose %in% 
      c("Cmo before receiving full dose",
        "Withdrew care after d5")              ~ date_of_first_bs_ab_full_dose
  ), .after = date_of_first_bs_ab_full_dose) %>% 
  # mutate(was_discontinued = case_when(
  #   date_of_disocntinuation_if_cotinues_enter_0 == 0     ~ "Continued",
  #   !is.na(date_of_disocntinuation_if_cotinues_enter_0)  ~ "Discontinued"
  # ), .after = ) %>%
  mutate(across(c("date_of_mm_diagnosis",
                  "date_of_bs_ab_step_up_d1",
                  "date_of_first_bs_ab_full_dose",
                  "date_of_max_ferritin",
                  "date_of_max_crp",
                  "date_of_onset_of_crs",
                  "date_of_max_crs",
                  "date_of_onset_of_icans",
                  "date_of_max_ican",
                  "date_of_infection_1st",
                  "date_of_infection_2nd",
                  "date_of_infection_3rd",
                  "date_of_infection_4th",
                  "anc_nadir_date", "hgb_nadir_date",
                  "plt_nadir_date", "date_of_ig_g_nadir",
                  "date_of_first_ivig_dose_following_bs_ab_administration",
                  "date_of_best_mrd_response_relative_to_bs_ab_start",
                  "date_of_pd",
                  "date_of_disocntinuation_if_cotinues_enter_0",
                  "date_of_death", "date_of_last_contact",
                  "date_of_first_exposure_to_prior_bcma_agent",
                  "date_of_last_exposure_to_prior_bcma_agent",
                  "date_last_exposure",
                  "while_on_bs_ab_date_of_first_documented_emd_response",
                  "date_of_best_documented_emd_response",
                  "date_of_last_imaging_before_data_cutoff_only_in_pts_who_are_still_responding"),
                ~ as.Date(., 
                          tryFormats = c("%m/%d/%y", "%m/%d/%Y"), 
                          origin = "1899-12-30"))) %>% 
  # Treatment
  mutate(bs_ab_name_tec_1_tal_2_elra_3 = case_when(
    bs_ab_name_tec_1_tal_2_elra_3 == 1        ~ "Teclistamab",
    bs_ab_name_tec_1_tal_2_elra_3 == 2        ~ "tal",
    bs_ab_name_tec_1_tal_2_elra_3 == 3        ~ "elra",
    TRUE                                      ~ as.character(bs_ab_name_tec_1_tal_2_elra_3)
  )) %>% 
  mutate(sex_0_male_1_female = case_when(
    sex_0_male_1_female == 0                  ~ "Male",
    sex_0_male_1_female == 1                  ~ "Female",
    TRUE                                      ~ as.character(sex_0_male_1_female)
  )) %>% 
  mutate(race_white_0_black_1_aapi_2_american_indian_3_other_4_unk_5 = case_when(
    race_white_0_black_1_aapi_2_american_indian_3_other_4_unk_5 == 0        ~ "White",
    race_white_0_black_1_aapi_2_american_indian_3_other_4_unk_5 == 1        ~ "Black",
    race_white_0_black_1_aapi_2_american_indian_3_other_4_unk_5 == 2        ~ "Asian American and Pacific Islander",
    race_white_0_black_1_aapi_2_american_indian_3_other_4_unk_5 == 3        ~ "American_indian",
    race_white_0_black_1_aapi_2_american_indian_3_other_4_unk_5 == 4        ~ "Other",
    race_white_0_black_1_aapi_2_american_indian_3_other_4_unk_5 == 5        ~ "Unknown",
    TRUE                                      ~ as.character(race_white_0_black_1_aapi_2_american_indian_3_other_4_unk_5)
  )) %>% 
  mutate(hispanic_1_if_hispanic_latin_x_or_spanish_origin_0_if_no_unknown = case_when(
    hispanic_1_if_hispanic_latin_x_or_spanish_origin_0_if_no_unknown == 0        ~ "No/Unknown",
    hispanic_1_if_hispanic_latin_x_or_spanish_origin_0_if_no_unknown == 1        ~ "Hispanic, latinX, spanish origin",
    TRUE                                      ~ as.character(hispanic_1_if_hispanic_latin_x_or_spanish_origin_0_if_no_unknown)
  )) %>%

  mutate(myeloma_involved_light_chain_kappa_lambda_biclonal_blank = case_when(
    myeloma_involved_light_chain_kappa_lambda_biclonal_blank == "K"        ~ "Kappa",
    TRUE                                      ~ as.character(myeloma_involved_light_chain_kappa_lambda_biclonal_blank)
  )) %>%

  mutate(pcl_5_percent_cp_cs_at_any_point_from_dx_through_last_pre_bs_ab_lo_t_0_no_1_secondary_pcl_not_at_dx_2_primary_pcl_at_dx = case_when(
    pcl_5_percent_cp_cs_at_any_point_from_dx_through_last_pre_bs_ab_lo_t_0_no_1_secondary_pcl_not_at_dx_2_primary_pcl_at_dx == 0        ~ "No",
    pcl_5_percent_cp_cs_at_any_point_from_dx_through_last_pre_bs_ab_lo_t_0_no_1_secondary_pcl_not_at_dx_2_primary_pcl_at_dx == 1        ~ "Secondary PCL (not at Dx)",
    pcl_5_percent_cp_cs_at_any_point_from_dx_through_last_pre_bs_ab_lo_t_0_no_1_secondary_pcl_not_at_dx_2_primary_pcl_at_dx == 2        ~ "Primary PCL (at Dx)",
    TRUE                                      ~ as.character(pcl_5_percent_cp_cs_at_any_point_from_dx_through_last_pre_bs_ab_lo_t_0_no_1_secondary_pcl_not_at_dx_2_primary_pcl_at_dx)
  )) %>%

  mutate(number_infection_1_bacterial_viral_fungal) %>%

  mutate(infection_severity_1st_severe_hospitalization_or_requiring_iv_antibiotics = case_when(
    infection_severity_1st_severe_hospitalization_or_requiring_iv_antibiotics == "N/a"        ~ NA_character_,
    TRUE       ~ infection_severity_1st_severe_hospitalization_or_requiring_iv_antibiotics
  )) %>%
  mutate(in_out_patient = case_when(
    is.na(if_inpatient_length_of_hospital_stay_total_days_including_all_admissions_0_if_done_outpt_with_no_admissions)
    ~ "Out",
    !is.na(if_inpatient_length_of_hospital_stay_total_days_including_all_admissions_0_if_done_outpt_with_no_admissions)
    ~ "In"
  ), .after = if_inpatient_length_of_hospital_stay_total_days_including_all_admissions_0_if_done_outpt_with_no_admissions) %>% 
  mutate(days_of_icu_admission = case_when(
    days_of_icu_admission == 0             ~ NA_real_,
    TRUE                                   ~ days_of_icu_admission
  )) %>%

  mutate(treatment_delays_during_first_cycle_infection_1_cytoepnias_2_crs_3_icans_4_other_5_explain_leave_blank_if_no_delays = str_replace(
    treatment_delays_during_first_cycle_infection_1_cytoepnias_2_crs_3_icans_4_other_5_explain_leave_blank_if_no_delays, 
    "1", "Infection")) %>% 
  mutate(treatment_delays_during_first_cycle_infection_1_cytoepnias_2_crs_3_icans_4_other_5_explain_leave_blank_if_no_delays = str_replace(
    treatment_delays_during_first_cycle_infection_1_cytoepnias_2_crs_3_icans_4_other_5_explain_leave_blank_if_no_delays, 
    "2", "Cytopenias")) %>% 
  mutate(treatment_delays_during_first_cycle_infection_1_cytoepnias_2_crs_3_icans_4_other_5_explain_leave_blank_if_no_delays = str_replace(
    treatment_delays_during_first_cycle_infection_1_cytoepnias_2_crs_3_icans_4_other_5_explain_leave_blank_if_no_delays, 
    "3", "CRS")) %>% 
  mutate(treatment_delays_during_first_cycle_infection_1_cytoepnias_2_crs_3_icans_4_other_5_explain_leave_blank_if_no_delays = str_replace(
    treatment_delays_during_first_cycle_infection_1_cytoepnias_2_crs_3_icans_4_other_5_explain_leave_blank_if_no_delays, 
    "4", "ICANS")) %>% 
  mutate(treatment_delays_during_first_cycle_infection_1_cytoepnias_2_crs_3_icans_4_other_5_explain_leave_blank_if_no_delays = str_replace(
    treatment_delays_during_first_cycle_infection_1_cytoepnias_2_crs_3_icans_4_other_5_explain_leave_blank_if_no_delays, 
    "5", "Other")) %>% 
  # mutate(treatment_delays_during_first_cycle_infection_1_cytoepnias_2_crs_3_icans_4_other_5_explain_leave_blank_if_no_delays = case_when(
    # treatment_delays_during_first_cycle_infection_1_cytoepnias_2_crs_3_icans_4_other_5_explain_leave_blank_if_no_delays == 1        ~ "Infection",
    # treatment_delays_during_first_cycle_infection_1_cytoepnias_2_crs_3_icans_4_other_5_explain_leave_blank_if_no_delays == 2        ~ "Cytopenias",
    # treatment_delays_during_first_cycle_infection_1_cytoepnias_2_crs_3_icans_4_other_5_explain_leave_blank_if_no_delays == 3        ~ "CRS",
    # treatment_delays_during_first_cycle_infection_1_cytoepnias_2_crs_3_icans_4_other_5_explain_leave_blank_if_no_delays == 4        ~ "ICANS",
    # treatment_delays_during_first_cycle_infection_1_cytoepnias_2_crs_3_icans_4_other_5_explain_leave_blank_if_no_delays == 5        ~ "Other",
  #   TRUE                                      ~ treatment_delays_during_first_cycle_infection_1_cytoepnias_2_crs_3_icans_4_other_5_explain_leave_blank_if_no_delays
  # )) %>%

  mutate(reason_for_discontinuation_pd_1_toxicity_2_ongoing_response_3_death_4_other_5_explain = case_when(
    reason_for_discontinuation_pd_1_toxicity_2_ongoing_response_3_death_4_other_5_explain == 1        ~ "PD",
    reason_for_discontinuation_pd_1_toxicity_2_ongoing_response_3_death_4_other_5_explain == 2        ~ "Toxicity",
    reason_for_discontinuation_pd_1_toxicity_2_ongoing_response_3_death_4_other_5_explain == 3        ~ "Ongoing response",
    reason_for_discontinuation_pd_1_toxicity_2_ongoing_response_3_death_4_other_5_explain == 4        ~ "Death",
    reason_for_discontinuation_pd_1_toxicity_2_ongoing_response_3_death_4_other_5_explain == 5        ~ "Other",
    TRUE                                      ~ reason_for_discontinuation_pd_1_toxicity_2_ongoing_response_3_death_4_other_5_explain
  )) %>%

  mutate(cause_of_death_pd_1_infection_while_of_bs_ab_2_crs_3_icans_4_other_5_explain = case_when(
    cause_of_death_pd_1_infection_while_of_bs_ab_2_crs_3_icans_4_other_5_explain == 1        ~ "PD",
    cause_of_death_pd_1_infection_while_of_bs_ab_2_crs_3_icans_4_other_5_explain == 2        ~ "infection_while_of_bs_ab",
    cause_of_death_pd_1_infection_while_of_bs_ab_2_crs_3_icans_4_other_5_explain == 3        ~ "CRS",
    cause_of_death_pd_1_infection_while_of_bs_ab_2_crs_3_icans_4_other_5_explain == 4        ~ "ICANS",
    cause_of_death_pd_1_infection_while_of_bs_ab_2_crs_3_icans_4_other_5_explain == 5        ~ "Others",
    TRUE                                      ~ as.character(cause_of_death_pd_1_infection_while_of_bs_ab_2_crs_3_icans_4_other_5_explain)
  )) %>%

  mutate(type_of_most_recent_bcma_agent_adc_1_cart_2_bs_ab_3 = case_when(
    type_of_most_recent_bcma_agent_adc_1_cart_2_bs_ab_3 == 1        ~ "ADC",
    type_of_most_recent_bcma_agent_adc_1_cart_2_bs_ab_3 == 2        ~ "CART",
    type_of_most_recent_bcma_agent_adc_1_cart_2_bs_ab_3 == 3        ~ "BsAb",
    TRUE                                      ~ as.character(type_of_most_recent_bcma_agent_adc_1_cart_2_bs_ab_3)
  )) %>%

  mutate(response_to_most_recent_bmca_agent_cr_1_vgpr_2_pr_3_sd_4_pd_5_ne_6 = case_when(
    response_to_most_recent_bmca_agent_cr_1_vgpr_2_pr_3_sd_4_pd_5_ne_6 == 1        ~ "CR",
    response_to_most_recent_bmca_agent_cr_1_vgpr_2_pr_3_sd_4_pd_5_ne_6 == 2        ~ "VGPR",
    response_to_most_recent_bmca_agent_cr_1_vgpr_2_pr_3_sd_4_pd_5_ne_6 == 3        ~ "PR",
    response_to_most_recent_bmca_agent_cr_1_vgpr_2_pr_3_sd_4_pd_5_ne_6 == 4        ~ "SD",
    response_to_most_recent_bmca_agent_cr_1_vgpr_2_pr_3_sd_4_pd_5_ne_6 == 5        ~ "PD",
    response_to_most_recent_bmca_agent_cr_1_vgpr_2_pr_3_sd_4_pd_5_ne_6 == 6        ~ "NE",
    TRUE                                      ~ as.character(response_to_most_recent_bmca_agent_cr_1_vgpr_2_pr_3_sd_4_pd_5_ne_6)
  )) %>%

  mutate(just_prior_to_bs_ab_initiation_is_there_true_emd_if_none_move_to_column_gk_0_none_1_paraskeletal_only_2_true_emd_paraskeletal = case_when(
    just_prior_to_bs_ab_initiation_is_there_true_emd_if_none_move_to_column_gk_0_none_1_paraskeletal_only_2_true_emd_paraskeletal == 0        ~ "None",
    just_prior_to_bs_ab_initiation_is_there_true_emd_if_none_move_to_column_gk_0_none_1_paraskeletal_only_2_true_emd_paraskeletal == 1        ~ "Paraskeletal only",
    just_prior_to_bs_ab_initiation_is_there_true_emd_if_none_move_to_column_gk_0_none_1_paraskeletal_only_2_true_emd_paraskeletal == 2        ~ "True EMD (+/-paraskeletal)",
    TRUE                                      ~ as.character(just_prior_to_bs_ab_initiation_is_there_true_emd_if_none_move_to_column_gk_0_none_1_paraskeletal_only_2_true_emd_paraskeletal)
  )) %>%

  mutate(total_number_true_emd_plasmacytomas_if_1_2_sites_write_1_or_2_if_2_plasmacytomas_just_write_3 = case_when(
    total_number_true_emd_plasmacytomas_if_1_2_sites_write_1_or_2_if_2_plasmacytomas_just_write_3 == 1        ~ "1 site",
    total_number_true_emd_plasmacytomas_if_1_2_sites_write_1_or_2_if_2_plasmacytomas_just_write_3 == 2        ~ "2 sites",
    total_number_true_emd_plasmacytomas_if_1_2_sites_write_1_or_2_if_2_plasmacytomas_just_write_3 == 3        ~ ">2 sites",
    TRUE                                      ~ as.character(total_number_true_emd_plasmacytomas_if_1_2_sites_write_1_or_2_if_2_plasmacytomas_just_write_3)
  )) %>%

  mutate(xrt_2_weeks_before_and_after_3_months_of_bs_ab_starting_yes_no_response = case_when(
    xrt_2_weeks_before_and_after_3_months_of_bs_ab_starting_yes_no_response == "N"         ~ "No",
    TRUE                                      ~ xrt_2_weeks_before_and_after_3_months_of_bs_ab_starting_yes_no_response
  )) %>%
  mutate(true_emd_relapse_yes_no = case_when(
    true_emd_relapse_yes_no == "Y"         ~ "Yes",
    is.na(true_emd_relapse_yes_no)         ~ NA_character_,
    TRUE                                   ~ as.character(true_emd_relapse_yes_no)
  )) %>%

  mutate(if_yes_what_location_0_same_site_1_new_site_2_both = case_when(
    if_yes_what_location_0_same_site_1_new_site_2_both == 0        ~ "Same site",
    if_yes_what_location_0_same_site_1_new_site_2_both == 1        ~ "New site",
    if_yes_what_location_0_same_site_1_new_site_2_both == 2        ~ "Both",
    TRUE                                      ~ as.character(if_yes_what_location_0_same_site_1_new_site_2_both)
  ))

# Fix the "<" and ">" to take the min or max or each column including numbers next to the signs----
# Also replace "Suboptimal" by NA
vec_col <- c("ig_g_nadir_during_first_6_cycles_mg_d_l",
             "baseline_ig_g_prior_to_bs_ab_initiation_mg_d_l"
             )
class(data1) <- "data.frame"

for (i in vec_col) {

  if (class(data1[, i]) == "character") {

    data1 <- data1 %>% 
      mutate(wo_sign = as.numeric(str_remove(data1[, i], ">|<"))) %>% 
      mutate(!!i := case_when(
        str_detect(data1[, i], ">")                 ~ max(wo_sign, na.rm = TRUE),
        str_detect(data1[, i], "<")                 ~ min(wo_sign, na.rm = TRUE),
        TRUE                                         ~ as.numeric(wo_sign)
      )) %>% 
      select(-c(wo_sign))
  }
}

```

```{r}
data2 <- data1 %>% 
  # Age categories----
  mutate(agecat = case_when(
    age_at_bs_ab_initiation < 70                                         ~ "< 70 years",
    age_at_bs_ab_initiation >= 70                                        ~ "≥ 70 years"
  ),
  agecat = factor(agecat, levels = c("< 70 years", "≥ 70 years")),
  .after = age_at_bs_ab_initiation
  ) %>% 
  # Sex categories----
  mutate(male_sex = case_when(
    sex_0_male_1_female == "Male"                                    ~ 1,
    sex_0_male_1_female == "Female"                                  ~ 0
  ), .after = sex_0_male_1_female) %>%

  # ECOG
  mutate(ecog_ps_at_bs_ab_initiation_0_5 = case_when(
    ecog_ps_at_bs_ab_initiation_0_5 == 0 |
      ecog_ps_at_bs_ab_initiation_0_5 == 1                                ~ "0-1",
    ecog_ps_at_bs_ab_initiation_0_5 %in% c(2:5)                           ~ "2-5"
  )) %>%
  # Cytogenetics ----
  mutate(cytogenetics = case_when(
    deletion_17p_prior_to_bs_ab_0_no_1_yes == "Yes" | 
      t_4_14_prior_to_bs_ab_0_no_1_yes == "Yes" | 
      t_14_16_prior_to_bs_ab_0_no_1_yes == "Yes"             ~ "Yes",
    deletion_17p_prior_to_bs_ab_0_no_1_yes == "No" &
      t_4_14_prior_to_bs_ab_0_no_1_yes == "No" &
      t_14_16_prior_to_bs_ab_0_no_1_yes == "No"               ~ "No",
    TRUE                                                           ~ NA_character_
  ), .before = deletion_17p_prior_to_bs_ab_0_no_1_yes) %>% 
  # Refractory status ----
  mutate(lenalidomide_status_exposed_0_vs_refractory_1_vs_naive_2 = case_when(
    str_detect(
      lenalidomide_status_exposed_0_vs_refractory_1_vs_naive_2, "Ref")     ~ "Refractory",
    TRUE           ~ lenalidomide_status_exposed_0_vs_refractory_1_vs_naive_2
  )) %>% 
  mutate(immuno = case_when(
    lenalidomide_status_exposed_0_vs_refractory_1_vs_naive_2 == "Refractory" |
      pomalidomide_status_exposed_0_vs_refractory_1_vs_naive_2 == "Refractory"           ~ "Yes",
    TRUE                                                           ~ "No"
  ), .after = pomalidomide_status_exposed_0_vs_refractory_1_vs_naive_2) %>% 
  mutate(proteasome = case_when(
    bortezomib_status_exposed_0_vs_refractory_1_vs_naive_2 == "Refractory" |
      carfilzomib_status_exposed_0_vs_refractory_1_vs_naive_2 == "Refractory"            ~ "Yes",
    TRUE                                                           ~ "No"
  ), .after = carfilzomib_status_exposed_0_vs_refractory_1_vs_naive_2) %>% 
  mutate(dara = case_when(
    daratumumab_or_isatuximab_status_exposed_0_vs_refractory_1_vs_naive_2 == "Refractory"            ~ "Yes",
    TRUE                                                           ~ "No"
  ), .after = daratumumab_or_isatuximab_status_exposed_0_vs_refractory_1_vs_naive_2) %>% 
  mutate(doubleref = case_when(
    proteasome == "Yes" &
      immuno == "Yes"                                              ~ "Yes",
    TRUE                                                           ~ "No"
  ), .after = proteasome) %>% 
  # mutate(triple_class_refractory_i_mi_d_pi_anti_cd38_no_0_y_1 = case_when(
  #   proteasome == "Yes" &
  #     immuno == "Yes" &
  #     daratumumab_or_isatuximab_status_exposed_0_vs_refractory_1_vs_naive_2 == "Refractory"          ~ "Yes",
  #   TRUE                                                           ~ "No"
  # )) %>% 
  # mutate(penta_refractory_2_p_is_2_i_mi_ds_anti_cd38_no_0_y_1 = case_when(
  #   bortezomib_status_exposed_0_vs_refractory_1_vs_naive_2 == "Refractory" & 
  #     carfilzomib_status_exposed_0_vs_refractory_1_vs_naive_2 == "Refractory" &
  #     lenalidomide_status_exposed_0_vs_refractory_1_vs_naive_2 == "Refractory" & 
  #     pomalidomide_status_exposed_0_vs_refractory_1_vs_naive_2 == "Refractory" & 
  #     daratumumab_or_isatuximab_status_exposed_0_vs_refractory_1_vs_naive_2 == "Refractory"          ~ "Yes",
  #   TRUE                                                           ~ "No"
  # )) %>% 
  mutate(number_of_prior_lines_of_therapy_before_bs_ab_cat = case_when(
    number_of_prior_lines_of_therapy_before_bs_ab <= 4                        ~ "≤ 4",
    number_of_prior_lines_of_therapy_before_bs_ab > 4                         ~ "> 4"
  ))
```

```{r toxicities}
cart2 <- data2 %>% 
# CRS ----
  mutate(CRS_any = ifelse(max_crs_grade_0_5 == 0, "No", "Yes"),
       crs_any_coded_0_1 = ifelse(CRS_any == "No", 0, 1),
       CRS_gradecat = case_when(
         max_crs_grade_0_5 == 0                             ~ "No CRS",
         max_crs_grade_0_5 == 1                             ~ "Grade 1 or 2",
         max_crs_grade_0_5 == 2                             ~ "Grade 1 or 2",
         max_crs_grade_0_5 >= 3                             ~ "Grade ≥3"),
       CRS_gradecat = factor(CRS_gradecat, levels = c("No CRS", "Grade 1 or 2", "Grade ≥3")),
       gr3_CRS = case_when(
         max_crs_grade_0_5 %in% c(0:2)                      ~ "Grade <3",
         max_crs_grade_0_5 %in% c(3:5)                      ~ "Grade ≥3",
         TRUE                                                   ~ NA_character_
       ), .after = max_crs_grade_0_5) %>% 
  mutate(gr2_CRS = case_when(
    max_crs_grade_0_5 %in% c(0:1)                      ~ "Grade <2",
    max_crs_grade_0_5 %in% c(2:5)                      ~ "Grade ≥2",
    TRUE                                               ~ NA_character_
  ), .after = CRS_any) %>%
  mutate(day_of_onset_crs_relative_to_initiation =
           interval(start = date_of_bs_ab_step_up_d1, end = date_of_onset_of_crs) /
           duration(n =1, unit = "days")
  ) %>%
  mutate(day_of_max_crs_relative_to_initiation =
           interval(start = date_of_bs_ab_step_up_d1, end = date_of_max_crs) /
           duration(n =1, unit = "days")
  ) %>%
  mutate(day_of_onset_icans_relative_to_initiation =
           interval(start = date_of_bs_ab_step_up_d1, end = date_of_onset_of_icans) /
           duration(n =1, unit = "days")
  ) %>%
  mutate(day_of_max_icans_grade_relative_to_initiation =
           interval(start = date_of_bs_ab_step_up_d1, end = date_of_max_ican) /
           duration(n =1, unit = "days")
  ) %>%
  # censored days > 100 to 100
  # mutate(across(c(day_of_onset_crs_relative_to_initiation,
  #                 day_of_max_crs_relative_to_initiation,
  #                 day_of_onset_icans_relative_to_initiation,
  #                 day_of_max_icans_grade_relative_to_initiation), ~ case_when(
  #   . > 100                                                       ~ 100,
  #   TRUE                                                          ~ as.numeric(.)
  # ))) %>%

  # ICANS ----
  mutate(ICANS_any = ifelse(max_icans_grade == 0, "No", "Yes"),
       ICANS_any2 = ifelse(ICANS_any == "No", 0, 1),
       ICANS_gradecat = case_when(
         max_icans_grade == 0                                ~ "No ICANS",
         max_icans_grade == 1                                ~ "Grade 1 or 2",
         max_icans_grade == 2                                ~ "Grade 1 or 2",
         max_icans_grade >= 3                                ~ "Grade ≥3"),
       ICANS_gradecat = factor(ICANS_gradecat, levels = c("No ICANS", "Grade 1 or 2", "Grade ≥3")),
       gr3_ICANS = case_when(
         max_icans_grade %in% c(0:2)                         ~ "Grade <3",
         max_icans_grade %in% c(3:5)                         ~ "Grade ≥3",
         TRUE                                          ~ NA_character_
       ), .after = max_icans_grade) %>% 
  mutate(gr2_ICANS = case_when(
    max_icans_grade %in% c(0:1)                              ~ "Grade <2",
    max_icans_grade %in% c(2:5)                              ~ "Grade ≥2",
    TRUE                                               ~ NA_character_
  ), .after = ICANS_any2)
  # BCMA ----

```

```{r survival}
cart3 <- cart2 %>% 
  # pfs
  mutate(pfs_event = case_when(
    !is.na(date_of_pd) | !is.na(date_of_death)         ~ 1,
    TRUE                                               ~ 0
  ), .after = death_yes_no) %>% 
  mutate(pfs_date = coalesce(date_of_pd, date_of_death, date_of_last_contact), 
         .after = death_yes_no) %>% 
  mutate(pfs_from_initiation_months = 
           interval(
             start = date_of_bs_ab_step_up_d1, 
             end = pfs_date)/
           duration(n=1, units = "months"), 
         .after = death_yes_no) %>% 
  # os 
  mutate(os_event = case_when(
    !is.na(date_of_death)                              ~ 1,
    TRUE                                               ~ 0
  ), 
  .after = death_yes_no) %>% 
  mutate(os_date = coalesce(date_of_death, date_of_last_contact), 
         .after = death_yes_no) %>% 
  mutate(os_from_initiation_months = 
           interval(
             start = date_of_bs_ab_step_up_d1, 
             end = os_date)/
           duration(n=1, units = "months"), 
         .after = death_yes_no)
```

```{r response}
cart4 <- cart3 %>%   
  # mutate(day_30_response_s_cr_cr_vgpr_pr_mr_sd_pd = case_when(
  #   str_detect(day_30_response_s_cr_cr_vgpr_pr_mr_sd_pd, "evaluable")   ~ NA_character_,
  #   TRUE           ~ day_30_response_s_cr_cr_vgpr_pr_mr_sd_pd
  # )) %>% 
  mutate(day30_response = case_when(
    (os_event == 1 | pfs_event == 1) & 
      pfs_from_initiation_months < 1                                   ~ "Died or progressed before day 30",
    str_detect(day_30_response_s_cr_cr_vgpr_pr_mr_sd_pd, "EVAL")        ~ NA_character_,
    str_detect(day_30_response_s_cr_cr_vgpr_pr_mr_sd_pd, "VGPR")        ~ "VGPR",
    str_detect(day_30_response_s_cr_cr_vgpr_pr_mr_sd_pd, "PR")          ~ "PR",
    # str_detect(day_30_response_s_cr_cr_vgpr_pr_mr_sd_pd, "UCR")	       ~ "UCR",
    str_detect(day_30_response_s_cr_cr_vgpr_pr_mr_sd_pd, "CR")          ~ "sCR or CR",
    str_detect(day_30_response_s_cr_cr_vgpr_pr_mr_sd_pd, "MR") |
      str_detect(day_30_response_s_cr_cr_vgpr_pr_mr_sd_pd, "SD")        ~ "SD",
    str_detect(day_30_response_s_cr_cr_vgpr_pr_mr_sd_pd, "PD")          ~ "PD",
    is.na(day_30_response_s_cr_cr_vgpr_pr_mr_sd_pd)                     ~ NA_character_,
    TRUE           ~ day_30_response_s_cr_cr_vgpr_pr_mr_sd_pd
  ), .after = day_30_response_s_cr_cr_vgpr_pr_mr_sd_pd) %>% 
  mutate(day30_response_v2 = case_when(
    str_detect(day30_response, "Died")                               ~ "PD",
    is.na(day30_response)                                            ~ NA_character_,
    TRUE           ~ day30_response
  ), .after = day30_response) %>% 
  mutate(ORR_30days = case_when(
    str_detect(day30_response_v2, "CR|PR")                           ~ "ORR",
    is.na(day30_response_v2)                                         ~ NA_character_,
    TRUE                                                             ~ "PD/SD/MR"
  ), .after = day30_response_v2) %>% 
  mutate(CRorbetter_30days = case_when(
    str_detect(day30_response_v2, "CR")	                             ~ "sCR or CR",
    is.na(day30_response_v2)                                         ~ NA_character_,
    TRUE                                                             ~ "Others"
  ), .after = ORR_30days) %>% 
  # Day 90 response ----
  mutate(mon3_response = case_when(
    day30_response_v2 == "PD"                                        ~ "PD",
    (os_event == 1 | pfs_event == 1) & 
      pfs_from_initiation_months < 3                                   ~ "Died or progressed before 3 months",
    str_detect(day_90_response_s_cr_cr_vgpr_pr_mr_sd_pd, "EVAL")      ~ NA_character_,
    str_detect(day_90_response_s_cr_cr_vgpr_pr_mr_sd_pd, "VGPR")      ~ "VGPR",
    str_detect(day_90_response_s_cr_cr_vgpr_pr_mr_sd_pd, "PR")        ~ "PR",
    str_detect(day_90_response_s_cr_cr_vgpr_pr_mr_sd_pd, "CR")	       ~ "sCR or CR",
    str_detect(day_90_response_s_cr_cr_vgpr_pr_mr_sd_pd, "MR") |
      str_detect(day_90_response_s_cr_cr_vgpr_pr_mr_sd_pd, "SD")      ~ "SD",
    str_detect(day_90_response_s_cr_cr_vgpr_pr_mr_sd_pd, "PD")       ~ "PD",
    is.na(day_90_response_s_cr_cr_vgpr_pr_mr_sd_pd)                     ~ NA_character_,
    TRUE           ~ day_90_response_s_cr_cr_vgpr_pr_mr_sd_pd
  ), .after = day_90_response_s_cr_cr_vgpr_pr_mr_sd_pd) %>% 
  mutate(mon3_response_v2 = case_when(
    str_detect(mon3_response, "Died")                              ~ "PD",
    is.na(mon3_response)                                           ~ NA_character_,
    TRUE           ~ mon3_response
  ), .after = mon3_response) %>% 
  mutate(ORR_mon3 = case_when(
    str_detect(mon3_response_v2, "CR|PR")                          ~ "ORR",
    is.na(mon3_response_v2)                                        ~ NA_character_,
    TRUE                                                            ~ "PD/SD/MR"
  ), .after = mon3_response_v2) %>% 
  mutate(CRorbetter_mon3 = case_when(
    mon3_response_v2 == "sCR or CR"                        	      ~ "sCR or CR",
    is.na(mon3_response_v2)                                        ~ NA_character_,
    TRUE                                                            ~ "Others"
  ), .after = ORR_mon3) %>%

    # Best response
  mutate(ORR_as_best_response = case_when(
    ORR_30days == "ORR" |
    ORR_mon3 == "ORR" #|
    # ORR_mon6 == "ORR" |
    # ORR_mon9 == "ORR" |
    # ORR_mon12 == "ORR"                                            
                                                                  ~ "ORR",
    is.na(ORR_30days) & 
      is.na(ORR_mon3) #& 
      # is.na(ORR_mon6) & 
      # is.na(ORR_mon9) &
      # is.na(ORR_mon12)                                            
                                                                  ~ NA_character_, 
     TRUE                                                         ~ "SD/PD"
  ), .after = CRorbetter_mon3) %>% 
  mutate(best_CRorbetter = case_when(
     CRorbetter_30days == "sCR or CR" | 
       CRorbetter_mon3 == "sCR or CR" #| 
       # CRorbetter_mon6 == "sCR or CR" | 
       # CRorbetter_mon9 == "sCR or CR" | 
       # CRorbetter_mon12 == "sCR or CR"                            
                                                                   ~ "sCR or CR",
     is.na(CRorbetter_30days) & 
       is.na(CRorbetter_mon3) #& 
       # is.na(CRorbetter_mon6) & 
       # is.na(CRorbetter_mon9) & 
       # is.na(CRorbetter_mon12) 
                                                                   ~ NA_character_, 
     TRUE                                                         ~ "< CR"
   ), .after = CRorbetter_mon3) %>%
  mutate(best_response = case_when(
    day30_response_v2 == "sCR or CR" |
      mon3_response_v2 == "sCR or CR" #|
      # mon6_response_v2 == "sCR or CR" |
      # mon9_response_v2 == "sCR or CR" |
      # mon12_response_v2 == "sCR or CR"                            
                                                                  ~ "sCR or CR",
    day30_response_v2 == "VGPR" |
      mon3_response_v2 == "VGPR" #|
      # mon6_response_v2 == "VGPR" |
      # mon9_response_v2 == "VGPR" |
      # mon12_response_v2 == "VGPR"                                 
                                                                  ~ "VGPR",
    day30_response_v2 == "PR" |
      mon3_response_v2 == "PR" #|
      # mon6_response_v2 == "PR" |
      # mon9_response_v2 == "PR" |
      # mon12_response_v2 == "PR"                                   
                                                                  ~ "PR",
    day30_response_v2 == "SD" |
      mon3_response_v2 == "SD" #|
      # mon6_response_v2 == "SD" |
      # mon9_response_v2 == "SD" |
      # mon12_response_v2 == "SD"                                   
                                                                  ~ "SD",
    day30_response_v2 == "PD" |
      mon3_response_v2 == "PD" #|
      # mon6_response_v2 == "PD" |
      # mon9_response_v2 == "PD" |
      # mon12_response_v2 == "PD"                                   
                                                                  ~ "PD",
    TRUE                                                          ~ NA_character_
  ), .after = best_CRorbetter) %>% 
  mutate(best_ORRcollapsed = case_when(
    str_detect(best_response, "CR|PR")                            ~ "PR or better", 
    str_detect(best_response, "SD|PD")                            ~ "< PR",
    TRUE                                                          ~ NA_character_
  ), .after = best_response) #%>% 
  # MRD----
  # mutate(across(c(contains("mrd_by_")),
  #           ~ case_when(
  #             . == "Positive"                                     ~ "Positive",
  #             . == "Negative"                                     ~ "Negative",
  #             TRUE                                                ~ NA_character_
  #           ))) %>%
  # mutate(MRD_day30 = coalesce(day_30_mrd_by_clono_seq_positive_vs_negative, 
  #                              day_30_mrd_by_flow_positive_vs_negative
  # )) %>%
  # mutate(MRD_mon3 = coalesce(x3_month_mrd_by_clono_seq_positive_vs_negative,
  #                              x3_month_mrd_by_flow_positive_vs_negative
  # )) %>% 
  # mutate(MRD_mon6 = coalesce(x6_month_mrd_by_clono_seq_positive_vs_negative,
  #                              x6_month_mrd_by_flow_positive_vs_negative
  # )) %>%
  # mutate(MRD_mon9 = coalesce(x9_month_mrd_by_clono_seq_positive_vs_negative,
  #                              x9_month_mrd_by_flow_positive_vs_negative
  # )) %>% 
  # mutate(MRD_mon12 = coalesce(x12_month_mrd_by_clono_seq_positive_vs_negative,
  #                              x12_month_mrd_by_flow_positive_vs_negative
  # )) %>% 
  # mutate(response_MRD_day30 = case_when(
  #   day30_response_v2 == "sCR or CR" & 
  #     MRD_day30 == "Negative"                                     ~ "sCR or CR, MRD-",
  #   day30_response_v2 == "sCR or CR" & 
  #     MRD_day30 == "Positive"                                     ~ "sCR or CR, MRD+",
  #   day30_response_v2 == "sCR or CR" & 
  #     is.na(MRD_day30)                                            ~ "sCR or CR, MRD unknown",
  #   TRUE ~ day30_response_v2
  # )) %>%
  # mutate(response_MRD_mon3 = case_when(
  #   mon3_response_v2 == "sCR or CR" & 
  #     MRD_mon3 == "Negative"                                      ~ "sCR or CR, MRD-",
  #   mon3_response_v2 == "sCR or CR" & 
  #     MRD_mon3 == "Positive"                                      ~ "sCR or CR, MRD+",
  #   mon3_response_v2 == "sCR or CR" & 
  #     is.na(MRD_mon3)                                             ~ "sCR or CR, MRD unknown",
  #   TRUE ~ mon3_response_v2
  # )) %>%
  # mutate(response_MRD_mon6 = case_when(
  #   mon6_response_v2 == "sCR or CR" & 
  #     MRD_mon6 == "Negative"                                      ~ "sCR or CR, MRD-",
  #   mon6_response_v2 == "sCR or CR" & 
  #     MRD_mon6 == "Positive"                                      ~ "sCR or CR, MRD+",
  #   mon6_response_v2 == "sCR or CR" & 
  #     is.na(MRD_mon6)                                             ~ "sCR or CR, MRD unknown",
  #   TRUE ~ mon6_response_v2
  # )) %>%
  # mutate(response_MRD_mon9 = case_when(
  #   mon9_response_v2 == "sCR or CR" & 
  #     MRD_mon9 == "Negative"                                      ~ "sCR or CR, MRD-",
  #   mon9_response_v2 == "sCR or CR" & 
  #     MRD_mon9 == "Positive"                                      ~ "sCR or CR, MRD+",
  #   mon9_response_v2 == "sCR or CR" & 
  #     is.na(MRD_mon9)                                             ~ "sCR or CR, MRD unknown",
  #   TRUE ~ mon9_response_v2
  # )) %>%
  # mutate(response_MRD_mon12 = case_when(
  #   mon12_response_v2 == "sCR or CR" & 
  #     MRD_mon12 == "Negative"                                      ~ "sCR or CR, MRD-",
  #   mon12_response_v2 == "sCR or CR" & 
  #     MRD_mon12 == "Positive"                                      ~ "sCR or CR, MRD+",
  #   mon12_response_v2 == "sCR or CR" & 
  #     is.na(MRD_mon12)                                             ~ "sCR or CR, MRD unknown",
  #   TRUE ~ mon12_response_v2
  # )) %>% 
  # mutate(best_response_MRD = case_when (
  #   response_MRD_day30 == "sCR or CR, MRD-" | 
  #     response_MRD_mon3 == "sCR or CR, MRD-" | 
  #     response_MRD_mon6 == "sCR or CR, MRD-" | 
  #     response_MRD_mon9 == "sCR or CR, MRD-" | 
  #     response_MRD_mon12 == "sCR or CR, MRD-"                      ~ "sCR or CR, MRD-",
  #    response_MRD_day30 == "sCR or CR, MRD+" | 
  #     response_MRD_mon3 == "sCR or CR, MRD+" | 
  #     response_MRD_mon6 == "sCR or CR, MRD+" | 
  #     response_MRD_mon9 == "sCR or CR, MRD+" | 
  #     response_MRD_mon12 == "sCR or CR, MRD+"                      ~ "sCR or CR, MRD+",
  #    response_MRD_day30 == "sCR or CR, MRD unknown" | 
  #     response_MRD_mon3 == "sCR or CR, MRD unknown" | 
  #     response_MRD_mon6 == "sCR or CR, MRD unknown" | 
  #     response_MRD_mon9 == "sCR or CR, MRD unknown" | 
  #     response_MRD_mon12 == "sCR or CR, MRD unknown"               ~ "sCR or CR, MRD unknown",
  #   TRUE ~ best_response
  # )) %>%
  # mutate(MRDneg_in_best_response = case_when(
  #   MRD_day30 ==  "Negative" |
  #     MRD_mon3 ==  "Negative" |
  #     MRD_mon6 ==  "Negative" |
  #     MRD_mon9 ==  "Negative" |
  #     MRD_mon12 ==  "Negative"                                    ~ "Negative",
  #   MRD_day30 ==  "Positive" |
  #   MRD_mon3 ==  "Positive" |
  #   MRD_mon6 ==  "Positive" |
  #   MRD_mon9 ==  "Positive" |
  #     MRD_mon12 ==  "Positive"                                    ~ "Positive",
#   TRUE                                                          ~ NA_character_
# )) %>% 
  # mutate(bridging_response = case_when(
  #   response_to_bridging_chemotherapy == "MR" |
  #     response_to_bridging_chemotherapy == "SD/MR" |
  #     response_to_bridging_chemotherapy == "SD" |
  #     response_to_bridging_chemotherapy == "PD" |
  #     response_to_bridging_chemotherapy == "No"               ~ "SD/PD",
  #   response_to_bridging_chemotherapy == "CR" |
  #     response_to_bridging_chemotherapy == "PR" |
  #     response_to_bridging_chemotherapy == "VGPR"             ~ "PR or better",
  #   response_to_bridging_chemotherapy == "Not Evaluable" |
  #     response_to_bridging_chemotherapy == "ND" |
  #     response_to_bridging_chemotherapy == "Unknown"          ~ NA_character_,
  #   TRUE                                                      ~ NA_character_ 
  # ))
```

```{r infection}
bsab_data <- cart4 %>%
  # Infection ----
  mutate(time_from_bsab_therapy_to_infection_days = 
           interval(
             start = date_of_bs_ab_step_up_d1, 
             end = date_of_infection_1st)/
           duration(n=1, units = "days"), 
         # no need of the following for this data but keep in case
         time_from_bsab_therapy_to_infection_days = case_when(
    time_from_bsab_therapy_to_infection_days >= 0     ~ time_from_bsab_therapy_to_infection_days,
    TRUE                                          ~ NA_real_
    ), .after = date_of_infection_1st) #%>% 
  # mutate(any_infection = case_when(
  #   time_from_bsab_therapy_to_infection_days >= 0     ~ infection_yes_1_no_0,
  #   TRUE                                          ~ "No"
  # )) %>%
  # mutate(any_sevinfection = case_when(
  #   time_from_bsab_therapy_to_infection_days >= 0     ~ severe_infection_need_for_iv_antibiotics_and_or_hospitalization_yes_1_no_0,
  #   TRUE                                          ~ "No"
  # )) %>% 
  # mutate(across(c("infection_type_details", "infection_type_bacterial_viral_fungal_parasitic",
  #                 "infection_type_cat"), ~ case_when(
  #   time_from_bsab_therapy_to_infection_days >= 0     ~ .,
  #   TRUE                                          ~ NA_character_
  # ))) %>% 
  # mutate(overall_hlh = case_when(
  #   hlh == "Yes" |
  #   hlh_hlh_like_toxicities == "Yes"              ~ "Yes",
  #   hlh == "No" &
  #   hlh_hlh_like_toxicities == "No"               ~ "No"
  # ))

# cart7 <- cart6 %>% 
#   mutate(infection = case_when(
#     notinfused=="YES" ~ NA_character_,
#     notinfused=="NO" & is.na(`Infection Type`) & is.na(`Infection (Yes, No)`) ~ NA_character_,
#     notinfused=="NO" & `Infection Type` == "No" ~ "No", 
#     notinfused=="NO" & `Infection Type` != "No" ~ "Yes",
#     notinfused=="NO" & is.na(`Infection Type`) ~ "No",
#     TRUE ~ NA_character_
#   )) %>%
#   mutate(infection_type = case_when(
#     infection=="No" | is.na(infection) ~ NA_character_,
#     infection=="Yes" & `Infection Details (specific infection, ie. E.coli bacteremia, etc.)` == "fungus was unidentified" ~ "Fungal",
#     infection== "Yes" & `Infection Details (specific infection, ie. E.coli bacteremia, etc.)` == "E. coli bacteremia and Salmonella GI infection treated with ceftriaxone x 7 days" ~ "Bacterial",
#     infection=="Yes" & `Infection Type`=="bacterial + viral" ~ "Bacterial + Viral",
#     infection=="Yes" & `Infection Type`=="viral" ~ "Viral",
#     infection=="Yes" & `Infection Type`=="All of the above" ~ "Bacterial, Viral, and Fungal",
#     infection=="Yes" & `Infection Type`=="Bacterial; Viral; Bacterial" ~ "Bacterial + Viral",
#     infection=="Yes" & `Infection Type`=="Viral; Bacterial; Bacterial" ~ "Bacterial + Viral",
#     TRUE ~ `Infection Type`
#   )) %>% 
#   mutate(infection_severity = case_when(
#      infection=="No" | is.na(infection) ~ NA_character_,
#      infection=="Yes" & is.na(`Infection Severity (Severe = hospitalization or requiring IV antibiotics)`) ~ NA_character_,
#      infection=="Yes" & `Infection Severity (Severe = hospitalization or requiring IV antibiotics)`=="severe" ~ "Severe",
#      infection=="Yes" & `Infection Severity (Severe = hospitalization or requiring IV antibiotics)`=="Yes" ~ "Severe",
#      infection=="Yes" & `Infection Severity (Severe = hospitalization or requiring IV antibiotics)`=="non-severe" ~ "Not severe",
#     infection=="Yes" & `Infection Severity (Severe = hospitalization or requiring IV antibiotics)`=="Non-Severe" ~ "Not severe",
#     TRUE ~ `Infection Severity (Severe = hospitalization or requiring IV antibiotics)`
#   )) %>%
#   mutate(infection_type = factor(infection_type, levels = c("Bacterial", "Fungal", "Viral", "Bacterial + Fungal", "Bacterial + Viral", "Bacterial, Viral, and Fungal")))
```

```{r Labeling}
## Labeling ----
var_label(bsab_data) <- list(age_at_bs_ab_initiation = "Patient age", agecat = "Patient age - Categories",

                             bs_ab_name_tec_1_tal_2_elra_3 = "Therapy name",
                             bs_ab_served_only_as_pre_car_t_bridging_0_no_1_yes = "Served as pre-CART bridging",
                             pfs_with_1st_line_mm_diagnosis_18_months_0_no_1_yes = "PFS 1st line",

                         male_sex = "Male Sex", sex_0_male_1_female = "Sex",
                         race_white_0_black_1_aapi_2_american_indian_3_other_4_unk_5 = "Race",
                         hispanic_1_if_hispanic_latin_x_or_spanish_origin_0_if_no_unknown = "Ethnicity",
                         # race_eth = "Race and ethnicity",
                         # extramedullary_disease_yes_no = "Extramedullary disease",
                         # circulating_plasma_cell = "UPDATED Circulating Plasma Cells",
                         high_marrow_burden_50_percent_p_cs_if_checked_within_30_days_before_bs_ab_initiation_n_0_y_1 = 
                           "High marrow burden (≥50%)",
                         ecog_ps_at_bs_ab_initiation_0_5 = "ECOG PS at therapy initiation",
                         # `ECOG > or = 2? (Yes, No)` ,
                         # r_iss_at_car_t_initiation = "R-ISS at cart initiation",
                         oligo_or_non_secretory_at_bs_ab_initiation_0_no_1_yes = "Oligo or non secretory at initiation",
                         myeloma_involved_heavy_chain_ig_g_ig_a_ig_m_ig_d_ig_e_or_blank = "Myeloma involved heavy chain",
                         myeloma_involved_light_chain_kappa_lambda_biclonal_blank = "Myeloma involved light chain",
                         prior_cart_0_no_1_yes = "Received prior CART",
                         amyloidosis_if_yes_fill_out_amyloid_sheet_as_well_please_once_youre_done_with_this_one_0_no_1_yes =
                           "Amyloidosis",
                         active_cns_mm_if_yes_fill_out_cns_sheet_as_well_once_done_with_this_one_0_no_1_yes = 
                           "Active CNS",
                         pcl_5_percent_cp_cs_at_any_point_from_dx_through_last_pre_bs_ab_lo_t_0_no_1_secondary_pcl_not_at_dx_2_primary_pcl_at_dx = 
                           "PCL (?5% CPCs) at any point from Dx through last pre-BsAb LoT?",
                         active_pcl_5_percent_cp_cs_within_30_days_prior_or_after_bs_ab_initiation_0_no_1_yes = 
                           "Active PCL (?5% CPCs) within 30 days prior or after BsAb initiation?",
                         percent_cp_cs_if_pb_checked_within_30_days_before_bs_ab_initiation = 
                           "% CPCs if PB checked within 30 days before BsAb initiation",
                         prophylactic_tocilizumab_n_0_y_1 = "Prophylactic Tocilizumab",

                         cytogenetics = "High risk cytogenetics (not include gain 1q)", 
                         deletion_17p_prior_to_bs_ab_0_no_1_yes = "Deletion 17p at initiation", 
                         t_4_14_prior_to_bs_ab_0_no_1_yes = "t(4:14) at initiation", 
                         t_14_16_prior_to_bs_ab_0_no_1_yes = "t(4:16) at initiation", 
                         gain_amp1q21_prior_to_bs_ab_0_no_1_yes = "Gain/amp 1q21",
                         amp_1_q_only_0_no_1_yes = "Amp 1q only",
                         deletion_1p_0_no_1_yes = "Deletion 1p",
                         number_of_prior_lines_of_therapy_before_bs_ab = "Number of prior lines of therapy",
                         number_of_prior_lines_of_therapy_before_bs_ab_cat = "Number of prior lines of therapy (cat)",
                         # bridging_therapy_yes_no = "Bridging Therapy",
                         prior_auto_sct_no_0_yes_1 = "Prior autoSCT",
                         # prior_allo_sct_within_6_months_before_apheresis = "Prior alloSCT",
                         prior_bcma_directed_therapy_no_0_y_1 = "Prior BCMA-directed therapy",
                         need_for_gcsf = "Granulocyte colony stimulating factor (G-CSF)",
                         ivig_use_yes_1_no_0 = "Intravenous immunoglobulin (IVIG)",
                         # cd34_stem_cell_boost_yes_1_no_0 = "CD34 stem cell boost",
                         # day_of_neutrophil_recovery = "Day of neutrophil recovery",
                         # transfusion_postCART = "Transfusion post-CAR T",
                         need_for_tpo_agonist = "Thrombopoietin (TPO) agonist",
                         # bmicat = "BMI at apheresis - Normal, Overweight, Obese", 
                         steroid_use_as_tx_n_0_y_1 = "Steroid Use", tocilizumab_as_tx_n_0_y_1 = "Tocilizumab Use", 
                         anakinra_n_0_y_1 = "Anakinra Use",
                         proteasome = "Refractory to Proteasome Inhibitors", 
                         lenalidomide_status_exposed_0_vs_refractory_1_vs_naive_2 = "Refractory to Lenalidomide",
                         bortezomib_status_exposed_0_vs_refractory_1_vs_naive_2 = "Refractory to Bortezomib", 
                         carfilzomib_status_exposed_0_vs_refractory_1_vs_naive_2 = "Refractory to Carfilzomib", 
                         pomalidomide_status_exposed_0_vs_refractory_1_vs_naive_2 = "Refractory to Pomalidomide",
                         dara = "Refractory to Daratumumab or Isatuximab",
                         immuno = "Refractory to Immunomodulators", doubleref = "Double refractory", 
                         triple_class_refractory_i_mi_d_pi_anti_cd38_no_0_y_1= "Triple refractory", 
                         penta_refractory_2_p_is_2_i_mi_ds_anti_cd38_no_0_y_1= "Penta refractory",

                         baseline_b2m_prior_to_bs_ab_initiation = "Baseline B2M prior to therapy",
                         baseline_ldh_prior_to_bs_ab_initiation_u_l = "Baseline LDH prior to therapy (u/l)",
                         baseline_albumin_prior_to_bs_ab_initiation_g_d_l = "Baseline albumin prior to therapy (g/dl)",
                         baseline_ferritin_prior_to_bs_ab_initiation_ng_m_l = "Baseline ferritin prior to therapy (ng/ml)",
                         max_ferritin = "Max ferritin",
                         baseline_crp_prior_to_bs_ab_initiation_mg_l = "Baseline CRP prior to therapy (mg/l)",
                         max_crp = "Max CRP",
                         baseline_ig_g_prior_to_bs_ab_initiation_mg_d_l = "Baseline Igg prior to therapy (mg/dl)",
                         baseline_anc_k_u_l_ie_100_200_etc = "Baseline ANC (k/ul) ie_100_200",
                         baseline_hgb = "Baseline HGB",
                         baseline_pl_ts = "baseline PLTs",
                         baseline_alc_k_u_l_ie_100_200_etc = "Baseline ALC (k/ul) ie_100_200",
                         alc_day_8 = "ALC day8",
                         alc_day_15 = "ALC day15",
                         alc_day_22 = "ALC day22",
                         alc_c2d1 = "ALC C2D1",
                         c2d1_ig_g_level_if_checked_mg_d_l = "C2D1 Igg (mg/dl)",
                         anc_nadir_value_during_first_6_cycles_k_u_l_ie_100_200_etc = "ANC during first 6 cycles (nadir)",
                         hgb_nadir_value_during_first_6_cycles = "HGB during first 6 cycles (nadir)",
                         plt_nadir_value_during_first_6_cycles = "PLT during first 6 cycles (nadir)",
                         ig_g_nadir_during_first_6_cycles_mg_d_l = "Igg during first 6 cycles (nadir)",
                         cr_cl_prior_to_step_up_dose_number_1_m_l_min = "CrCl prior to step up dose 1 (ml/min)",
                         cr_cl_at_d30 = "CrCl day 30",
                         cr_cl_at_d90 = "CrCl day 90",
                         hemodialysis_pre_bs_ab_y_1_n_0 = "Hemodialysis pre-BsAb",
                         hemodialysis_post_bs_ab_y_1_n_0 = "Hemodialysis post-BsAb",
                         number_of_prior_bcma_directed_agents = "Number of prior BCMA",
                         prior_adc_yes_no = "Prior ADC",
                         prior_adc_name = "Prior ADC name",
                         # prior_cart_yes_no = "Prior CART",
                         prior_cart_name = "Prior CART name",
                         prior_bs_ab_yes_no = "Prior BsAb",
                         prior_bs_ab_name = "Prior BsAb name",
                         prior_gprc5d_yes_no = "Prior gprc5d",
                         talquetamab_toxicities_oral_toxicities_e_g_dysgeusia_if_so_describe =
                           "Talquetamab toxicities",
                         just_prior_to_bs_ab_initiation_is_there_true_emd_if_none_move_to_column_gk_0_none_1_paraskeletal_only_2_true_emd_paraskeletal = 
                           "EMD prior to BsAb initiation",
                         total_number_true_emd_plasmacytomas_if_1_2_sites_write_1_or_2_if_2_plasmacytomas_just_write_3 =
                           "Total number true EMD",
                         total_number_paraskelatal_plasmacytomas = "Total number paraskelatal plasmacytomas",
                         true_emd_relapse_yes_no = "True EMD relapse",
                         best_documented_emd_response = "Best documented EMD response",
                         xrt_2_weeks_before_and_after_3_months_of_bs_ab_starting_yes_no_response =
                           "XRT 2 weeks before and after  3 months of BsAb",
                         # bmicat2 = "BMI at apheresis - Normal vs. Overweight/Obese", 
                         max_crs_grade_0_5 = "CRS Grade (0-5)",
                         CRS_any = "Any CRS", CRS_gradecat = "CRS Grade", 
                         gr2_CRS = "Grade2 CRS",
                         gr3_CRS = "Grade3 CRS",
                         recurrent_crs_n_0_y_1 = "Recurrent CRS",
                         max_icans_grade = "ICANS Grade (0-5)",
                         ICANS_any = "Any ICANS", ICANS_gradecat = "ICANS Grade", 
                         gr2_ICANS = "Grade 2 ICANS",
                         day_of_onset_crs_relative_to_initiation = "Day of onset of CRS relative to initiation",
                         day_of_max_crs_relative_to_initiation = "Day of Max CRS (relative to initiation)",
                         duration_of_crs_days = "Duration of CRS (days)", 
                         day_of_onset_icans_relative_to_initiation = "Day of onset of ICANS relative to initiation",
                         day_of_max_icans_grade_relative_to_initiation = "Day of Max ICANS (relative to initiation)",
                         duration_of_icans_days = "Duration of ICANS (days)", 
                         # delayed_neuro = "Delayed Neurotoxicity",
                         # parkinsonian_neurotoxicity_yes_no = "Parkinsonian Neurotoxicity",
                         # other_delayed_non_parkinsonian_neurologic_toxicity_yes_no = 
                         #   "Other/Delayed Non-Parkinsonian Neurologic Toxicity",
                         # delayed_neuro_type = "Delayed Neurotoxicity Type",
                         # hlh = "HLH", hlh_hlh_like_toxicities = "HLH/HLH-Like Toxicities?",
                         # overall_hlh = "Combined HLHs",
                         infection_n_0_y_1 = "Infection",
                         time_from_bsab_therapy_to_infection_days = "Time from BsAb to infection (days)",
                         number_of_infections = "Total number of infections",
                         number_infection_1_bacterial_viral_fungal = "Infection type (1st)",
                         infection_severity_1st_severe_hospitalization_or_requiring_iv_antibiotics = "Infection severity (1st)",
                         number_infection_2_bacterial_viral_fungal = "Infection type (2nd)",
                         infection_severity_2nd_severe_hospitalization_or_requiring_iv_antibiotics = "Infection severity (2nd)",
                         number_infection_3_bacterial_viral_fungal = "Infection type (3rd)",
                         infection_severity_3rd_severe_hospitalization_or_requiring_iv_antibiotics = "Infection severity (3rd)",
                         number_infection_4_bacterial_viral_fungal = "Infection type (4th)",
                         infection_severity_4th_severe_hospitalization_or_requiring_iv_antibiotics = "Infection severity (4th)",

                         treatment_delays_during_first_cycle_infection_1_cytoepnias_2_crs_3_icans_4_other_5_explain_leave_blank_if_no_delays =
                           "Cause for treatment delays during first cycle",
                         icu_admission_during_first_cycle_n_0_y_1 = "ICU admission during first cycle",
                         days_of_icu_admission = "Days of ICU admission",
                         in_out_patient = "In/Out patient",
                         if_inpatient_length_of_hospital_stay_total_days_including_all_admissions_0_if_done_outpt_with_no_admissions = 
                           "Length of hospital stay total (days)",
                         # tbv = "TBV", whole_blood_processed_m_l = "Whole blood processed (mL)",
                         # final_collection_volume_m_l = "Final collection volume (mL)", 
                         # tbv_processed = "TBV processed", 
                         # access_type = "Access type", total_bags = "Total bags",
                         # run_time_min = "Run time (min)",
                         # total_dose_volume_m_l = "Total dose volume (mL)", 
                         # cell_dose_million_cells = "Total cell dose (million cells)",
                         day_30_response_s_cr_cr_vgpr_pr_mr_sd_pd = "Raw day 30 Response",
                         day30_response = "Day 30 Response - Evaluable patients",
                         day30_response_v2 = "Day 30 Response - All patients",
                         ORR_30days = "Day 30 ORR",
                         CRorbetter_30days = "Day 30 CR or better - Evaluable patients",

                         day_90_response_s_cr_cr_vgpr_pr_mr_sd_pd = "Raw day 90 Response",
                         mon3_response = "Day 90 Response - Evaluable patients",
                         mon3_response_v2 = "Day 90 Response - All patients",
                         ORR_mon3 = "Day 90 ORR",
                         CRorbetter_mon3 = "Day 90 CR or better - Evaluable patients",

                         best_CRorbetter = "Best CR or better - Evaluable patients",
                         best_response = "Best Response - Evaluable patients",
                         best_ORRcollapsed = "Best ORR at ≤ 3 months",
                         ORR_as_best_response = "ORR as best response",
                         best_response_s_cr_cr_vgpr_pr_mr_sd_pd = "Raw best response",
                         best_response_mrd_flow_positive_vs_negative = "Raw best MRD response (flow)",
                         best_response_mrd_clonoseq_positive_vs_negative = "Raw best MRD response (clonoseq)",
                         progression_yes_no = "Progression",
                         pfs_event = "PFS events",
                         pfs_from_initiation_months = "PFS time (months)",
                         death_yes_no = "Death",
                         os_event = "OS events",
                         os_from_initiation_months = "Follow-up time from initiation (months)",
                         cause_of_death_pd_1_infection_while_of_bs_ab_2_crs_3_icans_4_other_5_explain =
                           "Cause of death",

                         discontinuation_yes_no = "Discontinuation",

                         eligible_for_corresponding_bs_ab_trial_on_first_dose_use_trial_elig_sheet_to_help_you_no_0_y_1 =
                           "Eligible for corresponding BsAb trial on first dose"#,
                         # best_response = "Best response in first 90 days"
)

```
<br>

# Tables
## Table 1. Patient characteristics
```{r Table 1}
bsab_data %>% 
  select(age_at_bs_ab_initiation, agecat, 
         sex_0_male_1_female, #male_sex, 
         race_white_0_black_1_aapi_2_american_indian_3_other_4_unk_5, 
         hispanic_1_if_hispanic_latin_x_or_spanish_origin_0_if_no_unknown, 
         ecog_ps_at_bs_ab_initiation_0_5,
         in_out_patient) %>%
  tbl_summary(by = in_out_patient,
    # type = list(
    #   c(agecat, sex, 
    #     extramedullary_disease_yes_no, high_marrow_burden_50_percent,
    #     cytogenetics, cytogenetics_incl_1q,
    #     deletion_17p_prior_to_bs_ab_0_no_1_yes,
    #     t_4_14_prior_to_bs_ab_0_no_1_yes,
    #     t_14_16_prior_to_bs_ab_0_no_1_yes, gain_amp1q21,
    #     bridging_therapy_yes_no, prior_auto_sct_no_0_yes_1,
    #     prior_bcma_directed_therapy_no_0_y_1,
    #     # `Received alkylating therapy with bridging (Yes=1, No=0)`,
    #     immuno, proteasome, dara, doubleref, triple_class_refractory_i_mi_d_pi_anti_cd38_no_0_y_1, penta_refractory_2_p_is_2_i_mi_ds_anti_cd38_no_0_y_1,
    #     did_patient_meet_eligibility_criteria_for_cartitude_1_on_the_day_of_apheresis_yes_no,
    #     ht_score, early_icaht_cat, any_early_icaht,
    #      gr3_early_icaht, 
    #      late_icaht_cat, any_late_icaht,
    #      gr3_late_icaht)
    #   ~"categorical"),

    statistic = list(all_continuous() ~ "{median} ({min}, {max})"), 
    digits = all_continuous() ~ 1
  ) %>% 
  bold_labels() %>% add_stat_label() %>% add_p() %>% bold_p(t = 0.05) %>% add_overall() %>% 
  modify_footnote(list(update = everything() ~ NA))
```

## Table 2. Disease characteristics
```{r Table 2}
bsab_data %>% 
  select(oligo_or_non_secretory_at_bs_ab_initiation_0_no_1_yes, myeloma_involved_heavy_chain_ig_g_ig_a_ig_m_ig_d_ig_e_or_blank, 
         myeloma_involved_light_chain_kappa_lambda_biclonal_blank, 
         high_marrow_burden_50_percent_p_cs_if_checked_within_30_days_before_bs_ab_initiation_n_0_y_1,
         cytogenetics, deletion_17p_prior_to_bs_ab_0_no_1_yes, 
         t_4_14_prior_to_bs_ab_0_no_1_yes, t_14_16_prior_to_bs_ab_0_no_1_yes, gain_amp1q21_prior_to_bs_ab_0_no_1_yes, 
         amp_1_q_only_0_no_1_yes, deletion_1p_0_no_1_yes,
         amyloidosis_if_yes_fill_out_amyloid_sheet_as_well_please_once_youre_done_with_this_one_0_no_1_yes, 
         active_cns_mm_if_yes_fill_out_cns_sheet_as_well_once_done_with_this_one_0_no_1_yes, 
         pcl_5_percent_cp_cs_at_any_point_from_dx_through_last_pre_bs_ab_lo_t_0_no_1_secondary_pcl_not_at_dx_2_primary_pcl_at_dx, 
         active_pcl_5_percent_cp_cs_within_30_days_prior_or_after_bs_ab_initiation_0_no_1_yes, 
         eligible_for_corresponding_bs_ab_trial_on_first_dose_use_trial_elig_sheet_to_help_you_no_0_y_1,
         in_out_patient) %>%
  tbl_summary(by = in_out_patient,
    statistic = list(all_continuous() ~ "{median} ({min}, {max})"), 
    digits = all_continuous() ~ 1
  ) %>% 
  bold_labels() %>% add_stat_label() %>% add_p() %>% bold_p(t = 0.05) %>% add_overall() %>% 
  modify_footnote(list(update = everything() ~ NA))
```

## Table 3. Hospital
```{r table 3}
bsab_data %>% 
  select(in_out_patient,
         if_inpatient_length_of_hospital_stay_total_days_including_all_admissions_0_if_done_outpt_with_no_admissions, 
         icu_admission_during_first_cycle_n_0_y_1, days_of_icu_admission
         ) %>%
  tbl_summary(type = list(days_of_icu_admission ~ "continuous"),
              statistic = list(all_continuous() ~ "{median} ({min}, {max})"), 
              digits = all_continuous() ~ 1
  ) %>% 
  bold_labels() %>% add_stat_label() %>%# add_p() %>% bold_p(t = 0.05) %>% add_overall() %>% 
  modify_footnote(list(update = everything() ~ NA))
```

## Table 4. Baseline
```{r table 4}
bsab_data %>% 
  select(percent_cp_cs_if_pb_checked_within_30_days_before_bs_ab_initiation,
         baseline_b2m_prior_to_bs_ab_initiation, 
         baseline_ldh_prior_to_bs_ab_initiation_u_l, baseline_albumin_prior_to_bs_ab_initiation_g_d_l, 
         baseline_ferritin_prior_to_bs_ab_initiation_ng_m_l, max_ferritin, 
         baseline_crp_prior_to_bs_ab_initiation_mg_l, max_crp, 
         baseline_ig_g_prior_to_bs_ab_initiation_mg_d_l,
         ig_g_nadir_during_first_6_cycles_mg_d_l,
         c2d1_ig_g_level_if_checked_mg_d_l, 
         baseline_anc_k_u_l_ie_100_200_etc, anc_nadir_value_during_first_6_cycles_k_u_l_ie_100_200_etc, 
         baseline_hgb, hgb_nadir_value_during_first_6_cycles, 
         baseline_pl_ts, plt_nadir_value_during_first_6_cycles, 
         baseline_alc_k_u_l_ie_100_200_etc, alc_day_8, alc_day_15, alc_day_22, alc_c2d1, 
         in_out_patient) %>%
  tbl_summary(by = in_out_patient,
    statistic = list(all_continuous() ~ "{median} ({min}, {max})"), 
    digits = all_continuous() ~ 1
  ) %>% 
  bold_labels() %>% add_stat_label() %>% add_p() %>% bold_p(t = 0.05) %>% add_overall() %>% 
  modify_footnote(list(update = everything() ~ NA))
```

## Table 5. Therapies
```{r table 5}
bsab_data %>% 
  select(bs_ab_name_tec_1_tal_2_elra_3, bs_ab_served_only_as_pre_car_t_bridging_0_no_1_yes,
         prior_cart_0_no_1_yes, #prior_cart_yes_no, 
         prior_cart_name, prior_bs_ab_yes_no, prior_bs_ab_name, 
         prior_adc_yes_no, prior_adc_name, 
         prior_bcma_directed_therapy_yes_no, prior_bcma_directed_therapy_no_0_y_1, 
         number_of_prior_bcma_directed_agents, 
         type_of_most_recent_bcma_agent_adc_1_cart_2_bs_ab_3, most_recent_bcma_agent_name, 
         response_to_most_recent_bmca_agent_cr_1_vgpr_2_pr_3_sd_4_pd_5_ne_6, prior_gprc5d_yes_no,
         prophylactic_tocilizumab_n_0_y_1,
         tocilizumab_as_tx_n_0_y_1, steroid_use_as_tx_n_0_y_1, anakinra_n_0_y_1,
         ivig_use_yes_1_no_0,

         need_for_gcsf, need_for_tpo_agonist,

         treatment_delays_during_first_cycle_infection_1_cytoepnias_2_crs_3_icans_4_other_5_explain_leave_blank_if_no_delays,
         number_of_prior_lines_of_therapy_before_bs_ab_cat,
         number_of_prior_lines_of_therapy_before_bs_ab, prior_auto_sct_no_0_yes_1,

         bortezomib_status_exposed_0_vs_refractory_1_vs_naive_2, carfilzomib_status_exposed_0_vs_refractory_1_vs_naive_2, 
         proteasome, 
         lenalidomide_status_exposed_0_vs_refractory_1_vs_naive_2, pomalidomide_status_exposed_0_vs_refractory_1_vs_naive_2, 
         immuno, daratumumab_or_isatuximab_status_exposed_0_vs_refractory_1_vs_naive_2, dara, 
         doubleref, 
         triple_class_refractory_i_mi_d_pi_anti_cd38_no_0_y_1, 
         penta_refractory_2_p_is_2_i_mi_ds_anti_cd38_no_0_y_1,

         talquetamab_toxicities_oral_toxicities_e_g_dysgeusia_if_so_describe, 
         talquetamab_only_target_dose_for_c2d1_0_4_q1w_0_8_q2w_or_other,
         x147_reason_for_stopping_talquetamab_if_not_for_pd_be_specific, 
         x148_tal_specifically_designed_as_pre_car_t_holding_or_bridging_y_n,
         in_out_patient) %>%
  tbl_summary(by = in_out_patient,
    statistic = list(all_continuous() ~ "{median} ({min}, {max})"), 
    digits = all_continuous() ~ 1
  ) %>% 
  bold_labels() %>% add_stat_label() %>% add_p() %>% bold_p(t = 0.05) %>% add_overall() %>% 
  modify_footnote(list(update = everything() ~ NA))
```

## Table 6. Toxicities
```{r Table 6}
bsab_data %>% 
  select(max_crs_grade_0_5, CRS_any, #crs_any_coded_0_1, 
         CRS_gradecat, 
         gr2_CRS, gr3_CRS, duration_of_crs_days, 
         recurrent_crs_n_0_y_1, 
         day_of_onset_crs_relative_to_initiation, 
         day_of_max_crs_relative_to_initiation, 
         max_icans_grade, ICANS_any, ICANS_any2, gr2_ICANS, ICANS_gradecat, 
         gr3_ICANS, duration_of_icans_days,
         day_of_onset_icans_relative_to_initiation, day_of_max_icans_grade_relative_to_initiation,
         cr_cl_prior_to_step_up_dose_number_1_m_l_min, cr_cl_at_d30, cr_cl_at_d90, 
         hemodialysis_pre_bs_ab_y_1_n_0, 
         hemodialysis_post_bs_ab_y_1_n_0,
         in_out_patient) %>%
  tbl_summary(by = in_out_patient,
              type = list(
                c(duration_of_crs_days, duration_of_icans_days,
                  day_of_onset_icans_relative_to_initiation, day_of_max_icans_grade_relative_to_initiation) ~ "continuous"),
    statistic = list(all_continuous() ~ "{median} ({min}, {max})"), 
    digits = all_continuous() ~ 1
  ) %>% 
  bold_labels() %>% add_stat_label() %>% add_p() %>% bold_p(t = 0.05) %>% add_overall() %>% 
  modify_footnote(list(update = everything() ~ NA))
```

## Table 7. Responses
```{r Table 7}
bsab_data %>% 
  select(day_30_response_s_cr_cr_vgpr_pr_mr_sd_pd, 
         day30_response, day30_response_v2, ORR_30days, CRorbetter_30days, day_90_response_s_cr_cr_vgpr_pr_mr_sd_pd, 
         mon3_response, mon3_response_v2, ORR_mon3, CRorbetter_mon3, best_CRorbetter, best_response, best_ORRcollapsed, 
         ORR_as_best_response, best_response_s_cr_cr_vgpr_pr_mr_sd_pd, best_response_mrd_flow_positive_vs_negative, 
         best_response_mrd_clonoseq_positive_vs_negative,
         just_prior_to_bs_ab_initiation_is_there_true_emd_if_none_move_to_column_gk_0_none_1_paraskeletal_only_2_true_emd_paraskeletal,
         total_number_true_emd_plasmacytomas_if_1_2_sites_write_1_or_2_if_2_plasmacytomas_just_write_3, 
         locations_of_true_emd_be_specific, total_number_paraskelatal_plasmacytomas, 
         xrt_2_weeks_before_and_after_3_months_of_bs_ab_starting_yes_no_response, best_documented_emd_response, 
         true_emd_relapse_yes_no, if_yes_what_location_0_same_site_1_new_site_2_both, concurrent_biochemical_relpase_yes_no, 
         concurrent_pmd_relapse_yes_no,
         in_out_patient) %>%
  tbl_summary(by = in_out_patient,
    statistic = list(all_continuous() ~ "{median} ({min}, {max})"), 
    digits = all_continuous() ~ 1
  ) %>% 
  bold_labels() %>% add_stat_label() %>% add_p() %>% bold_p(t = 0.05) %>% add_overall() %>% 
  modify_footnote(list(update = everything() ~ NA))
```

## Table 8. Clinical outcomes
```{r Table 8}
bsab_data %>% 
  select(progression_yes_no, pfs_from_initiation_months, #pfs_event, 
         death_yes_no, #os_event, 
         os_from_initiation_months, 
         cause_of_death_pd_1_infection_while_of_bs_ab_2_crs_3_icans_4_other_5_explain,
         in_out_patient) %>%
  tbl_summary(by = in_out_patient,
    statistic = list(all_continuous() ~ "{median} ({min}, {max})"), 
    digits = all_continuous() ~ 1
  ) %>% 
  bold_labels() %>% add_stat_label() %>% add_p() %>% bold_p(t = 0.05) %>% add_overall() %>% 
  modify_footnote(list(update = everything() ~ NA))
```

## Table 9. Infections
```{r table 9}
bsab_data %>% 
  select(infection_n_0_y_1, number_of_infections, number_infection_1_bacterial_viral_fungal, 
         infection_severity_1st_severe_hospitalization_or_requiring_iv_antibiotics, time_from_bsab_therapy_to_infection_days, 
         number_infection_2_bacterial_viral_fungal, infection_severity_2nd_severe_hospitalization_or_requiring_iv_antibiotics, 
         number_infection_3_bacterial_viral_fungal, infection_severity_3rd_severe_hospitalization_or_requiring_iv_antibiotics, 
         number_infection_4_bacterial_viral_fungal, infection_severity_4th_severe_hospitalization_or_requiring_iv_antibiotics,
         in_out_patient) %>%
  tbl_summary(by = in_out_patient,
              type = list(
                c(number_of_infections) ~ "continuous"),
    statistic = list(all_continuous() ~ "{median} ({min}, {max})"), 
    digits = all_continuous() ~ 1
  ) %>% 
  bold_labels() %>% add_stat_label() %>% add_p() %>% bold_p(t = 0.05) %>% add_overall() %>% 
  modify_footnote(list(update = everything() ~ NA))
```

# Survival
## OS
```{r Survival analysis - KM by in_out_patient}
ggsurvplot(survfit(Surv(os_from_initiation_months, os_event) ~ in_out_patient, 
                   data = bsab_data),
           # title = "PFS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(15, "bold", "black"),
           font.y = c(15, "bold", "black"),
           font.legend = c(14, "black"),
           font.tickslab = c(12, "bold", "black"),
           size = 1.5,

           xlab = "Time (months)",
           ylab = "OS (probability)",
           legend = "top",
           legend.title = "",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE,
           risk.table = TRUE, risk.table.y.text = TRUE,
           tables.theme = theme_survminer(#base_size = 5,
                                          font.main = c(14, "bold", "black"),
                                          # font.x = c(10, "bold", "black"),
                                          # font.y = c(16, "bold", "transparent"),
                                          # font.tickslab = c(19, "bold", "black")
                                          )
) %++% guides(colour = guide_legend(nrow = 1))

coxph(Surv(time = bsab_data$os_from_initiation_months, 
                   event = bsab_data$os_event) ~ in_out_patient, 
              data = bsab_data)  %>% 
  tbl_regression(exponentiate = TRUE) %>% 
  bold_p(t = .05) %>% 
  add_nevent(location = "level") %>% add_n(location = "level")
```

## PFS
```{r PFS analysis - KM by in_out_patient}
ggsurvplot(survfit(Surv(pfs_from_initiation_months, pfs_event) ~ in_out_patient, 
                   data = bsab_data),
           # title = "PFS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(15, "bold", "black"),
           font.y = c(15, "bold", "black"),
           font.legend = c(14, "black"),
           font.tickslab = c(12, "bold", "black"),
           size = 1.5,

           xlab = "Time (months)",
           ylab = "PSS (probability)",
           legend = "top",
           legend.title = "",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE,
           risk.table = TRUE, risk.table.y.text = TRUE,
           tables.theme = theme_survminer(#base_size = 5,
                                          font.main = c(14, "bold", "black"),
                                          # font.x = c(10, "bold", "black"),
                                          # font.y = c(16, "bold", "transparent"),
                                          # font.tickslab = c(19, "bold", "black")
                                          )
) %++% guides(colour = guide_legend(nrow = 1))

coxph(Surv(time = bsab_data$pfs_from_initiation_months, 
                   event = bsab_data$pfs_event) ~ in_out_patient, 
              data = bsab_data)  %>% 
  tbl_regression(exponentiate = TRUE) %>% 
  bold_p(t = .05) %>% 
  add_nevent(location = "level") %>% add_n(location = "level")
```